### 2014-05-06 Paper1/paper1.tex

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Title}
}
% Insert Author names, affiliations and corresponding author email.
\\
Justin Horowitz$^{1}$, 
James Patton$^{12\ast}$
\\
\bf{1} Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States
\\
\bf{2} SMPP, RIC, Chicago, Illinois, United States
\\
$\ast$ E-mail: Corresponding pattonj@uic.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
\textbf{Needs work.}\\
We derive and examine the parameter sensitivity of a novel modeling approach for extracting movement intent from a force-disturbed planar arm trajectory. While popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions. After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching. We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching. Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints. The robustness of this model holds promise for real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals. In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person’s intentions because action is decoupled from intention. Lacking a means to discern that intention and changes in intention leads to challenges in human machine interactions. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

It is important to distinguish our use of the term intent from other groups that consider \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}. In typical motor control studies, subjects are motivated to complete the experiment in a timely fashion and are usually explicitly provided targets to be reached. Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of actions (walking/standing/flexion/extension/etc). While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action when the actual movement is disturbed and hence no longer matches the intent.

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle’s equilibrium point. Supporters of the so-called lambda model\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement. Upon inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. We propose a revised method---intent extraction---that deduces intended trajectory, even in the face of disturbances. This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory. In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling. The underlying framework should be extensible to whole-body movement or even more abstract, psychological processes.

Collectively, the general problem of being able to best extract someone’s intent is critical to many fields, and hinges on our understanding of the underlying mechanisms through which control is accomplished. We have excellent models of the mechanical properties of arm and muscle because the physics that describe their gross behavior is so well-understood and the need to determine particular physical parameter values has long been clear. This wealth of previous work provides strong footing for our technique, which in turn should yield valuable information not previously available by other means. In fields where the process and actuators are not yet well-understood, this approach may help to structure the search for them.

Understanding the intent has many benefits, and one may be a better way to understand the composition of discrete movements into their subunits. Recent research has strived to understand the composition of movements intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements}, distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable. Flash and Henis \cite{flash1991arm} demonstrated that despite this apparent blending, cued corrections can only be produced at predictable places within the reach. They also showed that these modifications appear to take place by superimposing additional straight-line subunits onto the original movement. Novak et al. \cite{novak2002use} detected corrective subunits following disturbance during knob turning. These studies show planning and correction in discrete subunits, but did not analyze the actual intent and hence cannot be compared to a known, correct trajectory except in the trivial case that movement is undisturbed. A disturbed trajectory’s intention is hidden without proper inversion of the model plan for movement. Here we develop a method to extract intent and gain a clearer understanding of subunits from disturbed motions.

In this paper, we employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for a providing a better understanding of what a person means to do when they move and are disturbed.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
The sections below describe the theoretical method, and then present our evaluations. First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process. Next we conducted an experiment on humans and the process was then evaluated on human subject data.

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and stabilizes it with linear feedback of position and velocity error. More complex models can be used without any loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripal effects, and $E$ is any externally-applied torques that the system might interact with. Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q-q_d)+K_d(\dot{q}-\dot{q}_d)}_{\text{Impendance, Feedback Controller}}
\end{equation}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{paper1.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2014-05-12 Paper1/paper1.tex
Updated.

@@
 
-% Results and Discussion can be combined.
-\section*{Results}
-
-\subsection*{Subsection 1}
-
-\subsection*{Subsection 2}
-
-\section*{Discussion}
-
 % You may title this section "Methods" or "Models". 
@@
 The process begins by presuming a model of the controller.
-Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and stabilizes it with linear feedback of position and velocity error.
+Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation.
+Additionally, linear feedback rejects position and velocity error.
 More complex models can be used without any loss of generality.
@@
 \end{equation}
-Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripal effects, and $E$ is any externally-applied torques that the system might interact with.
+Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripetal effects, and $E$ is externally-applied torque that the system might interact with.
 Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles.
@@
 \end{equation}
+Where a hat ($\hat{ }$) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model.
+This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm.
+If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero.
+$K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$.
+This $q_d$ signifies the unknown desired trajectory that we seek to discover.
+
+For typical dynamic simulations, in order to determine the trajectory the system (i.e., the forward dynamics problem), this second order differential equation requires standard numerical integration to determine the solution to the initial value problem in time.
+This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory.
+The intent extraction approach modifies this method by instead solving for $\ddot{q}_d$:
+\begin{equation}
+\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q-q_d)+K_d(\dot{q}-\dot{q}_d)]\right\}
+\end{equation}  
+$\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$.
+Provided the model of plant and controller are precise, the initial conditions are available and accurate, the mass matrix estimate $\hat{M}$ is invertible, and the constants are accurate and do not vary, and externally-applied force can be accurately measured, then system yields an accurate estimate of the intent.
+
+In fact, this approach can be generalized to any system with an intention that governs the control law, and hence is agnostic to the form of controller.
+In order to prove this, we introduce one additional layer of complexity by separating the plant into its process and actuator components.
+Modeling the actuator explicitly allows us to compare our $x_d$ against lambda models \cite{feldman1995origin} and their hypothetical equilibrium trajectory, $x_e$.
+Additionally, this enables a more general derivation.
+We describe both as instantaneously linear differential equations.
+\begin{equation}
+\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{n=1}^N A_n(x-x_e)^{(n)}}^\text{Actuator}}_\text{Plant}+E=0
+\end{equation}
+Controlling the process as intended requires a specific relationship between the internal model and choice of actuator equilibria, $x_e$, and impedances, $A_n$:
+\begin{equation}
+\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{n=1}^N A_n(x_d-x_e)^{(n)}}^\text{Actuator}=0
+\end{equation}
+
+ Taking the system above and generalizing the extraction of intent trajectory $x_d(t)$ to any such feedforward system well-described by a differential equation up to some order N,
+Note that the impedance and feedback functions can take a variety of forms that include delays and other terms that might account for the actuation of many control processes.
+Similar to before, this can now be reformulated,
+
+    EQUATION HERE
+
+Where $P_N$ is the coefficient matrix associated with highest order N, and $P_(1...N)$ are all lower order coefficients.
+
+% Results and Discussion can be combined.
+\section*{Results}
+
+\subsection*{Subsection 1}
+
+\subsection*{Subsection 2}
+
+\section*{Discussion}
 
@@
 % The bibtex filename
-\bibliography{paper1.bib}
+\bibliography{paper1}
 

### 2014-05-12 matlab/final_fig45_combined.m

@@ -1,0 +1,152 @@
+clc
+clear all
+close all
+
+%Set ranges
+%For each range, show the bar plots with figure window titled
+%    set(gcf,'name',range)
+%Grab the mean and confidence interval from c=multcompare with output off
+%Plot mean and confidence interval against range
+%call that finalfig5
+
+NR=10;
+edges=linspace(0,100,NR+1); %sampling rate is 200 Hz = .005 s samples
+
+black=[0 0 0];
+red=[1 .3 .3];
+green=[.1 .7 .3];
+height=7;
+width=9;
+margin=1;
+figmargin=.1;
+
+figure(1)
+clf
+hold on
+set(gcf,'color',[1 1 1])
+set(gcf,'units','centimeters')
+set(gcf,'position',[4,8,figmargin+margin+width,figmargin+margin+height])
+
+for k=1:4
+    load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
+    load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
+    load(['../Data/Data_pulse/pulse',num2str(k),'U.mat'])
+
+    flag=0;
+
+    clear cleanStruct
+    f=find([trialInfo.clean]);
+    for t=2:length(f)
+        lt=length(trials(f(t)).t);
+        lty=length(undisturbed(f(t)).ty);
+        inds=(1+lt-lty):lt; %Onset of movement
+        cleanStruct(t).R2BX=getMaxPerp(trials(f(t)).x(inds,:));
+        cleanStruct(t).R2BY=getMaxPerp(undisturbed(f(t)).y);
+    end
+    subject(k).R2BX=vertcat(cleanStruct.R2BX);
+    subject(k).R2BY=vertcat(cleanStruct.R2BY);
+
+    clear triStruct
+
+    for R=1:NR
+        RANGE=edges(R):edges(R+1);
+        for t=1:length(trials)
+            if ~sum(trials(t).disturbcat==[1:4]) %Fast-ish skipping instead of indexing
+                continue
+            end
+            inds=trialInfo(t).forceinds(1)+RANGE;
+            onset=find(vecmag(trials(t).v)>.05,1,'first');
+            start=max(onset-35,1);
[…]

### 2014-05-13 Paper1/paper1.tex
kk, here.

@@
 \begin{equation}
-\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q-q_d)+K_d(\dot{q}-\dot{q}_d)}_{\text{Impendance, Feedback Controller}}
+\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
 \end{equation}
@@
 \begin{equation}
-\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q-q_d)+K_d(\dot{q}-\dot{q}_d)]\right\}
+\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})]\right\}
 \end{equation}  
@@
 Additionally, this enables a more general derivation.
-We describe both as instantaneously linear differential equations.
+Without loss of generality, we describe both as instantaneously linear differential equations in some space, $x$:
 \begin{equation}
-\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{n=1}^N A_n(x-x_e)^{(n)}}^\text{Actuator}}_\text{Plant}+E=0
+\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
 \end{equation}
-Controlling the process as intended requires a specific relationship between the internal model and choice of actuator equilibria, $x_e$, and impedances, $A_n$:
+Controlling the process as intended requires a specific relationship between the internal model, choice of actuator equilibria $x_e$, and impedances $A_n$:
 \begin{equation}
-\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{n=1}^N A_n(x_d-x_e)^{(n)}}^\text{Actuator}=0
+\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
 \end{equation}
+Because the actuator's reference trajectory is shared between both representations, we can solve both equations for $\sum_{m=1}^M A_mx_e^{(m)}$, set them equal, and collect the remaining actuator terms into a single feedback component:
+\begin{equation}
+\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
+\end{equation}
+We now recover our familiar model and learn its relationship with $x_e$.
+While a proper choice of $x_e$ was critical in controlling the actuator, a complete model relating actual and intended trajectory does not and need not reflect it.
+The actuator's equilibrium is not the process's equilibrium unless $x_d$ and all of its derivatives are zero.
+This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does represent reaching intent\cite{gomi1997human}.
 
- Taking the system above and generalizing the extraction of intent trajectory $x_d(t)$ to any such feedforward system well-described by a differential equation up to some order N,
-Note that the impedance and feedback functions can take a variety of forms that include delays and other terms that might account for the actuation of many control processes.
-Similar to before, this can now be reformulated,
+This general form reveals the conditions needed to solve for the intention of an arbitrary process.
+We can solve for $x_d$ if the actuator's instantaneous impedance is known and either the plant or its highest order term can be inverted.
+Moreover, the plant model does not need to be explicitly formulated.
+In cases like national security or athletic performance, the plant (a bomb in a crowded square or an opposing team) may be too complex to model.
+By substituting a lookup table made up of past state-outcome pairs for an explicit process model, it becomes possible to determine desired outcome even when the state is disturbed.
+It is not necessarily to allow a bomb to be planted or an opposing team to carry out a play in order to discover their plans.
 
-    EQUATION HERE
+\textcolor{red}{\textbf{This last bit needs work.
+Solving for the Feldman equilibrium addresses your prior objections and keeps everything in familiar forms throughout.
+I'm not sure if take-home points for the reader go here or in the discussion.
+In particular, this gets us back to skewering Feldman.}}
 
-Where $P_N$ is the coefficient matrix associated with highest order N, and $P_(1...N)$ are all lower order coefficients.
+\subsection*{Generation of synthetic data using simple dynamic simulation}
+We choose one prevailing arm model, that of Burdet et al.
+(2006), along with the anatomical mass distribution and geometry parameters of Winters (1990) to create a model for estimating intent.
+The model of Burdet et al (2006) was adapted to reflect measured physical parameters of the subject who provided the reach.
+Segment lengths were measured directly while gross body mass was self-reported.
+These values were converted to segment masses, centers of gravity, and moments of inertia inertia using anatomical landmarks and values from Dempster (1955) and Winters (1990).
+The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds).
+Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700ms starting and ending with zero velocity and acceleration.
+Simulated reaches were 15 cm long, beginning 38 cm out from and 5.7 cm left of the right shoulder then proceeding right.
+We used the following types of perturbing forces for each combination of distance and direction:
+Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
+Noise forces began once the subject moved 3 mm, and lasted for the duration of the motion.
+The forces were drawn from a white noise generator drawn at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.
+
+\subsection*{Error Metric}
+Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 1mm wide bins spaced evenly throughout the 15cm reach.
+Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis.
+Maximum unsigned error, also called maximum perpendicular distance, was used to measure reaching accuracy to avoid sensitivity to duration of inspection.
+
+\subsection*{Indices of variance-based sensitivity}
+Sobol-distributed matrices were generated using MATLAB’s sobolset() function and converted to parameter distributions (see Table 1) using inverse cumulative probability density.
+Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values.
+Combinations and calculations were made as prescribed by Saltelli et al.
+(2010) to arrive at direct and indirect sensitivity indices.
+
+\subsection*{Human Subjects}
+The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board.
+Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and was not compensated.
+Subjects’ arm segment lengths were directly measured in situ while body mass was self-reported.
+
+\subsection*{Apparatus}
+A planar manipulandum (described in Patton and Mussa-Ivaldi, 2004) was programmed to compensate and minimize any handle friction or mass.
+The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
+Visual feedback was performed at 60 Hz using OpenGL.
+Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was <8 ms, ensuring a visual delay less than one 60 Hz frame.
+
+\subsection*{Protocol}
+Subjects made 730 reaches in total, along a line parallel to their coronal plane and approximately 45 cm from their shoulder.
+Reaches were either 15 or 30 cm long, starting and ending at one of three points spaced 15 cm apart on the line.
+To prevent any learning effect, forces were presented intermittently with random frequency, but never less than 5 reaches apart.
+Forces were chosen pseudorandomly such that each type, direction, and distance combination mentioned above was presented 5 times.
+
+\subsection*{Statistical Analysis}
+The MATLAB statistics toolbox package (Natick, MA) was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences.
+Nonparametric testing was necessary due to floor effects from using unsigned measures.
 
@@
 
-\subsection*{Subsection 1}
+\subsection*{Model Viability}
+Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
+Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (figure 1) were extracted in the presence of 220,000 variations upon the parameters (table 1) according to the methods of Saltelli et al.
+(2011).
+Expected variation due to parameter uncertainty was 0.86 microns for pulse forces and 1.5 microns for filtered white forces.
+Variation in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was 2.93 microns.
+These variations both fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).
+In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.
 
-\subsection*{Subsection 2}
+\subsection*{Parameter Sensitivity}
+Simulated error due to direct parameter uncertainty (figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
+Indirect sensitivity, which reveals how errors from other parameters are affected by changes in a given parameter, was an order of magnitude lower than direct sensitivity and again implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
+Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.
+
+\subsection*{Human Intent Trajectories}
+Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3).
+In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.
+When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
+When disturbances occurred late in reaching, intent tended to continue to onto the target with change of intent occurring on or near the target (figure 3 lower plots).
+Patterns across all trials and subject revealed some systematic tendencies.
+For approximately 200 milliseconds after the onset of disturbance (figure 4 compare individual points).
+At this point intent significantly deviated from the baseline region of tendency.
+As the movement completed, the hand position and intent re-converged and become statistically indistinguishable at approximately 400ms after the onset of disturbance (figure 4 shaded areas).
 

### 2014-05-13 Paper1/paper1.tex
Updated paper, figs

@@
 This presumes that what people do is on average what they intended to do and that their goal remains unchanged.
-However, if there is a disturbance, it is unclear what happens to a person’s intentions because action is decoupled from intention.
+However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention.
 Lacking a means to discern that intention and changes in intention leads to challenges in human machine interactions.
@@
 Springs produce a force according to their impedance and stretch.
-By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle’s equilibrium point.
-Supporters of the so-called lambda model\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
+By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
+Supporters of the so-called $\lambda$ model\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
 Upon inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
@@
 
-Collectively, the general problem of being able to best extract someone’s intent is critical to many fields, and hinges on our understanding of the underlying mechanisms through which control is accomplished.
+Collectively, the general problem of being able to best extract someone's intent is critical to many fields, and hinges on our understanding of the underlying mechanisms through which control is accomplished.
 We have excellent models of the mechanical properties of arm and muscle because the physics that describe their gross behavior is so well-understood and the need to determine particular physical parameter values has long been clear.
@@
 These studies show planning and correction in discrete subunits, but did not analyze the actual intent and hence cannot be compared to a known, correct trajectory except in the trivial case that movement is undisturbed.
-A disturbed trajectory’s intention is hidden without proper inversion of the model plan for movement.
+A disturbed trajectory's intention is hidden without proper inversion of the model plan for movement.
 Here we develop a method to extract intent and gain a clearer understanding of subunits from disturbed motions.
@@
 
-In fact, this approach can be generalized to any system with an intention that governs the control law, and hence is agnostic to the form of controller.
-In order to prove this, we introduce one additional layer of complexity by separating the plant into its process and actuator components.
-Modeling the actuator explicitly allows us to compare our $x_d$ against lambda models \cite{feldman1995origin} and their hypothetical equilibrium trajectory, $x_e$.
-Additionally, this enables a more general derivation.
-Without loss of generality, we describe both as instantaneously linear differential equations in some space, $x$:
+It is important to distinguish our intended trajectory, $x_d$, from another variable often considered, the equilibrium point of the muscle actuators.
+Importantly, this has been claimed to be the intended movement.
+To consider this, we also generalize our approach to any system with a moving intention that governs the control law.
+This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $x_e$ (sometimes represented as $\lambda$).
+The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
+Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
 \begin{equation}
@@
 \end{equation}
-Controlling the process as intended requires a specific relationship between the internal model, choice of actuator equilibria $x_e$, and impedances $A_n$:
+The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ that $x$ will track a desired path $x_d$.
 \begin{equation}
@@
 \end{equation}
-Because the actuator's reference trajectory is shared between both representations, we can solve both equations for $\sum_{m=1}^M A_mx_e^{(m)}$, set them equal, and collect the remaining actuator terms into a single feedback component:
+We solve this for $x_e$ and substituting into eq 4 gives:
 \begin{equation}
@@
 \end{equation}
-We now recover our familiar model and learn its relationship with $x_e$.
-While a proper choice of $x_e$ was critical in controlling the actuator, a complete model relating actual and intended trajectory does not and need not reflect it.
-The actuator's equilibrium is not the process's equilibrium unless $x_d$ and all of its derivatives are zero.
-This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does represent reaching intent\cite{gomi1997human}.
 
-This general form reveals the conditions needed to solve for the intention of an arbitrary process.
-We can solve for $x_d$ if the actuator's instantaneous impedance is known and either the plant or its highest order term can be inverted.
-Moreover, the plant model does not need to be explicitly formulated.
-In cases like national security or athletic performance, the plant (a bomb in a crowded square or an opposing team) may be too complex to model.
-By substituting a lookup table made up of past state-outcome pairs for an explicit process model, it becomes possible to determine desired outcome even when the state is disturbed.
-It is not necessarily to allow a bomb to be planted or an opposing team to carry out a play in order to discover their plans.
+Note $x_e$ vanishes, recovering our familiar model.
+While a proper choice of $x_e$ is critical to control, our approach does not require a model relating equilibrium and intended trajectory.
+Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero.
+This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
 
-\textcolor{red}{\textbf{This last bit needs work.
-Solving for the Feldman equilibrium addresses your prior objections and keeps everything in familiar forms throughout.
-I'm not sure if take-home points for the reader go here or in the discussion.
-In particular, this gets us back to skewering Feldman.}}
+In summary, if the highest order coefficient of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
+This general form reveals the conditions needed to solve for the moving intent of an arbitrary dynamic process.
 
@@
 We choose one prevailing arm model, that of Burdet et al.
-(2006), along with the anatomical mass distribution and geometry parameters of Winters (1990) to create a model for estimating intent.
-The model of Burdet et al (2006) was adapted to reflect measured physical parameters of the subject who provided the reach.
+\cite{burdet2006stability}, along with the anatomical mass distribution and geometry parameters of Winter \cite{winter2009biomechanics} to create a model for estimating intent.
+The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach.
 Segment lengths were measured directly while gross body mass was self-reported.
-These values were converted to segment masses, centers of gravity, and moments of inertia inertia using anatomical landmarks and values from Dempster (1955) and Winters (1990).
+These values were converted to segment masses, centers of gravity, and moments of inertia inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}.
 The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds).
@@
 \subsection*{Indices of variance-based sensitivity}
-Sobol-distributed matrices were generated using MATLAB’s sobolset() function and converted to parameter distributions (see Table 1) using inverse cumulative probability density.
+Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
 Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values.
 Combinations and calculations were made as prescribed by Saltelli et al.
-(2010) to arrive at direct and indirect sensitivity indices.
+\cite{saltelli2010variance} to arrive at direct and indirect sensitivity indices.
 
@@
 Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and was not compensated.
-Subjects’ arm segment lengths were directly measured in situ while body mass was self-reported.
+Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
 
 \subsection*{Apparatus}
-A planar manipulandum (described in Patton and Mussa-Ivaldi, 2004) was programmed to compensate and minimize any handle friction or mass.
+A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass.
 The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
@@
 Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (figure 1) were extracted in the presence of 220,000 variations upon the parameters (table 1) according to the methods of Saltelli et al.
-(2011).
+\cite{saltelli2010variance}.
 Expected variation due to parameter uncertainty was 0.86 microns for pulse forces and 1.5 microns for filtered white forces.
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=4in]{fig1.eps}
+\end{center}
+\caption{
+{\bf Simulated data illustrating tautology of extraction absent parameter error across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial.
+Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory.
+Extraction to recover intention from arm and force trajectory follows.
+}
+\label{fig:synthetic}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=4in]{fig2.eps}
+\end{center}
+\caption{
+{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Direct sensitivity, in purple, shows how modeling error changes with the parameter.
+Indirect sensitivity, in teal, shows how modeling error from other parameters is affected by changes in the parameter.
+Variation in individual tests exceeds expected variance by two orders of magnitude.
+}
+\label{fig:sensitivity}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=4in]{fig3.eps}
+\end{center}
+\caption{
+{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied.
+This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
+}
+\label{fig:anecdotes}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=4in]{fig4.eps}
+\end{center}
+\caption{
+{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) deviate almost immediately from undisturbed hand errors (green dots, band).
+Movement intention (red dots, band) deviates at approximately 100ms after the onset of disturbance.
+The hand rejoins the intention approximately 500ms after the onset of disturbance.
+}
+\label{fig:grouptrends}
+\end{figure}
+
 \section*{Tables}
@@
 
+\begin{table}[!ht]
+\caption{
+\bf{Arm Parameter Values and Sensitivities}}
+\begin{tabular}{|c|c|c c|c c|}
+\hline
+Parameter Name &
+Units &
+Nominal &
+SD &
+$\frac{S}{\text{Total}}$ &
+$\frac{ST}{\text{Total}}$ \\ \hline
+Upper Arm Length ($L_1$) &
+m &
+0.33 &
+0.01 &
+0.07 &
+0 \\
+Forearm Length ($L_2$) &
+m &
+0.34 &
+0.01 &
+0.02 &
+0 \\
+Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
+1 &
+0.436 &
+0.0695 &
+0.01 &
+0 \\
+Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
+1 &
+0.682 &
+0.0431 &
+0.02 &
+0 \\
+Gross Body Mass ($m_g$) &
+kg &
+79.36 &
+3.1 &
+0.01 &
+0 \\
+Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
+1 &
+0.028 &
+0.0029 &
+0.01 &
+0 \\
+Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
+1 &
+0.022 &
+0.0025 &
+0.04 &
+0 \\
+Upper Arm Radius of Gyration Ratio &
+1 &
+0.322 &
+0.0161 &
+0.01 &
+0 \\
+Forearm Radius of Gyration Ratio &
+1 &
+0.468 &
+0.0234 &
+0.01 &
+0 \\
+Shoulder Parallel Coordinate &
+m &
+-0.057 &
+0.02 &
+0.13 &
+0.01 \\
+Shoulder Perpendicular Coordinate &
+m &
+0.88 &
+0.02 &
+0.05 &
+0 \\
+Force Sensor Miscalibration $x\text{-axis}$ &
+N &
+0 &
+0.231 &
+0.06 &
+0 \\
+Force Sensor Miscalibration $y\text{-axis}$ &
+N &
+0 &
+0.1067 &
+0.06 &
+0.02 \\
+Force Sensor Gaussian Noise SD $x\text{-axis}$ &
+N &
+0 &
+0.1653 &
+0.01 &
+0 \\
+Force Sensor Gaussian Noise SD $y\text{-axis}$ &
+N &
+0 &
+0.2869 &
+0.01 &
+0 \\
+Torque-Invariant Impedance Misestimation Ratio &
+1 &
+1 &
+0.15 &
+0.22 &
+0.04 \\
+Torque-Varying Impedance Misestimation Ratio &
+1 &
+1 &
+0.15 &
+0.11 &
+0.01 \\
+Damping-to-Stiffness Ratio ($k_d$) &
+$\text{sec}^{-1}$ &
+0.0833 &
+0.0125 &
+0.06 &
+0 \\
+Reflex Impedance Scale Factor &
+1 &
+0.02 &
+0.003 &
+0.03 &
+0 \\
+Reflex Damping to Stiffness Ratio ($g_d$) &
+$\text{sec}^{-1}$ &
+2 &
+0.3 &
+0.03 &
+0 \\ \hline
+\end{tabular}
+\begin{flushleft}Synthetic model parameters and their associated mean (nominal) values and standard deviations (SD) used to determine the sensitivity indices of Saltelli et al.
+\cite{saltelli2010variance}.
+Also shown are the resulting sensitivity indices.
+Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero.
+Total sensitivity was $1.8 \mu m$, showing very small average deviations when using this approach.
+\end{flushleft}
+\label{tab:parameters}
+\end{table}
+
 \end{document}

### 2014-05-13 matlab/final_fig45_combined.m

@@ -9,8 +9,8 @@ close all
 %Plot mean and confidence interval against range
 %call that finalfig5
 
-NR=10;
-edges=linspace(0,100,NR+1); %sampling rate is 200 Hz = .005 s samples
+NR=20;
+edges=linspace(0,180,NR+1); %sampling rate is 200 Hz = .005 s samples
 
 black=[0 0 0];
 red=[1 .3 .3];
@@ -75,7 +75,7 @@ for k=1:4
         Xlist=[CBX;CX;CY];
         Ylist=[subject(k).R2BX;subject(k).R2X;subject(k).R2Y];
 
-        DOTSIZE=4;
+        DOTSIZE=5;
         LINEWIDTH=5;
         binCenter=5*(edges(R)+edges(R+1))/2;
         drawnWidth=5*.25*(edges(R+1)-edges(R));
@@ -108,11 +108,11 @@ for k=1:NR
     kw(k).c=multcompare(stats,'display','off');
     
     [p,table,stats]=anova1(fullY,fullX,'off');
-    mids(k,:)=stats.means;
-%     for kk=1:3
-%         F=find(fullX==kk);
-%         mids(k,kk)=median(fullY(F));
-%     end
+    %mids(k,:)=stats.means;
+    for kk=1:3
+        F=find(fullX==kk);
+        mids(k,kk)=median(fullY(F));
+    end
     
     errors(k,:)=halfWidth(stats)';
     rangemids(k)=5*(edges(k)+edges(k+1))/2;
@@ -133,8 +133,8 @@ end
 plot([0,0],[0 5],'k-','linewidth',3)
 text(0,2.5,'Error, 5 cm','rotation',90,'horizontalalignment','center','verticalalignment','bottom')
 set(gca,'ycolor',[1 1 1])
-set(gca,'xtick',[25 475])
-set(gca,'xticklabels',{'0','500'})
+%set(gca,'xtick',[25 475])
+%set(gca,'xticklabels',{'0','500'})
 set(gca,'ytick',[])
 set(gca,'TickLength',[0 0]);
 set(gca,'linewidth',[1]);
@@ -146,6 +146,9 @@ set(gca,'position',[margin margin width height],'units','normalized')
 set(0,'defaulttextinterpreter','none')
 
 matlabfrag('figures/fig4raw','renderer','opengl','dpi',600);
+
+print figures/fig4.eps -depsc
+
 
 
 

### 2014-05-13 matlab/sensitivityPaper/final_fig1.m

@@ -126,4 +126,8 @@ text(.0005,.51,'10 N','rotation',90,'Horizontalalignment','center','Verticalalig
 
 set(0,'defaulttextinterpreter','none')
 
-laprint(gcf,'../figures/fig1raw','width',15,'scalefonts','off','factor',1)
+matlabfrag('../figures/fig1raw');
+%laprint(gcf,'../figures/fig1raw','width',15,'scalefonts','off','factor',1)
+
+set(0,'defaulttextinterpreter','latex')
+print ../figures/fig1.eps -depsc

### 2014-05-13 matlab/sensitivityPaper/final_fig2.m

@@ -64,9 +64,10 @@ for k=1:20
     S(k)=1/(2*N)*sum((allA-allAB(:,k)).^2);
     ST(k)=1/N*sum(allB.*(allAB(:,k)-allA));
 end
-v=[S' ST'];
+v=[S' ST']
+totalSensitivity=sum(sum(v))
 %v=v/var(allAB(:));
-v=v/sum(sum(v));
+v=v/totalSensitivity;
 [vals,order]=sort(v(:,1));
 %order=1:length(order)
 
@@ -111,8 +112,12 @@ set(gca,'position',[.5 margin 1-margin 1-margin],'units','normalized')
 
 set(0,'defaulttextinterpreter','none')
 set(gcf,'color',[1 1 1])
+matlabfrag('..figures/fig2raw');
 
-laprint(gcf,'../figures/fig2raw','scalefonts','off','asonscreen','on')
+print ../figures/fig2 -depsc
+
+
+%laprint(gcf,'../figures/fig2raw','scalefonts','off','asonscreen','on')
 
 for k=1:20
     disp([num2str(v(k,1),'%0.2f'),' ',num2str(v(k,2),'%0.2f'),' ',names{k}])

### 2014-05-14 Paper1/paper1.tex
Did this skip a commit?

@@
 \section*{Abstract}
-\textbf{Needs work.}\\
 We derive and examine the parameter sensitivity of a novel modeling approach for extracting movement intent from a force-disturbed planar arm trajectory.

### 2014-05-14 Paper1/paper1.tex
Old version of latex blows...

@@
 \section*{Discussion}
+We sought to determine the plausibility and robustness of an algorithm that determines a person’s intent of a motion, even if there disturbances.
+Sensitivity analysis on synthetic data revealed that the method was plausible and response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modeled.
+Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for ~100ms following the onset of disturbance.
+This test provides a method for testing and potentially falsifying our extraction.
+Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent.
+In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus a first look at how intent changes when faced with a force disturbance.
+
+$\lambda$ models seek to explain reaching intent as the muscles’ equilibrium trajectory.
+By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.
+Even as this rejects a major claim of those modeling approaches, it explains a key experimental underpinning: Bizzi’s et al.’s \cite{bizzi1984posture} discovery of a “virtual trajectory” in deafferented monkeys with a moving equilibrium point that progressed smoothly from the initial to final positions.
+Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass.
+
+Not only is intent slow to change, it does not necessarily change due to disturbance.
+While this work does not examine this directly, our observations are more compatible with intermittent control (or even submovements) than with continuous optimal feedback control.
+In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}.
+Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
+Interestingly, our data disturbances that come later in the movement appeared to not dramatically affect the intent (Figure 3, lower plot), suggesting that people may stubbornly “stick to their guns” when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
+In other words, response latencies may be  dependent on the time of the disturbance in the movement.
+Another explanation is that control is intermittent \cite{gawthrop2011intermittent}.
+It remains to be seen if and how change in intent might be triggered.
+
+This study might have large dividends in the study of error augmentation, which has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}.
+This relies on dictating the direction of reach.
+By instead scaling the magnitude of the difference between the desired and realized trajectory, error augmentation during undirected reaching and exploration becomes possible.
+This permits both error reduction and error magnification.
+As demonstrated by the success of the challenge-point framework \cite{guadagnoli2004challenge}, dynamic variation of augmentation as task learning progresses can be beneficial.
+With this extraction, error can be measured and augmented in real time even without an explicit task, potentially enabling wider utility.
+While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention.
+This filter incorporates the inversion of estimates of the actual human controller into the computation.
+While a large number of filters have already been proposed, we believe this to be a unique tool to be added to the arsenal of analytical methods for exploring human motor control.
+It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.
+The ability of such a filter to produce intended trajectories from force-disturbed movements allows for hypothesis testing that previously was not available.
+It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} “candidate model” of human control and create a filter that estimates the intended trajectory.
+By using tests such as the straight line reaching test employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
+Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting.
+Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions?
+Similarly, claims of composition of desired trajectories by submovements should now be directly testable.
+Finally, optimal feedback models can now gain direct access to an intended trajectory to better understand and fit their cost function.
+
+Nevertheless, there are some drawbacks and limitations to these methods.
+One limiting requirement of this method is that accurate interface force (within a few tenths of a newton) is needed.
+Such instrumentation is available, but must be used with great care in order to preserve the accuracy of the estimate.
+Our sensitivity analysis on synthetic data revealed that slight error can lead to a trajectory that accumulates a drift due to bias caused by the force term.
+More importantly, while modern force sensors can be highly accurate with very high signal to noise ratios, they tend to drift over time, leading to error that can grow if the device is not periodically tared.
+Our empirical approach used several methods to mitigate these effects, and hence the errors became small.
+It remains to be seen if there are methods to reduce the errors from force error, or better to eliminate their need altogether.
+
+Similarly, because our intent extraction approach involves integration, it  can be “re-zeroed” at any point in time to improve accuracy, provided new initial condition can be obtained.
+Our work in this study dealt with targeted point-to-point motions, and hence assumed that each motion’s initial position (prior to any motion) was equal to the initial intended position.
+In continuous motion tasks and other applications, it may be possible to obtain an indication from the operator to know that they are precisely where they intend to be at some moment in the activity, refining subsequent estimates.
+
+This study did not examine all disturbances.
+Although we have found and validated a potentially powerful tool for examining human intention as it changes (or fails to) in response to disturbance, other disturbances.
+Even as we demonstrated the power of this tool, we couldn’t examine all force disturbances.
+Impulses and Gaussian noise have convenient spectral properties that provide some generality to this claim, but the arm model examined here is not a linear, time-invariant system.
+The eigenfunctions of this nonlinear system are not known and it is unclear which force disturbances might best expose critical vulnerabilities.
+
+The approximation of stiffness appears to be the most critical factor in accuracy in estimating intent.
+While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
+Even as the model is insensitive to normal variations, halving or doubling this parameter due to contraction is plausible and has significant impact on accuracy.
+The limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction where or when it occurs.
+
+Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} may not be all that is needed.
+More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction.
+Even scaling the constant and torque-dependent stiffness matrixes separately would enable rotation of the extracted trajectory.
+Our results point to the possibility of the impedance changing during motion, perhaps also as a function of time.
+The system was also shown to be sensitive to uncertainties in impedance.
+While this may be perceived as a shortcoming, switching from a state-varying impedance to ongoing time-varying determination of impedance may overcome these issues.
+[summary and take home; Hypothesize /speculate what data suggest, are there new questions.  What contributions did this paper make? 
+In any case, the general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
+Moreover, the plant model does not need to be explicitly formulated.
+Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance.
+The plant (such as the actions of a crowd in a crowded square or the motions of the opposing team) may be too complex to model with simple linear transformations as we present above.
+However, by substituting some approximation or lookup table of input-outcome pairs, it is possible to determine in a variety of applications to obtain the intent, even when disturbed.
+It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand their plan behind the action.
 

### 2014-05-14 Paper1/paper1.tex
Better dempster reference, working on abstract

@@
 \\
-\bf{2} SMPP, RIC, Chicago, Illinois, United States
+\bf{2} Sensory Motor Performance Program, RIC, Chicago, Illinois, United States
 \\
@@
 \section*{Abstract}
-We derive and examine the parameter sensitivity of a novel modeling approach for extracting movement intent from a force-disturbed planar arm trajectory.
+We derive a novel modeling approach for extracting movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
 While popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions.
@@
 Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints.
-The robustness of this model holds promise for real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals.
+The robustness of this model allows us to apply it to human reaching data
+
+holds promise for real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals.
 In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.
@@
 With this extraction, error can be measured and augmented in real time even without an explicit task, potentially enabling wider utility.
+
 While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention.
@@
 It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.
+
 The ability of such a filter to produce intended trajectories from force-disturbed movements allows for hypothesis testing that previously was not available.
@@
 While this may be perceived as a shortcoming, switching from a state-varying impedance to ongoing time-varying determination of impedance may overcome these issues.
-[summary and take home; Hypothesize /speculate what data suggest, are there new questions.  What contributions did this paper make? 
+
 In any case, the general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
@@
 \begin{center}
-\includegraphics[width=4in]{fig1.eps}
+%\includegraphics[width=4in]{fig1.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4in]{fig2.eps}
+%\includegraphics[width=4in]{fig2.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4in]{fig3.eps}
+%\includegraphics[width=4in]{fig3.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4in]{fig4.eps}
+%\includegraphics[width=4in]{fig4.eps}
 \end{center}

### 2014-05-14 Paper1/paper1.tex
Nerfed the abstract

@@
 \\
-\bf{2} Sensory Motor Performance Program, RIC, Chicago, Illinois, United States
+\bf{2} Sensory Motor Performance Program, Rehabilitation Institute of Chicago, Chicago, Illinois, United States
 \\
-$\ast$ E-mail: Corresponding pattonj@uic.edu
+$\ast$ E-mail: pattonj@uic.edu
 \end{flushleft}
@@
 \section*{Abstract}
-We derive a novel modeling approach for extracting movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
-While popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions.
-After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching.
-We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching.
-Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints.
-The robustness of this model allows us to apply it to human reaching data
-
-holds promise for real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals.
-In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.
 

### 2014-05-20 Paper1/paper1.tex
Pushed and saved

@@
 \section*{Introduction}
-As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted or influenced by pathology.
+As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology.
 One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal.
@@
 However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention.
-Lacking a means to discern that intention and changes in intention leads to challenges in human machine interactions.
+Lacking a means to discern that intention and changes in intention leads to challenges in human-machine interactions.
 Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.
@@
 
-Understanding the intent has many benefits, and one may be a better way to understand the composition of discrete movements into their subunits.
-Recent research has strived to understand the composition of movements intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements}, distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable.
-Flash and Henis \cite{flash1991arm} demonstrated that despite this apparent blending, cued corrections can only be produced at predictable places within the reach.
-They also showed that these modifications appear to take place by superimposing additional straight-line subunits onto the original movement.
+Examination of movement intent during disturbance should allow us to test whether or not that intent is elemental or composed of subunits.
+Recent research has strived to understand the composition of movement intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements} where distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable.
+Past attempts to isolate subunits from healthy adult reaching are inconclusive.
+Flash and Henis \cite{flash1991arm} demonstrated that  cued corrections produce recoverable subunits with a predictable shape and structure, but Liu and Todorov were able to reproduce this without using subunits \cite{liu2007evidence}.
 Novak et al.
-\cite{novak2002use} detected corrective subunits following disturbance during knob turning.
-These studies show planning and correction in discrete subunits, but did not analyze the actual intent and hence cannot be compared to a known, correct trajectory except in the trivial case that movement is undisturbed.
-A disturbed trajectory's intention is hidden without proper inversion of the model plan for movement.
-Here we develop a method to extract intent and gain a clearer understanding of subunits from disturbed motions.
+\cite{novak2002use} detected corrective subunits following disturbance during knob turning, but lacking access to the intent they instead relied on kinematic features of the disturbed trajectory.
+Here we develop a method to extract intent from disturbed movement, which will allow us to observe how intent changes during uncued replanning and whether that change is discrete or continuous.
 
@@
 Segment lengths were measured directly while gross body mass was self-reported.
-These values were converted to segment masses, centers of gravity, and moments of inertia inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}.
+These values were converted to segment masses, centers of gravity, and moments of inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}.
 The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds).
@@
 Noise forces began once the subject moved 3 mm, and lasted for the duration of the motion.
-The forces were drawn from a white noise generator drawn at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.
+The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.
 
@@
 The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board.
-Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and was not compensated.
+Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated.
 Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
@@
 Visual feedback was performed at 60 Hz using OpenGL.
-Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was <8 ms, ensuring a visual delay less than one 60 Hz frame.
+Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 ms, ensuring a visual delay less than one 60 Hz frame.
 
@@
 
-\subsection*{Model Viability}
+\subsection*{Model Robustness}
 Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
@@
 \cite{saltelli2010variance}.
-Expected variation due to parameter uncertainty was 0.86 microns for pulse forces and 1.5 microns for filtered white forces.
-Variation in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was 2.93 microns.
-These variations both fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).
+Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces.
+Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$.
+These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).
 In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.
@@
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred late in reaching, intent tended to continue to onto the target with change of intent occurring on or near the target (figure 3 lower plots).
-Patterns across all trials and subject revealed some systematic tendencies.
-For approximately 200 milliseconds after the onset of disturbance (figure 4 compare individual points).
-At this point intent significantly deviated from the baseline region of tendency.
-As the movement completed, the hand position and intent re-converged and become statistically indistinguishable at approximately 400ms after the onset of disturbance (figure 4 shaded areas).
+When disturbances occurred late in reaching, intent tended to continue onto the target with intent changing on or near the target (figure 3 lower plots).
+Patterns across all disturbance timings and subjects revealed some systematic tendencies.
+Force-disturbed hand paths deviate from undisturbed reaching approximately 25 milliseconds after the onset of disturbance while intention remains comparable to undisturbed reaching for 100 milliseconds.
+While the hand returns within 500 milliseconds to undisturbed levels of error, intention is altered for an entire second (figure 4 shaded areas).
 
 \section*{Discussion}
-We sought to determine the plausibility and robustness of an algorithm that determines a person’s intent of a motion, even if there disturbances.
-Sensitivity analysis on synthetic data revealed that the method was plausible and response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modeled.
-Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for ~100ms following the onset of disturbance.
+We sought to determine the plausibility and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances.
+Sensitivity analysis on synthetic data revealed that the method was plausible.
+Response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modelled.
+Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100ms following the onset of disturbance.
 This test provides a method for testing and potentially falsifying our extraction.

### 2014-05-20 Paper1/paper1.tex
Updated .tex

@@
 \section*{Discussion}
-We sought to determine the plausibility and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances.
-Sensitivity analysis on synthetic data revealed that the method was plausible.
+We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances.
+Sensitivity analysis on synthetic data revealed that the method was suitable.
 Response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modelled.
-Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100ms following the onset of disturbance.
-This test provides a method for testing and potentially falsifying our extraction.
+Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100 milliseconds following the onset of disturbance.
+As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison.
+If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands.
+This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
+As demonstrated in figure 4, the method passes this test until 100 milliseconds after the onset of disturbance.
 Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent.
-In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus a first look at how intent changes when faced with a force disturbance.
+In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus supports a first look at how intent changes when faced with a force disturbance.
 
@@
 By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.
-Even as this rejects a major claim of those modeling approaches, it explains a key experimental underpinning: Bizzi’s et al.’s \cite{bizzi1984posture} discovery of a “virtual trajectory” in deafferented monkeys with a moving equilibrium point that progressed smoothly from the initial to final positions.
+Even as this rejects a major claim of those modeling approaches, it explains a key experimental underpinning: Bizzi’s et al.'s \cite{bizzi1984posture} discovery of a “virtual trajectory” in deafferented monkeys with a moving equilibrium point that progressed smoothly from the initial to final positions.
 Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass.
+This provides a striking and falsifiable prediction: human reaching intention cannot jump abruptly even if mass is wholly compensated.
 
 Not only is intent slow to change, it does not necessarily change due to disturbance.
-While this work does not examine this directly, our observations are more compatible with intermittent control (or even submovements) than with continuous optimal feedback control.
+While this work examines the time of onset of change, no statistical test could compellingly "fail to find" change in individual movements without markedly more data.
+Still, figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control.
 In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}.
 Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Interestingly, our data disturbances that come later in the movement appeared to not dramatically affect the intent (Figure 3, lower plot), suggesting that people may stubbornly “stick to their guns” when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
-In other words, response latencies may be  dependent on the time of the disturbance in the movement.
-Another explanation is that control is intermittent \cite{gawthrop2011intermittent}.
+Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly “stick to their guns” when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
+Response latencies may be dependent on the time of the disturbance in the movement.
+Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent.
 It remains to be seen if and how change in intent might be triggered.
 
-This study might have large dividends in the study of error augmentation, which has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}.
-This relies on dictating the direction of reach.
-By instead scaling the magnitude of the difference between the desired and realized trajectory, error augmentation during undirected reaching and exploration becomes possible.
+Determination of intent may facilitate may facilitate motor training and stroke recovery.
+Error augmentation, which presently relies on dictating the reaching intent, has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}.
+This static, dictated intent could be replaced By instead scaling the magnitude of the difference between the desired and realized trajectory, error augmentation during undirected reaching and exploration becomes possible.
 This permits both error reduction and error magnification.
+Partial error cancellation would allow learning to take place without harm to task goals.
 As demonstrated by the success of the challenge-point framework \cite{guadagnoli2004challenge}, dynamic variation of augmentation as task learning progresses can be beneficial.
-With this extraction, error can be measured and augmented in real time even without an explicit task, potentially enabling wider utility.
+With this extraction, error can be measured and augmented in real-time even without an explicit task, potentially enabling wider utility and application.
 
-While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention.
+\textbf{While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention.
 This filter incorporates the inversion of estimates of the actual human controller into the computation.
 While a large number of filters have already been proposed, we believe this to be a unique tool to be added to the arsenal of analytical methods for exploring human motor control.
-It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.
+It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.}
 
-The ability of such a filter to produce intended trajectories from force-disturbed movements allows for hypothesis testing that previously was not available.
-It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} “candidate model” of human control and create a filter that estimates the intended trajectory.
+Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible.
+It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} "candidate model" of the human arm and create a filter that estimates the intended trajectory.
 By using tests such as the straight line reaching test employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
@@
 Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions?
-Similarly, claims of composition of desired trajectories by submovements should now be directly testable.
-Finally, optimal feedback models can now gain direct access to an intended trajectory to better understand and fit their cost function.
+They do.
+Similarly, compositional claims, such as submovement decomposition, should now be directly testable.
+Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit their cost functions.
 
 Nevertheless, there are some drawbacks and limitations to these methods.
-One limiting requirement of this method is that accurate interface force (within a few tenths of a newton) is needed.
+One limiting requirement is that accurate interface force measurement (within a few tenths of a newton) is needed.
 Such instrumentation is available, but must be used with great care in order to preserve the accuracy of the estimate.
@@
 
-Similarly, because our intent extraction approach involves integration, it  can be “re-zeroed” at any point in time to improve accuracy, provided new initial condition can be obtained.
-Our work in this study dealt with targeted point-to-point motions, and hence assumed that each motion’s initial position (prior to any motion) was equal to the initial intended position.
-In continuous motion tasks and other applications, it may be possible to obtain an indication from the operator to know that they are precisely where they intend to be at some moment in the activity, refining subsequent estimates.
-
 This study did not examine all disturbances.
-Although we have found and validated a potentially powerful tool for examining human intention as it changes (or fails to) in response to disturbance, other disturbances.
+Although we have found and validated a potentially powerful tool for examining human intention as it changes (or fails to) in response to some disturbances, other disturbances.
 Even as we demonstrated the power of this tool, we couldn’t examine all force disturbances.

### 2014-05-20 Paper1/paper1.tex
Need to fix finalfig2 because of quirks with the s vs st. Not sure when I fixed that, but I fixed it spuriously.

@@
 {\Large
-\textbf{Title}
+\textbf{Intention can be Determined Despite Disturbances}
 }
@@
 \section*{Abstract}
+We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
+While we demonstrate that popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions.
+After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching.
+We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching.
+Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints.
+The robustness of this approach indicated that it was suitable for determining reaching intent from human point-to-point reaching movements disturbed by large force pulses.
+Intended trajectory showed no change from undisturbed reaching for 150 ms after the onset of disturbance.
+Beyond this point, stiffening of the arm, change in intent, or both usually lead to deviation from undisturbed reaching.
+Not all disturbed reaches showed evidence of change in intent.
+Unlike $\lambda$ models, the reaching intent determined by our model is physiologically plausible.
+This success allows real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals.
+In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.
+While not explicitly examined, this approach is also valid for determination intent is more complex situations like team sports and national security.
 
@@
 However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention.
-Lacking a means to discern that intention and changes in intention leads to challenges in human-machine interactions.
+Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned.
 Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.
@@
 
-Examination of movement intent during disturbance should allow us to test whether or not that intent is elemental or composed of subunits.
-Recent research has strived to understand the composition of movement intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements} where distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable.
-Past attempts to isolate subunits from healthy adult reaching are inconclusive.
-Flash and Henis \cite{flash1991arm} demonstrated that  cued corrections produce recoverable subunits with a predictable shape and structure, but Liu and Todorov were able to reproduce this without using subunits \cite{liu2007evidence}.
-Novak et al.
-\cite{novak2002use} detected corrective subunits following disturbance during knob turning, but lacking access to the intent they instead relied on kinematic features of the disturbed trajectory.
-Here we develop a method to extract intent from disturbed movement, which will allow us to observe how intent changes during uncued replanning and whether that change is discrete or continuous.
-
-In this paper, we employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction.
+Here we develop and analyze a method to extract intent from disturbed movement, which will allow us to observe if and how intent changes following disturbance.
+We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction.
 We then explore sensitivity and failure modes using standard model analysis techniques.
@@
 % authors may use "Analysis" 
-\section*{Materials and Methods}
+\section*{Methods}
 The sections below describe the theoretical method, and then present our evaluations.
@@
 Combinations and calculations were made as prescribed by Saltelli et al.
-\cite{saltelli2010variance} to arrive at direct and indirect sensitivity indices.
+\cite{saltelli2010variance} to arrive at direct and total sensitivity indices.
 
@@
 A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass.
-The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
+The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
 Visual feedback was performed at 60 Hz using OpenGL.
@@
 \subsection*{Statistical Analysis}
-The MATLAB statistics toolbox package (Natick, MA) was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences.
 Nonparametric testing was necessary due to floor effects from using unsigned measures.
@@
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred late in reaching, intent tended to continue onto the target with intent changing on or near the target (figure 3 lower plots).
-Patterns across all disturbance timings and subjects revealed some systematic tendencies.
-Force-disturbed hand paths deviate from undisturbed reaching approximately 25 milliseconds after the onset of disturbance while intention remains comparable to undisturbed reaching for 100 milliseconds.
-While the hand returns within 500 milliseconds to undisturbed levels of error, intention is altered for an entire second (figure 4 shaded areas).
+When disturbances occurred late in reaching, intent continued onto the target (figure 3 lower plots).
+We observed a systematic change of intent 150 milliseconds after the onset of force disturbance.
+While the hand returns within 600 milliseconds to undisturbed levels of error, intention is altered for almost 900 milliseconds (figure 4 shaded areas).
+
+As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison.
+If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands.
+This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
+As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance.
+Further
 
@@
 We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances.
-Sensitivity analysis on synthetic data revealed that the method was suitable.
-Response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modelled.
-Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100 milliseconds following the onset of disturbance.
-As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison.
-If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands.
-This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
-As demonstrated in figure 4, the method passes this test until 100 milliseconds after the onset of disturbance.
-Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent.
-In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus supports a first look at how intent changes when faced with a force disturbance.
+Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching.
+Significant misestimation could occur if the stiffness was inaccurately modelled.
+We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding.
+This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.
 
-$\lambda$ models seek to explain reaching intent as the muscles’ equilibrium trajectory.
+While the derivation of intent determination required many steps, the outcome is a filter whose use is straight-forward.
+Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated.
+Demonstrating the validity and plausibility of inverting those filters to determine intent from hand trajectory adds a novel method to the arsenal of analytical methods for exploring human motor control.
+
+This approach explains the discovery by Bizzi et al.
+\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$.
 By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.
-Even as this rejects a major claim of those modeling approaches, it explains a key experimental underpinning: Bizzi’s et al.'s \cite{bizzi1984posture} discovery of a “virtual trajectory” in deafferented monkeys with a moving equilibrium point that progressed smoothly from the initial to final positions.
 Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass.
-This provides a striking and falsifiable prediction: human reaching intention cannot jump abruptly even if mass is wholly compensated.
+This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.
 
-Not only is intent slow to change, it does not necessarily change due to disturbance.
-While this work examines the time of onset of change, no statistical test could compellingly "fail to find" change in individual movements without markedly more data.
+Not only is intent slow to change, it does not necessarily change following disturbance.
+While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement.
 Still, figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control.
@@
 Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly “stick to their guns” when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
-Response latencies may be dependent on the time of the disturbance in the movement.
+Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
+Time of response may be dependent on the time of the disturbance in the movement.
 Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent.
@@
 
-\textbf{While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention.
-This filter incorporates the inversion of estimates of the actual human controller into the computation.
-While a large number of filters have already been proposed, we believe this to be a unique tool to be added to the arsenal of analytical methods for exploring human motor control.
-It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.}
-
 Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible.
-It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} "candidate model" of the human arm and create a filter that estimates the intended trajectory.
-By using tests such as the straight line reaching test employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
+It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory.
+By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
 Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting.
@@
 They do.
-Similarly, compositional claims, such as submovement decomposition, should now be directly testable.
-Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit their cost functions.
+Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable.
+Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit hypothesized cost functions.
 
-Nevertheless, there are some drawbacks and limitations to these methods.
+The plant model does not need to be explicitly formulated.
+Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance.
+The plant (such as the actions of a crowd in a crowded square, the motions of the opposing team,) may be too complex to model with simple linear transformations as we present above.
+However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed.
+It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand the plan behind their action.
+
+Nevertheless, there are some limitations to these methods.
 One limiting requirement is that accurate interface force measurement (within a few tenths of a newton) is needed.
@@
 Our empirical approach used several methods to mitigate these effects, and hence the errors became small.
-It remains to be seen if there are methods to reduce the errors from force error, or better to eliminate their need altogether.
-
-This study did not examine all disturbances.
-Although we have found and validated a potentially powerful tool for examining human intention as it changes (or fails to) in response to some disturbances, other disturbances.
-Even as we demonstrated the power of this tool, we couldn’t examine all force disturbances.
-Impulses and Gaussian noise have convenient spectral properties that provide some generality to this claim, but the arm model examined here is not a linear, time-invariant system.
-The eigenfunctions of this nonlinear system are not known and it is unclear which force disturbances might best expose critical vulnerabilities.
+It remains to be seen if there are methods to further reduce the misestimation due to force error or eliminate the need for force estimation altogether.
 
@@
 While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
-Even as the model is insensitive to normal variations, halving or doubling this parameter due to contraction is plausible and has significant impact on accuracy.
-The limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction where or when it occurs.
+Even as the model is insensitive to normal variations, a human can volitionally halve or double their stiffness \cite{franklin2007endpoint}.
+Co-contraction is plausible and has significant impact on accuracy.
+These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations.
+Additionally, real-time determination of intent may allow real-time determination of stiffness.
 
-Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} may not be all that is needed.
-More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction.
-Even scaling the constant and torque-dependent stiffness matrixes separately would enable rotation of the extracted trajectory.
-Our results point to the possibility of the impedance changing during motion, perhaps also as a function of time.
-The system was also shown to be sensitive to uncertainties in impedance.
-While this may be perceived as a shortcoming, switching from a state-varying impedance to ongoing time-varying determination of impedance may overcome these issues.
+Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting.
+More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
+Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
+As mentioned above, while this may be perceived as a shortcoming, it was necessary and potentially facilitates real-time determination of stiffness in future work.
 
-In any case, the general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
-Moreover, the plant model does not need to be explicitly formulated.
-Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance.
-The plant (such as the actions of a crowd in a crowded square or the motions of the opposing team) may be too complex to model with simple linear transformations as we present above.
-However, by substituting some approximation or lookup table of input-outcome pairs, it is possible to determine in a variety of applications to obtain the intent, even when disturbed.
-It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand their plan behind the action.
+The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
 
@@
 \caption{
-{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Direct sensitivity, in purple, shows how modeling error changes with the parameter.
-Indirect sensitivity, in teal, shows how modeling error from other parameters is affected by changes in the parameter.
-Variation in individual tests exceeds expected variance by two orders of magnitude.
+{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1.
+First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value.
+Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
 }
@@
 \caption{
-{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) deviate almost immediately from undisturbed hand errors (green dots, band).
-Movement intention (red dots, band) deviates at approximately 100ms after the onset of disturbance.
+{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) begin to deviate from undisturbed hand errors (green dots, band) beginning 95 ms after the onset of disturbance.
+Movement intention (red dots, band) deviates at 150ms.
 The hand rejoins the intention approximately 500ms after the onset of disturbance.
@@
 SD &
-$\frac{S}{\text{Total}}$ &
-$\frac{ST}{\text{Total}}$ \\ \hline
+$S$ &
+$S_T$ \\ \hline
 Upper Arm Length ($L_1$) &
@@
 0.01 &
-0.07 &
-0 \\
+0 &
+0.10 \\
 Forearm Length ($L_2$) &
@@
 0.01 &
-0.02 &
-0 \\
+0 &
+0.03 \\
 Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
@@
 0.0695 &
-0.01 &
-0 \\
+0 &
+0.02 \\
 Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
@@
 0.0431 &
-0.02 &
-0 \\
+0 &
+0.02 \\
 Gross Body Mass ($m_g$) &
@@
 3.1 &
-0.01 &
-0 \\
+0 &
+0.02 \\
 Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
@@
 0.0029 &
-0.01 &
-0 \\
+0 &
+0.01 \\
 Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
@@
 0.0025 &
-0.04 &
-0 \\
+0 &
+0.06 \\
 Upper Arm Radius of Gyration Ratio &
@@
 0.0161 &
-0.01 &
-0 \\
+0 &
+0.01 \\
 Forearm Radius of Gyration Ratio &
@@
 0.0234 &
-0.01 &
-0 \\
+0 &
+0.01 \\
 Shoulder Parallel Coordinate &
@@
 0.02 &
-0.13 &
-0.01 \\
+0.02 &
+0.18 \\
 Shoulder Perpendicular Coordinate &
@@
 0.02 &
-0.05 &
-0 \\
+0 &
+0.08 \\
 Force Sensor Miscalibration $x\text{-axis}$ &
@@
 0.231 &
-0.06 &
-0 \\
+0 &
+0.08 \\
 Force Sensor Miscalibration $y\text{-axis}$ &
@@
 0.1067 &
-0.06 &
-0.02 \\
+0.03 &
+0.08 \\
 Force Sensor Gaussian Noise SD $x\text{-axis}$ &
@@
 0.1653 &
-0.01 &
-0 \\
+0 &
+0.01 \\
 Force Sensor Gaussian Noise SD $y\text{-axis}$ &
@@
 0.2869 &
-0.01 &
-0 \\
+0 &
+0.01 \\
 Torque-Invariant Impedance Misestimation Ratio &
@@
 0.15 &
-0.22 &
-0.04 \\
+0.06 &
+0.31 \\
 Torque-Varying Impedance Misestimation Ratio &
@@
 0.15 &
-0.11 &
-0.01 \\
+0.0 &
+0.15 \\
 Damping-to-Stiffness Ratio ($k_d$) &
@@
 0.0125 &
-0.06 &
-0 \\
+0 &
+0.08 \\
 Reflex Impedance Scale Factor &
@@
 0.003 &
-0.03 &
-0 \\
+0 &
+0.04 \\
 Reflex Damping to Stiffness Ratio ($g_d$) &
@@
 0.3 &
-0.03 &
-0 \\ \hline
+0 &
+0.04 \\ \hline
 \end{tabular}
@@
 Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero.
-Total sensitivity was $1.8 \mu m$, showing very small average deviations when using this approach.
+Total sensitivity was $1.3 \mu m^2$, showing very small average deviations when using this approach.
+First order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$ can be interpreted as the fraction left after correcting all other factors.
 \end{flushleft}

### 2014-05-20 matlab/final_fig45_combined.m

@@ -9,7 +9,7 @@ close all
 %Plot mean and confidence interval against range
 %call that finalfig5
 
-NR=20;
+NR=10;
 edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
 
 black=[0 0 0];
@@ -110,11 +110,97 @@ for k=1:4
     end
 end
 
+NR=200;
+edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
+
+for k=1:4
+    load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
+    load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
+    load(['../Data/Data_pulse/pulse',num2str(k),'U.mat'])
+
+    flag=0;
+
+    clear cleanStruct
+    f=find([trialInfo.clean]);
+    for t=2:length(f)
+        lt=length(trials(f(t)).t);
+        lty=length(undisturbed(f(t)).ty);
+        inds=(1+lt-lty):lt; %Onset of movement
+        cleanStruct(t).R2BX=getMaxPerp(trials(f(t)).x(inds,:));
+        cleanStruct(t).R2BY=getMaxPerp(undisturbed(f(t)).y);
+    end
+    subject(k).R2BX=vertcat(cleanStruct.R2BX);
+    subject(k).R2BY=vertcat(cleanStruct.R2BY);
+
+    clear triStruct
+
+    for R=1:NR+1
+        RANGE=edges(R):edges(R+1);
+        
+        for t=1:length(trials)
+            if ~sum(trials(t).disturbcat==[1:4]) %Fast-ish skipping instead of indexing
+                continue
+            end
+            inds=trialInfo(t).forceinds(1)+RANGE;
+            onset=find(vecmag(trials(t).v)>.05,1,'first');
+            start=max(onset-35,1);
+            indsy=max(1,trialInfo(t).forceinds(1)-start)+RANGE;
+
+            try
+                triStruct(t).R2X=getR2(trials(t).x(inds,:));
+                triStruct(t).R2Y=getR2(trials(t).y(indsy,:));
+            catch
+                st1=size(trials(t).x,1);
[…]

### 2014-05-20 matlab/finalfig3.m

@@ -1,11 +1,16 @@
 clc
 clear all
 
-anecdote(1).sub=1;
-anecdote(1).SE=[1 3];
+anecdote(1).sub=4;
+anecdote(1).SE=[3 1];
 anecdote(1).dcat=2;
 anecdote(1).rot=0;
 anecdote(1).yoff=0;
+%anecdote(1).sub=1;
+%anecdote(1).SE=[1 3];
+%anecdote(1).dcat=2;
+%anecdote(1).rot=0;
+%anecdote(1).yoff=0;
 anecdote(2).sub=4;
 anecdote(2).SE=[3 1];
 anecdote(2).dcat=4;

### 2014-05-21 Paper1/paper1.tex
Mild update

@@
 \section*{Abstract}
-We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
-While we demonstrate that popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions.
+We derive a novel modeling approach for determining movement intent from force-disturbed arm movement, examine its parameter sensitivity, and validate it on human reaching data.
+We demonstrate that inverting popular arm modeling approaches algebraically allows determination of movement intent from recordings of force disturbances and arm movement.
+We then extend this to a more general approach reveals strengths, limitations, and underlying assumptions.
 After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching.

### 2014-05-27 Paper1/paper1.tex
Same
Updated to reflect weekend edits

@@
 We demonstrate that inverting popular arm modeling approaches algebraically allows determination of movement intent from recordings of force disturbances and arm movement.
-We then extend this to a more general approach reveals strengths, limitations, and underlying assumptions.
+We then extend this to a more general approach, which reveals strengths, limitations, and underlying assumptions.
 After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching.
@@
 \end{equation}
-Where a hat ($\hat{ }$) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model.
+Where a hat ( $\hat{ }$ ) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model.
 This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm.
@@
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
+In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
 
-In summary, if the highest order coefficient of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
-This general form reveals the conditions needed to solve for the moving intent of an arbitrary dynamic process.
+While we treat both the process and actuator as dynamic models in our derivation, both the derivation and the technique are agnostic to the process model.
+Instead of a dynamic model, it could be a lookup table.
+For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
+In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance.
+As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined.
+This allows determination of intention in many situations of interest, even where the process is irreducibly complex.
 
@@
 \subsection*{Human Subjects}
-The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board.
+The human data trajectories analyzed here are drawn from four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board.
 Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated.
@@
 As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance.
-Further
 
@@
 Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching.
-Significant misestimation could occur if the stiffness was inaccurately modelled.
+Significant misestimation could occur if the stiffness was inaccurately modeled.
 We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding.
@@
 
-This approach explains the discovery by Bizzi et al.
+This modeling approach explains the discovery by Bizzi et al.
 \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$.
-By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.
+By deriving the technique in a general form, we discovered that our $x_d$ should be equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.
+$x_d$ and $\lambda$ are only equivalent when $x_d$ is not changing.
 Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass.
@@
 
-Determination of intent may facilitate may facilitate motor training and stroke recovery.
+Determination of intent may facilitate motor training and stroke recovery.
 Error augmentation, which presently relies on dictating the reaching intent, has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}.
-This static, dictated intent could be replaced By instead scaling the magnitude of the difference between the desired and realized trajectory, error augmentation during undirected reaching and exploration becomes possible.
-This permits both error reduction and error magnification.
+Augmentation of difference from static, dictated intent could be replaced with scaling of the magnitude of the difference between the desired and realized trajectory.
+This makes error augmentation during undirected reaching and exploration possible, including both error reduction and error magnification.
 Partial error cancellation would allow learning to take place without harm to task goals.
@@
 Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible.
-It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory.
+It is important to reiterate that because the human arm's mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory.
 By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
@@
 
-The plant model does not need to be explicitly formulated.
+The process model does not need to be explicitly formulated.
 Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance.
-The plant (such as the actions of a crowd in a crowded square, the motions of the opposing team,) may be too complex to model with simple linear transformations as we present above.
+The plant (such as the actions of a crowd in a crowded square or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
 However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed.
-It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand the plan behind their action.
+It is not necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.
 
@@
 
-The approximation of stiffness appears to be the most critical factor in accuracy in estimating intent.
+The approximation of stiffness appears to be the most critical factor to accuracy in estimating intent.
 While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
-Even as the model is insensitive to normal variations, a human can volitionally halve or double their stiffness \cite{franklin2007endpoint}.
-Co-contraction is plausible and has significant impact on accuracy.
+Even as the model is insensitive to normal variations, a human can halve or double their stiffness \cite{franklin2007endpoint} significantly reducing determination accuracy.
 These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations.
@@
 Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
-As mentioned above, while this may be perceived as a shortcoming, it was necessary and potentially facilitates real-time determination of stiffness in future work.
+As mentioned above, while this may be perceived as a shortcoming, it was necessary and can potentially be overcome in future work.
 
 The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
+Sensitivity analysis for this approach applied to the human arm demonstrated favorable conditions for determination of intent during force-disturbed reaching.
+Reaching intent became significantly different from baseline reaching only after enough time had elapsed for the disturbance to be processed and descending motor commands to change.
+This window into intent should allow advances in arm modeling, motor training, and human-machine interaction.
 
@@
 \begin{center}
-%\includegraphics[width=4in]{fig1.eps}
+\includegraphics[width=4in]{fig1.eps}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig2.eps}
+\includegraphics[width=4in]{fig2.eps}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig3.eps}
+\includegraphics[width=4in]{fig3.eps}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig4.eps}
+\includegraphics[width=4in]{fig4.eps}
 \end{center}
@@
 Also shown are the resulting sensitivity indices.
-Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero.
-Total sensitivity was $1.3 \mu m^2$, showing very small average deviations when using this approach.
-First order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$ can be interpreted as the fraction left after correcting all other factors.
+Sensitivity indices less than one-thousandth were reported as zero.
+Total variance was $1.3 \mu m^2$, showing very small average deviations when using this approach.
+First-order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
 \end{flushleft}

### 2014-05-28 Paper1/paper1.tex
Updated for paper ready-ish to go out.

@@
 {\Large
-\textbf{Intention can be Determined Despite Disturbances}
+\textbf{Determining the Intent of Action Despite Disturbances}
 }
@@
 Justin Horowitz$^{1}$, 
-James Patton$^{12\ast}$
+James Patton$^{1,2\ast}$
 \\
@@
 \section*{Abstract}
-We derive a novel modeling approach for determining movement intent from force-disturbed arm movement, examine its parameter sensitivity, and validate it on human reaching data.
-We demonstrate that inverting popular arm modeling approaches algebraically allows determination of movement intent from recordings of force disturbances and arm movement.
-We then extend this to a more general approach, which reveals strengths, limitations, and underlying assumptions.
-After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching.
-We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching.
-Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints.
-The robustness of this approach indicated that it was suitable for determining reaching intent from human point-to-point reaching movements disturbed by large force pulses.
-Intended trajectory showed no change from undisturbed reaching for 150 ms after the onset of disturbance.
-Beyond this point, stiffening of the arm, change in intent, or both usually lead to deviation from undisturbed reaching.
+We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
+Popular modeling approaches can be inverted in a general approach, and we show here the strengths, limitations, and underlying assumptions.
+We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen in human targeted reaching.
+Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness.
+We also tested the robustness of this approach on human point-to-point reaching movements disturbed by large force pulses.
+As expected, the intended trajectory showed no change from undisturbed reaching for 150 ms after the disturbance onset.
+Beyond this point, intent change, arm stiffening, or both usually lead to deviation from undisturbed reaching.
 Not all disturbed reaches showed evidence of change in intent.
@@
 In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.
-While not explicitly examined, this approach is also valid for determination intent is more complex situations like team sports and national security.
+While not explicitly examined, this approach is also valid for determining intent in more complex situations such as human-machine interaction, team sports, or national security.
 
@@
 Additionally, linear feedback rejects position and velocity error.
-More complex models can be used without any loss of generality.
+This model is chosen to be very simple to help illuminate the approach, but it is limited in how it captures all of the complexity of motor control.
+Nevertheless later we will discuss how this might be used on more complex models without loss of generality.
 The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
@@
 
-For typical dynamic simulations, in order to determine the trajectory the system (i.e., the forward dynamics problem), this second order differential equation requires standard numerical integration to determine the solution to the initial value problem in time.
+For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation requires standard numerical integration to determine the solution to the initial value problem in time.
 This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory.
@@
 $\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$.
-Provided the model of plant and controller are precise, the initial conditions are available and accurate, the mass matrix estimate $\hat{M}$ is invertible, and the constants are accurate and do not vary, and externally-applied force can be accurately measured, then system yields an accurate estimate of the intent.
+Provided the model of plant and controller are precise, the initial conditions are available and accurate, the mass matrix estimate $\hat{M}$ is invertible, and the constants are accurate and do not vary, and externally-applied force can be accurately measured, then the system yields an accurate estimate of the intent.
 
@@
 Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
-\begin{equation}
+\begin{equation} \label{eq:generalPlant}
 \underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
@@
 \end{equation}
-We solve this for $x_e$ and substituting into eq 4 gives:
+We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives:
 \begin{equation}
@@
 The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds).
-Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700ms starting and ending with zero velocity and acceleration.
-Simulated reaches were 15 cm long, beginning 38 cm out from and 5.7 cm left of the right shoulder then proceeding right.
+Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.
+
+Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.
 We used the following types of perturbing forces for each combination of distance and direction:
-Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
-Noise forces began once the subject moved 3 mm, and lasted for the duration of the motion.
+pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
+Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion.
 The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.
@@
 \subsection*{Error Metric}
-Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 1mm wide bins spaced evenly throughout the 15cm reach.
+Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15centimeters reach.
 Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis.
-Maximum unsigned error, also called maximum perpendicular distance, was used to measure reaching accuracy to avoid sensitivity to duration of inspection.
+Signed error relative to the direction of disturbance, also called perpendicular deviation, was used to measure reaching accuracy in order to elucidate direction of any corrections.
 
@@
 \subsection*{Human Subjects}
-The human data trajectories analyzed here are drawn from four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board.
+The human data trajectories analyzed here are drawn from four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
 Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated.
@@
 Visual feedback was performed at 60 Hz using OpenGL.
-Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 ms, ensuring a visual delay less than one 60 Hz frame.
+Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
 
 \subsection*{Protocol}
-Subjects made 730 reaches in total, along a line parallel to their coronal plane and approximately 45 cm from their shoulder.
-Reaches were either 15 or 30 cm long, starting and ending at one of three points spaced 15 cm apart on the line.
+Subjects made 730 reaches in total, along a line parallel to their coronal plane and approximately 45 centimeters from their shoulder.
+Reaches were either 15 or 30 centimeters long, starting and ending at one of three points spaced 15 centimeters apart on the line.
 To prevent any learning effect, forces were presented intermittently with random frequency, but never less than 5 reaches apart.
@@
 The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences.
-Nonparametric testing was necessary due to floor effects from using unsigned measures.
+Nonparametric testing was necessary due to differences in the shapes of the force-disturbed and undisturbed hand error distributions.
 
@@
 Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
-Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (figure 1) were extracted in the presence of 220,000 variations upon the parameters (table 1) according to the methods of Saltelli et al.
+Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al.
 \cite{saltelli2010variance}.
@@
 \subsection*{Parameter Sensitivity}
-Simulated error due to direct parameter uncertainty (figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
+Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
 Indirect sensitivity, which reveals how errors from other parameters are affected by changes in a given parameter, was an order of magnitude lower than direct sensitivity and again implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
@@
 \subsection*{Human Intent Trajectories}
-Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3).
+Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3).
 In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred late in reaching, intent continued onto the target (figure 3 lower plots).
-We observed a systematic change of intent 150 milliseconds after the onset of force disturbance.
-While the hand returns within 600 milliseconds to undisturbed levels of error, intention is altered for almost 900 milliseconds (figure 4 shaded areas).
+When disturbances occurred late in reaching, intent continued onto the target (Figure 3 lower plots).
+We observed no significant change of intent during force disturbance relative to undisturbed reaching.
+Despite statistical insignificance, intention showed some systematic deflection opposite the direction of disturbance beginning around 275 milliseconds after the onset of force disturbance.
+While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to over-correct and is altered for almost 900 milliseconds (Figure 4 shaded areas).
 
@@
 This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
-As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance.
+As demonstrated in Figure 4, the method passes this test.
 
 \section*{Discussion}
-We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances.
+We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there are disturbances.
 Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching.
 Significant misestimation could occur if the stiffness was inaccurately modeled.
-We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding.
+We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds.
 This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.
@@
 
-Not only is intent slow to change, it does not necessarily change following disturbance.
+Intent is slow to change if it changes at all following disturbance.
 While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement.
-Still, figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control.
+Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control.
 In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}.
 Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
+Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
 Time of response may be dependent on the time of the disturbance in the movement.
@@
 Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions?
-They do.
+Not always.
 Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable.
@@
 However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed.
-It is not necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.
+In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.
 
@@
 \section*{Acknowledgments}
+Research supported by NIH NS053606.
+We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
+We also the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
 
@@
 \begin{center}
-\includegraphics[width=4in]{fig4.eps}
+%\includegraphics[width=4in]{fig4.eps}
 \end{center}
 \caption{
-{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) begin to deviate from undisturbed hand errors (green dots, band) beginning 95 ms after the onset of disturbance.
-Movement intention (red dots, band) deviates at 150ms.
+{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching.} Hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ simultaneous 95\% confidence interval on the mean) begin to deviate from undisturbed hand deviation (green band) beginning 95 ms after the onset of disturbance.
+Movement intention (red dots, band) does not significantly deviate.
 The hand rejoins the intention approximately 500ms after the onset of disturbance.
@@
 Total variance was $1.3 \mu m^2$, showing very small average deviations when using this approach.
-First-order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
+First-order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor.
+Total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
 \end{flushleft}

### 2014-05-28 matlab/final_fig45_combined.m

@@ -15,10 +15,10 @@ edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
 black=[0 0 0];
 red=[1 .3 .3];
 green=[.1 .7 .3];
-height=7;
-width=9;
+width=12;
+height=1.618*width;
 lmargin=.5;
-bmargin=1;
+bmargin=0;
 figmargin=.4;
 
 figure(1)
@@ -41,8 +41,8 @@ for k=1:4
         lt=length(trials(f(t)).t);
         lty=length(undisturbed(f(t)).ty);
         inds=(1+lt-lty):lt; %Onset of movement
-        cleanStruct(t).R2BX=getMaxPerp(trials(f(t)).x(inds,:));
-        cleanStruct(t).R2BY=getMaxPerp(undisturbed(f(t)).y);
+        cleanStruct(t).R2BX=getR2(trials(f(t)).x(inds,:));
+        cleanStruct(t).R2BY=getR2(undisturbed(f(t)).y);
     end
     subject(k).R2BX=vertcat(cleanStruct.R2BX);
     subject(k).R2BY=vertcat(cleanStruct.R2BY);
@@ -61,102 +61,21 @@ for k=1:4
             start=max(onset-35,1);
             indsy=max(1,trialInfo(t).forceinds(1)-start)+RANGE;
 
+            sPeak=sign(trials(t).x(trialInfo(t).forceinds(1)+50,2)-.5);
+            
             try
-                triStruct(t).R2X=getR2(trials(t).x(inds,:));
-                triStruct(t).R2Y=getR2(trials(t).y(indsy,:));
+                triStruct(t).R2X=sPeak*getR2(trials(t).x(inds,:));
+                triStruct(t).R2Y=sPeak*getR2(trials(t).y(indsy,:));
             catch
                 st1=size(trials(t).x,1);
-                triStruct(t).R2X=getR2(trials(t).x(inds(1):st1,:));
+                triStruct(t).R2X=sPeak*getR2(trials(t).x(inds(1):st1,:));
                 st1=size(trials(t).x,1);
-                triStruct(t).R2Y=getR2(trials(t).y(indsy(1):st1,:));
+                triStruct(t).R2Y=sPeak*getR2(trials(t).y(indsy(1):st1,:));
             end
+            
         end
[…]

### 2014-05-28 matlab/finalfig3.m

@@ -1,16 +1,16 @@
 clc
 clear all
 
-anecdote(1).sub=4;
-anecdote(1).SE=[3 1];
+% anecdote(1).sub=4;
+% anecdote(1).SE=[3 1];
+% anecdote(1).dcat=2;
+% anecdote(1).rot=0;
+% anecdote(1).yoff=0;
+anecdote(1).sub=1;
+anecdote(1).SE=[1 3];
 anecdote(1).dcat=2;
 anecdote(1).rot=0;
 anecdote(1).yoff=0;
-%anecdote(1).sub=1;
-%anecdote(1).SE=[1 3];
-%anecdote(1).dcat=2;
-%anecdote(1).rot=0;
-%anecdote(1).yoff=0;
 anecdote(2).sub=4;
 anecdote(2).SE=[3 1];
 anecdote(2).dcat=4;

### 2014-05-28 matlab/halfWidth.m

@@ -1,4 +1,4 @@
-function out = halfWidth(stats)
+function out = halfWidth(stats,varargin)
 
 %% Code from multcompare.m
 gmeans = stats.means(:);
@@ -7,7 +7,11 @@ s = stats.s;
 ng = sum(n>0);
 df = stats.df;
 
-alpha=.01;
+if nargin>1
+    alpha=varargin{1};
+else
+    alpha=.01;
+end
 
 ctype='tukey-kramer';
 crit = getcrit(ctype, alpha, df, ng);

### 2014-05-28 matlab/sensitivityPaper/final_fig2.m

@@ -82,9 +82,9 @@ figure(1)
 clf
 hold on
 set(gcf,'color',[1 1 1])
-set(gcf,'units','centimeters')
-set(gcf,'position',[4,8,figmargin+lmargin+width,figmargin+bmargin+height])
-set(gcf,'units','normalized')
+%set(gcf,'units','centimeters')
+%set(gcf,'position',[4,8,figmargin+lmargin+width,figmargin+bmargin+height])
+%set(gcf,'units','normalized')
 
 spacer=.1;
 NPLOT=1000;
@@ -120,17 +120,15 @@ text(labN+.5,TIGHT,'1 mm','Horizontalalignment','center','verticalalignment','to
 
 margin=.05;
 
-set(gcf,'color',[1 1 1])
-set(gca,'units','centimeters')
-set(gca,'position',[lmargin bmargin width height])
-set(0,'defaulttextinterpreter','none')
+set(gcf,'position',[250 300 625 400])
+set(gca,'position',[.5 margin 1-margin 1-margin],'units','normalized')
 
-matlabfrag('..figures/fig2raw','epspad',5000*[1 1 1 1]);
+set(0,'defaulttextinterpreter','none')
+set(gcf,'color',[1 1 1])
+
+laprint(gcf,'../figures/fig2raw','scalefonts','off','asonscreen','on')
 
 %print ../figures/fig2 -depsc
-
-
-%laprint(gcf,'../figures/fig2raw','scalefonts','off','asonscreen','on')
 
 for k=1:20
     disp([num2str(v(k,1),'%0.2f'),' ',num2str(v(k,2),'%0.2f'),' ',names{k}])

### 2014-05-29 matlab/final_fig45_combined.m
One last time?

@@ -9,6 +9,8 @@ close all
 %Plot mean and confidence interval against range
 %call that finalfig5
 
+SUBS=2;
+
 NR=10;
 edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
 
@@ -28,7 +30,7 @@ set(gcf,'color',[1 1 1])
 set(gcf,'units','centimeters')
 set(gcf,'position',[4,8,figmargin+lmargin+width,figmargin+bmargin+height])
 
-for k=1:4
+for k=SUBS
     load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
     load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
     load(['../Data/Data_pulse/pulse',num2str(k),'U.mat'])
@@ -117,7 +119,7 @@ end
 NR=200; %200
 edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
 
-for k=1:4
+for k=SUBS
     load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
     load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
     load(['../Data/Data_pulse/pulse',num2str(k),'U.mat'])
@@ -155,11 +157,13 @@ for k=1:4
             indsy=max(1,trialInfo(t).forceinds(1)-start)+RANGE;
             
             sPeak=sign(trials(t).x(trialInfo(t).forceinds(1)+50,2)-.5);
+            if sPeak > 10
+                continue
+            end
 
             try
                 triStruct(t).R2X=sPeak*getR2(trials(t).x(inds,:));
-                %triStruct(t).R2Y=sPeak*getR2(trials(t).y(indsy,:));
-                triStruct(t).R2Y=getR2(trials(t).y(indsy,:));
+                triStruct(t).R2Y=sPeak*getR2(trials(t).y(indsy,:));
             catch
                 thishappened=1
                 st1=size(trials(t).x,1);
@@ -254,13 +258,13 @@ set(gca,'ytick',[])
 set(gca,'xcolor',[1 1 1])
 set(gca,'xtick',[])
 set(gca,'TickLength',[0 0]);
-set(gca,'linewidth',1);
+%set(gca,'linewidth',1);
 xlim([-40 1020])
 ylim([-13 15])
 for k=0:100:1000
[…]

### 2014-05-30 Paper1/paper1.tex
Updated for time-varying fit

@@
 The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach.
+A scalar multiple of feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation}.
 Segment lengths were measured directly while gross body mass was self-reported.
@@
 When disturbances occurred late in reaching, intent continued onto the target (Figure 3 lower plots).
-We observed no significant change of intent during force disturbance relative to undisturbed reaching.
-Despite statistical insignificance, intention showed some systematic deflection opposite the direction of disturbance beginning around 275 milliseconds after the onset of force disturbance.
-While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to over-correct and is altered for almost 900 milliseconds (Figure 4 shaded areas).
+We observed no significant change of intent during force disturbance until 315 milliseconds after the onset of disturbance.
+Despite statistical insignificance, intention showed some systematic deflection toward the direction of disturbance beginning around 275 milliseconds after the onset of force disturbance.
+While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for almost 900 milliseconds (Figure 4 shaded areas).
 

### 2014-05-30 matlab/armdynamicsInvertedBurdet.m

@@ -1,6 +1,6 @@
 function [dqi,torque_outside]=armdynamicsInvertedBurdet(t,qi)
 
-global measuredVals measuredTime kp0gain kp1gain
+global measuredVals measuredTime tFit kpFit
 
 lqi=length(qi);
 
@@ -21,7 +21,9 @@ omega_desired=qi(lqi/2+1:end);
 kp0=[10.8 2.83; 2.51 8.67];
 joint_torques=abs(D_real*alpha_real+C_real+torque_outside);
 kp1=[3.18 2.15; 2.34 6.18];
-kp=kp0gain*kp0+kp1gain*kp1*diag(joint_torques);
+kp=kp0+kp1*diag(joint_torques);
+
+kp=kp*twoNearestNeighbor(kpFit,tFit,t);
 
 torque_fb=kp*((theta_real-theta_desired) + (1/12)*(omega_real-omega_desired));
 

### 2014-05-30 matlab/armdynamicsInvertedBurdetReflexes.m

@@ -1,6 +1,6 @@
 function [dqi,torque_fb,torque_inertial]=armdynamicsInvertedBurdetReflexes(t,qi)
 
-global measuredVals measuredTime errorVals errorTime kp0gain kp1gain
+global measuredVals measuredTime errorVals errorTime tFit kpFit
 
 interped=twoNearestNeighbor(measuredVals,measuredTime,t);
 theta_real=interped(1:2)';
@@ -23,7 +23,9 @@ omega_desired=qi(3:4);
 kp0=[10.8 2.83; 2.51 8.67];
 joint_torques=abs(D_real*alpha_real+C_real+torque_outside);
 kp1=[3.18 2.15; 2.34 6.18];
-kp=kp0gain*kp0+kp1gain*kp1*diag(joint_torques);
+kp=kp0+kp1*diag(joint_torques);
+
+kp=kp*twoNearestNeighbor(kpFit,tFit,t);
 
 torque_fb=kp*((theta_real-theta_desired) + (1/12)*(omega_real-omega_desired))+(kp/50)*(reflexE+2*reflexV);
 

### 2014-05-30 matlab/extractPulses.m

@@ -1,6 +1,6 @@
 function extractPulses(k)
 
-global kp0gain kp1gain
+global tFit kpFit
 
 figure(k)
 clf
@@ -8,14 +8,18 @@ hold on
 
 load(['../Data/Data_pulse/pulse',num2str(k),'.mat'])
 load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
-kp0gain=W(end,1);
-kp1gain=W(end,1);
-set2dGlobals(params.l1,params.l2,params.origin,params.shoulder,params.mass)
 
-offsetForce=onset;
+W(end,1)
+lowerlim=0;
+Wd(Wd<lowerlim)=lowerlim;
+
+set2dGlobals(params.l1,params.l2,params.origin,params.shoulder,params.mass)
 
 dcats=[trials.disturbcat];
 F=find((dcats>0)&(dcats<5)); %Just like title implies, only the pulses
+
+avemeK=zeros(length(F),200);
+avemeS=zeros(length(F),200);
 
 for c=1:length(F)
     c/length(F)
@@ -28,23 +32,48 @@ for c=1:length(F)
     xvaf2=[trials(kk+1).x trials(kk+1).v trials(kk+1).a trials(kk+1).f];
     
     t=[trials(kk).t(start:end); trials(kk+1).t(1:onset2)]';
+    t=t(1:min(length(t),trialInfo(kk).forceinds(1)+201));
     xvaf=[xvaf1(start:end,:); xvaf2(1:onset2,:)];
-    y=extract(t,xvaf,'reflex');
-    trials(kk).y=y;
+    
+    tFit=trials(kk).t(trialInfo(kk).forceinds(1))+(0:.005:.005*length(Wd));
+    kpFit=Wd;
+    yK=extract(t,xvaf,'reflex');
+    trials(kk).yK=yK;
     trials(kk).ty=t;
     
-    first=trialInfo(kk).forceinds(1);
-    tp=[trials(kk).t(first:end); trials(kk+1).t(1:onset2)]';
-    xvafp=[xvaf1(first:end,:); xvaf2(1:onset2,:)];
-    yp=extract(tp,xvafp,'reflex');
-    trials(kk).yp=yp;
-    trials(kk).typ=tp;
+    tFit=1;
+    kpFit=.8*W(end,1);
[…]

### 2014-05-30 matlab/finalfig3.m

@@ -1,16 +1,16 @@
 clc
 clear all
 
-anecdote(1).sub=4;
-anecdote(1).SE=[3 1];
+anecdote(1).sub=1;
+anecdote(1).SE=[1 3];
 anecdote(1).dcat=2;
 anecdote(1).rot=0;
 anecdote(1).yoff=0;
-%anecdote(1).sub=1;
-%anecdote(1).SE=[1 3];
-%anecdote(1).dcat=2;
-%anecdote(1).rot=0;
-%anecdote(1).yoff=0;
+% anecdote(1).sub=1;
+% anecdote(1).SE=[1 3];
+% anecdote(1).dcat=2;
+% anecdote(1).rot=0;
+% anecdote(1).yoff=0;
 anecdote(2).sub=4;
 anecdote(2).SE=[3 1];
 anecdote(2).dcat=4;

### 2014-05-30 matlab/findKpGain.m

@@ -125,8 +125,8 @@ set2dGlobals(params.l1, params.l2, params.origin, params.shoulder, params.mass)
 figure(5)
 clf
 hold on
-fitlower=0;
-fitupper=.25;
+fitlower=.0;
+fitupper=.5;
 for U=1:length(urc)
     f=find((dcats>0)&(dcats<5)&(reachcat'==(urc(U))));
     catx=[catme(U,:).x]';
@@ -172,17 +172,27 @@ figure(2666)
 clf
 hold on
 for kk=1:length(time)
-    plot(Eb(kk),Ts(kk),'.','markersize',1,'color',[.3,time(kk)/.25,0])
+    plot(Eb(kk),Ts(kk),'.','markersize',1,'color',[.3,kk/length(time),0])
 end
-for kk=0:.01:fitupper
-    X=Eb(time<=kk);
-    Y=Ts(time<=kk);
-    W=X\Y;
+
+spans=fitlower:.005:fitupper;
+Wd=zeros(length(spans),1);
+for kk=1:length(spans)-1
+    inds=(time>=spans(kk))&(time<=spans(kk+1));
+    X=Eb(inds);
+    Y=Ts(inds);
+    Wd(kk)=X\Y;
     minX=min(X);
     maxX=max(X);
-    plot([minX,maxX],W*[minX,maxX],'-','color',[0 kk/fitupper 1])
-    text(maxX,W*maxX,num2str(kk))
+    plot([minX,maxX],Wd(kk)*[minX,maxX],'-','color',[0 kk/length(spans) 1])
+    %text(maxX,Wd(kk)*maxX,[num2str(kk),' ',num2str(Wd(kk),2)])
 end
+Wd(end)=Wd(end-1);
+
+figure(27)
+clf
+plot(spans,Wd)
+drawnow
 
 nT=ceil(log2(length(u)));
 W=zeros(nT+2,3);
@@ -205,12 +215,12 @@ clf
 subplot(2,1,1)
 plot(-1:U,W(:,1),'-')
 ylabel('Fit Feedback Gain, unitless')
-set(gca,'xticklabels',[])
+set(gca,'xticklabel',[])
 subplot(2,1,2)
 plot(-1:U,W(:,2),'-b',-1:U,W(:,3),'-r')
 ylabel('Mean Error, Nm')
 xlabel('Trials Used to Determine Weight')
-set(gca,'xticklabels',labels)
+set(gca,'xticklabel',labels)
 
 baselineCatme=catme;
[…]

### 2014-05-30 matlab/getR2.m

@@ -1,3 +1,5 @@
-function out=getMaxPerp(X)
+function out=getR2(X)
 
-out=max(abs(X(:,2)-.5))*100;
+x=(X(:,2)-.5)*100;
+
+out=mean(x);

### 2014-06-11 Paper1/paper1.tex
Added stiffness derivation and what appear to be minor tweaks.

@@
 {\Large
-\textbf{Determining the Intent of Action Despite Disturbances}
+\textbf{I meant to do that: Extracting the intentions of action in the face of disturbances}
 }
@@
 We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
-Popular modeling approaches can be inverted in a general approach, and we show here the strengths, limitations, and underlying assumptions.
+Popular feedforward control models     approaches can be inverted in a general approach, and we show here the strengths, limitations, and underlying assumptions.
 We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen in human targeted reaching.
@@
 As expected, the intended trajectory showed no change from undisturbed reaching for 150 ms after the disturbance onset.
-Beyond this point, intent change, arm stiffening, or both usually lead to deviation from undisturbed reaching.
-Not all disturbed reaches showed evidence of change in intent.
-Unlike $\lambda$ models, the reaching intent determined by our model is physiologically plausible.
-This success allows real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals.
-In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent.
-While not explicitly examined, this approach is also valid for determining intent in more complex situations such as human-machine interaction, team sports, or national security.
+Beyond this point, intent change, arm stiffening, or both usually lead to deviation from undisturbed reaching after an average of 300 milliseconds after the onset of the disturbance, even when the intention was near the target.
+Unlike the well-known $\lambda$ models, this intent determined by considering a both feedforward (predictive) and feedback controls.
+Knowing the intent along with the action should allow real-time determination of error, making it possible to enhance error, and also directly inspect a neurally-impaired patient's intent for deficits.
+This approach may also be scaled for determining intent in complex systems of team sports, crowds, or swarms.
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLoS ONE authors please skip this step. 
+% Author Summary not valid for PLoS ONE submissions.   
+\section*{Author Summary}
+In our everyday struggles with bumpy roads and bustling crowds we often make movements in ways other than our plan.
+Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do.
+We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model.
+Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight.
+Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance.
+This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people.
+It is also a new tool for understanding and treating impaired movement plans in neural injury.
 
@@
 As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology.
-One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal.
+One can assume that a person who is familiar with a particular task will try to do the same thing each time, provided all other conditions are equal.
 This presumes that what people do is on average what they intended to do and that their goal remains unchanged.
@@
 While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves.
-Particularly of interest is the intended course of action when the actual movement is disturbed and hence no longer matches the intent.
+Particularly of interest is the intended course of action, even when the actual movement is disturbed and hence no longer matches the intent.
 
@@
 Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis.
-Signed error relative to the direction of disturbance, also called perpendicular deviation, was used to measure reaching accuracy in order to elucidate direction of any corrections.
+Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.
 
@@
 \subsection*{Statistical Analysis}
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences.
-Nonparametric testing was necessary due to differences in the shapes of the force-disturbed and undisturbed hand error distributions.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
 
@@
 Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
-Indirect sensitivity, which reveals how errors from other parameters are affected by changes in a given parameter, was an order of magnitude lower than direct sensitivity and again implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
+Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
 Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.
@@
 This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
-As demonstrated in Figure 4, the method passes this test.
+We test this in Figure 4.
+
+Following disturbance, we compare deviation in the hand and intended paths against the maximum deviation of undisturbed movements in order to determine at what point in time changes in hand path or intent can no longer be explained by normal variation and inaccuracy.
+Subjects' force-disturbed hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance.
+By contrast, intent did not change significantly for at least 300 milliseconds after the onset of disturbance (Table 2).
+This passes the test described above.
+While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this is not significant.
+An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.
 
@@
 Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting.
-More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
+More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
 Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
@@
 We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
-We also the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
+We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
 
@@
 \caption{
-{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching.} Hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ simultaneous 95\% confidence interval on the mean) begin to deviate from undisturbed hand deviation (green band) beginning 95 ms after the onset of disturbance.
-Movement intention (red dots, band) does not significantly deviate.
-The hand rejoins the intention approximately 500ms after the onset of disturbance.
+{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ half of the critical value for simultaneous comparisons at 95\% confidence interval) can no longer be explained by undisturbed hand deviation (green band is mean maximum deviation).
+Movement intention (red dots, band is mean deviation) does not significantly deviate until at least 300 ms after the onset of disturbance.
 }
@@
 0.15 &
-0.0 &
+0 &
 0.15 \\
@@
 
+\begin{table}[!ht]
+\caption{
+\bf{Time Elapsed until Error in Undisturbed and Disturbed Reaching are Significantly Different}}
+\begin{tabular}{|c|c|c|}
+\hline
+Subject Number &
+Hand Change (milliseconds) &
+Intent Change (milliseconds) \\ \hline
+1 &
+105 &
+305 \\
+2 &
+105 &
+405 \\
+3 &
+135 &
+350 \\
+4 &
+105 &
+300 \\ \hline
+\end{tabular}
+\begin{flushleft} ANOVA with Tukey-Kramer post-\textit{hoc} test ($p=0.05$) was used to simultaneously compare undisturbed hand maximum deviations to force-disturbed hand and intent deviations at 5 millisecond intervals.
+When  a significant difference was first detected, this was noted as the onset of change.
+\end{flushleft}
+\label{tab:onsets}
+\end{table}
+
 \end{document}

### 2014-06-11 Paper2/derivingStiffness.tex
Added stiffness derivation and what appear to be minor tweaks.

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}
\usepackage{color}

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\pagestyle{myheadings}

\begin{document}

% Title must be 150 characters or less
Define terms: $x$ is hand, $y$ is intent. Capital $F$, $X$ and $Y$ are Laplace transforms. $M$, $B$, and $K$ are mass, damping, and stiffness. $f$ is some externally applied force. Hats ( $\hat{ }$ ) are estimates from an internal model.
\begin{equation}
M\ddot{x}+B\dot{x}+Kx+f=\hat{M}\ddot{y}+\hat{B}\dot{y}+\hat{K}y
\end{equation}
Drop the hats because mass and damping are more or less correctly estimated and we're going to re-represent stiffness misestimation later. Also, switch to the Laplace domain:
\begin{equation}
(Ms^2+Bs+K)X+F=(Ms^2+Bs+K)Y
\end{equation}  
\begin{equation}
F=(Ms^2+Bs+K)(Y-X)
\end{equation}  
Define the quantity, $E=Y-X$, which is essentially error:
\begin{equation}
F=(Ms^2+Bs+K)E
\end{equation} 
Since X and F are presumably matters of record, Y and the arm parameters are the only unknowns, so we can extract. Now, extract with some stiffness estimation error, $\epsilon$. This means that the stiffness used in the extraction, $K_e=K+\epsilon$.
\begin{equation}
F=(Ms^2+Bs+K+\epsilon)E_e
\end{equation}
So the extracted error, $E_e$ is \textit{not} the same as $E$ because of $\epsilon$. But they're related through $F$:
\begin{equation}
(Ms^2+Bs+K)E=(Ms^2+Bs+K+\epsilon)E_e
\end{equation} 
\begin{equation}
(Ms^2+Bs+K)(E-E_e)=\epsilon E_e
\end{equation} 
\begin{equation}
\frac{E-E_e}{E_e}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
Generally speaking, this next equation is probably a fair place to stop:
\begin{equation}
\frac{E}{E_e}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}
But we can re-write it a bit:
\begin{equation}
\frac{Y-X}{Y_e-X}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
\begin{equation}
Y-Y_e=\frac{\epsilon}{Ms^2+Bs+K}(Y_e-X)
\end{equation}
\begin{equation}
\frac{Y-Y_e}{Y_e-X}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}
Stiffness misestimation creates a transient in the extraction whose frequency response is entirely defined by the true dynamics of the system. Because we know $E_e$, the sum of $\epsilon$ and $K$, we have two unknowns: $Y$ and $\epsilon$. A difference of two different extractions using equation 11 does not contain this term:
\begin{equation}
(Y-Y_{1e})-(Y-Y_{2e})=(\frac{\epsilon_1}{Ms^2+Bs+K}(Y_{1e}-X))-(\frac{\epsilon_2}{Ms^2+Bs+K}(Y_{2e}-X))
\end{equation}
\begin{equation}
Y_{2e}-Y_{1e}=\frac{\epsilon_1 Y_{1e}-\epsilon_2 Y_{2e}+(\epsilon_2-\epsilon_1)X}{Ms^2+Bs+K}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- \epsilon_2 \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because we know the difference in $K_e$ values used for the two extractions ($K_{1e}$ and $K_{2e}$), we know the difference in $\epsilon$ values already.
\begin{equation}
Y_{2e}-Y_{1e}=\frac{\epsilon_1 Y_{1e}-\epsilon_2 Y_{2e}+(\epsilon_2-\epsilon_1)X}{Ms^2+Bs+K}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- (\epsilon_1+K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because $Y_{1e}$, $Y_{2e}$, and $X$ are known, we can solve for our unknown. In order to deal with the dynamics in the denominators, a slightly lagged value of $K$ should be fine (1 samples = 1 ms, $K_{next}=K_{now}-\epsilon_1$). This implies that dynamic tracking of stiffness should be possible in real time. Convergence properties remain unexamined.
\end{document}

### 2014-06-11 matlab/final_fig45_combined.m

@@ -9,10 +9,11 @@ close all
 %Plot mean and confidence interval against range
 %call that finalfig5
 
-SUBS=2;
+SUBS=1:4;
+distCats=[3 4];
 
 NR=10;
-edges=[0 linspace(0,200,NR+1)]; %sampling rate is 200 Hz = .005 s samples
+edges=[0 linspace(0,200,NR+1) 200]; %sampling rate is 200 Hz = .005 s samples
 
 black=[0 0 0];
 red=[1 .3 .3];
@@ -22,6 +23,9 @@ height=1.618*width;
 lmargin=.5;
 bmargin=0;
 figmargin=.4;
+
+yoffSpan=8;
+plotOrder=[1 4 2 3];
 
 figure(1)
 clf
@@ -33,29 +37,23 @@ set(gcf,'position',[4,8,figmargin+lmargin+width,figmargin+bmargin+height])
 for k=SUBS
     load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
     load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
-    load(['../Data/Data_pulse/pulse',num2str(k),'U.mat'])
-
-    flag=0;
 
     clear cleanStruct
     f=find([trialInfo.clean]);
     for t=2:length(f)
-        lt=length(trials(f(t)).t);
-        lty=length(undisturbed(f(t)).ty);
-        inds=(1+lt-lty):lt; %Onset of movement
-        cleanStruct(t).R2BX=getR2(trials(f(t)).x(inds,:));
-        cleanStruct(t).R2BY=getR2(undisturbed(f(t)).y);
+        X=trials(f(t)).x(:,2)-.5;
+        [ma,indX]=max(abs(X));
+        cleanStruct(t).R2BX=X(indX)*100;
     end
     subject(k).R2BX=vertcat(cleanStruct.R2BX);
-    subject(k).R2BY=vertcat(cleanStruct.R2BY);
 
     clear triStruct
 
-    for R=1:NR+1
+    for R=1:NR+2
         RANGE=edges(R):edges(R+1);
-        
+
         for t=1:length(trials)
-            if ~sum(trials(t).disturbcat==[1:4]) %Fast-ish skipping instead of indexing
+            if ~sum(trials(t).disturbcat==[distCats]) %Fast-ish skipping instead of indexing
                 continue
             end
             inds=trialInfo(t).forceinds(1)+RANGE;
[…]

### 2014-06-11 matlab/finalfig3.m

@@ -11,7 +11,7 @@ anecdote(1).yoff=0;
 % anecdote(1).dcat=2;
 % anecdote(1).rot=0;
 % anecdote(1).yoff=0;
-anecdote(2).sub=4;
+anecdote(2).sub=1;
 anecdote(2).SE=[3 1];
 anecdote(2).dcat=4;
 anecdote(2).rot=pi;

### 2014-06-11 matlab/halfWidth.m

@@ -4,13 +4,13 @@ function out = halfWidth(stats,varargin)
 gmeans = stats.means(:);
 n = stats.n(:);
 s = stats.s;
-ng = sum(n>0);
+ng = sum(n>0)-1;
 df = stats.df;
 
 if nargin>1
     alpha=varargin{1};
 else
-    alpha=.01;
+    alpha=.1;
 end
 
 ctype='tukey-kramer';

### 2014-06-12 Paper1/paper1.tex
Shouldn't fig5 be deleted?

@@
 \section*{Abstract}
-We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data.
-Popular feedforward control models     approaches can be inverted in a general approach, and we show here the strengths, limitations, and underlying assumptions.
-We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen in human targeted reaching.
+We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test the approach using human reaching data.
+We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions.
+We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen when humans reach to visual targets.
 Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness.
-We also tested the robustness of this approach on human point-to-point reaching movements disturbed by large force pulses.
-As expected, the intended trajectory showed no change from undisturbed reaching for 150 ms after the disturbance onset.
-Beyond this point, intent change, arm stiffening, or both usually lead to deviation from undisturbed reaching after an average of 300 milliseconds after the onset of the disturbance, even when the intention was near the target.
-Unlike the well-known $\lambda$ models, this intent determined by considering a both feedforward (predictive) and feedback controls.
-Knowing the intent along with the action should allow real-time determination of error, making it possible to enhance error, and also directly inspect a neurally-impaired patient's intent for deficits.
-This approach may also be scaled for determining intent in complex systems of team sports, crowds, or swarms.
+We also tested the robustness of this approach on human point-to-point reaching movements disturbed by force pulses.
+As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset.
+Beyond this point, we detected change in intent even when the intention was near the target.
+Unlike well-known $\lambda$ models, this intent was determined by considering both feedforward (predictive) and feedback controls.
+Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits.
+This approach may also be scaled for determining intent during human-machine interaction and in complex systems like team sports, crowds, or swarms.
 
@@
 \section*{Author Summary}
-In our everyday struggles with bumpy roads and bustling crowds we often make movements in ways other than our plan.
+In our everyday struggles with bumpy roads and bustling crowds we often make movements that do not match what we had planned.
 Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do.
@@
 \section*{Introduction}
-As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology.
+There are many situations, such as disturbances, distractions, and pathologies, that interfere with execution of an intended movement.
 One can assume that a person who is familiar with a particular task will try to do the same thing each time, provided all other conditions are equal.
@@
 By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the so-called $\lambda$ model\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
-Upon inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
+Supporters of the "$\lambda$ model"\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
+Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
 Therefore, it could not well-represent the intent of a simple reaching movement.
@@
 In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling.
-The underlying framework should be extensible to whole-body movement or even more abstract, psychological processes.
+The mathematics underlying intent determination are sufficiently general to be extensible to whole-body movement or even more abstract, psychological processes.
 
-Collectively, the general problem of being able to best extract someone's intent is critical to many fields, and hinges on our understanding of the underlying mechanisms through which control is accomplished.
-We have excellent models of the mechanical properties of arm and muscle because the physics that describe their gross behavior is so well-understood and the need to determine particular physical parameter values has long been clear.
-This wealth of previous work provides strong footing for our technique, which in turn should yield valuable information not previously available by other means.
-In fields where the process and actuators are not yet well-understood, this approach may help to structure the search for them.
-
-Here we develop and analyze a method to extract intent from disturbed movement, which will allow us to observe if and how intent changes following disturbance.
+Here we develop and analyze a method to extract intent from disturbed movement.
+This method then allows us to observe if and how intent changes following disturbance.
 We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction.
@@
 We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching.
-We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for a providing a better understanding of what a person means to do when they move and are disturbed.
+We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed.
 
@@
 \section*{Methods}
-The sections below describe the theoretical method, and then present our evaluations.
+The sections below describe the theoretical method and the means by which we evaluate it.
 First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process.
-Next we conducted an experiment on humans and the process was then evaluated on human subject data.
+Next we conducted an experiment on humans to evaluate the process using human subject data.
 
@@
 Additionally, linear feedback rejects position and velocity error.
-This model is chosen to be very simple to help illuminate the approach, but it is limited in how it captures all of the complexity of motor control.
-Nevertheless later we will discuss how this might be used on more complex models without loss of generality.
+This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches.
+Nevertheless later we will discuss how this approach can be used on more complex models without loss of generality.
 The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
@@
 \end{equation}
-Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripetal effects, and $E$ is externally-applied torque that the system might interact with.
-Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles.
+Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque.
 The motion behavior changes with the addition of feedforward and/or feedback controllers:
@@
 \end{equation}
-Where a hat ( $\hat{ }$ ) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model.
-This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm.
+Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{kawato1999internal}.
+This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance.
 If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero.
@@
 
-For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation requires standard numerical integration to determine the solution to the initial value problem in time.
+For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation is typically solved by numerical integration to determine the solution to the initial value problem in time.
 This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory.
-The intent extraction approach modifies this method by instead solving for $\ddot{q}_d$:
+Intent determination takes the novel approach of instead solving for $\ddot{q}_d$:
 \begin{equation}
@@
 $\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$.
-Provided the model of plant and controller are precise, the initial conditions are available and accurate, the mass matrix estimate $\hat{M}$ is invertible, and the constants are accurate and do not vary, and externally-applied force can be accurately measured, then the system yields an accurate estimate of the intent.
+This relies on many assumptions: The model of plant and controller must be accurate and precise.
+The initial conditions must be available and accurate.
+The mass matrix estimate, $\hat{M}$, must be invertible.
+Externally-applied force must be precisely and accurately measured.
+If all of these conditions are met, then the system yields an accurate estimate of the intent.
 
-It is important to distinguish our intended trajectory, $x_d$, from another variable often considered, the equilibrium point of the muscle actuators.
+It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators.
 Importantly, this has been claimed to be the intended movement.
 To consider this, we also generalize our approach to any system with a moving intention that governs the control law.
-This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $x_e$ (sometimes represented as $\lambda$).
+This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $q_e$ (sometimes represented as $\lambda$).
 The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
@@
 Note $x_e$ vanishes, recovering our familiar model.
-While a proper choice of $x_e$ is critical to control, our approach does not require a model relating equilibrium and intended trajectory.
+While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory.
 Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero.
@@
 
-While we treat both the process and actuator as dynamic models in our derivation, both the derivation and the technique are agnostic to the process model.
+While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
 Instead of a dynamic model, it could be a lookup table.
@@
 \subsection*{Generation of synthetic data using simple dynamic simulation}
-We choose one prevailing arm model, that of Burdet et al.
-\cite{burdet2006stability}, along with the anatomical mass distribution and geometry parameters of Winter \cite{winter2009biomechanics} to create a model for estimating intent.
-The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach.
-A scalar multiple of feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation}.
-Segment lengths were measured directly while gross body mass was self-reported.
-These values were converted to segment masses, centers of gravity, and moments of inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}.
-The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds).
+While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models.
+Appropriate plant modeling is well-understand: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
+Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
+For the actuator, we choose the model of Burdet et al.
+\cite{burdet2006stability}.
+A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged.
+As a consequence of choosing this model, muscle torque must be used to calculate stiffness even though it arises in part from stiffness.
+Like Burdet et al.
+\cite{burdet2006stability}, we approximated muscle torque by looking backward in time a small amount (500 microseconds).
+This approximation is not necessary during intent extraction, which leads to a small (sub-micrometer mean unsigned error) discrepancy.
 Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.
 
+\subsection*{Experimental Design}
 Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.
@@
 \subsection*{Error Metric}
-Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15centimeters reach.
+Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach.
 Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis.
@@
 Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
-Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values.
+Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range.
 Combinations and calculations were made as prescribed by Saltelli et al.
@@
 \subsection*{Apparatus}
-A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass.
+A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass.
 The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
@@
 \subsection*{Model Robustness}
-Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
+While some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.
+Sensitivity analysis revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
 Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al.
@@
 These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).
-In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.
 
@@
 Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
-Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
+Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
 Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.
@@
 Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3).
-In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.
+In both pulse timing conditions, the intended trajectory remained straight towards the target, supporting the approach.
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred late in reaching, intent continued onto the target (Figure 3 lower plots).
-We observed no significant change of intent during force disturbance until 315 milliseconds after the onset of disturbance.
-Despite statistical insignificance, intention showed some systematic deflection toward the direction of disturbance beginning around 275 milliseconds after the onset of force disturbance.
-While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for almost 900 milliseconds (Figure 4 shaded areas).
+When disturbances occurred late in reaching, intent continued onto the target but then corrected anyways (Figure 3 lower plots).
+We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance.
+Intention showed some systematic deflection toward the direction of disturbance peaking around 150 milliseconds after the onset of force disturbance, but this effect did not leave the range of error found in undisturbed movement.
+While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for as much as one second (Figure 4 shaded areas).
 
-As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison.
+We argue that the approach here represents the first sound method for looking at intent following disturbance, so we have only the undisturbed intention for comparison.
 If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands.
@@
 This passes the test described above.
-While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this is not significant.
+While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this residual disturbance does not exceed the statistical threshold for significance.
 An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.
@@
 While the derivation of intent determination required many steps, the outcome is a filter whose use is straight-forward.
-Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated.
-Demonstrating the validity and plausibility of inverting those filters to determine intent from hand trajectory adds a novel method to the arsenal of analytical methods for exploring human motor control.
+Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated \cite{shadmehr1994adaptive, franklin2003adaptation, burdet2006stability}.
+Demonstrating the validity and plausibility of mathematically transforming those filters to instead covert measured hand trajectories to intent adds a novel method to the arsenal of analytical methods for exploring human motor control.
 
-This modeling approach explains the discovery by Bizzi et al.
+The proposed approach could explain the discovery by Bizzi et al.
 \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$.
@@
 
-Intent is slow to change if it changes at all following disturbance.
+Our analysis suggests that intent is slow to change if it changes at all following disturbance.
 While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement.
@@
 
-Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible.
-It is important to reiterate that because the human arm's mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory.
+This novel determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible.
+This approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory.
 By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.
 Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting.
-Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions?
-Not always.
-Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable.
-Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit hypothesized cost functions.
+Direct access to the intended trajectory allows optimal feedback control models to better understand and fit hypothesized cost functions and their parameters.
+In addition, mid-movement replanning became falsifiable: we were able to show that intended trajectories from disturbed motions differ from those undisturbed motions and how they differed.
 
@@
 Nevertheless, there are some limitations to these methods.
-One limiting requirement is that accurate interface force measurement (within a few tenths of a newton) is needed.
+Accurate interface force measurement (within a few tenths of a newton) is needed.
 Such instrumentation is available, but must be used with great care in order to preserve the accuracy of the estimate.
@@
 
-The approximation of stiffness appears to be the most critical factor to accuracy in estimating intent.
+Our sensitivity analysis showed that approximation of stiffness is the most critical factor to accuracy in estimating intent.
 While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
@@
 Subject Number &
-Hand Change (milliseconds) &
-Intent Change (milliseconds) \\ \hline
+Hand Deviation (milliseconds) &
+Intent Deviation (milliseconds) \\ \hline
 1 &
@@
 \begin{flushleft} ANOVA with Tukey-Kramer post-\textit{hoc} test ($p=0.05$) was used to simultaneously compare undisturbed hand maximum deviations to force-disturbed hand and intent deviations at 5 millisecond intervals.
-When  a significant difference was first detected, this was noted as the onset of change.
+When  a significant difference was first detected, this was noted as the onset of deviation.
 \end{flushleft}

### 2014-06-16 Paper2/derivingStiffness.tex
Deal with intern stuff tomorrow

@@
 \end{equation} 
-Since X and F are presumably matters of record, Y and the arm parameters are the only unknowns, so we can extract.
-Now, extract with some stiffness estimation error, $\epsilon$.
+Since $M$, $B$, $X$ and $F$ are presumably matters of record, $Y$ and $K$ are the only unknowns.
+We can extract, but we do so with some as yet unknown stiffness estimation error, $\epsilon$.
 This means that the stiffness used in the extraction, $K_e=K+\epsilon$.
@@
 Stiffness misestimation creates a transient in the extraction whose frequency response is entirely defined by the true dynamics of the system.
-Because we know $E_e$, the sum of $\epsilon$ and $K$, we have two unknowns: $Y$ and $\epsilon$.
-A difference of two different extractions using equation 11 does not contain this term:
+Because we know the value of $K$ used for each extraction, $K_e=\epsilon+K$, we have two unknowns: $Y$ and $\epsilon$.
+A difference of two different extractions using equation 11 does not contain the $Y$ term:
 \begin{equation}
-(Y-Y_{1e})-(Y-Y_{2e})=(\frac{\epsilon_1}{Ms^2+Bs+K}(Y_{1e}-X))-(\frac{\epsilon_2}{Ms^2+Bs+K}(Y_{2e}-X))
-\end{equation}
-\begin{equation}
-Y_{2e}-Y_{1e}=\frac{\epsilon_1 Y_{1e}-\epsilon_2 Y_{2e}+(\epsilon_2-\epsilon_1)X}{Ms^2+Bs+K}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- \epsilon_2 \frac{Y_{2e}-X}{Ms^2+Bs+K}
+(Y-Y_{1e})-(Y-Y_{2e})=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- \epsilon_2 \frac{Y_{2e}-X}{Ms^2+Bs+K}
 \end{equation}
@@
 \begin{equation}
-Y_{2e}-Y_{1e}=\frac{\epsilon_1 Y_{1e}-\epsilon_2 Y_{2e}+(\epsilon_2-\epsilon_1)X}{Ms^2+Bs+K}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- (\epsilon_1+K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
+Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}-(\epsilon_1+K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
+\end{equation}
+\begin{equation}
+Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-Y_{2e}}{Ms^2+Bs+K}-(K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
 \end{equation}
 Because $Y_{1e}$, $Y_{2e}$, and $X$ are known, we can solve for our unknown.
-In order to deal with the dynamics in the denominators, a slightly lagged value of $K$ should be fine (1 samples = 1 ms, $K_{next}=K_{now}-\epsilon_1$).
+In order to deal with the dynamics in the denominators, a slightly lagged value of $K$ should be fine (1 sample = 1 ms delay, $K_{next}=K_{now}-\epsilon_1$).
 This implies that dynamic tracking of stiffness should be possible in real time.
-Convergence properties remain unexamined.
+Convergence properties remain unexamined but this might just start and stay at your initial guess (and probably will).
+\begin{equation}
+(Ms^2+Bs+K)(Y_{2e}-Y_{1e})=\epsilon_1 (Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs)(Y_{2e}-Y_{1e})=(K+\epsilon_1)(Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K+\epsilon_1)(Y_{2e}-Y_{1e})=(K_{2e}-K_{1e})(X-Y_{2e})
+\end{equation}
+\begin{equation}
+\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
+\end{equation}
+\begin{equation}
+Y_{1e}-Y_{2e}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)
+\end{equation}
+Back to zero unknowns, which basically dooms equation 15. Notice that this is equivalent to eq 12.
+\begin{equation}
+\frac{Y-Y_{2e}}{Y_{2e}-X}-\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
+\end{equation}
+\begin{equation}
+\frac{Y-Y_{1e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
+\end{equation}
+
+\newpage
+What if you know that $Y=720/s^6-360/s^5+60/s^4$ because of minimum jerk submovements?
+\begin{equation}
+Y_e=720/s^6-360/s^5+60/s^4-\frac{K_e-K}{Ms^2+Bs+K_e}(Y_e-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_e)Y_e=720(M/s^4+B/s^5+K_e/s^6)-360(M/s^3+B/s^4+K/s^5)+60(M/s^2+B/s^3+K/s^4)+(K-K_e)(Y_e-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_e)Y_e=720K_e/s^6+(720K-(M/s^4+B/s^5+K_e/s^6)-360(M/s^3+B/s^4+K/s^5)+60(M/s^2+B/s^3+K/s^4)+(K-K_e)(Y_e-X)
+\end{equation}
+
+\newpage
+Show really quickly why you can't solve eq 16 for K:
+\begin{equation}
+(Ms^2+Bs+K)(Y_{2e}-Y_{1e})=(K_{1e}-K) (Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs)(Y_{2e}-Y_{1e})+K(Y_{2e}-Y_{1e})+K(Y_{1e}-Y_{2e})=K_{1e}(Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs)(Y_{2e}-Y_{1e})=K_{1e}(Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+As seen in eq 19, K actually vanishes.
+Because extractions are linearly dependent on one another, sums and differences contain no information you don't already have.
+Differences wipe out your two unknowns while sums contain them.
+There's an interest sum involving equal and opposite $\epsilon$ though.
+Snagging eq 11:
+\begin{equation}
+Y-Y_e=\frac{\epsilon}{Ms^2+Bs+K}(Y_e-X)
+\end{equation}
+\begin{equation}
+Y_e=Y-\frac{\epsilon}{Ms^2+Bs+K}(Y_e-X)
+\end{equation}
+\begin{equation}
+Y_1+Y_2=2Y-\frac{\epsilon_1}{Ms^2+Bs+K}(Y_1-X)-\frac{\epsilon_2}{Ms^2+Bs+K}(Y_2-X)
+\end{equation}
+Choose $\epsilon_1=-\epsilon_2$:
+\begin{equation}
+Y_1+Y_2=2Y+\frac{\epsilon_1}{Ms^2+Bs+K}(Y_2-X)-\frac{\epsilon_1}{Ms^2+Bs+K}(Y_1-X)
+\end{equation}
+\begin{equation}
+Y_1+Y_2=2Y+\epsilon_1\frac{(Y_2-X)-(Y_1-X)}{Ms^2+Bs+K}
+\end{equation}
+\begin{equation}
+Y_1+Y_2=2Y+\epsilon_1\frac{Y_2-Y_1}{Ms^2+Bs+K}
+\end{equation}
+\begin{equation}
+Y_1+Y_2=2Y+\epsilon_1((K_{1e}-K) (Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X))
+\end{equation}
+
+\newpage
+Jumping off from the difference of two extractions containing zero unknowns, let's add a new unknown, $\alpha$.
+So the setup we want is $Y_{1e}-\alpha(Y_{2e}-Y_{1e})=Y+...$ where that ellipsis is a pile of knowns and $\alpha$.
+Because this is a fudge factor and not a physical parameter, it might allow us to not make weird assumptions about $Y$.
+\begin{equation}
+Y_{1e}-Y_{2e}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+Y_{1e}=Y-\frac{K_{1e}-K}{Ms^2+Bs+K_{1e}}(Y_{1e}-X)
+\end{equation}
+\begin{equation}
+Y_{1e}-\alpha(Y_{1e}-Y_{2e})=Y-\frac{K_{1e}-K}{Ms^2+Bs+K_{1e}}(Y_{1e}-X)-\alpha\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_{1e})(Y_{1e}-\alpha(Y_{1e}-Y_{2e})-Y)=-(K_{1e}-K)(Y_{1e}-X)-\alpha(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+Going off on a tangent to entertain $\alpha=1$, what happens?
+\begin{equation}
+(Ms^2+Bs+K_{1e})(Y_{2e}-Y)=-(K_{1e}-K)(Y_{1e}-X)-(K_{2e}-K_{1e})(Y_{2e}-X)
+\end{equation}
+It looks like we've achieved a situation where $Y$ and $K$ are not multiplying one another.
+\begin{equation}
+(Ms^2+Bs+K_{1e})Y_{2e}+(K_{2e}-K_{1e})(Y_{2e}-X)+K_{1e}(Y_{1e}-X)=(Y_{1e}-X)K+(Ms^2+Bs+K_{1e})Y
+\end{equation}
+So this is exactly what we've been looking for: knowns and constants on the left, $K$ and $Y$ on the right and multiplied only by knowns.
+
+I messed up the math, try this route again later.
+
+\newpage
+\begin{equation}
+Y=Y_e+\frac{\epsilon}{Ms^2-Bs+K_e-\epsilon}(Y_e-X)
+\end{equation} 
+Three equations and three unknowns $Y(t=-1)$, $Y(t=1)$, $\epsilon$
+\begin{subequations}
+\begin{align}
+Y(t=-1)&=Y_e(t=-1)+\frac{\epsilon}{Ms^2-Bs+K_e-\epsilon}(Y_e(t=-1)-X(t=-1)) \\
+\frac{Y(t=1)+Y(t=-1)}{2}&=Y_e(t=0)+\frac{\epsilon}{Ms^2-Bs+K_e-\epsilon}(Y_e(t=0)-X(t=0)) \\
+Y(t=1)&=Y_e(t=1)+\frac{\epsilon}{Ms^2-Bs+K_e-\epsilon}(Y_e(t=1)-X(t=1))
+\end{align}
+\end{subequations}
+
+That's three equations in three unknowns, but what if you go with two instead?
+\begin{equation}
+Y=Y_1+\frac{\epsilon}{Ms^2-Bs+K_e-\epsilon}(Y_e-X)
+\end{equation}
+
+\newpage
+
+\begin{equation}
+F=(Ms^2+Bs+K_e-\epsilon)(Y-X)
+\end{equation}
+\begin{equation}
+F=(Ms^2+Bs+K_e)(Y_e-X)
+\end{equation}
+
+\newpage
+\textbf{Tactic: why does some weighted difference of three extractions work?}
+
+Quickly define $D=Ms^2+Bs+K$ and take the difference of three extractions.
+\begin{equation}
+(Y-X)-((Y-X)-(Y-X))=Y_1-X+\frac{\epsilon_1}{D}-((Y_2-X+\frac{\epsilon_2}{D})-(Y_3-X+\frac{\epsilon_3}{D}))
+\end{equation} 
+\begin{equation}
+Y=Y_1-Y_2+Y_3+\frac{\epsilon_1(Y_1-X)-\epsilon_2(Y_2-X)+\epsilon_3(Y_3-X)}{D}
+\end{equation} 
+\begin{equation}
+Y=Y_1-Y_2+Y_3+\frac{\epsilon_1(Y_1-X)-(K_2-K_1+\epsilon_1)(Y_2-X)+(K_3-K_1+\epsilon_1)(Y_3-X)}{D}
+\end{equation} 
+\begin{equation}
+(Ms^2+Bs+K_1-\epsilon_1)Y=(Ms^2+Bs+K_1-\epsilon_1)(Y_1-Y_2+Y_3)+\epsilon_1(Y_1-X)-(K_2-K_1+\epsilon_1)(Y_2-X)+(K_3-K_1+\epsilon_1)(Y_3-X)
+\end{equation}
+
+\begin{equation}
+Y=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)+Y_3+\frac{\epsilon_1(Y_1-X)-(K_2-K_1+\epsilon_1)(Y_2-X)+(K_3-K_1+\epsilon_1)(Y_3-X)}{D}
+\end{equation}
+
+\begin{equation}
+Y-Y_e=\frac{\epsilon}{Ms^2+Bs+K_e-\epsilon}(Y_e-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_e-\epsilon)(Y-Y_e)=\epsilon(Y_e-X)
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_e)Y_e-(Ms^2+Bs+K_e-\epsilon)Y=\epsilon X
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_e)(Y_e-Y)=\epsilon(X+Y)
+\end{equation}
+
+\newpage
+\textbf{It wasn't this:}
+
+Maybe it's three equations in three unknowns ($\epsilon_1$, $\epsilon_2$, and $Y$)?
+\begin{subequations}
+\begin{align}
+K_{2e}-K_{1e} &= \epsilon_2 - \epsilon_1 \\
+Y_{1e}&=Y-\frac{\epsilon_1}{Ms^2+Bs+K_{1e}-\epsilon_1}(Y_{1e}-X) \\
+Y_{2e}&=Y-\frac{\epsilon_2}{Ms^2+Bs+K_{2e}-\epsilon_2}(Y_{2e}-X)
+\end{align}
+\end{subequations}
+If so, can we somehow rearrange our way out of this?
+\begin{subequations}
+\begin{align}
+K_{2e}-K_{1e} &= \epsilon_2 - \epsilon_1 \\
+(Ms^2+Bs+K_{1e})Y_{1e}&=(Ms^2+Bs+K_{1e}-\epsilon_1)Y-\epsilon_1 X \\
+(Ms^2+Bs+K_{2e})Y_{2e}&=(Ms^2+Bs+K_{2e}-\epsilon_2)Y-\epsilon_2 X
+\end{align}
+\end{subequations}
+\begin{subequations}
+\begin{align}
+K_{2e}-K_{1e} &= \epsilon_2 - \epsilon_1 \\
+(Ms^2+Bs+K_{1e})Y_{1e}&=(Ms^2+Bs+K_{1e}-\epsilon_1)Y-\epsilon_1 X \\
+(Ms^2+Bs+K_{2e})Y_{2e}&=(Ms^2+Bs+K_{1e}-\epsilon_1)Y-\epsilon_2 X
+\end{align}
+\end{subequations}
+Down to two equations in two unknowns:
+\begin{subequations}
+\begin{align}
+(Ms^2+Bs+K_{1e})Y_{1e}&=(Ms^2+Bs+K_{1e}-\epsilon_1)Y-\epsilon_1 X \\
+(Ms^2+Bs+K_{2e})Y_{2e}&=(Ms^2+Bs+K_{1e}-\epsilon_1)Y-(K_{2e}-K_{1e}+\epsilon_1) X
+\end{align}
+\end{subequations}
+\begin{equation}
+\frac{(Ms^2+Bs+K_{1e})Y_{1e}+\epsilon_1 X}{(Ms^2+Bs+K_{1e}-\epsilon_1)}=Y
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_{2e})Y_{2e}=(Ms^2+Bs+K_{1e}-\epsilon_1)(\frac{(Ms^2+Bs+K_{1e})Y_{1e}+\epsilon_1 X}{(Ms^2+Bs+K_{1e}-\epsilon_1)})-(K_{2e}-K_{1e}+\epsilon_1) X
+\end{equation}
+\begin{equation}
+(Ms^2+Bs+K_{2e})Y_{2e}=(Ms^2+Bs+K_{1e})Y_{1e}-(K_{2e}-K_{1e}) X
+\end{equation}
+
 \end{document}

### 2014-06-16 internStuff/addSubject.m

@@ -1,0 +1,50 @@
+function addSubject(k)
+
+name=num2str(k);
+
+input=load('./input.dat');
+output=load(['./output',name,'.dat']);
+params=getSubjectParams(name);
+
+%outStream << trial TAB now-zero TAB position.X() TAB position.Y() TAB velocity.X() TAB velocity.Y() TAB accel.X() TAB accel.Y() TAB force.X() TAB force.Y() TAB desposition.X() TAB desposition.Y() TAB xpcTime << endl;
+
+traw=output(:,2);
+xvafraw=output(:,3:10);
+trial=output(:,1);
+lastTrial=max(trial)-1;
+f=find(trial==lastTrial,1,'last');
+samprate=1/mean(gradient(traw))
+if samprate<300 %our filtering method apparently has nonlinear algorithmic complexity and goes from seconds to hours+.
+    inds=1:f;
+else
+    inds=1:5:f;
+end
+trials=generalAddSubject(name,traw(inds),xvafraw(inds,:),trial(inds),params);
+
+y=output(:,11:12);
+y=y(inds,:);
+trial=trial(inds);
+
+%Because trial 1 starts from an unknown location/where the robot is
+%turned on
+
+cats=[1 0 2];
+dists=[.15 .3];
+
+[a,b,c]=unique(sum(input(:,2:3),2));
+remap=[1 0 2 3];
+targs=remap(c);
+
+
+for c=1:length(trials)
+    trials(c).disturbance=input(c,6);
+    if c==1
+        trials(c).orig=trials(c).x(1,:);
+    else
+        trials(c).orig=targs(c-1).targ;
+    end
+    trials(c).targ=targs(c);
+    trials(c).y=y(trial==c,:);
+end
+
+save(['./intern',num2str(k),'.mat'],'trials','params')

### 2014-06-16 internStuff/fkinRobot.m

@@ -1,0 +1,13 @@
+function [x,x1]=fkinRobot(theta)
+
+l1=.46251;
+l2=.33521;
+
+x=[l1*cos(theta(1))+l2*cos(theta(1)+theta(2));
+   l1*sin(theta(1))+l2*sin(theta(1)+theta(2))];
+
+if nargout>1
+    x1=[l1*cos(theta(1));
+        l1*sin(theta(1))];
+end
+

### 2014-06-16 internStuff/ikinRobot.m

@@ -1,0 +1,11 @@
+function theta=ikinRobot(x)
+
+%x0=[0, 0]
+l1=.46251;
+l2=.33521;
+
+theta2=-2*atan(sqrt(((l1+l2)^2-(x(1)^2+x(2)^2))/(x(1)^2+x(2)^2-(l1-l2)^2)));  %+/- at the beginning of this is elbow up/down
+theta1=atan2(x(2),x(1))-atan2(l2*sin(theta2),l1+l2*cos(theta2));
+
+
+theta=[theta1; theta2];

### 2014-06-16 internStuff/makeInput.m

@@ -1,0 +1,39 @@
+clc
+clear all
+
+dist=.15;
+
+center=[0 .48];
+angles=(180+[30 150 270])'*pi/180;
+
+targets=[center;center;center]+dist*[cos(angles) sin(angles)];
+
+% #Each direction, force mag, EA gain
+phases=[16 0 1;
+    16 1.2 1;
+    16 1.2 0;
+    16 1.2 1;
+    16 0 1];
+
+q=0;
+for P=1:size(phases,1)
+    %Make a list of #1s, 2s and 3s
+    deck=sort(mod(1:phases(P,1)*3,3)+1);
+    deck=deck(randperm(phases(P,1)*3));
+    for k=1:length(deck)
+        q=q+1;
+        REACHES(q).vals=[q targets(deck(k),1) targets(deck(k),2) 0 0 phases(P,2) -1 1 1 phases(P,3)];
+        q=q+1;
+        REACHES(q).vals=[q center(1) center(2) 0 0 phases(P,2) -1 1 1 phases(P,3)];
+
+    end
+end
+
+reaches=vertcat(REACHES.vals)
+
+fid=fopen('internInput.dat','w');
+[stringy,ermsg]=sprintf('%d\t%g\t%g\t%g\t%g\t%g\t%d\t%d\t%d\t%g\n',reaches')
+fprintf(fid,'%d\t%g\t%g\t%g\t%g\t%g\t%d\t%d\t%d\t%g\n',reaches')
+fclose(fid);
+
+%temptrial,&tempx,&tempy,&tempEarly,&tempLate,&tempWhite,&tempshape,&tempShowCursor, &tempAcquisitionsNeeded)

### 2014-06-16 internStuff/vmlt.m

@@ -1,0 +1,3 @@
+function out=vmlt(in,x)
+
+out=sum(in.^2,2)<(x^2);

### 2014-07-01 Paper1/derivingStiffness.tex

--- 2014-06-16 Paper2/derivingStiffness.tex
@@
 \end{equation}
+
 Stiffness misestimation creates a transient in the extraction whose frequency response is entirely defined by the true dynamics of the system.

### 2014-07-01 Paper2/manuscript/paper2.tex
Started in on paper 2

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Grammar and Vocabulary of Human Reaching Intent}
}
% Insert Author names, affiliations and corresponding author email.
\\
Justin Horowitz$^{1}$, 
James Patton$^{1,2\ast}$
\\
\bf{1} Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States
\\
\bf{2} Sensory Motor Performance Program, Rehabilitation Institute of Chicago, Chicago, Illinois, United States
\\
$\ast$ E-mail: pattonj@uic.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}
blah blah blah

We hypothesize that human reaching intention is built from submovements that constitute a fixed vocabulary and grammar. Were this hypothesis true, we should find it everywhere we look and without exception. We test this hypothesis by comparing the composition of movement in stroke survivors and force-disturbed control subjects whose reaching intent has been recovered by a novel computational technique. In particular, we presume a grammar based on the work of Flash and Henis \cite{flash1991arm} and a vocabulary to be recovered from the stroke survivors whose submovements can be easily recovered given the presumed vocabulary \cite{rohrer2004submovements}. If this grammar and vocabulary are accurate and complete, they should fully explain force-disturbed reaching intent in control subjects.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}
We set out to test the hypothesis that both stroke survivors and neurologically intact subjects exposed to force disturbance share a compact grammar and vocabulary for constructing movement intent. The grammar, or rules for combination, was that submovements were combined by superposition (vector addition) and no more than two were present at any point in time. The vocabulary was the minimum-jerk, $5^{th}$ order polynomial starting and ending at at zero speed and acceleration. This grammar and vocabulary recovered from stroke survivors was able to completely explain reaching intent in neurologically intact subjects despite imposing significant restrictions. While some might argue that these submovements are an artifact of stroke or an artifact of intent extraction, it is extremely unlikely that both situations should share these features unless they're fundamental to human movement. Moreover, we were able to detect predictable changes in arm stiffness that were well-correlated with the onset of submovements. This suggests that reaching intent may have an additional volitional impedance component not detectable in undisturbed reaching.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
When performing intent extraction, we're faced with many significant unknowns. Certain arm parameters might change volitionally while others might diverge from cadaver anatomy as studied in the 1950s. We derive a relationship between extracted intent and parameter misestimation in the hope of finding components of the extraction that relate primarily to estimation error in some parameter value. With the ``signature'' of these errors in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.

We begin by noting that the arm is described by four sets of parameters at any point in time: two in our model ($M_e, B_e, K_e$ and $M_d, B_d, K_d$), one in reality ($M, B, K$), and one estimated by the brain ($\hat{M}, \hat{B}, \hat{K}$). These are related through a pair of force balances where $f$ is measured force, $x$ is measured arm position, $y$ is intent, and $y_e$ is estimated intent:
\begin{equation}
M\ddot{x}+B\dot{x}+Kx+f=\hat{M}\ddot{y}+\hat{B}\dot{y}+\hat{K}y
\end{equation}
\begin{equation}
M_e\ddot{x}+B_e\dot{x}+K_ex+f=M_d\ddot{y}_e+B_d\dot{y}_e+K_dy_e
\end{equation}
We move to the Laplace domain for ease of representation and solve each equation for F.
\begin{equation}
F=(\hat{M}s^2+\hat{B}s+\hat{K})Y-(Ms^2+Bs+K)X
\end{equation}
\begin{equation}
F=(M_ds^2+B_ds+K_d)Y_e-(M_es^2+B_es+K_e)X
\end{equation}
Set those equal:
\begin{equation}
(\hat{M}s^2+\hat{B}s+\hat{K})Y-(Ms^2+Bs+K)X=(M_ds^2+B_ds+K_d)Y_e-(M_es^2+B_es+K_e)X
\end{equation}
Define a pair of error terms two terms whose sole purpose is to make our notation more convenient for addressing the question at hand. First we define the difference in arm inverse models:
\begin{equation}
\epsilon_i=(\hat{M}s^2+\hat{B}s+\hat{K})-(M_ds^2+B_ds+K_d)
\end{equation}
Next we define the difference in arm forward models:
\begin{equation}
\epsilon_f=(Ms^2+Bs+K)-(M_es^2+B_es+K_e)
\end{equation}
Collect terms:
\begin{equation}
(\hat{M}s^2+\hat{B}s+\hat{K})(Y-Y_e)+\epsilon_i Y_e=\epsilon_f X
\end{equation}
And rearrange:
\begin{equation}
Y_e=Y+\frac{\epsilon_i Y_e-\epsilon_f X}{\hat{M}s^2+\hat{B}s+\hat{K}}
\end{equation}
While on first glance this appears to contain a tremendous number of unknowns, $y_e$ is locally-in-time linear in each unknown. Therefore, the problem simplifies to a linear regression where the term that varies with no parameters represents the true intent, $y(t)$. We can therefore learn the true intent and parameter values simultaneously by extracting the intent many times with a carefully chosen variety of parameters. Moreover, it appears that this should be true of any system. This requires that some assumptions be made about $y$, however. In particular, the problem is trivial if $y$ is stationary and more-or-less orthogonal to each parameter error's time trajectory.

While we can perfectly know change in each component of estimation error, we cannot yet know which ground truth parameter value it has been varied with respect to. ($\epsilon_i=\hat{K}-K_d$ but $\Delta \epsilon_i=-\Delta K_d$) It is therefore desirable to take this one step further. After performing the linear regression suggested above at a given point in time across a variety of parameter estimates, we find a set of linear weights $e$ and an offset term which contains the sum of $y$ and each weight multiplying its respective ground-truth parameter value. Each of these weights represents a scaled version of the error due to misestimation in its respective parameter. Because we know each weight as a time series and we know $y_e$, we can use QRD-LSL (a Kalman workalike with superior numerical properties on floating point hardware) to recover the time-varying contribution of each to $y_e(t)$ in addition to a component, $y(t)$, not explained by parameter estimation error.

This suggests a process with several stages:
\begin{enumerate}
  \item Create many candidate estimations of intent with parameter values based on a Bayesian prior.
  \item Decompose the \textit{a priori} maximum likelihood intent into submovements using the vocubulary and grammar found in stroke.
  \item Use linear regression as suggested by eq. 9 to determine the time-varying error signature that would be caused by misestimation in each parameter.
  \item Use Kalman-like filtering (QRD-LSL) to find the time-varying contribution of each error and submovement to the maximum likelihood intent and recover a more accurate estimate of the intent.
  \item Correct each parameter using the schedule determined above to produce a superior estimation of reaching intent.
\end{enumerate}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{paper2.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2014-07-03 Paper1/paper1.tex
Updates from home

@@
 Beyond this point, we detected change in intent even when the intention was near the target.
-Unlike well-known $\lambda$ models, this intent was determined by considering both feedforward (predictive) and feedback controls.
+Unlike well-known lambda ($\lambda$) models, this intent was determined by considering both feedforward (predictive) and feedback controls.
 Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits.
@@
 This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people.
-It is also a new tool for understanding and treating impaired movement plans in neural injury.
+It is also a new tool for understanding and treating impaired movement plans following neural injury.
 
@@
 While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves.
-Particularly of interest is the intended course of action, even when the actual movement is disturbed and hence no longer matches the intent.
+Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.
 
@@
 Therefore, it could not well-represent the intent of a simple reaching movement.
-We propose a revised method---intent extraction---that deduces intended trajectory, even in the face of disturbances.
+We propose a revised method --- \textit{intent extraction} --- that deduces intended trajectory, even in the face of disturbances.
 This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory.
-In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling.
-The mathematics underlying intent determination are sufficiently general to be extensible to whole-body movement or even more abstract, psychological processes.
+In human reaching, these initial ingredients are available: the process is a dynamic model of the arm, the actuators (muscles) are assumed to act as viscoelastic torque generators, the human interface forces are measured by a sensor, and the realized trajectory is measured.
 
-Here we develop and analyze a method to extract intent from disturbed movement.
-This method then allows us to observe if and how intent changes following disturbance.
-We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction.
-We then explore sensitivity and failure modes using standard model analysis techniques.
-We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances.
-We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching.
-We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed.
+Here we develop and analyze the mathematics underlying intent extraction, a method to determine one's intent.
+We focused, for simplicity in this initial study, on planar targeted arm reaching actions.
+First, we produce synthetic reaching data using a forward model so that we can examine the quality and uncertainty of intent extraction.
+We then explore sensitivity and failure modes, demonstrating the accuracy and utility of this method.
+We show efficacy on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances.
+Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intent may change following a disturbance.
+
+% M E T H O D S
 
@@
 \end{equation}
-Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{kawato1999internal}.
+Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}.
 This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance.
@@
 Instead of a dynamic model, it could be a lookup table.
-For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
+For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
 In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance.
 As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined.
-This allows determination of intention in many situations of interest, even where the process is irreducibly complex.
+This allows determination of intent in many situations of interest, even where the process is irreducibly complex.
 
@@
 While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models.
-Appropriate plant modeling is well-understand: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
+Here we show the results from our choice of Burdet and colleagues \cite{burdet2006stability}.
+Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
 Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
+Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
 For the actuator, we choose the model of Burdet et al.
@@
 A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged.
-As a consequence of choosing this model, muscle torque must be used to calculate stiffness even though it arises in part from stiffness.
+In this  model, musculoskeletal stiffness was linearly related to torque.
 Like Burdet et al.
-\cite{burdet2006stability}, we approximated muscle torque by looking backward in time a small amount (500 microseconds).
-This approximation is not necessary during intent extraction, which leads to a small (sub-micrometer mean unsigned error) discrepancy.
-Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.
+\cite{burdet2006stability}, we approximated muscle torque by looking backward in time 500 microseconds.
+Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.
 
 \subsection*{Experimental Design}
-Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.
-We used the following types of perturbing forces for each combination of distance and direction:
-pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
-Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion.
-The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.
+We chose 15 centimeters long simulated reaches of the right arm, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then first proceeding to the right.
+We used the following two types of perturbing forces for each combination of distance and direction:
+(1) pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
+(2) Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion.
+The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ rad/s.
 
-\subsection*{Error Metric}
-Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach.
+Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target.
+This was evaluated across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach.
 Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis.
@@
 \subsection*{Indices of variance-based sensitivity}
-Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
+To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
 Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range.
@@
 Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
+Because force sensors tend to drift, we performed a linear re-zeroing procedure between each motion to assure unbiased measurements.
 
@@
 Forces were chosen pseudorandomly such that each type, direction, and distance combination mentioned above was presented 5 times.
+We used these disturbed, intermittent trials for the analysis.
 
@@
 The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
+We used an alpha level of 0.05 for declaring statistical significance.
+
+% R E S U L T S
 
@@
 \section*{Results}
+As expected, the model was able to recover the original intended trajectory even when disturbed (Figure 1).
+However this idealized analysis cannot reveal any vulnerabilities to unexamined model parameters or to inaccuracy in the structure of the model itself as discussed in the next sections.
 
-\subsection*{Model Robustness}
-While some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.
-Sensitivity analysis revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.
+\subsection*{Noise Robustness}
+While some inversion processes might be highly sensitive to noise, our process for recovering intent from action did not appear to be vulnerable.
+Our position sensors are very accurate (no detectable drift, RMS noise 1.1 micrometers), but the force sensors are less so.
+Even after correcting for drift in our force measurements, the force measurements reported during reaching can be expected to have noticeable bias and further noise on the order of that bias as reported in Table 1.
+Taken together, force measurement errors account for 21\% of all sensitivity.
+We also examined noise in sensor signals explicitly in the more comprehensive  sensitivity analysis, discussed below.
+
+\subsection*{Parameter Sensitivity}
+Variance in intended trajectories due to estimated uncertainty in model parameters was lower than the natural variation in undisturbed motions.
 Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al.
@@
 
-\subsection*{Parameter Sensitivity}
-Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial.
+Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to mis-estimation of stiffness and changes in shoulder position from trial to trial.
 Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
@@
 Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3).
-In both pulse timing conditions, the intended trajectory remained straight towards the target, supporting the approach.
+If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands.
+We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay.
+In both pulse timing conditions, the intended trajectory remained straight towards the target, even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.
+
+These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances.
+Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance.
+We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance (Table 2).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred late in reaching, intent continued onto the target but then corrected anyways (Figure 3 lower plots).
-We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance.
-Intention showed some systematic deflection toward the direction of disturbance peaking around 150 milliseconds after the onset of force disturbance, but this effect did not leave the range of error found in undisturbed movement.
-While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for as much as one second (Figure 4 shaded areas).
+When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3 lower plots).
+Intention showed systematic deflections that counteracted the direction of disturbance about 200 milliseconds after the onset of the force disturbance.
+This counteraction did not depart from the range of undisturbed movements, however (Figure 4, green region).
+The movements completed with the hand returning to undisturbed levels of error within 600 milliseconds, intention differed from actual hand path for as long as one second after the onset of motion.
 
-We argue that the approach here represents the first sound method for looking at intent following disturbance, so we have only the undisturbed intention for comparison.
-If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands.
-This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay.
-We test this in Figure 4.
-
-Following disturbance, we compare deviation in the hand and intended paths against the maximum deviation of undisturbed movements in order to determine at what point in time changes in hand path or intent can no longer be explained by normal variation and inaccuracy.
-Subjects' force-disturbed hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance.
-By contrast, intent did not change significantly for at least 300 milliseconds after the onset of disturbance (Table 2).
-This passes the test described above.
-While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this residual disturbance does not exceed the statistical threshold for significance.
-An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.
+% D I S C U S S I O N
 
@@
 We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there are disturbances.
-Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching.
-Significant misestimation could occur if the stiffness was inaccurately modeled.
-We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds.
+Sensitivity analysis on synthetic data revealed that errors in response to parameter variations were smaller than the trial-to-trial variance commonly observed in human reaching.
+Mis-estimation was largest if the stiffness was inaccurately modeled.
+We tested this on human reaching and found that in spite of the hand being disturbed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds.
 This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.
@@
 The proposed approach could explain the discovery by Bizzi et al.
-\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$.
-By deriving the technique in a general form, we discovered that our $x_d$ should be equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.
-$x_d$ and $\lambda$ are only equivalent when $x_d$ is not changing.
-Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass.
-This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.
+\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging.
+By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.
 
-Our analysis suggests that intent is slow to change if it changes at all following disturbance.
-While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement.
+Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibration.
+Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of the mass term in the intention formulation -- reaching intent cannot jump abruptly.
+The intent trajectories certainly follow this pattern, smoothly responding to disturbances at a latency of about 200 milliseconds.
+While this work examines the tendencies of the onset of change, statistical testing cannot be used to detect the absence of change in an individual movement.
 Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control.
-In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}.
+In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}.
 Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
+Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly persist with the same intent when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
 Time of response may be dependent on the time of the disturbance in the movement.
@@
 The process model does not need to be explicitly formulated.
-Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance.
-The plant (such as the actions of a crowd in a crowded square or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
-However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed.
-In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.
+Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance.
+The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
+However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
+In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to uncover the plan behind their action.
 
@@
 
-Our sensitivity analysis showed that approximation of stiffness is the most critical factor to accuracy in estimating intent.
-While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
+A key assumption in this study is that the joint impedance (stiffness) is linearly related to torque.
+While other approaches tend to construct paradigms that indicate or presume that the intention is static in order to derive impedance -- the so called family of system identification applications, this study makes a constrained assumption of stiffness in order to determine intent.
+In a sense, this approach widens the scope to include movement.
+Sensitivity analyses showed that approximation of stiffness is the most critical factor to accuracy in estimating intent.
+While serious trial-to-trial stiffness mis-estimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach.
 Even as the model is insensitive to normal variations, a human can halve or double their stiffness \cite{franklin2007endpoint} significantly reducing determination accuracy.
@@
 
-Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting.
+The simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting.
 More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
 Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
-As mentioned above, while this may be perceived as a shortcoming, it was necessary and can potentially be overcome in future work.
 

### 2014-07-03 Paper2/compactVersion.tex
Updates from home

--- 2014-07-01 Paper2/manuscript/paper2.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
 \documentclass[10pt]{article}
@@
 \usepackage{graphicx}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
 \usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
 
@@
 
-% Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
 \pagestyle{myheadings}
-%% ** EDIT HERE **
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
 
@@
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{The Grammar and Vocabulary of Human Reaching Intent}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Justin Horowitz$^{1}$, 
-James Patton$^{1,2\ast}$
-\\
-\bf{1} Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States
-\\
-\bf{2} Sensory Motor Performance Program, Rehabilitation Institute of Chicago, Chicago, Illinois, United States
-\\
-$\ast$ E-mail: pattonj@uic.edu
-\end{flushleft}
+When performing intent extraction, we're faced with many significant unknowns.
+Certain arm parameters can be changed volitionally (humans can co-contract their muscles, raising the arm's stiffness) while others might diverge from cadaver anatomy as studied in the 1950s.
+We derive the relationship between extractions with different parameters in the hope of finding components of the extraction that relate primarily to estimation error in some parameter value.
+With such ``signature'' in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
 
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
-
-\section*{Introduction}
-blah blah blah
-
-We hypothesize that human reaching intention is built from submovements that constitute a fixed vocabulary and grammar.
-Were this hypothesis true, we should find it everywhere we look and without exception.
-We test this hypothesis by comparing the composition of movement in stroke survivors and force-disturbed control subjects whose reaching intent has been recovered by a novel computational technique.
-In particular, we presume a grammar based on the work of Flash and Henis \cite{flash1991arm} and a vocabulary to be recovered from the stroke survivors whose submovements can be easily recovered given the presumed vocabulary \cite{rohrer2004submovements}.
-If this grammar and vocabulary are accurate and complete, they should fully explain force-disturbed reaching intent in control subjects.
-
-% Results and Discussion can be combined.
-\section*{Results}
-
-\subsection*{Subsection 1}
-
-\subsection*{Subsection 2}
-
-\section*{Discussion}
-We set out to test the hypothesis that both stroke survivors and neurologically intact subjects exposed to force disturbance share a compact grammar and vocabulary for constructing movement intent.
-The grammar, or rules for combination, was that submovements were combined by superposition (vector addition) and no more than two were present at any point in time.
-The vocabulary was the minimum-jerk, $5^{th}$ order polynomial starting and ending at at zero speed and acceleration.
-This grammar and vocabulary recovered from stroke survivors was able to completely explain reaching intent in neurologically intact subjects despite imposing significant restrictions.
-While some might argue that these submovements are an artifact of stroke or an artifact of intent extraction, it is extremely unlikely that both situations should share these features unless they're fundamental to human movement.
-Moreover, we were able to detect predictable changes in arm stiffness that were well-correlated with the onset of submovements.
-This suggests that reaching intent may have an additional volitional impedance component not detectable in undisturbed reaching.
-
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\section*{Materials and Methods}
-When performing intent extraction, we're faced with many significant unknowns.
-Certain arm parameters might change volitionally while others might diverge from cadaver anatomy as studied in the 1950s.
-We derive a relationship between extracted intent and parameter misestimation in the hope of finding components of the extraction that relate primarily to estimation error in some parameter value.
-With the ``signature'' of these errors in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
-
-We begin by noting that the arm is described by four sets of parameters at any point in time: two in our model ($M_e, B_e, K_e$ and $M_d, B_d, K_d$), one in reality ($M, B, K$), and one estimated by the brain ($\hat{M}, \hat{B}, \hat{K}$).
+We begin by noting that the arm has four sets of parameters at any point in time: two in our model ($M_e, B_e, K_e$ and $M_d, B_d, K_d$), one in reality ($M, B, K$), and one estimated by the brain ($\hat{M}, \hat{B}, \hat{K}$).
 These are related through a pair of force balances where $f$ is measured force, $x$ is measured arm position, $y$ is intent, and $y_e$ is estimated intent:
@@
 \end{equation}
-We move to the Laplace domain for ease of representation and solve each equation for F.
+We move to the LaPlace domain for ease of representation and solve each equation for F.
 \begin{equation}
@@
 \end{equation}
-Set those equal:
+Set these equal:
 \begin{equation}
@@
 \end{equation}
-Define a pair of error terms two terms whose sole purpose is to make our notation more convenient for addressing the question at hand. First we define the difference in arm inverse models:
+Since we desire to find the difference of $y$ and $y_e$, we collect terms and define two terms whose sole purpose is to make our notation more convenient for addressing the question at hand.
+First we define the difference in estimated arm models:
 \begin{equation}
@@
 \end{equation}
-Next we define the difference in arm forward models:
+Next we define the difference in arm models:
 \begin{equation}
@@
 
-While we can perfectly know change in each component of estimation error, we cannot yet know which ground truth parameter value it has been varied with respect to.
-($\epsilon_i=\hat{K}-K_d$ but $\Delta \epsilon_i=-\Delta K_d$) It is therefore desirable to take this one step further.
-After performing the linear regression suggested above at a given point in time across a variety of parameter estimates, we find a set of linear weights $e$ and an offset term which contains the sum of $y$ and each weight multiplying its respective ground-truth parameter value.
-Each of these weights represents a scaled version of the error due to misestimation in its respective parameter.
-Because we know each weight as a time series and we know $y_e$, we can use QRD-LSL (a Kalman workalike with superior numerical properties on floating point hardware) to recover the time-varying contribution of each to $y_e(t)$ in addition to a component, $y(t)$, not explained by parameter estimation error.
+While you can perfectly know how you've varied each component of each $\epsilon$, this does not reveal which ground truth parameter value it has been varied with respect to.
+For example: $\epsilon_K=\hat{K}-K_d$ but $\Delta \epsilon_K=-\Delta K_d$.
+It is therefore desirable to take this one step further and ask what structure might be inherent in $y$.
+In particular, we have reason to suspect that $y$ might be composed of a sum of minimum submovements.
+Since these are defined by an initial and final position along with a span in time, we can fully both intent and parameter values by linear regression so long as we can determine the extent in time of each submovement.
 
@@
 \begin{enumerate}
-  \item Create many candidate estimations of intent with parameter values based on a Bayesian prior.
-  \item Decompose the \textit{a priori} maximum likelihood intent into submovements using the vocubulary and grammar found in stroke.
-  \item Use linear regression as suggested by eq. 9 to determine the time-varying error signature that would be caused by misestimation in each parameter.
-  \item Use Kalman-like filtering (QRD-LSL) to find the time-varying contribution of each error and submovement to the maximum likelihood intent and recover a more accurate estimate of the intent.
-  \item Correct each parameter using the schedule determined above to produce a superior estimation of reaching intent.
+  \item Extract an estimate of the intent with parameter values based on some \textbf{prior}.
+  \item Decompose the intent into submovements by using prior knowledge of their form and combination rules.
+  \item Create many candidate estimations of the intent using the Sobol set to carefully span a reasonable parameter space.
+    \item Use linear regression on the outcome of steps 2 and 3 to determine the parameter values as they vary in time.
+      \item Perform a "final" best-estimate extraction and decomposition based on the steps above.
 \end{enumerate}
 
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgments}
-
-%\section*{References}
-% The bibtex filename
-\bibliography{paper2.bib}
-
-\section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
-
-\section*{Tables}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
 \end{document}

### 2014-07-03 Paper2/updatedExtract/innerproductspace.m

@@ -1,0 +1,115 @@
+clc
+clear all
+close all
+
+%% Set them up
+
+%Presume the shortest human submovement is 150ms and the longest is 450ms.
+tspan=[.15 .45];
+
+%Presume minima and maxima in speed, too
+mspan=[.01 .05];
+
+Nsubs=20;
+ts=zeros(Nsubs,2);
+dirs=zeros(Nsubs,2);
+mags=zeros(Nsubs,1);
+
+for k=1:Nsubs
+    if k==1
+        ts(k,1)=0;
+    elseif k==2
+        ts(k,1)=rand*(tspan(2)-tspan(1))+tspan(1);
+    else
+        if rand>.3
+            ts(k,1)=ts(k-2,2);
+        else
+            ts(k,1)=ts(k-2,2)+rand*tspan(1);
+        end
+    end
+    ts(k,2)=ts(k,1)+rand*(tspan(2)-tspan(1))+tspan(1);
+    
+    mags(k)=rand*(mspan(2)-mspan(1))+mspan(1);
+    dirs(k,:)=2*rand(1,2)-1;
+    dirs(k,:)=dirs(k,:)/norm(dirs(k,:));
+end
+
+figure(1)
+clf
+subplot(3,1,1)
+hold on
+plot(ts(:,1),zeros(Nsubs,1),'bo',ts(:,2),zeros(Nsubs,1),'bx')
+
+%% Get a speed profile.
+
+tstep=.001;
+t=(0:tstep:max(max(ts)))';
+v=zeros(length(t),2);
+
+for k=1:Nsubs
+    fLow=find(t>=ts(k,1),1,'first');
+    if isempty(fLow)
+        fLow=1;
+    end
+    fHigh=find(t<=ts(k,2),1,'last');
+    if isempty(fHigh)
+        fHigh=length(t);
+    end
+    
+    ta=(t(fLow:fHigh)-ts(k,1))/(ts(k,2)-ts(k,1));
+    vd=(30*ta.^2-60*ta.^3+30*ta.^4)*(mags(k)/(ts(k,2)-ts(k,1)))*dirs(k,:);
+    c=.9*rand(3,1)+.1;
+    plot(t(fLow:fHigh),sqrt(sum(vd.^2,2)),'-','color',c,'linewidth',2)
+    v(fLow:fHigh,:)=v(fLow:fHigh,:)+vd;
+end
+
+plot(t,sqrt(sum(v.^2,2)),'k')
[…]

### 2014-07-03 Paper1/paper1.tex
Updated with minor stuff

@@
 We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions.
-We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen when humans reach to visual targets.
+We found that sensitivity to parameter inaccuracies was on the order of square microns and hence much lower than the inaccuracies seen when humans reach to visual targets.
 Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness.
@@
 Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance.
-This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people.
+This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from small organisms to crowds of people.
 It is also a new tool for understanding and treating impaired movement plans following neural injury.
@@
 By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the "$\lambda$ model"\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
+Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
 Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
@@
 First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process.
-Next we conducted an experiment on humans to evaluate the process using human subject data.
+Next, we conducted an experiment on humans to evaluate the process using human subject data.
 
@@
 This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches.
-Nevertheless later we will discuss how this approach can be used on more complex models without loss of generality.
+Nevertheless, later we will discuss how this approach can be used on more complex models without loss of generality.
 The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:

### 2014-07-15 Paper1/paper1.tex
Updated paper 1, updating toy extract. Need matlab class

@@
 
-For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation is typically solved by numerical integration to determine the solution to the initial value problem in time.
+For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation is solved by numerical integration to determine the solution to the initial value problem in time.
 This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory.
@@
 \end{equation}
-The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ that $x$ will track a desired path $x_d$.
+The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ such that $x$ will track a desired path $x_d$.
 \begin{equation}
@@
 Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated \cite{shadmehr1994adaptive, franklin2003adaptation, burdet2006stability}.
-Demonstrating the validity and plausibility of mathematically transforming those filters to instead covert measured hand trajectories to intent adds a novel method to the arsenal of analytical methods for exploring human motor control.
+Demonstrating the validity and plausibility of mathematically transforming those filters to instead convert measured hand trajectories to intent adds a novel method to the arsenal of analytical methods for exploring human motor control.
 
@@
 Direct access to the intended trajectory allows optimal feedback control models to better understand and fit hypothesized cost functions and their parameters.
-In addition, mid-movement replanning became falsifiable: we were able to show that intended trajectories from disturbed motions differ from those undisturbed motions and how they differed.
+In addition, mid-movement replanning became falsifiable: we were able to show that intended trajectories from disturbed motions differ from those of undisturbed motions and how they differed.
 

### 2014-07-15 Paper2/LMSExtract/rlsNexample.m

@@ -11,7 +11,7 @@ x=randn(M,N);
 xt=x;
 %xt(2,:)=sin(xt(2,:));
 Y=wReal*xt;
-intent=cos(12.7*(1:N))+5;
+intent=rand(1,N); %cos(12.7*(1:N))+5;
 y=Y+S2N*std(Y)*randn(1,N)+intent;
 linRegress=(x'\y')' %#ok<*NOPTS>
 

### 2014-08-12 Paper1/paper1.tex
updated paper and paper 2 math
Signed-off-by: Justin Horowitz <justin.horowitz@gmail.com>

@@
 \usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
+\usepackage[noperiod]{jabbrv}
 % Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
+\bibliographystyle{jneurophys}
 
@@
 
-% Title must be 150 characters or less
+\begin{titlepage}
+\begin{center}
+{\Large
+\textbf{I meant to do that: determining the intentions of action in the face of disturbances}
+}\\[1cm]
+\end{center}
+
 \begin{flushleft}
-{\Large
-\textbf{I meant to do that: Extracting the intentions of action in the face of disturbances}
-}
 % Insert Author names, affiliations and corresponding author email.
-\\
 Justin Horowitz$^{1}$, 
@@
 \\
-$\ast$ E-mail: pattonj@uic.edu
+$\ast$ Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States. E-mail: pattonj@uic.edu
+\\[1cm]
+Running Head: \\
+Determining the intention of action despite disturbances
 \end{flushleft}
+\end{titlepage}
 
@@
 \section*{Abstract}
-We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test the approach using human reaching data.
-We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions.
-We found that sensitivity to parameter inaccuracies was on the order of square microns and hence much lower than the inaccuracies seen when humans reach to visual targets.
-Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness.
-We also tested the robustness of this approach on human point-to-point reaching movements disturbed by force pulses.
+Our actions often do not match our intentions when there are external disturbances.
+We derive a novel modeling approach for determining motor intent from a reaching motion that is disturbed by an unexpected force.
+We first show how to mathematically invert both feedforward (predictive) and feedback controls to obtain intent.
+We next examine the model’s sensitivity to parameter uncertainty, and test the approach using human reaching data.
+We found that sensitivity to parameter inaccuracies was on the order of square microns, and hence much lower than the inaccuracies seen in actual human motions.
+Further, the largest sensitivity arose mainly from uncertainty in our assumptions of joint stiffness.
+We also tested the hypothesis that this would predict straight-line intent in human targeted reaching movements, even when disturbed by various pulses of force.
 As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset.
-Beyond this point, we detected change in intent even when the intention was near the target.
-Unlike well-known lambda ($\lambda$) models, this intent was determined by considering both feedforward (predictive) and feedback controls.
-Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits.
-This approach may also be scaled for determining intent during human-machine interaction and in complex systems like team sports, crowds, or swarms.
+Beyond this point, however, we detected a change in intent, even when the intention was near the target.
+Knowing such an intent signal is broadly applicable in areas such as in enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and in any complex system that embodies planning, such as brain machine interfaces, team sports, crowds, or swarms.
+In addition, observing intent as it changes might act as a window into the mechanisms of planning, correction, and learning.
+
+\section*{Keywords}
+Intention; Motor Control; Visually Guided Reaching; On-line Control
 
@@
 
+\section*{Grants} 
+Supported by US NIH NINDS1 R01 NS053606
+
 %\section*{References}

### 2014-08-12 Paper2/krlsProofsAndToys/slmj5op.m

@@ -1,0 +1,27 @@
+function [kern,fop]=slmj5op(t,tc,ts)
+
+if size(t,2)>size(t,1)
+    t=t';
+end
+
+if size(tc,2)>size(tc,1)
+    tc=tc';
+end
+
+ts2=ts/2;
+
+td=t-tc;
+ta=(td+ts2)/ts;
+
+kern=(30*ta.^2-60*ta.^3+30*ta.^4)/ts;
+kern(td<=-ts2)=0;
+kern(td>=ts2)=0;
+
+if nargout>1
+    fop=[10*ta.^3-15*ta.^4+6*ta.^5,kern,(60*ta-180*ta.^2+120*ta.^3)/(ts^2)];
+    fop(td<=-ts2,:)=0;
+    fop(td>=ts2,1)=1;
+    fop(td>=ts2,2:3)=0;
+end
+
+

### 2014-08-26 Paper2/compactVersion.tex
Added start of gradient descent

@@
 \begin{document}
+\section*{Why Endure the Math?}
+Humanity's search for meaning is well-evidenced in the archaeological record.
+By using the mathematics of advanced control theory to relate experience, intent, and action, we derive from first principles the structure of learning and planning.
+The early derivations contain few interesting insights, but produce a scaffold from which we can develop the more interesting frameworks later on.
+These frameworks produce an overwhelming number of explanations and predictions.
+Full exploration of their implications is beyond the scope of this work.
+Even limiting ourselves to the bare essentials, we are forced to confront explanations to very old questions: What is meaning?
+Is there an objective morality?
+What do we know and how do we know it?
+What is the best way of knowing?
+Following the math enables a reader to tackle these questions from the standpoint of a formal language, a language in which statements are clearly true or false.
+While we provide interpretation along the way in markedly less formal language, many implications are counter-intuitive and disappointing.
+It is our sobering duty to remind our readers that unless an error in the math itself can be found, the answers to these questions are fundamental and inherent to be accepted as laws that cannot be fought or changed.
 
-When performing intent extraction, we're faced with many significant unknowns.
+\subsection*{TL;DR}
+Math is hard, but the outcomes are numerous and generally useful well outside the original question.
+
+\section*{Time-varying Linear Processes Governed by Intent}
+It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators.
+Importantly, this has been claimed to be the intended movement.
+To consider this, we also generalize our approach to any system with a moving intention that governs the control law.
+This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $q_e$ (sometimes represented as $\lambda$).
+The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
+Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
+\begin{equation} \label{eq:generalPlant}
+\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
+\end{equation}
+The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ such that $x$ will track a desired path $x_d$.
+\begin{equation}
+\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
+\end{equation}
+We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives:
+\begin{equation}
+\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
+\end{equation}
+
+Note $x_e$ vanishes, recovering our familiar model.
+While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory.
+Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero.
+In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
+
+While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
+Instead of a dynamic model, it could be a lookup table.
+For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
+In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance.
+As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined.
+This allows determination of intent in many situations of interest, even where the process is irreducibly complex.
+
+\subsection*{TL;DR}
+So long as we can use a series representation (you almost always can) and intent matches action when disturbance isn't present (ie.
+the control system is competent), we can back out intent from action by knowing that series representation.
+
+\section*{Linear Independence of Intent and Process Uncertainty}
+Whether turning intent into action or determining it from action, we're faced with many significant unknowns.
 Certain arm parameters can be changed volitionally (humans can co-contract their muscles, raising the arm's stiffness) while others might diverge from cadaver anatomy as studied in the 1950s.
 We derive the relationship between extractions with different parameters in the hope of finding components of the extraction that relate primarily to estimation error in some parameter value.
-With such ``signature'' in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
+With such a ``signature'' in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
 
@@
 \end{equation}
+Finally, simplify by assuming uniform misestimation: $\epsilon_i=\epsilon_f=\epsilon$
+\begin{equation}
+Y_e=Y+\epsilon\frac{Y_e-X}{\hat{M}s^2+\hat{B}s+\hat{K}}
+\end{equation}
 While on first glance this appears to contain a tremendous number of unknowns, $y_e$ is locally-in-time linear in each unknown.
@@
 We can therefore learn the true intent and parameter values simultaneously by extracting the intent many times with a carefully chosen variety of parameters.
-Moreover, it appears that this should be true of any system.
-This requires that some assumptions be made about $y$, however.
-In particular, the problem is trivial if $y$ is stationary and more-or-less orthogonal to each parameter error's time trajectory.
-
-While you can perfectly know how you've varied each component of each $\epsilon$, this does not reveal which ground truth parameter value it has been varied with respect to.
-For example: $\epsilon_K=\hat{K}-K_d$ but $\Delta \epsilon_K=-\Delta K_d$.
-It is therefore desirable to take this one step further and ask what structure might be inherent in $y$.
-In particular, we have reason to suspect that $y$ might be composed of a sum of minimum submovements.
-Since these are defined by an initial and final position along with a span in time, we can fully both intent and parameter values by linear regression so long as we can determine the extent in time of each submovement.
-
-This suggests a process with several stages:
+Moreover, it appears that this should be true of any system designed to drive a process along a reference trajectory.
+\\\\
+This suggests a process with two stages:
 \begin{enumerate}
-  \item Extract an estimate of the intent with parameter values based on some \textbf{prior}.
-  \item Decompose the intent into submovements by using prior knowledge of their form and combination rules.
 \item Create many candidate estimations of the intent using the Sobol set to carefully span a reasonable parameter space.
-    \item Use linear regression on the outcome of steps 2 and 3 to determine the parameter values as they vary in time.
-      \item Perform a "final" best-estimate extraction and decomposition based on the steps above.
+\item Use adaptive filtering to recover intent and parameter values as they vary in time.
 \end{enumerate}
 
+\subsection*{TL;DR}
+So long as we can use a time-varying series representation for the process (you almost always can) and intent matches action when disturbance isn't present (ie.
+the control system is competent), we can back out intent from action even without knowing that series representation.
+
+\section*{Intent as a Sum of Kernels}
+Based on the results of Flash and Hogan and Flash and Henis, one might reasonably set out to model intent as the sum of minimum-jerk, $5^{th}$-order polynomial submovements.
+For the sake of a simple narrative, we make our first assumption:
+\begin{quote}
+\textbf{Hypothesis: Intent is composed of vector-summed subunits that evolve in time as minimum-jerk, $5^{th}$ order polynomials with no more than two contributing at any point in time.}
+\end{quote}
+Putting the hypothesis into mathematics is straight-forward: the kernel's magnitude at some point in time is related to the kernel's center and width like so:
+\begin{equation}
+t_n=\frac{t-t_c+\frac{t_s}{2}}{t_s}
+\end{equation}
+\begin{equation}
+\kappa(t,t_c,t_s)=
+  \begin{cases}
+  0 & : t-t_c \leq -\frac{t_s}{2}\\
+  10t_n^3-15t_n^4+6t_n^5 & : -\frac{t_s}{2} < t-t_c < \frac{t_s}{2}\\
+  0 & : t-t_c \geq \frac{t_s}{2}
+  \end{cases}
+\end{equation}
+And with this expression in hand, we express some intent y(t) like so:
+\begin{equation}
+\boldsymbol{y}(t)=\sum_n \boldsymbol{\omega}_n \kappa(t,t_c[n],t_s[n])
+\end{equation}
+While the dimensionality of our signal did not previously matter, we note here that $\boldsymbol{y}$, the intent, and $\boldsymbol{\omega}$, the direction and magnitude of a submovement, are of some dimensionality while $\kappa$ is a scalar inner product.
+
+This has some immediate implications: minimum-jerk, $5^{th}$-order polynomial submovements form a reproducing kernel Hilbert space (RKHS) in the velocity domain.
+While we've made this assumption to simplify the narrative and the coming derivations, we could safely immediately retract it and instead claim that intent is well-modeled in some sparse RKHS without any loss of generality or any assumptions at all.
+General approximations generally approximate.
+
+While general approximations generally approximate, this becomes a testable hypothesis so long as the total number of submovements present at any point in time is very low as we claim.
+Past examination by our group and others generally stopped here: while this is a testable hypothesis, it is a question of "fossil rabbits in the pre-Cambrian." No single compelling test is available.
+
+While the composition of movement and planning has no known compelling test, so long as we claim that we know this form for $\dot{\boldsymbol{y}}(t)$ something wonderful emerges:
+\begin{equation}
+\dot{\kappa}(t,t_c,t_s)=
+  \begin{cases}
+  0 & : t-t_c \leq -\frac{t_s}{2}\\
+  \frac{30t_n^2-60t_n^3+30t_n^4}{t_s} & : -\frac{t_s}{2} < t-t_c < \frac{t_s}{2}\\
+  0 & : t-t_c \geq \frac{t_s}{2}
+  \end{cases}
+\end{equation}
+\begin{equation}
+\dot{\boldsymbol{y}}(t)=\sum_n \boldsymbol{\omega}_n \dot{\kappa}(t,t_c[n],t_s[n])
+\end{equation}
+\begin{equation}
+M\dot{\boldsymbol{y}}^2(t)=M \sum_n \boldsymbol{\omega}^2_n \dot{\kappa}^2(t,t_c[n],t_s[n])
+\end{equation}
+In plainer language: kinetic energy is also represented in a RKHS because the square of a symmetric, positive definite function is also symmetric and positive definite.
+This immediately implies a mechanism for learning and planning in an energy or action domain.
+Since total energy will be conserved, the change in potential energy is the negative of the change in kinetic energy.
+In other words, if the world is represented as a virtual potential field in an RKHS, the intended trajectory is nothing more than a direct traversal of the Lagrangian action.
+Lagrangian optimization falls out automatically.
+Moreover, learning is also trivial since one need only know their kinetic energy history to map out a potential energy field relative to some notion of state.
+
+Even stranger, because this is encoded in an RKHS, we can actually describe equivalent maps from (perceived) external state, $x$ to internal representation of virtual potential, $v$ in terms of some interpretation $I$.
+This relies on the dual representation of the RKHS both in time and state (which are equally valid) and its finite support in both domains.
+Telling the story backwards to make it more computable and then proving it...
+	If potential energy is defined as a function of state can you consider the present potential energy as the sum of the kernel inner product of all known potential energy/state pairs and the current state?
+If so, this is implies that experience forms a \textit{dictionary} and because of the kernel trick and its symmetry, it implies that that dictionary can actually be computed in reverse.
+Instead of considering the current state with respect to a dictionary of centers, you can instead consider those centers with respect to the current state.
+
+\begin{equation}
+v=\kappa(\langle I,x \rangle)
+\end{equation}
+\begin{equation}
+v=\langle\varphi(I),\varphi(x)\rangle
+\end{equation}
+
+\section*{T}
+Stuff
+
 \end{document}

### 2014-08-26 Paper2/krlsProofsAndToys/main.m

@@ -72,7 +72,7 @@ H_new=H(end);
 H0=slmj5op(0,0,ks);
 lH=length(H);
 
-sH2=3.5/sum(H);
+sH2=3.75/sum(H);
 H_reg=H;
 H_reg(1:(lH-1)/2)=0; %The trailing edge should not cause learning
 H_reg=H_reg/sum(H_reg);
@@ -148,3 +148,18 @@ plot(t,atan2(W(subset,2),W(subset,1)))
 
 figure(3)
 plot(t,gradient(atan2(W(subset,2),W(subset,1))))
+
+%% Step 4: Paper 3 in theory comes down to whether or not gradient descent recovers the original submovements.
+
+figure(4)
+Wg=W;
+
+c=0;
+while c<10
+    c=c+1;
+end
+%gradient of kern with respect to ts
+(45*(t - tc)^2)/ts^4 - (150*(t - tc)^4)/ts^6 - 15/(8*ts^2)
+
+    %You can no longer use the same tricks as above because
+

### 2014-08-26 Paper2/krlsProofsAndToys/main_descent.m

@@ -1,0 +1,81 @@
+clc
+clear all
+close all
+
+
+tfinal=.7;
+firstspan=tfinal/2;
+
+tc=[firstspan/2 tfinal/2 tfinal-firstspan/2]
+ts=[firstspan 5*tfinal/6 firstspan];
+
+cols=.8*rand(3,length(tc));
+
+ti=tc-ts/2;
+tf=tc+ts/2;
+
+x0=[1 1 zeros(1,4)];
+dx=[.02 .02;.1 0;.02 -.02];
+
+t_step=.005;
+t=0:t_step:1;
+
+[yg,kerns]=supMJP(x0,dx,ti,tf,t);
+
+figure(1)
+clf
+subplot(2,1,1)
+plot(yg(:,1),yg(:,2),'g-','linewidth',5)
+axis equal
+
+subplot(2,1,2)
+hold on
+plot(t,sqrt(sum(yg(:,3:4).^2,2)),'g-','linewidth',5)
+
+%% Part 2 is a perfect expansion because why not?
+ks=.25; %So at any given time, we're carrying around ks/.005=30 kernels. Wide span risks overfitting, but probably gets the direction right
+lambda=0.01; %No reason to think we should punish large velocities, but let's preempt spikes to infinity.
+
+%H in time is weird because you already perfectly know FUTURE input samples, but not output samples.
+%Oddly this implies that H never changes because of symmetry.
+bound=floor(ks/(2*t_step))*t_step;
+H=slmj5op(0,-bound:t_step:bound,ks);
+H_old=H(1:end-1);
+H_new=H(end);
+H0=slmj5op(0,0,ks);
+lH=length(H);
+
+Hdiv=H/sum(H);
+
+sH2=3.5/sum(H);
+H_reg=H;
+H_reg(1:(lH-1)/2)=0; %The trailing edge should not cause learning
+H_reg=H_reg/sum(H_reg);
+H_reg((lH-1)/2+1:end)=H_reg(end:-1:(lH-1)/2+1);
+
+%In the "degenerate" case, we're already in motion but don't have learned
+%coefficients. Ignore this case because we're never forced to deal with
+%this in practice. Presume we start from rest.
+Q=zeros(lH); %(lambda+H0)*eye(lH);
+
[…]

### 2014-08-26 Paper2/compactVersion.tex
Overwrite this

--- 2014-07-03 Paper2/compactVersion.tex
@@
 \begin{document}
+\section*{Why Endure the Math?}
+By making two assumptions, we're able to derive human learning and planning from first principles. Moreover, we develop the tools needed to verifying our predictions incidentally along the way.
 
+\section*{Linear Independence of Intent and Process Uncertainty}
 When performing intent extraction, we're faced with many significant unknowns.
@@
 We derive the relationship between extractions with different parameters in the hope of finding components of the extraction that relate primarily to estimation error in some parameter value.
-With such ``signature'' in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
+With such a ``signature'' in hand, we can subtract a scaled version of this signature off our estimated intent to recover a better estimate of the true intent.
 
@@
 We can therefore learn the true intent and parameter values simultaneously by extracting the intent many times with a carefully chosen variety of parameters.
-Moreover, it appears that this should be true of any system.
-This requires that some assumptions be made about $y$, however.
-In particular, the problem is trivial if $y$ is stationary and more-or-less orthogonal to each parameter error's time trajectory.
+Moreover, it appears that this should be true of any system designed to drive a process along a reference trajectory.
 
-While you can perfectly know how you've varied each component of each $\epsilon$, this does not reveal which ground truth parameter value it has been varied with respect to.
-For example: $\epsilon_K=\hat{K}-K_d$ but $\Delta \epsilon_K=-\Delta K_d$.
-It is therefore desirable to take this one step further and ask what structure might be inherent in $y$.
-In particular, we have reason to suspect that $y$ might be composed of a sum of minimum submovements.
-Since these are defined by an initial and final position along with a span in time, we can fully both intent and parameter values by linear regression so long as we can determine the extent in time of each submovement.
-
-This suggests a process with several stages:
+This suggests a process with two stages:
 \begin{enumerate}
-  \item Extract an estimate of the intent with parameter values based on some \textbf{prior}.
-  \item Decompose the intent into submovements by using prior knowledge of their form and combination rules.
 \item Create many candidate estimations of the intent using the Sobol set to carefully span a reasonable parameter space.
-    \item Use linear regression on the outcome of steps 2 and 3 to determine the parameter values as they vary in time.
-      \item Perform a "final" best-estimate extraction and decomposition based on the steps above.
+\item Use adaptive filtering to recover intent and parameter values as they vary in time.
 \end{enumerate}
 
+\section*{Intent as a Sum of Kernels}
+Based on the results of Flash and Hogan and Flash and Henis, one might reasonably set out to model intent as the sum of $5^{th}$-order polynomial submovements. We make our first assumption:
+\begin{quote}
+\textbf{Intent is well-modeled as vector-summed subunits that evolve in time as minimum-jerk, $5^{th}$ order polynomials.}
+\end{quote}
+This has some immediate implications.
+
+\section*{T}
+Stuff
+
 \end{document}

### 2014-09-22 Paper1/paper1.tex
Overwrite this

@@
 % Use the PLoS provided bibtex style
-\bibliographystyle{jneurophys}
+\bibliographystyle{plos2009}
 
@@
 In human reaching, these initial ingredients are available: the process is a dynamic model of the arm, the actuators (muscles) are assumed to act as viscoelastic torque generators, the human interface forces are measured by a sensor, and the realized trajectory is measured.
+
+If intent could be known during disturbance, we would be able to determine its composition through hypothesis testing.
+Shadmehr and Mussa-Ivaldi demonstrated that an unchanging intent passed through a simple arm model was sufficient to recreate human-like reaching movements in the presence of force disturbances\cite{shadmehr1994adaptive}.
+To determine whether or how intent changes following disturbance requires an infinite number of \textit{candidate} intents vying to best reproduce human motion over an infinite number of statistical tests.
+Optimal feedback control is presently the best and changes intent over the course of a reach\cite{todorov2002optimal}.
+This approach has many free parameters that need to be tuned to reproduce a reach.
+By fitting the parameters directly to a library of known intent and state trajectories, a single, maximum-likelihood model could be found.
+This also represents a means of potentially falsifying optimal feedback control by showing that the maximum likelihood model explains little variance or requires a number of free parameters that grows with the library.
+Knowing intent would therefore represents a significant advance because it eliminates the need to compare many candidate models.
 

### 2014-10-01 Paper1/paper1.tex
Updated-ish

--- 2014-08-12 Paper1/paper1.tex
@@
 \usepackage{cite}
+\usepackage[section]{placeins}
 
@@
 \usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-\usepackage[noperiod]{jabbrv}
+%\usepackage[noperiod]{jabbrv}
 % Use the PLoS provided bibtex style
-\bibliographystyle{jneurophys}
+\bibliographystyle{plos2009}
 
@@
 Therefore, it could not well-represent the intent of a simple reaching movement.
-We propose a revised method --- \textit{intent extraction} --- that deduces intended trajectory, even in the face of disturbances.
+We propose a revised and more general method, \textit{intent extraction}, which calculates the intended trajectory even in the face of disturbances by relating environmental interaction and state to intent.
+This method affords new ways to 1) test and falsify candidate models of motor control, 2) study the timing and composition of alterations to intent during movement, and 3) understand how intent might be altered by pathology after neural insult.
 This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory.
@@
 
-\section*{Figure Legends}
+\pagebreak
+\section*{Figures and Figure Legends}
 %\begin{figure}[!ht]
@@
 %\end{figure}
-
 \begin{figure}[!ht]
 \begin{center}
-%\includegraphics[width=4in]{fig1.eps}
+\includegraphics[width=6in]{fig1.png}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig2.eps}
+\includegraphics[width=6in]{fig2.png}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig3.eps}
+\includegraphics[width=6in]{fig3.png}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=4in]{fig4.eps}
+\includegraphics[width=4in]{fig4.png}
 \end{center}

### 2014-10-01 Paper 3/loadDataForVPF.m

@@ -1,0 +1,47 @@
+clc
+clear all
+
+c=dir('./org_data');
+
+cMin=[0 1 0];
+cMax=[1 0 1];
+
+STEP=5;
+
+maxUts(2)=0.1970;
+
+for S=2 %[2 7 8];
+    figure(S)
+    clf
+    hold on
+    for k=1:length(c)
+        r=regexp(c(k).name,['outputsS',num2str(S,'%1.2i')]);
+        if ~isempty(r)
+            st=open(['./org_data/',c(k).name]);
+
+            X=st.Robot_arm.smooth_pos;
+            V=st.Robot_arm.vel;
+            t=st.Robot_arm.time;
+
+            A=[gradient(V(:,1),t) gradient(V(:,2),t) gradient(V(:,3),t)];
+
+            [Ut,Ux,Xgrid]=makeVPF(t,X,V,A);
+
+            Ut=Ut-min(Ut); %Just a reference anyways
+            mUt=max(Ut);
+            nUt=Ut/maxUts(S);
+
+            lnUt=log(nUt+.01);
+            
+            if mUt>maxUts(S)
+                maxUts(S)=mUt;
+            end
+            
+            for kk=STEP+1:length(t)
+                plot3(X(kk-STEP:kk,1),X(kk-STEP:kk,2),X(kk-STEP:kk,3),'color',cMin*(1-lnUt(kk))+cMax*lnUt(kk))
+            end
+        end
+    end
+    
+    axis equal
+end

### 2014-10-01 Paper2/krlsProofsAndToys/main.m

@@ -159,5 +159,5 @@ Wm=W(26:226,:);
 dt=t_step;
 yfit=summed(:,3:4);
 
-P=fminunc(@supMJPfit,[Wm(:);ks*ones(length(t),1)],optimoptions(@fminunc,'GradObj','on'));
+P=fminunc(@supMJPfit,[Wm(:);ks*ones(length(t),1)],optimoptions('GradObj','on'));
 

### 2014-10-01 matlab/finalfig3.m

@@ -21,7 +21,7 @@ lw=1.5;
 
 SKIP=4;
 qscale=.002;
-fade=0 %.7;
+fade=.7;
 
 gray=.5*[1 1 1];
 green=[.1 .7 .3];
@@ -47,11 +47,11 @@ for A=1:length(anecdote)
     m=means(anecdote(A).SE(1));
     cosa=cos(anecdote(A).rot);
     sina=sin(anecdote(A).rot);
-    rotmat=[cosa sina;-sina cosa]
+    rotmat=[cosa sina;-sina cosa];
     for c=1:length(f)
         k=f(c);
         X=[trials(k).x(:,1)-m,trials(k).x(:,2)-.5]*rotmat;
-        %plot(X(:,1),X(:,2)+anecdote(A).yoff,'color',green,'linewidth',.1)
+        plot(X(:,1),X(:,2)+anecdote(A).yoff,'color',green,'linewidth',.1)
         %plot(undisturbed(k).y(:,1),undisturbed(k).y(:,2),'-','linewidth',.0000001,'color',[1 fade fade])
     end
 
@@ -77,23 +77,23 @@ set(gcf,'color',[1 1 1])
 axis equal
 
 %annotate(h);
-% p=[.17 -.105;.17 .006;.15 -.152;.123 .025];
-% d=[1 -1; -1 -1;1 1;-1 -1];
-% colors=[gray; green; red; 0 0 0];
-% labs{1}='Force Disturbance';
-% labs{2}='Undisturbed Hand Trajectory';
-% labs{3}='Extracted Desired Hand Trajectory';
-% labs{4}='Hand Trajectory';
-% 
-% h=annotate(p,d,labs,colors,.015);
-% 
-% 
-% plot([.01 .04],-.01*[1 1],'k','linewidth',3)
-% text(.025,-.011,'3 cm','horizontalalignment','center','verticalalignment','top')
-% 
-% A=.008*[1 1];
-% arrow(A,A+[0 qscale*10],gray,.3,2)
-% text(.006,(.008+qscale*5),'10 N','rotation',90,'Horizontalalignment','center','Verticalalignment','bottom')
+p=[.17 -.105;.17 .006;.15 -.152;.123 .025];
+d=[1 -1; -1 -1;1 1;-1 -1];
+colors=[gray; green; red; 0 0 0];
+labs{1}='Force Disturbance';
+labs{2}='Undisturbed Hand Trajectory';
[…]

### 2014-10-27 Paper1/paper1.tex

@@
 We next examine the model’s sensitivity to parameter uncertainty, and test the approach using human reaching data.
-We found that sensitivity to parameter inaccuracies was on the order of square microns, and hence much lower than the inaccuracies seen in actual human motions.
+We found that sensitivity to parameter inaccuracies was below one square millimeter, and hence much lower than the inaccuracies seen in actual human motions.
 Further, the largest sensitivity arose mainly from uncertainty in our assumptions of joint stiffness.
@@
 Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned.
-Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.
+Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as crowd control, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.
 
-It is important to distinguish our use of the term intent from other groups that consider \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
+It is important to distinguish intended trajectory from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
 In typical motor control studies, subjects are motivated to complete the experiment in a timely fashion and are usually explicitly provided targets to be reached.
@@
 By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement.
+Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not consider mass as a feedforward controller would.
 Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
@@
 
-Here we develop and analyze the mathematics underlying intent extraction, a method to determine one's intent.
+Here we develop and analyze the mathematics underlying intent extraction, a method to determine one's intended trajectory.
 We focused, for simplicity in this initial study, on planar targeted arm reaching actions.
@@
 We show efficacy on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances.
-Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intent may change following a disturbance.
+Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intended trajectory may change following a disturbance.
 
@@
 \end{equation}  
-$\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$.
+$\ddot{q}_d$ can then be integrated numerically using a differential equation solver to determine the intended state trajectory $q_d(t)$.
 This relies on many assumptions: The model of plant and controller must be accurate and precise.
@@
 
-\subsection*{Generation of synthetic data using simple dynamic simulation}
-While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models.
-Here we show the results from our choice of Burdet and colleagues \cite{burdet2006stability}.
+\subsection*{Dynamic simulation of arm and intended trajectories}
+While the derivation above is general, testing its application to human reaching requires choosing plant and actuator models, which themselves require physical parameters.
 Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
 Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
-Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
-For the actuator, we choose the model of Burdet et al.
+For the actuator, we choose the model and parameters of Burdet et al.
 \cite{burdet2006stability}.
 A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged.
+Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
 In this  model, musculoskeletal stiffness was linearly related to torque.
@@
 To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
+Nominal parameters values were mostly taken from taken Dempster \cite{dempster1955space} and Burdet et al.
+\cite{burdet2006stability} with typical values where subject-specific parameters (height and weight) and direct where necessary (force sensor noise and drift).
+Standard deviations reflect expected variation in repeated measurement (ie.
+height and weight) or measured variance (ie.
+trial-to-trial variation in sensor noise or uncertainty in mass ratios reported by Dempster \cite{dempster1955space}).
 Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range.
@@
 The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
-Visual feedback was performed at 60 Hz using OpenGL.
+Visual feedback of hand position was performed at 60 Hz using OpenGL.
 Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
@@
 \subsection*{Statistical Analysis}
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons of maximum deviation among subjects and trajectory sources (hand and deviation) using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
 We used an alpha level of 0.05 for declaring statistical significance.
@@
 \cite{saltelli2010variance}.
-Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces.
-Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$.
-These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).
+Expected variance due to parameter uncertainty was $0.86  mm^2$ for pulse forces and $1.5 mm^2$ for filtered white forces.
+Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 mm^2$.
+These variances describe mean unsigned error, but do not address maximum error.
 

### 2014-11-05 Paper1/paper1.tex
Updated manuscript

@@
 % graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
+% include graphics with the command \includegraphicsf
 \usepackage{graphicx}
@@
 By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not consider mass as a feedforward controller would.
+Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would.
 Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
@@
 \bf{Arm Parameter Values and Sensitivities}}
-\begin{tabular}{|c|c|c c|c c|}
+\begin{tabular}{|c|c|c c|c c|c c|c|}
 \hline
@@
 Nominal &
+Source &
 SD &
+Source &
 $S$ &
@@
 0.33 &
+TAM &
 0.01 &
+RM &
 0 &
@@
 0.34 &
+TAM &
 0.01 &
+RM &
 0 &
@@
 0.436 &
+\cite{winter2009biomechanics} &
 0.0695 &
+RM &
 0 &
@@
 0.682 &
+\cite{winter2009biomechanics} &
 0.0431 &
+RM &
 0 &
@@
 79.36 &
+TAM &
 3.1 &
+\cite{stunkard1981accuracy} &
 0 &
@@
 0.028 &
+\cite{winter2009biomechanics} &
 0.0029 &
+\cite{dempster1955space} &
 0 &
@@
 0.022 &
+\cite{winter2009biomechanics} &
 0.0025 &
+\cite{dempster1955space} &
 0 &
@@
 0.322 &
+\cite{winter2009biomechanics} &
 0.0161 &
+5\% &
 0 &
@@
 0.468 &
+\cite{winter2009biomechanics} &
 0.0234 &
+5\% &
 0 &
@@
 -0.057 &
+RM &
 0.02 &
+RM &
 0.02 &
@@
 0.88 &
+RM &
 0.02 &
+RM &
 0 &
@@
 0 &
+RM &
 0.231 &
+RM &
 0 &
@@
 0 &
+RM &
 0.1067 &
+RM &
 0.03 &
@@
 0 &
+RM &
 0.1653 &
+RM &
 0 &
@@
 0 &
+MR &
 0.2869 &
@@
 1 &
+\cite{burdet2006stability} &
 0.15 &
@@
 0.0833 &
+\cite{burdet2006stability} &
 0.0125 &
@@
 0.02 &
+\cite{burdet2006stability} &
 0.003 &
@@
 2 &
+\cite{burdet2006stability} &
 0.3 &
+15\% &
 0 &
@@
 Total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
+While most parameter values were available in the literature, some were chosen to correspond to a typical adult male (TAM), empirically through repeated measurements (RM), or as a percentage of the mean.
 \end{flushleft}

### 2014-11-05 Paper 3/2Dfreeexploration/batch_boltzmann_stroke.m

@@ -4,7 +4,7 @@ close all
 
 outs=zeros(10,4);
 
-for k=2 %1:10
+for k=1 %:20
     subname=num2str(k,'%2.2d');
     fname=['free_exp_',subname,'stroke.mat'];
     

### 2014-11-05 Paper 3/3Dfreeexploration/boltzmannsubjectplots.m

@@ -1,261 +1,165 @@
-function ofinterest=boltzmannsubjectplots(t,x,v,a,fnumber)
+function ofinterest=boltzmannsubjectplots(t,x,v,a)
 
 global fitmecounts fitmebins
 
 %% Set constants
 numbins=40;
-lpctile=10;
-upctile=99.5; %Throw away the craziest outliers to help hist not find zeros
+lpctile=0; %This needs to be fixed instead by throwing out "stops"
+upctile=99; %Throw away the craziest outliers because of Farnaz's coding quirks
 
 sWidth=2;
-sHeight=4;
+sHeight=2;
 
-%% Compute Ut, Tt
-Pt=dot(v',a')';
-Ut=cumtrapz(t,Pt);
-Ut=Ut-min(Ut); %Flush to zero just for ease of reference
-Tt=.5*(v(:,1).^2+v(:,2).^2);
-ofinterest.maxT=max(Tt);
+%% Compute Pt, Tt
+m=150*0.453592*.0597; %Dempster says arm is 5.97% of tbm, assume 150 lbs. Obviously effective mass is different and subject specific.
+Pt=m*dot(v',a')';
+Tt=.5*m*dot(v',v');
 
-%% Bin and count Ut levels, plot this as a Boltzmann distribution
-
-figure(fnumber)
-clf
+%% Bin and count Pt levels, plot raw and fit
 subplot(sWidth,sHeight,1)
-[R2P,W,aPbins,aPcounts,Pfitcounts]=histBoltzmann(abs(Pt),lpctile,upctile,1);
-ofinterest.R2P=R2P;
-ofinterest.W=W;
-title(['Boltzmann Distribution Fit R^2 = ',num2str(R2P,3)])
-xlabel('|Work|/Mass, (m/s)^2')
-ylabel(['Fraction of Time Points (N=',num2str(length(t)),')'])
-legend('Data',['Boltzmann Fit, W=',num2str(W,2)])
+hold on
 
-% fitmecounts=aPcounts(aPbins<.025);
-% fitmebins=aPbins(aPbins<.25);
-% xfit=fmincon(@bruteforceexp,1,-1,0)
-% 
-% fitcounts=1/xfit*exp(-aPbins/xfit);
-% fitcounts=fitcounts/sum(fitcounts);
-% figure(20)
[…]

### 2014-11-10 Paper1/paper1.tex
Updated to get paper1 back out

@@
 m &
-0.33 &
-TAM &
-0.01 &
-RM &
+0.353 &
+\cite{dempster1955space} &
+0.017 &
+\cite{dempster1955space} &
 0 &
@@
 m &
-0.34 &
-TAM &
-0.01 &
-RM &
+0.363 &
+\cite{dempster1955space} &
+0.011 &
+\cite{dempster1955space} &
 0 &
@@
 0.0695 &
-RM &
+15\% &
 0 &
@@
 0.0431 &
-RM &
+15\% &
 0 &
@@
 kg &
-79.36 &
-TAM &
+88.4 &
+\cite{fryar2012anthropometric} &
 3.1 &
@@
 0.0161 &
-5\% &
+15\% &
 0 &
@@
 0.0234 &
-5\% &
+15\% &
 0 &
@@
 0 &
-MR &
+RM &
 0.2869 &
+RM &
 0 &
@@
 0.15 &
+15\% &
 0.06 &
@@
 0.15 &
+15\% &
 0 &
@@
 0.0125 &
+15\% &
 0 &
@@
 0.003 &
+15\% &
 0 &
@@
 Total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
-While most parameter values were available in the literature, some were chosen to correspond to a typical adult male (TAM), empirically through repeated measurements (RM), or as a percentage of the mean.
+Most parameter values come from the literature, but instrument-specific parameters were determined empirically through repeated measurements (RM).
+Where parameter values were unavailable, 15\% of the mean was used as a conservative estimate
 \end{flushleft}

### 2014-11-10 matlab/sensitivityPaper/annotate.m

@@ -1,0 +1,41 @@
+function h=annotate(p,d,labs,colors,alength)
+
+if nargin==1
+    for k=1:length(p)
+        delete(p(k).t)
+        delete(p(k).a(1))
+        delete(p(k).a(2))
+    end
+    h=[];
+    return
+end
+
+latexscale=1.2;
+if length(alength)==1
+    alength=ones(size(colors,1),1)*alength;
+end
+
+for k=1:length(labs)
+    x=p(k,1);
+    y=p(k,2);
+    l=-alength(k)*d(k,:)/norm(d(k,:));
+    h(k).a=arrow([x y]+latexscale*l,[x y],colors(k,:),.3,2);
+    
+    if d(k,1)>0
+        ha='right';
+    elseif d(k,1)<0
+        ha='left';
+    else
+        ha='center';
+    end
+    
+    if d(k,2)>0
+        va='top';
+    elseif d(k,2)<0
+        va='bottom';
+    else
+        va='middle';
+    end
+    
+    h(k).t=text(x+l(1), y+l(2),labs{k},'horizontalalignment',ha,'Verticalalignment',va,'color',colors(k,:));
+end

### 2014-11-10 matlab/sensitivityPaper/fig1_uncertainty.m

@@ -1,0 +1,84 @@
+clc
+clear all
+
+qscale=.001;
+SKIP=2;
+yoff=.04;
+
+gray=.5*[1 1 1];
+green=[.1 .15 .7];
+red=[1 .5 .1];
+pink=[1 .8 .8];
+blue=[.8 .8 1];
+black=[0 0 0];
+offset=[0 -.06 -.1];
+
+setGlobals(paramsPopulator('burdet'))
+
+figure(1)
+clf
+hold on
+batches=[7 8 2];
+
+coeff=calcminjerk([0 .5],[.15 .5],[0 0],[0 0],[0 0],[0 0],0,.7);
+
+for k=1:3
+    load(['BATCH',num2str(batches(k)),'.mat'])
+    
+    rp=randperm(1000);
+    for kk=1:50
+        plot(simA(rp(kk)).y(:,1),simA(rp(kk)).y(:,2)+offset(k),'Color',pink,'linewidth',.5)
+    end
+    
+    y=extract(t,xvaf,'reflex');
+    
+    tcalc=t;
+    tcalc(t>=.7)=.7;
+    x=minjerk(coeff,tcalc)';
+    
+    plot(xvaf(:,1),xvaf(:,2)+offset(k),'-','Color',black,'Linewidth',2)
+    X=[xvaf(1:SKIP:end,1) xvaf(1:SKIP:end,2)+offset(k)];
+    Y=X+qscale*xvaf(1:SKIP:end,7:8);
+    arrow(X,Y,gray,.3)
+    plot(x(1:SKIP:end,1),x(1:SKIP:end,2)+offset(k),'o-','Color',green,'markerfacecolor',green)
+    plot(y(1:SKIP:end,1),y(1:SKIP:end,2)+offset(k),'.-','Color',red,'markersize',12)
+    
+    
+end
+
+margin=.02;
+set(gcf,'position',[250 300 625 305])
+set(gca,'position',[margin margin 1-margin 1-margin],'units','normalized')
+set(gcf,'color',[1 1 1])
+
+axis equal
+axis off
+
+latexscale=1.2;
+
+text(0,.54,'Early Pulse','horizontalalignment','left','Verticalalignment','top','color','k')
+text(0,.48,'Late Pulse','horizontalalignment','left','Verticalalignment','top','color','k')
+text(0,.4325,'Filtered Gaussian','horizontalalignment','left','Verticalalignment','top','color','k')
+
+%annotate(h)
+colors=[gray;black;green;red;red];
[…]

### 2014-11-10 matlab/sensitivityPaper/final_fig2.m

@@ -1,10 +1,8 @@
 clc
 clear all
 
-if ~exist('justMUE4fig2.mat','file')
-    whites=6:10;
-    whites=whites(randperm(5));
-    loadme=[whites(1:4) 11:14];
+if ~exist('justMUE4fig2.mat','file')|1
+    loadme=1:8;
 
     bins=0:.005:.15;
 
@@ -13,7 +11,7 @@ if ~exist('justMUE4fig2.mat','file')
     for k=1:length(loadme)
         load(['BATCH',num2str(loadme(k)),'.mat'])
         yex=extract(t,xvaf,'reflex');
-
+        
         [trash,ref]=getMUE(bins,0*bins,yex);
 
         mueA=zeros(1000,1);
@@ -21,15 +19,18 @@ if ~exist('justMUE4fig2.mat','file')
         mueAB=zeros(1000,20);
 
         for c=1:1000
+            if ~mod(c,100)
+                plot(simA(c).y(:,1),simA(c).y(:,2),'g')
+            end
             mueA(c)=getMUE(bins,ref,simA(c).y);
             mueB(c)=getMUE(bins,ref,simB(c).y);
             for cc=1:20
-                mueAB(c,cc)=1000*getMUE(bins,ref,simAB(cc,c).y);
+                mueAB(c,cc)=getMUE(bins,ref,simAB(cc,c).y);
             end
         end
-        justMUE(k).mueA=mueA;
-        justMUE(k).mueB=mueB;
-        justMUE(k).mueAB=mueAB;
+        justMUE(k).mueA=1000*mueA;
+        justMUE(k).mueB=1000*mueB;
+        justMUE(k).mueAB=1000*mueAB;
     end
 
     save('justMUE4fig2.mat','justMUE')

### 2014-11-10 matlab/sensitivityPaper/main.m

@@ -48,72 +48,75 @@ xvafsm=[xsimSM f];
 
 save('baselines.mat','yex','yexsm')
 
+
 %% Step 3a: OAT - Burdet
-
-clear OAT
-
-names=paramsPopulator('names');
-dat=paramsPopulator('burdet');
-f=find(dat(:,3));
-
-params=paramsPopulator(f(1));
-sp1=size(params,1);
-lf=length(f);
-
-vals=zeros(lf,sp1);
-
-for k=1:lf
-    params=paramsPopulator(f(k));
-    if k==1
-        OAT=repeatedSim(params,t,xvaf,'reflex',0,lf);
-    else
-        OAT(k,:)=repeatedSim(params,t,xvaf,'reflex',0,lf-k+1);
+if 0
+    clear OAT
+    
+    names=paramsPopulator('names');
+    dat=paramsPopulator('burdet');
+    f=find(dat(:,3));
+    
+    params=paramsPopulator(f(1));
+    sp1=size(params,1);
+    lf=length(f);
+    
+    vals=zeros(lf,sp1);
+    
+    for k=1:lf
+        params=paramsPopulator(f(k));
+        if k==1
+            OAT=repeatedSim(params,t,xvaf,'reflex',0,lf);
+        else
+            OAT(k,:)=repeatedSim(params,t,xvaf,'reflex',0,lf-k+1);
+        end
+        for kk=1:sp1
+            vals(k,kk)=params(kk,f(k));
+        end
     end
-    for kk=1:sp1
-        vals(k,kk)=params(kk,f(k));
-    end
-end
-
-for k=1:lf
-    for kk=1:sp1
-        OAT(k,kk).name=names{f(k)};
-        OAT(k,kk).val=vals(k,kk);
+    
+    for k=1:lf
+        for kk=1:sp1
+            OAT(k,kk).name=names{f(k)};
+            OAT(k,kk).val=vals(k,kk);
+        end
     end
 end
 
 %% Step 3b: OAT - ShadMuss
 
-clear OATSM
-dat=paramsPopulator('shadmuss');
-f=find(dat(:,3));
-
-params=paramsPopulator(f(1));
-sp1=size(params,1);
-lf=length(f);
-
-vals=zeros(lf,sp1);
-
-for k=1:lf
-    params=paramsPopulator(f(k));
-    if k==1
[…]

### 2014-11-10 matlab/sensitivityPaper/main_postsim.m

@@ -1,7 +1,7 @@
 load baselines.mat
-load OAT_KICK.mat
+%load OAT_KICK.mat
 load KICK.mat
-load KICKsm.mat
+%load KICKsm.mat
 
 %% fig 1
 t=0:.005:2;

### 2014-11-10 Paper1/paper1.tex
Figs 1 and 2 with some updates to the .tex for paper1.

@@
 \caption{
-{\bf Simulated data illustrating tautology of extraction absent parameter error across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial.
+{\bf Simulated data illustrating tautology of extraction across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial.
 Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory.
 Extraction to recover intention from arm and force trajectory follows.
+Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity.
+Examining the distribution of trajectories from this analysis reveals that this sensitivity is not uniformly distributed throughout a reach.
 }

### 2014-11-10 matlab/sensitivityPaper/final_fig2.m

@@ -1,7 +1,7 @@
 clc
 clear all
 
-if ~exist('justMUE4fig2.mat','file')|1
+if ~exist('justMUE4fig2.mat','file')
     loadme=1:8;
 
     bins=0:.005:.15;

### 2014-11-10 Paper1/paper1.tex
updated table 1

@@
 \usepackage{cite}
+
+% PLoS doesn't support/use the next package, but we get issues if we don't use this to generate the PDF. Comment this line out when submitting.
 \usepackage[section]{placeins}
@@
 \cite{dempster1955space} &
-0 &
-0.10 \\
+0.03 &
+0.17 \\
 Forearm Length ($L_2$) &
@@
 15\% &
-0 &
-0.02 \\
+0.01 &
+0.07 \\
 Gross Body Mass ($m_g$) &
@@
 15\% &
-0 &
-0.01 \\
+0.01 &
+0.06 \\
 Forearm Radius of Gyration Ratio &
@@
 0 &
-0.01 \\
+0.04 \\
 Shoulder Parallel Coordinate &
@@
 RM &
-0.02 &
-0.18 \\
+0.01 &
+0.14 \\
 Shoulder Perpendicular Coordinate &
@@
 RM &
-0 &
-0.08 \\
+0.02 &
+0.06 \\
 Force Sensor Miscalibration $x\text{-axis}$ &
@@
 RM &
-0 &
-0.08 \\
+0.01 &
+0.05 \\
 Force Sensor Miscalibration $y\text{-axis}$ &
@@
 15\% &
-0.06 &
-0.31 \\
+0.05 &
+0.19 \\
 Torque-Varying Impedance Misestimation Ratio &
@@
 15\% &
-0 &
-0.15 \\
+0.05 &
+0.19 \\
 Damping-to-Stiffness Ratio ($k_d$) &
@@
 15\% &
-0 &
+0.01 &
 0.08 \\
@@
 15\% &
-0 &
+0.01 &
 0.04 \\
@@
 Sensitivity indices less than one-thousandth were reported as zero.
-Total variance was $1.3 \mu m^2$, showing very small average deviations when using this approach.
+Total variance was $1.56 mm^2$, showing very small average deviations when using this approach.
 First-order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor.
@@
 Most parameter values come from the literature, but instrument-specific parameters were determined empirically through repeated measurements (RM).
-Where parameter values were unavailable, 15\% of the mean was used as a conservative estimate
+Where parameter values were unavailable, 15\% of the mean was used as a conservative estimate.
 \end{flushleft}

### 2014-11-19 Paper1/paper1B.tex

--- 2014-11-10 Paper1/paper1.tex
@@
 % Template for PLoS
-% Version 1.0 January 2009
+% Version 2.0 July 2014
 %
@@
 % dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% Be advised that this is merely a template 
+% designed to facilitate accurate translation of manuscript content 
+% into our production files. 
+%
+% This template contains extensive comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template 
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication and enters production, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
+% Packages included in this template are intentionally
+% limited and basic in order to reduce the possibility
+% of issues during our production process.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text.
+For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
+%
+% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
+% For bolding characters in mathmode, please use $\mathbf{}$ 
+%
+% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
+%
+% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
+%
+% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
+%
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphicsf
-\usepackage{graphicx}
-
 % cite package, to clean up citations in the main text. Do not remove.
@@
 
-% PLoS doesn't support/use the next package, but we get issues if we don't use this to generate the PDF. Comment this line out when submitting.
-\usepackage[section]{placeins}
+\usepackage{hyperref}
 
-\usepackage{color}
+% line numbers
+\usepackage{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% rotating package for sideways tables
+%\usepackage{rotating}
+
+% If you wish to include algorithms, please use one of the packages below.
+Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
+%\usepackage{algorithmic}
+%\usepackage{algorithmicx}
 
@@
 \usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-%\usepackage[noperiod]{jabbrv}
-% Use the PLoS provided bibtex style
+
+% Use the PLoS provided BiBTeX style
 \bibliographystyle{plos2009}
@@
 \pagestyle{myheadings}
-%% ** EDIT HERE **
 
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
+%% Include all macros below. Please limit the use of macros.
 
@@
 
-\begin{titlepage}
-\begin{center}
+% Title must be 150 characters or less
+\begin{flushleft}
 {\Large
 \textbf{I meant to do that: determining the intentions of action in the face of disturbances}
-}\\[1cm]
-\end{center}
-
-\begin{flushleft}
+}
 % Insert Author names, affiliations and corresponding author email.
+\\
 Justin Horowitz$^{1}$, 
@@
 \\
-$\ast$ Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States. E-mail: pattonj@uic.edu
-\\[1cm]
-Running Head: \\
-Determining the intention of action despite disturbances
+$\ast$ E-mail: pattonj@uic.edu
 \end{flushleft}
-\end{titlepage}
 
@@
 We derive a novel modeling approach for determining motor intent from a reaching motion that is disturbed by an unexpected force.
-We first show how to mathematically invert both feedforward (predictive) and feedback controls to obtain intent.
+We first show how to mathematically invert both feedforward (predictive) and feedback controls to obtain an intended trajectory.
 We next examine the model’s sensitivity to parameter uncertainty, and test the approach using human reaching data.
-We found that sensitivity to parameter inaccuracies was below one square millimeter, and hence much lower than the inaccuracies seen in actual human motions.
+We found that sensitivity to parameter inaccuracies was 1.6 $\mathrm{mm}^2$, and hence lower than the inaccuracies seen in actual human motions.
 Further, the largest sensitivity arose mainly from uncertainty in our assumptions of joint stiffness.
@@
 
-\section*{Keywords}
-Intention; Motor Control; Visually Guided Reaching; On-line Control
-
 % Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
-In our everyday struggles with bumpy roads and bustling crowds we often make movements that do not match what we had planned.
-Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do.
-We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model.
-Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight.
-Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance.
-This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from small organisms to crowds of people.
-It is also a new tool for understanding and treating impaired movement plans following neural injury.
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+%\section*{Author Summary}
 
@@
 Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intended trajectory may change following a disturbance.
-
-% M E T H O D S
 
@@
 \cite{burdet2006stability}.
-A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged.
-Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
+Impedance magnitude is known to be task-dependent \cite{gomi1998task}.
+We fit a subject-specific constant scale factor relative to the impedance model of Burdet et al.
+\cite{burdet2006stability} using linear regression, as in Fraon the task-perpendicular data from the first 150 ms of movement in their noise-disturbed trials.
+
+This IS the linear regression from Franklin and Milner 2003.
+It's jut that we're using task-perpendicular since that's the only part we really care about and so we have to use a constant scale instead of a whole matrix because the subject is unambiguously in motion.
+
+Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 ms starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
 In this  model, musculoskeletal stiffness was linearly related to torque.
 Like Burdet et al.
-\cite{burdet2006stability}, we approximated muscle torque by looking backward in time 500 microseconds.
+\cite{burdet2006stability}, we approximated muscle torque by looking backward in time $500 \mathrm{\mu s}$.
 Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.
@@
 We used the following two types of perturbing forces for each combination of distance and direction:
-(1) pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
-(2) Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion.
+\begin{enumerate}
+\item Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
+\item Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion.
 The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ rad/s.
-
+\end{enumerate}
 Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target.
@@
 To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
-Nominal parameters values were mostly taken from taken Dempster \cite{dempster1955space} and Burdet et al.
-\cite{burdet2006stability} with typical values where subject-specific parameters (height and weight) and direct where necessary (force sensor noise and drift).
+Nominal parameters values were mostly taken from Dempster \cite{dempster1955space} and Burdet et al.
+\cite{burdet2006stability} with typical values where subject-specific parameters (height and weight) and direct measurements where necessary (force sensor noise and drift).
 Standard deviations reflect expected variation in repeated measurement (ie.
@@
 \subsection*{Statistical Analysis}
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons of maximum deviation among subjects and trajectory sources (hand and deviation) using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
+To determine when the intended trajectory's deviation exceeded the deviation that could be explained by undisturbed reaching, we compared the maximum deviation of the undisturbed reaches to the instantaneous deviation of the pulse-disturbed reaches at 5 ms intervals.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
 We used an alpha level of 0.05 for declaring statistical significance.
-
-% R E S U L T S
 
@@
 Even after correcting for drift in our force measurements, the force measurements reported during reaching can be expected to have noticeable bias and further noise on the order of that bias as reported in Table 1.
-Taken together, force measurement errors account for 21\% of all sensitivity.
+Taken together, force measurement errors account for 12\% of all sensitivity.
 We also examined noise in sensor signals explicitly in the more comprehensive sensitivity analysis, discussed below.
@@
 \cite{saltelli2010variance}.
-Expected variance due to parameter uncertainty was $0.86  mm^2$ for pulse forces and $1.5 mm^2$ for filtered white forces.
-Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 mm^2$.
+Expected variance due to parameter uncertainty was $2.24 \mathrm{mm}^2$ for pulse forces and $.3 \mathrm{mm}^2$ for filtered white forces.
+Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mathrm{mm}^2$.
 These variances describe mean unsigned error, but do not address maximum error.
@@
 The movements completed with the hand returning to undisturbed levels of error within 600 milliseconds, intention differed from actual hand path for as long as one second after the onset of motion.
-
-% D I S C U S S I O N
-
 \section*{Discussion}
@@
 
-% Do NOT remove this, even if you are not including acknowledgments
+% Do NOT remove this, even if you are not including acknowledgments.
 \section*{Acknowledgments}
-Research supported by NIH NS053606.
+Research supported by US NIH NINDS1 R01 NS053606.
 We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
@@
 
-\section*{Grants} 
-Supported by US NIH NINDS1 R01 NS053606
+\section*{References}
+\bibliography{paper1}
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% OR
+%
+% Compile your BiBTeX database using our plos2009.bst
+% style file and paste the contents of your .bbl file
+% here.
+%
 
-%\section*{References}
-% The bibtex filename
-\bibliography{paper1}
-
-\pagebreak
-\section*{Figures and Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
+\section*{Figure Legends}
+% This section is for figure legends only, do not include
+% graphics in your manuscript file.
+%
+%\begin{figure}
 %\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
+%{\bf Bold the first sentence.}  Rest of figure caption.  
 %}
@@
 %\end{figure}
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6in]{fig1.png}
-\end{center}
-\caption{
-{\bf Simulated data illustrating tautology of extraction across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial.
-Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory.
-Extraction to recover intention from arm and force trajectory follows.
-Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity.
-Examining the distribution of trajectories from this analysis reveals that this sensitivity is not uniformly distributed throughout a reach.
-}
-\label{fig:synthetic}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6in]{fig2.png}
-\end{center}
-\caption{
-{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1.
-First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value.
-Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
-}
-\label{fig:sensitivity}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6in]{fig3.png}
-\end{center}
-\caption{
-{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied.
-This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
-}
-\label{fig:anecdotes}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=4in]{fig4.png}
-\end{center}
-\caption{
-{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ half of the critical value for simultaneous comparisons at 95\% confidence interval) can no longer be explained by undisturbed hand deviation (green band is mean maximum deviation).
-Movement intention (red dots, band is mean deviation) does not significantly deviate until at least 300 ms after the onset of disturbance.
-}
-\label{fig:grouptrends}
-\end{figure}
 
 \section*{Tables}
+% 
+% See introductory notes if you wish to include sideways tables.
+%
+% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
+Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
+%
 %\begin{table}[!ht]
@@
 % \end{table}
-
 \begin{table}[!ht]
@@
 
+%\section*{Supporting Information Legends}
+%
+% Please enter your Supporting Information captions below in the following format:
+%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
+% 
+%\begin{description}
+%\item {\bf}
+%\item {\bf}
+%\end{description}
+
 \end{document}

### 2014-11-20 Paper1/paper1.tex
Does not matter, stylistic change

@@
 trial-to-trial variation in sensor noise or uncertainty in mass ratios reported by Dempster \cite{dempster1955space}).
-Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range.
+Distributions were clamped to the range of $\pm 3$ standard deviations in order to limit parameters to a realistic range.
 Combinations and calculations were made as prescribed by Saltelli et al.
@@
 Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity.
-Examining the distribution of trajectories from this analysis reveals that this sensitivity is not uniformly distributed throughout a reach.
+Examining the distribution of deviation in trajectories from this analysis reveals that sensitivity is not uniformly distributed throughout a reach.
 }

### 2014-11-20 Paper1/paper1B.tex

@@
 \cite{burdet2006stability}.
-Impedance magnitude is known to be task-dependent \cite{gomi1998task}.
-We fit a subject-specific constant scale factor relative to the impedance model of Burdet et al.
-\cite{burdet2006stability} using linear regression, as in Fraon the task-perpendicular data from the first 150 ms of movement in their noise-disturbed trials.
-
-This IS the linear regression from Franklin and Milner 2003.
-It's jut that we're using task-perpendicular since that's the only part we really care about and so we have to use a constant scale instead of a whole matrix because the subject is unambiguously in motion.
+We use a subject-specific constant scale factor, $c$, relative to the model of Burdet et al.
+\cite{burdet2006stability} to account for task-dependence of the subjects' impedance\cite{gomi1998task}.
+\begin{equation}
+\tau_{tp}=c(K_p\Delta q+K_d \Delta \dot{q})
+\end{equation}
+This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that the task-perpendicular data from the first 150 ms of movement in their noise-disturbed trials is used in place of a stationary task.
 
@@
 \subsection*{Statistical Analysis}
-To determine when the intended trajectory's deviation exceeded the deviation that could be explained by undisturbed reaching, we compared the maximum deviation of the undisturbed reaches to the instantaneous deviation of the pulse-disturbed reaches at 5 ms intervals.
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.
-We used an alpha level of 0.05 for declaring statistical significance.
+To determine when the intended trajectory's deviation exceeded the deviation that could be explained by undisturbed reaching, we compared the maximum deviation of the undisturbed reaches to the instantaneous deviation of the pulse-disturbed reaches during 100 ms intervals.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform one-tailed T-tests with an alpha level of 0.05.
+This also provided a confidence interval on the difference.
 

### 2014-11-24 Paper1/paper1B.tex
Mild updates

@@
 \pagestyle{myheadings}
+
+% PLoS doesn't support/use the next package, but we get issues if we don't use this to generate the PDF. Comment this line out when submitting.
+\usepackage[section]{placeins}
 
@@
 % 
-
+\pagebreak
 \section*{Figure Legends}
@@
 %\end{figure}
+\begin{figure}
+\caption{
+{\bf Simulated data illustrating tautology of extraction across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial.
+Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory.
+Extraction to recover intention from arm and force trajectory follows.
+Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity.
+Examining the distribution of deviation in trajectories from this analysis reveals that sensitivity is not uniformly distributed throughout a reach.
+}
+\label{fig:synthetic}
+\end{figure}
+
+\begin{figure}
+\caption{
+{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1.
+First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value.
+Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
+}
+\label{fig:sensitivity}
+\end{figure}
+
+\begin{figure}
+\caption{
+{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied.
+This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
+}
+\label{fig:anecdotes}
+\end{figure}
+
+\begin{figure}
+\caption{
+{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ half of the critical value for simultaneous comparisons at 95\% confidence interval) can no longer be explained by undisturbed hand deviation (green band is mean maximum deviation).
+Movement intention (red dots, band is mean deviation) does not significantly deviate until at least 300 ms after the onset of disturbance.
+}
+\label{fig:grouptrends}
+\end{figure}
+\pagebreak
 

### 2015-01-07 Paper1/paper1B.tex

@@
 {\Large
-\textbf{I meant to do that: determining the intentions of action in the face of disturbances}
+\textbf{I Meant to Do That: Determining the Intentions of Action in the Face of Disturbances}
 }
@@
 \section*{Abstract}
-Our actions often do not match our intentions when there are external disturbances.
-We derive a novel modeling approach for determining motor intent from a reaching motion that is disturbed by an unexpected force.
-We first show how to mathematically invert both feedforward (predictive) and feedback controls to obtain an intended trajectory.
-We next examine the model’s sensitivity to parameter uncertainty, and test the approach using human reaching data.
-We found that sensitivity to parameter inaccuracies was 1.6 $\mathrm{mm}^2$, and hence lower than the inaccuracies seen in actual human motions.
-Further, the largest sensitivity arose mainly from uncertainty in our assumptions of joint stiffness.
-We also tested the hypothesis that this would predict straight-line intent in human targeted reaching movements, even when disturbed by various pulses of force.
-As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset.
-Beyond this point, however, we detected a change in intent, even when the intention was near the target.
-Knowing such an intent signal is broadly applicable in areas such as in enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and in any complex system that embodies planning, such as brain machine interfaces, team sports, crowds, or swarms.
+Our actions often do not match our intentions when there are external disturbances such as turbulence.
+We derived a novel modeling approach for determining this motor intent from targeted reaching motions that are disturbed by an unexpected force.
+First, we demonstrated how to mathematically invert both feedforward (predictive) and feedback controls to obtain an intended trajectory.
+We next examined the model's sensitivity to a realistic range of parameter uncertainties, and found that the expected inaccuracy due to all possible parameter mis-estimations was less than typical movement-to-movement variations seen when humans reach to similar targets.
+The largest sensitivity arose mainly from uncertainty in joint stiffnesses.
+Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 ms following the onset of a disturbance.
+As expected, the intended trajectory showed no change from undisturbed reaching for more than 300 ms after the disturbance onset.
+Beyond this point, however, we detected a change in intent, surprisingly even when the hand is already near the target.
+Knowing such an intent signal is broadly applicable: enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and complex systems that embody planning such as brain machine interfaces, team sports, crowds, or swarms.
 In addition, observing intent as it changes might act as a window into the mechanisms of planning, correction, and learning.
@@
 \section*{Introduction}
-There are many situations, such as disturbances, distractions, and pathologies, that interfere with execution of an intended movement.
-One can assume that a person who is familiar with a particular task will try to do the same thing each time, provided all other conditions are equal.
-This presumes that what people do is on average what they intended to do and that their goal remains unchanged.
-However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention.
-Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned.
-Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as crowd control, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.
+Disturbances, distractions, and pathologies can interfere in many situations and prevent our actions from matching our intent.
+A pilot, for example, may fail to complete a maneuver because of turbulence.
+Such challenges are broadly present in many human-machine interactions.
+One may speculate on how we might use these very same machines to elucidate the underlying intention.
+Such a possibility would be broadly useful in any area where intended actions might be thwarted by disturbances.
 
-It is important to distinguish intended trajectory from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
-In typical motor control studies, subjects are motivated to complete the experiment in a timely fashion and are usually explicitly provided targets to be reached.
-Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of actions (walking/standing/flexion/extension/etc).
+It is important to distinguish intended action from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
+In typical motor control studies, subjects are motivated to complete an experiment in a timely fashion, and are usually explicitly provided targets to reach to.
+Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of discrete actions (walking/standing/flexion/extension/etc).
 While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves.
@@
 Therefore, it could not well-represent the intent of a simple reaching movement.
-We propose a revised and more general method, \textit{intent extraction}, which calculates the intended trajectory even in the face of disturbances by relating environmental interaction and state to intent.
-This method affords new ways to 1) test and falsify candidate models of motor control, 2) study the timing and composition of alterations to intent during movement, and 3) understand how intent might be altered by pathology after neural insult.
-This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory.
-In human reaching, these initial ingredients are available: the process is a dynamic model of the arm, the actuators (muscles) are assumed to act as viscoelastic torque generators, the human interface forces are measured by a sensor, and the realized trajectory is measured.
 
-Here we develop and analyze the mathematics underlying intent extraction, a method to determine one's intended trajectory.
-We focused, for simplicity in this initial study, on planar targeted arm reaching actions.
-First, we produce synthetic reaching data using a forward model so that we can examine the quality and uncertainty of intent extraction.
-We then explore sensitivity and failure modes, demonstrating the accuracy and utility of this method.
-We show efficacy on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances.
-Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intended trajectory may change following a disturbance.
+Here we propose a revised and more general method, \textit{intent extraction}, which calculates the intended trajectory even in the face of disturbances by relating environmental interaction and state to intent.
+The method is a class of filters that can infer the intended trajectory from the disturbance (turbulence), process (dynamic equations), actuator (muscles), and motion.
+The method affords new ways to study models of motor control, understand the timing and composition of intent, and how it might be altered by disturbances and injuries.
+Below we first present the mathematics underlying intent extraction.
+Then, for simplicity in this initial study, we create synthetic reaching data to examine the quality and uncertainty of this method.
+Finally, we demonstrate its effectiveness extracting intended action from real human reaches that have been disrupted by unpredictable disturbances.
+We demonstrate this effectiveness by testing the hypothesis that disturbance does not change intended action for at least 120 milliseconds.
+This early experiment also provides some new insights on how intended trajectories can change in reaction to disturbances.
 
@@
 \section*{Methods}
-The sections below describe the theoretical method and the means by which we evaluate it.
-First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process.
-Next, we conducted an experiment on humans to evaluate the process using human subject data.
+The sections below describe the theoretical method used and the means by which we evaluated it.
+First, we present an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process.
+Next, we present an experimental study on humans that evaluated success in a simple test of the approach's ability to extract straight intentions even when disturbed.
 
@@
 Additionally, linear feedback rejects position and velocity error.
-This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches.
-Nevertheless, later we will discuss how this approach can be used on more complex models without loss of generality.
+This model was chosen to help illuminate the approach, but later we present how to generalize this approach to more complex models without loss of generality.
 The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
@@
 For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation is solved by numerical integration to determine the solution to the initial value problem in time.
-This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory.
+This entails algebraic manipulation to solve for $\ddot{q}$, followed by integration to determine the state trajectory.
 Intent determination takes the novel approach of instead solving for $\ddot{q}_d$:
@@
 It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators.
-Importantly, this has been claimed to be the intended movement.
+Importantly, this has been claimed to be the intended movement\cite{feldman1995origin}.
 To consider this, we also generalize our approach to any system with a moving intention that governs the control law.
@@
 The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
-Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
+Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some action space, $x$:
 \begin{equation} \label{eq:generalPlant}
@@
 In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance.
-As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined.
+As long as some bidirectional relationship  exists between state and outcome (even if determined empirically), intended outcome can be determined.
 This allows determination of intent in many situations of interest, even where the process is irreducibly complex.
 
-\subsection*{Dynamic simulation of arm and intended trajectories}
-While the derivation above is general, testing its application to human reaching requires choosing plant and actuator models, which themselves require physical parameters.
-Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
-Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
-For the actuator, we choose the model and parameters of Burdet et al.
-\cite{burdet2006stability}.
-We use a subject-specific constant scale factor, $c$, relative to the model of Burdet et al.
-\cite{burdet2006stability} to account for task-dependence of the subjects' impedance\cite{gomi1998task}.
-\begin{equation}
-\tau_{tp}=c(K_p\Delta q+K_d \Delta \dot{q})
-\end{equation}
-This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that the task-perpendicular data from the first 150 ms of movement in their noise-disturbed trials is used in place of a stationary task.
-
-Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 ms starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
-In this  model, musculoskeletal stiffness was linearly related to torque.
-Like Burdet et al.
-\cite{burdet2006stability}, we approximated muscle torque by looking backward in time $500 \mathrm{\mu s}$.
-Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.
-
 \subsection*{Experimental Design}
-We chose 15 centimeters long simulated reaches of the right arm, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then first proceeding to the right.
+We chose 15 centimeters long simulated reaches of the right arm, beginning at 38 centimeters out from and 5.7 centimeters left of the right shoulder and ending at 38 centimeters out from and 10.7 centimeters right of the right shoulder.
 We used the following two types of perturbing forces for each combination of distance and direction:
@@
 
+\subsection*{Dynamic simulation of arm and intended trajectories}
+While the derivation of intent determination is general, testing its application to human reaching requires choosing plant and actuator models, which themselves require physical parameters.
+Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms.
+Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
+For the actuator, we choose the model and parameters of Burdet et al.
+\cite{burdet2006stability}.
+We fit subject-specific constant scale factors for body mass, $m$, and impedance, $c$, relative to this model to account for atypical mass distribution and any task-dependence of the subjects' impedance\cite{gomi1998task}.
+\begin{equation}
+E=c(K_p\Delta q+K_d \Delta \dot{q})+m(I(q)\Delta\ddot{q})
+\end{equation}
+This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that task-perpendicular torque and displacement from the first 150 ms of movement from noise-disturbed trials is used in place of a stationary task.
+
+Desired trajectory in time was idealized as a typical minimum jerk, 5th order polynomial of duration 700 ms starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
+In this  model, musculoskeletal stiffness was linearly related to torque.
+Like Burdet et al.
+\cite{burdet2006stability}, we approximated muscle torque by looking backward in time $500 \mathrm{\mu s}$.
+Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.
+
 \subsection*{Indices of variance-based sensitivity}
 To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
-Nominal parameters values were mostly taken from Dempster \cite{dempster1955space} and Burdet et al.
-\cite{burdet2006stability} with typical values where subject-specific parameters (height and weight) and direct measurements where necessary (force sensor noise and drift).
+Nominal parameter values were mostly taken from Dempster \cite{dempster1955space} and Burdet et al.
+\cite{burdet2006stability} with typical values for subject-specific parameters (height and weight) and direct measurements for device-specific parameters (force sensor noise and drift).
 Standard deviations reflect expected variation in repeated measurement (ie.
@@
 \subsection*{Statistical Analysis}
-To determine when the intended trajectory's deviation exceeded the deviation that could be explained by undisturbed reaching, we compared the maximum deviation of the undisturbed reaches to the instantaneous deviation of the pulse-disturbed reaches during 100 ms intervals.
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform one-tailed T-tests with an alpha level of 0.05.
-This also provided a confidence interval on the difference.
+Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories.
+To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used.
 
@@
 Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mathrm{mm}^2$.
-These variances describe mean unsigned error, but do not address maximum error.
+These variances describe mean unsigned error, but this uncertainty is not distributed evenly in space or time.
 
@@
 These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances.
-Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance.
-We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance (Table 2).
+Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 135 milliseconds after the onset of disturbance.
+We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance (Figure 4).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3 lower plots).
-Intention showed systematic deflections that counteracted the direction of disturbance about 200 milliseconds after the onset of the force disturbance.
-This counteraction did not depart from the range of undisturbed movements, however (Figure 4, green region).
-The movements completed with the hand returning to undisturbed levels of error within 600 milliseconds, intention differed from actual hand path for as long as one second after the onset of motion.
+When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3, bottom two plots).
+Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 4, orange region).
+Hand movements returned to undisturbed levels of error within 600 milliseconds, but intention did not always return to undisturbed levels within one second.
+
 \section*{Discussion}
@@
 Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly persist with the same intent when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target.
 Time of response may be dependent on the time of the disturbance in the movement.
@@
 However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
-In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to uncover the plan behind their action.
+In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.
 
@@
 Our empirical approach used several methods to mitigate these effects, and hence the errors became small.
-It remains to be seen if there are methods to further reduce the misestimation due to force error or eliminate the need for force estimation altogether.
+It remains to be seen if there are methods to further reduce the mis-estimation due to force error or eliminate the need for force estimation altogether.
 
@@
 Research supported by US NIH NINDS1 R01 NS053606.
-We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
+We thank Felix Huang, Yazan Majeed, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of our preliminary manuscripts.
 We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
 
-\section*{References}
-\bibliography{paper1}
+%\section*{References}
+\bibliography{paper1.bib}
 % Either type in your references using
@@
 \caption{
-{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied.
-This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
+{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied.
+This subject's extracted desired trajectories, in orange, do not significantly differ from baseline during force application, but show some reaction afterwards.
 }
@@
 \caption{
-{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ half of the critical value for simultaneous comparisons at 95\% confidence interval) can no longer be explained by undisturbed hand deviation (green band is mean maximum deviation).
-Movement intention (red dots, band is mean deviation) does not significantly deviate until at least 300 ms after the onset of disturbance.
+{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ 95\% confidence interval on the mean) can no longer be explained by undisturbed hand deviation (blue band).
+The specific time at which this occurs is marked with a green line for each subject.
+Movement intention (orange dots, orange band is mean $\pm$ 95\% confidence interval on the mean) does not significantly deviate until at least 300 ms after the onset of disturbance.
+The specific time at which this occurs is marked with a red line for each subject.
 }
@@
 0.01 \\
-Torque-Invariant Impedance Misestimation Ratio &
+Torque-Invariant Impedance Mis-estimation Ratio &
 1 &
@@
 0.19 \\
-Torque-Varying Impedance Misestimation Ratio &
+Torque-Varying Impedance Mis-estimation Ratio &
 1 &
@@
 
-\begin{table}[!ht]
-\caption{
-\bf{Time Elapsed until Error in Undisturbed and Disturbed Reaching are Significantly Different}}
-\begin{tabular}{|c|c|c|}
-\hline
-Subject Number &
-Hand Deviation (milliseconds) &
-Intent Deviation (milliseconds) \\ \hline
-1 &
-105 &
-305 \\
-2 &
-105 &
-405 \\
-3 &
-135 &
-350 \\
-4 &
-105 &
-300 \\ \hline
-\end{tabular}
-\begin{flushleft} ANOVA with Tukey-Kramer post-\textit{hoc} test ($p=0.05$) was used to simultaneously compare undisturbed hand maximum deviations to force-disturbed hand and intent deviations at 5 millisecond intervals.
-When  a significant difference was first detected, this was noted as the onset of deviation.
-\end{flushleft}
-\label{tab:onsets}
-\end{table}
-
 %\section*{Supporting Information Legends}

### 2015-01-21 RealtimeFeedbackPaper/realtimefeedback.tex
Updated for paper and newest plots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                         % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Reaching is Easier When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence
}

\author{Justin Horowitz$^{1}$, interns, and James Patton$^{1,2*}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
       University of Twente, 7500 AE Enschede, The Netherlands
       {\tt\small albert.author@papercept.net}}%
\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
       Dayton, OH 45435, USA
       {\tt\small b.d.researcher@ieee.org}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Turbulence is an everyday problem that affects piloting and input devices. We derive a method of cancelling its effects that can be performed in realtime. Subjects completed a standard reaching task that required no learning with a turbulent force environment. During the reaching task, subjects were able to view their actual hand positions and their intended hand positions in real time.

To see how this affects performance and effort, we place subjects in a standard reaching task with a turbulent force environment. Replacing feedback of hand position with intended position requires no learning to use, reaching accuracy does not degrade, and arm stiffness drops significantly. This therefore represents a promising new feedback modality.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Humans often interact with uncertain and complicated environments, such as turbulence, crowds and traffic. Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning. High frequency disturbances, such as turbulence, that arise while piloting cars and airplanes are particularly difficult to learn. Worse, delays in our responses and reflexes can give to rise to feedback instabilities that cause crashes and deaths. Modern cars and airplanes incorporate electronic controllers to detect these situations and take control from the operator to prevent harm. Operators sometimes turn off these systems, perhaps due to a cognitive wherein humans feel irrationally safer when they are ``in control." Instead of taking the operator out of the control loop, we propose an algorithm that can determine an operator's intent and deliver it while rejecting any external disturbance and the operators's now-unnecessary response to it.

The word \textit{intent} is ambiguous and must be distinguished from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, and \textit{goal selection decisions}\cite{ziebart2010modeling}. While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of an arm or vehicle) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.

Past attempts to study motor intent have focused on either finding a cost function to generate intended trajectories using \textit{optimal feedback control} or infering the moving \textit{equilibrium point} that results from the spring-like properties of human muscles. Optimal feedback control has been very successful at replicating bulk statistics of outcomes, but it cannot be used to determine the intent underlying any particular action. The $\lambda$ version of the \textit{equilibrium point hypothesis} claimed it could determine the intent underlying any action through a relationship with measured force, impedance, and position of muscles. Gomi and Kawato\cite{gomi1997human} performed these measurements and reported highly complex and often not anatomically realizable trajectories that could not well-represent the intent of a simple reaching movement.

Can't know intent, but at least we know there's a correlation between stiffness and error. We do know stiffness and "instability" and noise.

Even though we lack an accepted means of deducing intent,

If indeed we have deduced the motor intent, we should expect a few properties. First and foremost, no learning should required to use one?s own intent. By definition, there is no error to respond to when you get exactly what you want. Second, this intent should look like an undisturbed movement: a straight-line reach to a target, perhaps changing midreach, perhaps not. Third, the intent should be as accurate as the hand when the human is shown their intent if not moreso. Finally, we know that co-contraction occurs in response to disturbances. If the human is getting what they want, co-contraction is unnecessary and we should see measures of stiffness return to undisturbed levels. If these conditions are met, we can reasonably declare provisional success.

\section{Methods}

\subsection{Intent Extraction}
The well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} relates arm trajectory, $q$, to desired arm trajectory, $q_d$, and any external disturbance, $E$ using physical parameters of the arm. To show how this model can be algebraically inverted to instead describe desired arm trajectory as a function of arm trajectory and external disturbance, we write it as a torque balance:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=
\underbrace{\overbrace{\tau_{ff}}^{\text{Feedforward}}+\overbrace{\tau_{fb}}^{\text{Feedback}}}_\text{Controller}
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, and $G$ contains both Coriolis and centripetal effects. Typical applications solve this torque balance for $\ddot{q}$ and use a numerical differential equation solver to predict arm trajectory in the context of a disturbance of interest, a feedback model, and feedforward torques determined by inverse dynamics. Rather than test hypotheses regarding the learning, production, or composition of this feedforward torque, we instead solve for it:
\begin{equation}
\tau_{ff}=M(q)\ddot{q}+G(q,\dot{q})+E-\tau_{fb}
\end{equation}
Then we note that feedforward torque can have a one-to-one correspondence with desired acceleration, $\ddot{q}_d$:
\begin{equation}
\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)+\hat{E}=\tau_{ff}
\end{equation}
Hats ($\hat{}$) denote the nervous system's best estimate of a physical quantity. This relationship presumes that any torque produced to compensate for expected disturbances,$\hat{E}$ is known. When there is no history of disturbance or that disturbance cannot be predicted or learned (as in the case of turbulence), we can presume that $\hat{E}=0$. Inversion also requires that $\hat{M}(q_d)$ be invertible, but we can ensure this through our choice of workspace. Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent and to vary over the course of a reach. With no prior knowledge of arm impedance for this task-disturbance combination, we use the model of Shadmehr and Mussa-Ivaldi unchanged.

Under these conditions, desired trajectory can be solved numerically and given to a person as feedback because it is a function of its own history, measurable quantities, and feedback torque.

Arm impedance is task-dependent and can be volitionally controlled, so it is unclear what feedback model is appropriate when subjects are shown their estimated intent. In experiment 1, we presume the feedback torque model of Shadmehr and Mussa-Ivaldi in the hope that it is sufficiently accurate or easy to learn. After determining that subjects' arm impedance was much lower than modeled, we lowered our impedance estimate to the smallest value that was numerically stable for experiment 2.

subsection*{Experimental Design}
We chose 15 centimeters long, center-out reaches of the right arm to targets chosen at 60 degree intervals about that central point. Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each 96 reach block. (Protocol figure) In the absence of disturbing forces, the intended trajectory is nearly identical to the hand trajectory, but reflects force sensor error. Blocks are designed to test the consequences of disturbing forces with feedback of hand position versus feedback of determined intended position. Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target.

\subsection*{Dynamic simulation of arm and intended trajectories}
Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. Viscosity parameters were taken from ShadMuss. Stiffness parameters were either taken from ShadMuss or estimated for experiment 2 (mat2(8,2,2,5)). Numerical simulation was performed in real time using the GNU Scientific Library?s odeiv2 driver with Runge-Kutta 4,5 stepping.

\subsection*{Human Subjects}
The human data trajectories analyzed here are drawn from nine subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki. Seven male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback of hand position was performed at 60 Hz using OpenGL. Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame. Because force sensors tend to drift, we performed a linear re-zeroing procedure between each motion to assure unbiased measurements.

\subsection*{Statistical Analysis}
Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories. To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories. The MATLAB statistics toolbox package \cite{MATLAB:2008} was used.

\section{Results}

Need to cover a few figures:
A learning curve (that?s flat)
straightness (which might be an issue)
error tradeoff figure
stiffness figure

\subsection{Headings, etc}

\section{CONCLUSIONS}

\section*{ACKNOWLEDGMENT}

%\begin{thebibliography}{99}

%\end{thebibliography}

\end{document}

### 2015-01-21 internStuff/doRotations.m

@@ -11,20 +11,122 @@ for S=subs
         prefix='output';
     end
     load(['./Data/',prefix,num2str(S),'.mat'])
-    if ~isfield(trials,'xrot')|1
+    if ~isfield(trials,'xrot')||1
+        coeffs=zeros(480,4);
         for T=2:480
-            if (T>(2*96))&&(T<=(96*3))
-                p=trials(T).y;
-            else
-                p=trials(T).x;
-            end
+            %             if (T>(2*96))&&(T<=(96*3))
+            %                 p=trials(T).y;
+            %             else
+            %                 p=trials(T).x;
+            %             end
+            p=trials(T).x;
             x=[p(:,1)-trials(T).orig(1), p(:,2)-trials(T).orig(2)];
-            trials(T).start=max(find(vecmag(x)>.02,1,'first'),1);
+            trials(T).start=find(vecmag(x)>.02,1,'first');
             
             trials(T).xrot=rotateProgressError(trials(T).x,trials(T).orig,trials(T).targ);
             trials(T).frot=rotateProgressError(trials(T).f,trials(T).orig,trials(T).targ);
             trials(T).yrot=rotateProgressError(trials(T).y,trials(T).orig,trials(T).targ);
+            reginds=trials(T).start+(0:39);
+            mgt=mean(gradient(trials(T).t(reginds)));
+            
+            v_error=gradient(trials(T).xrot(reginds,2))/mgt;
+            a_error=gradient(v_error)/mgt;
+            %trials(T).regressX=[trials(T).xrot(reginds,2) v_error a_error]; %#ok<*SAGROW>
+            trials(T).regressX=[trials(T).xrot(reginds,2) v_error a_error ones(size(a_error))]; %#ok<*SAGROW>
+            
+            fshift=-16; %80 ms
+            if reginds(1)+fshift>=1
+                trials(T).regressF=trials(T).frot(reginds+fshift,2);
+            else
+                undershoot=reginds(1)+fshift;
[…]

### 2015-01-21 RealtimeFeedbackPaper/realtimefeedback.tex
Updated text

@@
 During the reaching task, subjects were able to view their actual hand positions and their intended hand positions in real time.
-
-To see how this affects performance and effort, we place subjects in a standard reaching task with a turbulent force environment.
-Replacing feedback of hand position with intended position requires no learning to use, reaching accuracy does not degrade, and arm stiffness drops significantly.
-This therefore represents a promising new feedback modality.
 \end{abstract}
@@
 Optimal feedback control has been very successful at replicating bulk statistics of outcomes, but it cannot be used to determine the intent underlying any particular action.
-The $\lambda$ version of the \textit{equilibrium point hypothesis} claimed it could determine the intent underlying any action through a relationship with measured force, impedance, and position of muscles.
+The $\lambda$ version of the equilibrium point hypothesis claimed it could determine the intent underlying any action through a relationship with measured force, impedance, and position of muscles.
 Gomi and Kawato\cite{gomi1997human} performed these measurements and reported highly complex and often not anatomically realizable trajectories that could not well-represent the intent of a simple reaching movement.
 
-Can't know intent, but at least we know there's a correlation between stiffness and error. We do know stiffness and "instability" and noise.
-
-Even though we lack an accepted means of deducing intent,
-
-If indeed we have deduced the motor intent, we should expect a few properties.
-First and foremost, no learning should required to use one?s own intent.
-By definition, there is no error to respond to when you get exactly what you want.
-Second, this intent should look like an undisturbed movement: a straight-line reach to a target, perhaps changing midreach, perhaps not.
-Third, the intent should be as accurate as the hand when the human is shown their intent if not moreso.
-Finally, we know that co-contraction occurs in response to disturbances.
-If the human is getting what they want, co-contraction is unnecessary and we should see measures of stiffness return to undisturbed levels.
-If these conditions are met, we can reasonably declare provisional success.
+Since lack an accepted model of intent or an accepted means of deducing it, we turn to arm stiffness and its well-studied relationships with instability and uncertainty.
+If subjects reach in an uncertain and unstable force environment, we should expect their arm stiffness to increase.
+If subjects are then asked to reach using their intent, even if our estimate is not perfect, the decrease in uncertainty and instability should lead to a decrease in stiffness.
 

### 2015-01-26 RealtimeFeedbackPaper/realtimefeedback.tex
Updated. Hope it works

@@
 \title{\LARGE \bf
-Reaching is Easier When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence
+Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence
 }
 
-\author{Justin Horowitz$^{1}$, interns, and James Patton$^{1,2*}$% <-this % stops a space
-\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-       University of Twente, 7500 AE Enschede, The Netherlands
-       {\tt\small albert.author@papercept.net}}%
-\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
-       Dayton, OH 45435, USA
-       {\tt\small b.d.researcher@ieee.org}}%
+\author{Justin Horowitz$^{1}$, Tejas Madhavan$^{1}$, Christine Massie$^{1}$, and James Patton$^{1,2*}$% <-this % stops a space
+\thanks{*Ask Jim Who Funded This}% <-this % stops a space
+\thanks{$^{1}$All authors are with the Deparment of Bioengineering Science, University of Illinois at Chicago, 
+		Chicago, IL 60607, USA 
+		}
+\thanks{$^{2}$James Patton is with the Sensory Motor Performance Program, Rehabiliation Institute of Chicago,
+       Chicago, IL 60611, USA
+       {\tt\small PattonJ@uic.edu}}%
 }
@@
 Turbulence is an everyday problem that affects piloting and input devices.
-We derive a method of cancelling its effects that can be performed in realtime.
-Subjects completed a standard reaching task that required no learning with a turbulent force environment.
-During the reaching task, subjects were able to view their actual hand positions and their intended hand positions in real time.
+We derive a method of cancelling its effects to arrive at a human operator's intent that can be performed in realtime.
+We then explore subjects' ability to make targeted reaching motions using this estimate of their intent as a cursor while experiencing turbulent forces.
+All eight subjects were capable of doing so.
+Reaching performance during turbulence was significantly better using the estimated intent than the hand and arm stiffness was significantly lower.
+As arm stiffness is known to correlate with perceived instability and uncertainty, this may indicate a decline in either or both.
+In this first exploration we used a classic arm model, but even better performance might be achieved with more accurate modeling.
 \end{abstract}
@@
 Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning.
-High frequency disturbances, such as turbulence, that arise while piloting cars and airplanes are particularly difficult to learn.
-Worse, delays in our responses and reflexes can give to rise to feedback instabilities that cause crashes and deaths.
-Modern cars and airplanes incorporate electronic controllers to detect these situations and take control from the operator to prevent harm.
-Operators sometimes turn off these systems, perhaps due to a cognitive wherein humans feel irrationally safer when they are ``in control." Instead of taking the operator out of the control loop, we propose an algorithm that can determine an operator's intent and deliver it while rejecting any external disturbance and the operators's now-unnecessary response to it.
+Worse, delays in our responses and reflexes can give to rise to feedback instabilities that cause crashes and fatal mistakes\cite{mcruer1995pilot}.
+Modern cars incorporate electronic controllers to detect instability and assist in regaining control\cite{lie2006effectiveness}, but such a system for airplanes remains elusive\cite{newman2012thirty}.
+Even where technology is available and effective, operators sometimes turn off these systems\cite{itoh2013evaluation}, likely due to a tendency to overestimate their knowledge and abilities as well as the precision of their information\cite{del2012decision}.
+Instead of taking the operator out of the control loop, we propose an algorithm that can determine an operator's intent and deliver it while rejecting any external disturbance and the operators's now-unnecessary response to it.
 
-The word \textit{intent} is ambiguous and must be distinguished from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, and \textit{goal selection decisions}\cite{ziebart2010modeling}.
+The word \textit{intent} is ambiguous and must be distinguished from \textit{motivation}\cite{mcclelland1985motives}\cite{rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal}\cite{flash1985coordination}, and \textit{goal selection decisions}\cite{ziebart2010modeling}.
 While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of an arm or vehicle) taken in service of goals and motives and not the goals nor the motives themselves.
@@
 Past attempts to study motor intent have focused on either finding a cost function to generate intended trajectories using \textit{optimal feedback control} or infering the moving \textit{equilibrium point} that results from the spring-like properties of human muscles.
-Optimal feedback control has been very successful at replicating bulk statistics of outcomes, but it cannot be used to determine the intent underlying any particular action.
-The $\lambda$ version of the equilibrium point hypothesis claimed it could determine the intent underlying any action through a relationship with measured force, impedance, and position of muscles.
+Optimal feedback control has been very successful at replicating bulk statistics of outcomes, but it cannot be used to determine the intent underlying any particular action\cite{todorov2002optimal}.
+The $\lambda$ version of the equilibrium point hypothesis claimed it could determine the intent underlying any action through a relationship with measured force, impedance, and position of muscles\cite{feldman1995origin}.
 Gomi and Kawato\cite{gomi1997human} performed these measurements and reported highly complex and often not anatomically realizable trajectories that could not well-represent the intent of a simple reaching movement.
 
-Since lack an accepted model of intent or an accepted means of deducing it, we turn to arm stiffness and its well-studied relationships with instability and uncertainty.
-If subjects reach in an uncertain and unstable force environment, we should expect their arm stiffness to increase.
-If subjects are then asked to reach using their intent, even if our estimate is not perfect, the decrease in uncertainty and instability should lead to a decrease in stiffness.
+Since we lack an accepted model of intent or an accepted means of deducing it, we turn to reaching accuracy and stiffness to test whether or not our deduced intended trajectory is reasonable.
+First, if this is indeed an estimate of intent, subjects should be able to perform targeted reaching using it even in the presence of a force-based disturbance.
+Second, we should expect that during an unpredictable disturbance, such as white noise, the intended trajectory will deviate less on average than the hand path during the pre-planned, early phase of movement.
+Finally, arm stiffness is known to increase during exposure to instability\cite{franklin2003adaptation} and uncertainty\cite{takahashi2001impedance}.
+If subjects reach in an uncertain and unstable force environment, we should expect their arm stiffness to increase from baseline.
+If subjects are then asked to reach using visual feedback of their intent instead of their hand, even if our estimate is not perfect, the decrease in uncertainty and instability should lead to a decrease in stiffness.
 
@@
 \end{equation}
-Hats ($\hat{}$) denote the nervous system's best estimate of a physical quantity.
-This relationship presumes that any torque produced to compensate for expected disturbances,$\hat{E}$ is known.
-When there is no history of disturbance or that disturbance cannot be predicted or learned (as in the case of turbulence), we can presume that $\hat{E}=0$.
-Inversion also requires that $\hat{M}(q_d)$ be invertible, but we can ensure this through our choice of workspace.
-Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent and to vary over the course of a reach.
-With no prior knowledge of arm impedance for this task-disturbance combination, we use the model of Shadmehr and Mussa-Ivaldi unchanged.
-
-Under these conditions, desired trajectory can be solved numerically and given to a person as feedback because it is a function of its own history, measurable quantities, and feedback torque.
-
-Arm impedance is task-dependent and can be volitionally controlled, so it is unclear what feedback model is appropriate when subjects are shown their estimated intent.
-In experiment 1, we presume the feedback torque model of Shadmehr and Mussa-Ivaldi in the hope that it is sufficiently accurate or easy to learn.
-After determining that subjects' arm impedance was much lower than modeled, we lowered our impedance estimate to the smallest value that was numerically stable for experiment 2.
-
-subsection*{Experimental Design}
-We chose 15 centimeters long, center-out reaches of the right arm to targets chosen at 60 degree intervals about that central point.
-Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each 96 reach block.
-(Protocol figure) In the absence of disturbing forces, the intended trajectory is nearly identical to the hand trajectory, but reflects force sensor error.
-Blocks are designed to test the consequences of disturbing forces with feedback of hand position versus feedback of determined intended position.
-Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target.
-
-\subsection*{Dynamic simulation of arm and intended trajectories}
-Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
-Viscosity parameters were taken from ShadMuss.
-Stiffness parameters were either taken from ShadMuss or estimated for experiment 2 (mat2(8,2,2,5)).
-Numerical simulation was performed in real time using the GNU Scientific Library?s odeiv2 driver with Runge-Kutta 4,5 stepping.
-
-\subsection*{Human Subjects}
-The human data trajectories analyzed here are drawn from nine subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
-Seven male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
-Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
+Hats ( $\hat{}$ ) denote the nervous system's best estimate of a physical quantity. Combining these expressions, suppressing state dependencies, and solving for $\ddot{q}_d$:
+\begin{equation}
+\ddot{q}_d=\hat{M}^{-1} \{M\ddot{q}+G-\hat{G}+E-\hat{E}-\tau_{fb}\}
+\end{equation}
+In this form, a differential equation solver can determine $q_d$ as it evolves in time if a few assumptions are made and conditions are met.
+First, $\hat{E}$ must be modeled or assumed, so we choose $\hat{E}=0$.
+In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures\cite{scheidt2001learning}.
+Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{} and to vary over the course of a reach\cite{shapiro}.
+With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
+The experiment was repeated with a lower stiffness estimate (experiment 2) to explore any dependence on this assumption.
 
@@
 The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
-Visual feedback of hand position was performed at 60 Hz using OpenGL.
+Visual feedback of cursor position was performed at 60 Hz using OpenGL.
 Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
-Because force sensors tend to drift, we performed a linear re-zeroing procedure between each motion to assure unbiased measurements.
 
-\subsection*{Statistical Analysis}
-Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories.
-To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used.
+\subsection*{Human Subjects}
+The human data trajectories analyzed here are drawn from nine subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
+Seven male and one female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
+Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
+
+\subsection*{Experimental Design}
+We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at 60 degree intervals about that central point.
+Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of 5 blocks of 96 reaches each.
+During blocks 2 through 4, subjects experienced filtered white noise forces drawn from a white noise generator at 1000 Hertz with flat power spectral density of 1 Newton, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ radians per second.
+During block 3, cursor position indicated deduced intent.
+During all other blocks, the cursor position indicated hand position (Figure 2, top panel).
+
+\subsection*{Dynamic simulation of arm and intended trajectories}
+Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
+Viscosity parameters were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
+Stiffness parameters were either taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} ($K_{P1}$, experiment 1) or estimated ($K_{P2}$, experiment 2).
+Expressed in Newton-Meters per Radian:
+\begin{equation}
+K_{P1}=
+\begin{bmatrix}
+15 & 6 \\
+6 & 16
+\end{bmatrix}
+\quad
+K_{P2}=
+\begin{bmatrix}
+8 & 2 \\
+2 & 5
+\end{bmatrix}
+\end{equation}
+Numerical simulation was performed in real time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\cite{gough2009gnu}.
+
+\subsection*{Metrics and Statistical Analysis}
+Trajectories and forces were rotated such that movement and force parallel to the line connecting the previous target (the reach origin) and the presented target lie along a \textit{progress} axis while perpendicular movement and force lie along an \textit{error} axis.
+Reach onset was detected as the moment the cursor's distance from the center of the previous target first exceeded 1 centimeter.
+Maximum perpendicular error for a trajectory was the largest error magnitude within 250 milliseconds of reach onset.
+A scalar stiffness, $k$, was calculated for the \textit{error} axis during this same 250 millisecond time span by linear regression:
+\begin{equation}
+F_e=m\ddot{e}+b\dot{e}+ke+F_O
+\end{equation}
+Force ($F_e$) and state ($\ddot{e}$, $\dot{e}$, $e$) are known.
+Mass ($m$), viscosity ($b$), and stiffness offset ($F_0$) terms were calculated, but discarded.
+While stiffness is better described as a matrix, a scalar was calculated to facilitate statistical comparison between treatments and blocks.
+The paired Wilcoxon signed rank test was used to detect differences in maximum perpendicular error and stiffness between blocks and treatments at the 5\% significance level using the MATLAB statistics toolbox package \cite{MATLAB:2014}.
 
 \section{Results}
+As expected, the model was able to deduce an intended trajectory and all subjects were able to utilize this estimate of their intent to perform targeted reaching while experiencing turbulent forces (Figure 1).
+The following sections address the quality of this reaching.
 
-Need to cover a few figures:
-A learning curve (that?s flat)
-straightness (which might be an issue)
-error tradeoff figure
-stiffness figure
+\subsection*{Reaching Accuracy}
+Even if subjects are capable of performing a reaching task using their estimated intent, a poor estimate should lead to poor performance.
+Indeed, the hand significantly outperforms our intent estimate when the force disturbance is not present in block 1 (Figure 2B).
+Model and sensor inaccuracies degrade performance and result in an error of approximately 5 millimeters.
+In the presence of a force disturbance in block 2, the hand loses its advantage as model and sensor inaccuracies are balanced by some degree of cancellation of the disturbance.
+In block 3 where visual feedback of the intent is given instead of hand, the intent significantly outperforms the hand.
+For this intent estimate to be useful, subjects should perform better when using feedback of the intent than when using feedback of the hand.
+Comparing the hand's performance in block 2 to the intent's performance in block 3, subjects perform significantly better using their estimated intent.
+Finally, no significant differences were present across the two experiments.
+Performance appears relatively insensitive to this change in the stiffness model used to deduce the intent.
 
-\subsection{Headings, etc}
+\subsection*{Endpoint Stiffness}
+While accuracy is desirable, it can be attained by cocontracting the muscles to increase arm stiffness at the cost of metabolic effort \cite{gribble2003role}\cite{takahashi2001impedance}.
+Subjects stiffen significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3B).
+Subjects' stiffness significantly decreases when feedback of the deduced intent replaces feedback of the hand as seen by comparing blocks 2 and 3.
+This decrease does not return stiffness to undisturbed levels and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
+Stiffness tends to decrease with duration of exposure to a forced disturbance\cite{o}
 
-\section{CONCLUSIONS}
+ If feedback of intent increases accuracy and decreases stiffness,
+
+\section{Discussion}
+Long story short, white noise still causes people to stiffen up and intent feedback alleviates that.
+
+In theory, our intended trajectories can be used to better fit cost functions for optimal control and this is more a refinement of the equilibrium point hypothesis that sort of gets physics right.
+
+While we could can uniquely determine muscle torque through a torque balance, we assumed a feedback torque model to arrive at feedforward torque.
+While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness and this seems to be the case during the ballistic phase of movement.
+Difficulties acquiring and holding the target support claims that different mechanisms are at work during that stage of movement\cite{shapiro}.
+
+With a feed-forward torque in hand, we made a second assumption: no force learning.
+Without presumed limitations or a model of force learning to constrain our choice of $\hat{E}$, we could choose it to obtain an arbitrary value for $\ddot{q}_d$.
+If we were to presume $\ddot{q}_d$, perhaps most defensibly as a minimum-jerk straight line reach\cite{flash1985coordination}, we could instead determine $\hat{E}$.
+This suggests a third treatment group in future work: cancelling the target-perpendicular component of movement entirely.
+If movement intent is always a straight-line reach to the target, our technique represents unnecessary complication.
+Conversely, the planned reach may not always be a straight line.
+This \textit{straight-line hypothesis} could be readily tested by examining whether or not the determined $\hat{E}$ is symmetrically distributed.
+
+Subjects got more of what they seemingly wanted using our feedback than without it. This represents a promising new treatment, but further study is needed.
 
@@
 %\begin{thebibliography}{99}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,realtimefeedback.bib}
 

### 2015-01-26 internStuff/doRotations.m

@@ -1,90 +1,95 @@
 clc
 clear all
-close all
 
-subs=[2:8 23:29];
 
-for S=subs
-    if S<20
-        prefix='intern';
-    else
-        prefix='output';
-    end
-    load(['./Data/',prefix,num2str(S),'.mat'])
-    if ~isfield(trials,'xrot')||1
-        coeffs=zeros(480,4);
-        for T=2:480
-            %             if (T>(2*96))&&(T<=(96*3))
-            %                 p=trials(T).y;
-            %             else
-            %                 p=trials(T).x;
-            %             end
-            p=trials(T).x;
-            x=[p(:,1)-trials(T).orig(1), p(:,2)-trials(T).orig(2)];
-            trials(T).start=find(vecmag(x)>.02,1,'first');
-            
-            trials(T).xrot=rotateProgressError(trials(T).x,trials(T).orig,trials(T).targ);
-            trials(T).frot=rotateProgressError(trials(T).f,trials(T).orig,trials(T).targ);
-            trials(T).yrot=rotateProgressError(trials(T).y,trials(T).orig,trials(T).targ);
-            reginds=trials(T).start+(0:39);
-            mgt=mean(gradient(trials(T).t(reginds)));
-            
-            v_error=gradient(trials(T).xrot(reginds,2))/mgt;
-            a_error=gradient(v_error)/mgt;
-            %trials(T).regressX=[trials(T).xrot(reginds,2) v_error a_error]; %#ok<*SAGROW>
-            trials(T).regressX=[trials(T).xrot(reginds,2) v_error a_error ones(size(a_error))]; %#ok<*SAGROW>
-            
-            fshift=-16; %80 ms
-            if reginds(1)+fshift>=1
-                trials(T).regressF=trials(T).frot(reginds+fshift,2);
-            else
-                undershoot=reginds(1)+fshift;
-                trials(T).regressF=[-trials(T-1).frot(end+undershoot:end,2);trials(T).frot(1:(reginds(end)+fshift),2)];
-            end
-            coeffs(T,:)=trials(T).regressX\trials(T).regressF;
+subs=[2:9 23:30];
+
+if ~exist('stiffness.mat','file')
+    for S=subs
[…]

### 2015-01-26 internStuff/errorplot.m

@@ -1,0 +1,120 @@
+clc
+clear all
+
+nsubs=8;
+subs=[2:9 23:30];
+tlist=(1:480)';
+b=floor((tlist-1)/96)+1;
+
+if ~exist('jerror.mat','file')
+    for S=subs
+        prefix='output';
+        
+        load(['./Data/',prefix,num2str(S),'.mat'])
+        
+        for T=5:480
+            if trials(T).targcat~=0
+                inds=trials(T).start:min(trials(T).start+49,length(trials(T).t));
+                tr(T).jerrs=max([abs(trials(T).xrot(inds,2)) abs(trials(T).yrot(inds,2))]);
+            end
+        end
+        
+        
+        for blocks=1:5
+            f=find((tlist>5)&mod(tlist,2)&(b==blocks));
+            v=vertcat(tr(f).jerrs);
+            if S==5
+            jerror1(S,blocks)=mean(v(v(:,1)<.05,1));
+            jerror2(S,blocks)=mean(v(v(:,2)<.05,2));    
+            else
+            m=mean(v);
+            jerror1(S,blocks)=m(1);
+            jerror2(S,blocks)=m(2);
+            end
+        end
+    end
+    save('jerror.mat','jerror1','jerror2')
+else
+    load('jerror.mat')
+end
+
+jerror1=jerror1*100;
+jerror2=jerror2*100;
+
+fh=figure(18);
+clf
+
+subplot(3,4,[2:4 6:8])
+hold on
+exp1=2:9;
+exp2=23:30;
+rn=.1*(rand(nsubs,1)-.5);
+
+spacer=5;
+
+for s=1:nsubs
+    plot((1:5)+rn(s),jerror1(exp1(s),:)+spacer,'r-')
+    plot((1:5)+rn(s),jerror1(exp1(s),:)+spacer,'r.')
+    plot((1:5)+rn(s),jerror1(exp2(s),:)+spacer,'b-')
+    plot((1:5)+rn(s),jerror1(exp2(s),:)+spacer,'b.')
+    
+    plot((1:5)+rn(s),jerror2(exp1(s),:),'r-')
+    plot((1:5)+rn(s),jerror2(exp1(s),:),'r.')
+    plot((1:5)+rn(s),jerror2(exp2(s),:),'b-')
[…]

### 2015-01-26 internStuff/finddelay.m

@@ -1,0 +1,57 @@
+clc
+clear all
+
+S=26 %why not
+if S<20
+    prefix='intern';
+else
+    prefix='output';
+end
+
+load(['./Data/',prefix,num2str(S),'.mat'])
+
+a=vertcat(trials(3:479).a);
+f=vertcat(trials(3:479).f);
+
+l=size(a,1);
+
+figure(42)
+clf
+hold on
+
+for blah=3:12
+ind=floor(blah*l/16):floor((blah+1)*l/16);
+pairs=zeros(31,2);
+for k=-30:30
+    mdl=fitlm(a(ind,1),f(ind+k,1));
+    pairs(31+k,:)=[k*.005 mdl.Rsquared.Ordinary];
+end
+[v,i]=max(pairs(:,2));
+plot(1000*pairs(:,1),pairs(:,2))
+plot(1000*pairs(i,1),pairs(i,2),'rx')
+end
+xlabel('Force Sensor Lag, ms')
+ylabel('R^2, regression of acceleration and force')
+
+return 
+data=load(['./Data/',prefix,num2str(S),'.dat']);
+
+x=data(:,3:4);
+v=[gradient(x(:,1)) gradient(x(:,2))]/.001;
+a=[gradient(v(:,1)) gradient(v(:,2))]/.001;
+f=data(:,9:10);
+
+l=size(a,1);
+
+for blah=3:12
+ind=floor(blah*l/16):floor((blah+1)*l/16);
+pairs=zeros(31,2);
+for k=-30:30
+    mdl=fitlm(a(ind,1),f(ind+k,1));
+    pairs(31+k,:)=[k*.005 mdl.Rsquared.Ordinary];
+end
+[v,i]=max(pairs(:,2));
+plot(1000*pairs(:,1),pairs(:,2),'k')
+plot(1000*pairs(i,1),pairs(i,2),'gx')
+end
+

### 2015-01-27 RealtimeFeedbackPaper/realtimefeedback.tex
updated

@@
 \author{Justin Horowitz$^{1}$, Tejas Madhavan$^{1}$, Christine Massie$^{1}$, and James Patton$^{1,2*}$% <-this % stops a space
-\thanks{*Ask Jim Who Funded This}% <-this % stops a space
-\thanks{$^{1}$All authors are with the Deparment of Bioengineering Science, University of Illinois at Chicago, 
+\thanks{*Funded By NIH R01-NS053606 and NIDRR H133E120010}% <-this % stops a space
+\thanks{$^{1}$All authors are with the Deparment of Bioengineering, University of Illinois at Chicago, 
 		Chicago, IL 60607, USA 
@@
 \begin{abstract}
-Turbulence is an everyday problem that affects piloting and input devices.
-We derive a method of cancelling its effects to arrive at a human operator's intent that can be performed in realtime.
-We then explore subjects' ability to make targeted reaching motions using this estimate of their intent as a cursor while experiencing turbulent forces.
-All eight subjects were capable of doing so.
-Reaching performance during turbulence was significantly better using the estimated intent than the hand and arm stiffness was significantly lower.
-As arm stiffness is known to correlate with perceived instability and uncertainty, this may indicate a decline in either or both.
-In this first exploration we used a classic arm model, but even better performance might be achieved with more accurate modeling.
+Turbulence is an everyday problem that affects piloting and input devices in human-machine interaction.
+Here we experimentally evaluate an alternative to turbulence-cancellation, which estimates and displays the human operator's underlying intent in real-time.
+Subjects performed targeted reaching motions while seeing only their intent estimates while they experienced turbulent forces rendered by a robotic handle.
+We found that reaching performance during turbulence was significantly better using the estimated intent.
+Interestingly, when viewing intent, the subjects' estimated arm stiffness became lower, indicating a more relaxed state of operation when viewing their estimated intent.
+In the future, this method could be abstracted and applied to situations where the human arm is disturbed, such as in vehicle turbulence, tremor, and other noisy human-machine interactions.
 \end{abstract}
@@
 Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning.
-Worse, delays in our responses and reflexes can give to rise to feedback instabilities that cause crashes and fatal mistakes\cite{mcruer1995pilot}.
+Worse, delays in our responses and reflexes can give rise to feedback instabilities that cause crashes and fatal mistakes\cite{mcruer1995pilot}.
 Modern cars incorporate electronic controllers to detect instability and assist in regaining control\cite{lie2006effectiveness}, but such a system for airplanes remains elusive\cite{newman2012thirty}.
 Even where technology is available and effective, operators sometimes turn off these systems\cite{itoh2013evaluation}, likely due to a tendency to overestimate their knowledge and abilities as well as the precision of their information\cite{del2012decision}.
-Instead of taking the operator out of the control loop, we propose an algorithm that can determine an operator's intent and deliver it while rejecting any external disturbance and the operators's now-unnecessary response to it.
+Often the operator has new information (such as a change in priorities or plan) that may still require high-level piloting commands.
+Instead of taking the operator out of the control loop, we recently developed an algorithm \cite{horowitz2015determining} that can determine the operator's underlying intent in spite of disturbances.
+One question might be if the operator could see their own instantaneous intent, would they be able to recognize, appreciate, and make use of it?
 
-The word \textit{intent} is ambiguous and must be distinguished from \textit{motivation}\cite{mcclelland1985motives}\cite{rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal}\cite{flash1985coordination}, and \textit{goal selection decisions}\cite{ziebart2010modeling}.
-While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of an arm or vehicle) taken in service of goals and motives and not the goals nor the motives themselves.
-Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.
-
-Past attempts to study motor intent have focused on either finding a cost function to generate intended trajectories using \textit{optimal feedback control} or infering the moving \textit{equilibrium point} that results from the spring-like properties of human muscles.
+Past attempts to study motor intent have focused on either finding a cost function to generate intended trajectories using \textit{optimal feedback control} or inferring the moving \textit{equilibrium point} that results from the spring-like properties of human muscles.
 Optimal feedback control has been very successful at replicating bulk statistics of outcomes, but it cannot be used to determine the intent underlying any particular action\cite{todorov2002optimal}.
@@
 
-Since we lack an accepted model of intent or an accepted means of deducing it, we turn to reaching accuracy and stiffness to test whether or not our deduced intended trajectory is reasonable.
+Since we lack a fully accepted model, we turn to reaching accuracy and stiffness to test whether or not our deduced intended trajectory is reasonable.
 First, if this is indeed an estimate of intent, subjects should be able to perform targeted reaching using it even in the presence of a force-based disturbance.
-Second, we should expect that during an unpredictable disturbance, such as white noise, the intended trajectory will deviate less on average than the hand path during the pre-planned, early phase of movement.
+Second, we should expect that during an unpredictable disturbance, such as turbulent forces, the intended trajectory will deviate less on average than the hand path during the pre-planned, early phase of movement.
 Finally, arm stiffness is known to increase during exposure to instability\cite{franklin2003adaptation} and uncertainty\cite{takahashi2001impedance}.
@@
 \end{equation}
-Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, and $G$ contains both Coriolis and centripetal effects.
+Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocities, $\ddot{q}$ is joint angular accelerations, and $G$ contains both Coriolis and centripetal effects.
 Typical applications solve this torque balance for $\ddot{q}$ and use a numerical differential equation solver to predict arm trajectory in the context of a disturbance of interest, a feedback model, and feedforward torques determined by inverse dynamics.
@@
 Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
-Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{} and to vary over the course of a reach\cite{shapiro}.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{gomi1998task} and may vary over the course of a reach\cite{niu2010temporal}.
 With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
@@
 During block 3, cursor position indicated deduced intent.
-During all other blocks, the cursor position indicated hand position (Figure 2, top panel).
+During all other blocks, the cursor position indicated hand position (Figure 2A).
 
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics{error.eps}      
+	\caption{This is a figure caption.}
+	\label{error}
+\end{figure}
+
 \subsection*{Reaching Accuracy}
@@
 This decrease does not return stiffness to undisturbed levels and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
-Stiffness tends to decrease with duration of exposure to a forced disturbance\cite{o}
+Stiffness may decrease with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
+Subjects do not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
 
- If feedback of intent increases accuracy and decreases stiffness,
+\begin{figure}[h]
+	\centering
+	\includegraphics{stiffness.eps}      
+	\caption{This is a figure caption.}
+	\label{stiffness}
+\end{figure}
 
 \section{Discussion}
-Long story short, white noise still causes people to stiffen up and intent feedback alleviates that.
+Exposure to turbulent forces damaged subjects' reaching accuracy and increased their arm stiffness, but replacing feedback of a subject's hand with feedback of their estimated intent alleviated accuracy and stiffness.
+In the absence of disturbing forces, estimated intent accuracy was lower than that of the hand, which suggests the need for model improvements.
+While we could uniquely determine muscle torque through a torque balance, we assumed a feedback torque model to arrive at feedforward torque.
+While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness, but this does not seem to be the case.
+Moreover, difficulties acquiring and holding at the target during intent feedback support claims that different mechanisms are at work during that stage of movement\cite{niu2010temporal}.
+Improving modeling assumptions to better estimate limb mechanics, force learning, and impedance may improve our estimate in future work.
 
-In theory, our intended trajectories can be used to better fit cost functions for optimal control and this is more a refinement of the equilibrium point hypothesis that sort of gets physics right.
-
-While we could can uniquely determine muscle torque through a torque balance, we assumed a feedback torque model to arrive at feedforward torque.
-While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness and this seems to be the case during the ballistic phase of movement.
-Difficulties acquiring and holding the target support claims that different mechanisms are at work during that stage of movement\cite{shapiro}.
-
-With a feed-forward torque in hand, we made a second assumption: no force learning.
-Without presumed limitations or a model of force learning to constrain our choice of $\hat{E}$, we could choose it to obtain an arbitrary value for $\ddot{q}_d$.
-If we were to presume $\ddot{q}_d$, perhaps most defensibly as a minimum-jerk straight line reach\cite{flash1985coordination}, we could instead determine $\hat{E}$.
-This suggests a third treatment group in future work: cancelling the target-perpendicular component of movement entirely.
-If movement intent is always a straight-line reach to the target, our technique represents unnecessary complication.
-Conversely, the planned reach may not always be a straight line.
-This \textit{straight-line hypothesis} could be readily tested by examining whether or not the determined $\hat{E}$ is symmetrically distributed.
-
-Subjects got more of what they seemingly wanted using our feedback than without it. This represents a promising new treatment, but further study is needed.
+Future directions include extending this modeling approach to more complex plants such as motor vehicles, touch screen devices, and reaching movements in three dimensions.
+This may potentially improve human-robot interaction by giving roboticists a real-time estimate of human movement intent despite disturbances and incongruous interaction on the part of the robot.
+In order to extend this work to situations in which learning occurs, a learning model will be necessary or additional assumptions regarding the form and production of intent will be required.
+Intent estimation may be a useful tool for revealing previously unknown features of planning and adaptation.
 
 \section*{ACKNOWLEDGMENT}
+The authors wish to thank Yazan Abdel Majeed for his review of the preliminary manuscript.
+We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
 
-%\begin{thebibliography}{99}
 \bibliographystyle{IEEEtran}
@@
 
-%\end{thebibliography}
-
 \end{document}

### 2015-02-16 Paper 3/2Dfreeexploration/summarynew.m
Work machine updates. Better .gitignore

@@ -1,0 +1,45 @@
+clc
+clear all
+
+load stroke.mat
+stroke=outs;
+
+load healthy.mat
+healthy=outs;
+
+figure(1)
+clf
+
+fields={'W','n','T'};
+
+nsp=length(fields);
+
+for k=1:length(fields)
+    subplot(1,nsp,k)
+    nh=[healthy.(fields{k})];
+    ns=[stroke.(fields{k})];
+    plot([ones(length(nh),1); 2*ones(length(ns),1)],[nh ns]','.')
+    xlim([.5 2.5])
+    set(gca,'xtick',[1 2])
+    set(gca,'xticklabels',{'Healthy','Stroke'})
+    title(regexprep(fields(k),'_',' '))
+end
+
+figure(2)
+clf
+
+fields={'power_symmetry','power_rmse','kinetic_rmse'};
+
+nsp=length(fields);
+
+for k=1:length(fields)
+    subplot(1,nsp,k)
+    nh=[healthy.(fields{k})];
+    ns=[stroke.(fields{k})];
+    plot([ones(length(nh),1); 2*ones(length(ns),1)],[nh ns]','.')
+    xlim([.5 2.5])
+    set(gca,'xtick',[1 2])
+    set(gca,'xticklabels',{'Healthy','Stroke'})
+    title(regexprep(fields(k),'_',' '))
+end
+

### 2015-02-16 Paper 3/main.m

@@ -1,0 +1,70 @@
+clc
+clear all
+
+c=dir('./org_data');
+
+%Specify subject numbers
+
+%Each subject comes "with" a library of movements
+%--Yazan owes Justin a means of getting this part right.
+%--Reaches, to the center, with explicit encouragement to be bimanual with
+%  NO external force disturbances
+
+%For each movement, load its shit into a structure (and compute A), while doing that get Xrange
+
+%Compute Xgrid
+
+%For each movement, compute Ux, Ut
+
+%From all movements, compile Ux as a statistical object
+
+%Test P(x|movement)=P(x|U,T) for some T
+
+%Use the confidence limit from the above in order to test Uhealthy=Uimpaired
+
+%Plot voxels that fail the test as translucent cubes with spaghetti drawn on top
+%--Maybe plot degree of difference as transparent-to-opaque
+
+
+cMin=[0 1 0];
+cMax=[1 0 1];
+
+STEP=5;
+
+maxUts(2)=0.1970;
+
+for S=2 %[2 7 8];
+    figure(S)
+    clf
+    hold on
+    for k=1:length(c)
+        r=regexp(c(k).name,['outputsS',num2str(S,'%1.2i')]);
+        if ~isempty(r)
+            st=open(['./org_data/',c(k).name]);
+
+            X=st.Robot_arm.smooth_pos;
+            V=st.Robot_arm.vel;
+            t=st.Robot_arm.time;
+
+            A=[gradient(V(:,1),t) gradient(V(:,2),t) gradient(V(:,3),t)];
+
+            [Ut,Ux,Xgrid]=makeVPF(t,X,V,A);
+
+            Ut=Ut-min(Ut); %Just a reference anyways
+            mUt=max(Ut);
+            nUt=Ut/maxUts(S);
+
+            lnUt=log(nUt+.01);
+            
+            if mUt>maxUts(S)
+                maxUts(S)=mUt;
+            end
+            
+            for kk=STEP+1:length(t)
+                plot3(X(kk-STEP:kk,1),X(kk-STEP:kk,2),X(kk-STEP:kk,3),'color',cMin*(1-lnUt(kk))+cMax*lnUt(kk))
+            end
+        end
+    end
+    
+    axis equal
+end

### 2015-02-16 Paper 3/model.m

@@ -1,0 +1,6 @@
+load('forYaz.mat','W','R2','bb')
+
+y = bb;
+x = [W R2];
+
+model =

### 2015-02-17 Paper1 (copy)/paper1B.tex
Work machine updates. Better .gitignore
huh

--- 2015-01-07 Paper1/paper1B.tex
@@
 The largest sensitivity arose mainly from uncertainty in joint stiffnesses.
-Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 ms following the onset of a disturbance.
-As expected, the intended trajectory showed no change from undisturbed reaching for more than 300 ms after the disturbance onset.
+Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 milliseconds following the onset of a disturbance.
+As expected, the intended trajectory showed no change from undisturbed reaching for more than 300 milliseconds after the disturbance onset.
 Beyond this point, however, we detected a change in intent, surprisingly even when the hand is already near the target.
@@
 \subsection*{Human Subjects}
-The human data trajectories analyzed here are drawn from four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
+The human data trajectories analyzed here are drawn from four subjects who gave written informed consent after reading a consent form approved by Northwestern University Institutional Review Board, which specifically approved this study following the principles expressed in the Declaration of Helsinki.
 Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated.
@@
 Research supported by US NIH NINDS1 R01 NS053606.
-We thank Felix Huang, Yazan Majeed, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of our preliminary manuscripts.
+We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
 We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.

### 2015-03-04 Paper1Editted/paper1Beditted.tex
Updated for edits to paper

--- 2015-01-07 Paper1/paper1B.tex
@@
 Therefore, it could not well-represent the intent of a simple reaching movement.
+Unlike the equilibrium point of muscle, the equilibrium point of the entire arm matches undisturbed movement.
+In the presence of disturbance, the arm might be deflected from its equilibrium.
+It is presently unknown if or how the arm's equilibrium point might change during external disturbances.
 
@@
 This model was chosen to help illuminate the approach, but later we present how to generalize this approach to more complex models without loss of generality.
-The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
+The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form,
 \begin{equation}
@@
 \end{equation}
-Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque.
-The motion behavior changes with the addition of feedforward and/or feedback controllers:
+where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque.
+The motion behavior changes with the addition of feedforward and/or feedback controllers,
 \begin{equation}
@@
 \end{equation}
-Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}.
+where terms with hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}.
 This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance.

### 2015-03-13 RealtimeFeedbackPaper/realtimefeedback.tex
Updated on the mac for plots

--- 2015-01-26 RealtimeFeedbackPaper/realtimefeedback.tex
@@
 \thanks{*Ask Jim Who Funded This}% <-this % stops a space
-\thanks{$^{1}$All authors are with the Deparment of Bioengineering Science, University of Illinois at Chicago, 
+\thanks{$^{1}$All authors are with the Deparment of Bioengineering, University of Illinois at Chicago, 
 		Chicago, IL 60607, USA 
@@
 Reaching performance during turbulence was significantly better using the estimated intent than the hand and arm stiffness was significantly lower.
-As arm stiffness is known to correlate with perceived instability and uncertainty, this may indicate a decline in either or both.
+As arm stiffness is known to correlate with instability and uncertainty, this may indicate a decline in either or both.
 In this first exploration we used a classic arm model, but even better performance might be achieved with more accurate modeling.
+In the future, this method could be abstracted and applied to plants other than the human arm such as automobiles, boats, airplanes, touch screen devices.
 \end{abstract}
@@
 This decrease does not return stiffness to undisturbed levels and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
-Stiffness tends to decrease with duration of exposure to a forced disturbance\cite{o}
-
- If feedback of intent increases accuracy and decreases stiffness,
+Stiffness may decrease with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
+Subjects do not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
 
 \section{Discussion}
-Long story short, white noise still causes people to stiffen up and intent feedback alleviates that.
+Exposure to turbulent forces damaged subjects' reaching accuracy and increasing their arm stiffness, but replacing feedback of a subject's hand with feedback of their estimated intent alleviated accuracy and stiffness.
+In the absence of disturbing forces estimated intent accuracy was lower than that of the hand, which suggests the need for model improvements.
+While we could can uniquely determine muscle torque through a torque balance, we assumed a feedback torque model to arrive at feedforward torque.
+While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness, but this does not seems to be the case.
+Moreover, difficulties acquiring and holding the target during intent feedback support claims that different mechanisms are at work during that stage of movement\cite{niu2010temporal}.
+Improving modeling assumptions address limb mechanics, force learning, and impedance can be improved in future work.
 
-In theory, our intended trajectories can be used to better fit cost functions for optimal control and this is more a refinement of the equilibrium point hypothesis that sort of gets physics right.
+Future directions include extending this modeling approach to more complex plants such as motor vehicles, touch screen devices, and reaching movements in three dimensions.
+This may potentially improve human-robot interaction by giving roboticists a real-time estimate of human movement intent despite disturbances and incongruous interaction on the part of the robot.
+In order to extend this work to situations in which learning occurs, a learning model will be necessary or additional assumptions regarding the form and production of intent will be required.
+Intent estimation may be useful tool for revealing previously unknown features of planning and adaptation.
 
-While we could can uniquely determine muscle torque through a torque balance, we assumed a feedback torque model to arrive at feedforward torque.
-While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness and this seems to be the case during the ballistic phase of movement.
-Difficulties acquiring and holding the target support claims that different mechanisms are at work during that stage of movement\cite{shapiro}.
+\begin{figure}[thpb]
+      \centering
+      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
+}}
+      %\includegraphics[scale=1.0]{figurefile}
+      \caption{Typical subjects' center-out targeted reaching motions under experimentally varied force and feedback conditions.
+Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
+While both were recorded under all conditions, only one was shown on screen.
+Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
+The force disturbance was designed to be turbulent and unpredictable to minimize any effect of learning.}
+      \label{typicalsubject}
+   \end{figure}
 
-With a feed-forward torque in hand, we made a second assumption: no force learning.
-Without presumed limitations or a model of force learning to constrain our choice of $\hat{E}$, we could choose it to obtain an arbitrary value for $\ddot{q}_d$.
-If we were to presume $\ddot{q}_d$, perhaps most defensibly as a minimum-jerk straight line reach\cite{flash1985coordination}, we could instead determine $\hat{E}$.
-This suggests a third treatment group in future work: cancelling the target-perpendicular component of movement entirely.
-If movement intent is always a straight-line reach to the target, our technique represents unnecessary complication.
-Conversely, the planned reach may not always be a straight line.
-This \textit{straight-line hypothesis} could be readily tested by examining whether or not the determined $\hat{E}$ is symmetrically distributed.
+   \begin{figure}[thpb]
+      \centering
+      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
+}}
+      %\includegraphics[scale=1.0]{figurefile}
+      \caption{Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback.
+Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement reveals that turbulent force disturbance degrade reaching performance.
+}
+      \label{figurelabel}
+   \end{figure}
 
-Subjects got more of what they seemingly wanted using our feedback than without it. This represents a promising new treatment, but further study is needed.
+   \begin{figure}[thpb]
+      \centering
+      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
+}}
+      %\includegraphics[scale=1.0]{figurefile}
+      \caption{Inductance of oscillation winding on amorphous
+       magnetic core versus DC bias magnetic field}
+      \label{figurelabel}
+   \end{figure}
 
 \section*{ACKNOWLEDGMENT}
-
-%\begin{thebibliography}{99}
 \bibliographystyle{IEEEtran}

### 2015-03-13 Paper 3/2Dfreeexploration/smokinggun.m

@@ -7,4 +7,6 @@ en=dot(v',v');
 
 figure(75)
 clf
-plot3(x(:,1),x(:,2),en,'.')
+sx1=size(x,1);
+ds=1:150:sx1;
+plot3(x(ds,1),x(ds,2),en(ds),'.')

### 2015-03-13 internStuff/errorplot.m

@@ -41,80 +41,165 @@ end
 jerror1=jerror1*100;
 jerror2=jerror2*100;
 
-fh=figure(18);
+%% New stuff
+msize=8;
+
+f=figure(3);
+set(f,'units','in');
+width=6;
+lmargin=1;
+rmargin=.05;
+
+set(f,'Position',[5,5,3,5.5])
 clf
 
-subplot(3,4,[2:4 6:8])
+margin=.1;
+ploth=1;
+
+sp=subplot('Position',[0 .9 1 .05]);
+set(sp,'Units','in','Position',[lmargin 4.75 3-rmargin-lmargin .5])
+hold on
+plot([1.5 4.5],3*[1 1],'k')
+plot([2.5 3.5],1*[1 1],'k')
+plot([.5 2.5],2*[1 1],'k',[3.5 5.5],2*[1 1],'k')
+
+set(gca,'xtick',[])
+set(gca,'ytick',1:3)
+ylabs={'Cursor=Intent','Cursor=Hand','Forces On'};
+
+set(gca,'yticklabel',ylabs)
+set(gca,'xcolor','w')
+
+ylim([.5 3.5])
+
+yl=ylim;
+xl=xlim;
+xrat=-.9/1.95;
+yrat=1+.1/.5;
+text(xrat*xl(2)+(1-xrat)*xl(1),yrat*yl(2)+(1-yrat)*yl(1),'A','FontWeight','Bold')
+
+%% Hand
+
+sp1=subplot('Position',[0 .4 1 .2])
+set(sp1,'Units','in','Position',[lmargin 3.5 3-rmargin-lmargin 1])
 hold on
 exp1=2:9;
 exp2=23:30;
-rn=.1*(rand(nsubs,1)-.5);
+rn=.1*linspace(-1,1,8);
 
-spacer=5;
+for s=1:8
+    plot((1:5)-rn(s),jerror1(exp1(s),:),'r-')
+    plot((1:5)-rn(s),jerror1(exp1(s),:),'r.')
+    plot((1:5)-rn(s),jerror1(exp2(s),:),'b-')
+    plot((1:5)-rn(s),jerror1(exp2(s),:),'b.')
+end
+xlim([.5 5.5])
+set(gca,'xtick',[])
+%set(gca,'xcolor','w')
+ylabel('Hand Error (cm)')
+yl=ylim;
+xl=xlim;
+xrat=-.9/1.95;
+yrat=1+.1/1;
[…]

### 2015-03-13 Paper1Editted/paper1Beditted.tex
almost back to working.

@@
 This entails algebraic manipulation to solve for $\ddot{q}$, followed by integration to determine the state trajectory.
-Intent determination takes the novel approach of instead solving for $\ddot{q}_d$:
+Intent determination takes the novel approach of instead solving for $\ddot{q}_d$,
 \begin{equation}
@@
 \end{equation}  
-$\ddot{q}_d$ can then be integrated numerically using a differential equation solver to determine the intended state trajectory $q_d(t)$.
+such that $\ddot{q}_d$ can then be integrated numerically using a differential equation solver to determine the intended state trajectory $q_d(t)$.
 This relies on many assumptions: The model of plant and controller must be accurate and precise.

### 2015-03-13 Paper 3/2Dfreeexploration/batch_boltzmann.m

@@ -4,7 +4,14 @@ close all
 
 sublets={'JL','JT','LP','MF','YM'};
 
-for k=1 %:length(sublets)
+figure(4000)
+clf
+subplot(2,1,1)
+hold on
+subplot(2,1,2)
+hold on
+
+for k=1:length(sublets)
     fname=['free_exp_',sublets{k},'.mat'];
     
     if ~exist(fname,'file')
@@ -44,13 +51,19 @@ for k=1 %:length(sublets)
     
     figure(k)
     clf
-    outs(k)=boltzmannsubjectplots(t,x,v,a)
+    [outs(k),P,E]=boltzmannsubjectplots(t,x,v,a)
 
     suplabel(['Healthy Subject ',sublets{k}],'t')
-    set(findall(gcf,'type','text'),'fontname','Times','fontSize',48)
+    set(findall(gcf,'type','text'),'fontname','Times','fontSize',12)
     set(gcf,'position',[76 11 1195 925])
     print('-dtiff','-r300',['summary',sublets{k},'.tiff'])
     print('-dpng','-r300',['summary',sublets{k},'.png'])
+    
+    figure(4000)
+    subplot(2,1,1)
+    plot(P(:,2),P(:,1),'b')
+    subplot(2,1,2)
+    plot(E(:,2),E(:,1),'b')
 end
 
 save('healthy.mat','outs')

### 2015-03-13 Paper 3/2Dfreeexploration/batch_boltzmann_stroke.m

@@ -2,7 +2,7 @@ clc
 clear all
 close all
 
-for k=1:10
+for k=5 %1:10
     subname=num2str(k,'%2.2d');
     fname=['free_exp_',subname,'stroke.mat'];
     

### 2015-03-13 Paper 3/2Dfreeexploration/boltzmannsubjectplots.m

@@ -1,4 +1,4 @@
-function ofinterest=boltzmannsubjectplots(t,x,v,a)
+function [ofinterest,Pnormecdf,Enormecdf]=boltzmannsubjectplots(t,x,v,a)
 
 global fitmecounts fitmebins
 
@@ -36,18 +36,14 @@ bar(centers,counts)
 fitmecounts=counts;
 fitmebins=edges;
 
-W=expfit(abs(Pt));
+W=fminbnd(@fitexpcdf,0,10);
 
 ofinterest.W=W;
 
+Pnormecdf=[counts centers/W];
+
 cdf=expcdf(fitmebins,W);
 fit=cdf(2:end)-cdf(1:end-1);
-error
-[trash,ofinterest.power_ksP]=kstest(aPt,'CDF',[aPt,expcdf(aPt,W)]);
-figure(20)
-clf
-[ex,ec,lo,up]=ecdf(aPt);
-plot(aPt,expcdf(aPt,W),'b',ec,ex,'r',ec,lo,'r--',ec,up,'r--')
 ofinterest.power_rmse=sqrt(mean((fit-counts).^2));
 plot(centers,fit,'r','linewidth',3)
 
@@ -98,9 +94,7 @@ bar(centers,counts)
 
 fitmecounts=counts;
 fitmebins=edges;
-nT=gamfit(Tt);
-ofinterest.n=nT(1);
-ofinterest.T=nT(2);
+nT=fmincon(@fitgamcdf,[1.5; 1],-eye(2),-[1; 0.01]);
 
 cdf=gamcdf(fitmebins,nT(1),nT(2));
 fit=cdf(2:end)-cdf(1:end-1);
@@ -136,10 +130,18 @@ bar(centers,counts)
 fitmecounts=counts;
 fitmebins=.5*m*edges.^2;
 
+nT=fmincon(@fitgamcdf,[1.5; 1],-eye(2),-[1; 0.01]);
 cdf=gamcdf(fitmebins,nT(1),nT(2));
 fit=cdf(2:end)-cdf(1:end-1);
-error
-[trash,ofinterest.speed_ksP]=kstest(speed,[.5*m*speed.^2,gamcdf(speed,nT(1),nT(2))]);
+ofinterest.n=nT(1);
+ofinterest.T=nT(2);
+ofinterest.speed_rmse=sqrt(mean((fit-counts).^2));
+
+Enormecdf=[counts centers/nT(2)];
+
[…]

### 2015-03-13 Paper 3/2Dfreeexploration/fitexpcdf.m

@@ -1,7 +1,7 @@
 function [cost,fit]=fitexpcdf(lambda)
 
-global binedges bincounts
+global fitmecounts fitmebins
 
-ecdf=expcdf(binedges,lambda);
+ecdf=expcdf(fitmebins,lambda);
 fit=ecdf(2:end)-ecdf(1:end-1);
-cost=sum((bincounts-fit).^2);
+cost=sum((fitmecounts-fit).^2);

### 2015-03-13 Paper 3/2Dfreeexploration/fitgamcdf.m

@@ -1,12 +1,12 @@
-function [cost,fit]=fitgamcdf(params)
+function cost=fitgamcdf(params)
 
-global binedges bincounts
+global fitmecounts fitmebins
 
 n=params(1);
 T=params(2);
 
-gcdf=gamcdf(binedges,n,T);
+gcdf=gamcdf(fitmebins,n,T);
 fit=gcdf(2:end)-gcdf(1:end-1);
 
-cost=sum((bincounts-fit).^2);
+cost=sum((fitmecounts-fit).^2);
 

### 2015-03-13 Paper 3/2Dfreeexploration/summarynew.m

@@ -28,7 +28,7 @@ end
 figure(2)
 clf
 
-fields={'power_symmetry','power_rmse','kinetic_rmse'};
+fields={'power_symmetry','power_rmse','speed_rmse'};
 
 nsp=length(fields);
 
@@ -36,10 +36,13 @@ for k=1:length(fields)
     subplot(1,nsp,k)
     nh=[healthy.(fields{k})];
     ns=[stroke.(fields{k})];
-    plot([ones(length(nh),1); 2*ones(length(ns),1)],[nh ns]','.')
+    plot([ones(length(nh),1); 2*ones(length(ns),1)],100*[nh ns]','.')
     xlim([.5 2.5])
     set(gca,'xtick',[1 2])
     set(gca,'xticklabels',{'Healthy','Stroke'})
     title(regexprep(fields(k),'_',' '))
+    if k==1
+        ylabel('RMS Percent Error')
+    end
 end
 

### 2015-03-13 matlab/final_fig45_combinedT.m

@@ -2,7 +2,7 @@ clc
 clear all
 close all
 
-SUBS=1:4;
+SUBS=1:8;
 distCats=[3 4];
 
 black=[0 0 0];
@@ -15,7 +15,7 @@ bmargin=0;
 figmargin=.4;
 
 yoffSpan=8;
-plotOrder=[1 4 2 3];
+plotOrder=1:8 %[1 4 2 3 5 6 7 8];
 
 figure(1)
 clf
@@ -26,7 +26,7 @@ set(gcf,'position',[4,8,figmargin+lmargin+width,figmargin+bmargin+height])
 
 %% Compute everything we'll need later
 
-if ~exist('fig4dotsNmeans.mat','file')
+if ~exist('fig4dotsNmeans.mat','file')|1
     for k=SUBS
         load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
         load(['../Data/Data_pulse/pulse',num2str(k),'Y.mat'])
@@ -57,7 +57,7 @@ if ~exist('fig4dotsNmeans.mat','file')
                 catch
                     st1=size(trials(t).x,1);
                     triStruct(t).dX=sPeak*getR2(trials(t).x(inds(1):st1,:));
-                    st1=size(trials(t).x,1);
+                    st1=size(trials(t).y,1);
                     triStruct(t).dY=sPeak*getR2(trials(t).y(indsy(1):st1,:));
                 end
                 
@@ -102,7 +102,7 @@ if ~exist('fig4dotsNmeans.mat','file')
                     thishappened=1
                     st1=size(trials(t).x,1);
                     triStruct(t).dX=sPeak*getR2(trials(t).x(inds(1):st1,:));
-                    st1=size(trials(t).x,1);
+                    st1=size(trials(t).y,1);
                     triStruct(t).dY=sPeak*getR2(trials(t).y(indsy(1):st1,:));
                 end
             end
@@ -222,7 +222,7 @@ for S=SUBS
             end
         end
     end
-    
+    differences(S)=onset-handonset
 
     plot([0,1000],yoff+[0 0],'-','linewidth',1,'color',color)
     
@@ -237,6 +237,8 @@ for S=SUBS
         set(h(k),'EdgeAlpha',ALPHA,'FaceAlpha',ALPHA);
     end
 end
+
+return
 
 plot(-30+[0,0],[0 5],'-','linewidth',1,'color',color)
[…]

### 2015-03-13 matlab/findWhiteKpGain.m

@@ -7,6 +7,9 @@ global kpgain massgain
 kpgain=1;
 massgain=1;
 
+params
+params.mass=0.453592*190 %*params.mass
+
 GWR=2000;
 
 %% Categorize by start/end pair
@@ -65,7 +68,7 @@ for U=1:length(urc)
     rp=randperm(length(f));
     for krp=1:length(rp)
         fk=rp(krp);
-        if krp<=20
+        if 1 %krp<=20
             start=onsetDetector(trials(f(fk)));
             time=trials(f(fk)).t-trials(f(fk)).t(start);
             x=trials(f(fk)).x;

### 2015-03-13 matlab/fullWorkFlow.m

@@ -14,11 +14,14 @@ addpath ../../DeOpt/
 %     'NEVERMATCH','finalfig4'
 %     };
 
-workflow={'','addSubjectPulse';
-    'W','findWhiteKpGain';
+% workflow={'','addSubjectPulse';
+%     'W','findWhiteKpGain';
+%     'Y','extractPulses'
+%     };
+
+workflow={'W','findWhiteKpGain';
     'Y','extractPulses'
     };
-
 
 mfiles=dir('.');
 datafiles=dir('../Data/Data_pulse/');

### 2015-03-13 matlab/onsetDetector.m

@@ -1,4 +1,4 @@
 function start=onsetDetector(trial)
 
-onset=find(vecmag(trial.v)>.05,1,'first');
+onset=find(vecmag(trial.v)>.1,1,'first');
 start=max(onset-35,1);

### 2015-03-20 Paper1Editted/paper1Beditted.tex
edits

@@
 Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 ms following the onset of a disturbance.
-As expected, the intended trajectory showed no change from undisturbed reaching for more than 300 ms after the disturbance onset.
-Beyond this point, however, we detected a change in intent, surprisingly even when the hand is already near the target.
+As expected, the intended trajectory showed no change from undisturbed reaching for more than 150 ms after the disturbance onset.
+Beyond this point, however, we sometimes detected a change in intent, surprisingly even when the hand is already near the target.
 Knowing such an intent signal is broadly applicable: enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and complex systems that embody planning such as brain machine interfaces, team sports, crowds, or swarms.
@@
 In the presence of disturbance, the arm might be deflected from its equilibrium.
-It is presently unknown if or how the arm's equilibrium point might change during external disturbances.
+It is presently unknown if or how the arm's equilibrium point might change due to disturbances.
 
@@
 The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
-Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some action space, $x$:
+Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some action space, $x$,
 \begin{equation} \label{eq:generalPlant}
@@
 \end{equation}
+where $x_e$ is the equilibrium trajectory of the actuator.
 The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ such that $x$ will track a desired path $x_d$.
@@
 \end{equation}
-We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives:
+We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives
 \begin{equation}
@@
 \end{equation}
-
-Note $x_e$ vanishes, recovering our familiar model.
+wherein $x_e$ vanishes, recovering our familiar model.
 While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory.
@@
 While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
-Instead of a dynamic model, it could be a lookup table.
+Even if the process model was a reverberating loop \cite{fortney2012computational} driven by a reference intent, we might still recover that reference intent by inverting the loop.
+We are not limited to dynamic models as this might be extended to a lookup table.
 For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
@@
 \cite{burdet2006stability}.
-We fit subject-specific constant scale factors for body mass, $m$, and impedance, $c$, relative to this model to account for atypical mass distribution and any task-dependence of the subjects' impedance\cite{gomi1998task}.
+We fit a subject-specific constant scale factor for impedance, $c$, relative to this model to account for any task-dependence of the subjects' impedance\cite{gomi1998task}.
 \begin{equation}
-E=c(K_p\Delta q+K_d \Delta \dot{q})+m(I(q)\Delta\ddot{q})
+E=c(K_p\Delta q+K_d \Delta \dot{q})+I(q)\Delta\ddot{q}
 \end{equation}
@@
 To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
-The MATLAB statistics toolbox package \cite{MATLAB:2008} was used.
+The spans between departure of the hand and estimated intent were compared against 120 milliseconds using the Wilcoxon signed rank test.
+The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.
 
@@
 \subsection*{Human Intent Trajectories}
-Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3).
+Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3A).
 If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands.
@@
 These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances.
-Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 135 milliseconds after the onset of disturbance.
-We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance (Figure 4).
+Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance.
+While some subjects' intents showed no significant from baseline despite disturbance, when intent deviated it did so 150-255 milliseconds after the hand deviated (Figure 3B).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3, bottom two plots).
+When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3A).
 Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 4, orange region).
@@
 Mis-estimation was largest if the stiffness was inaccurately modeled.
-We tested this on human reaching and found that in spite of the hand being disturbed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds.
+We tested this on human reaching and found that in spite of the hand being disturbed by forces, the intended movement remained on-course to the target, possibly changing after 150 milliseconds.
 This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.
@@
 In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}.
-Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}.
-Time of response may be dependent on the time of the disturbance in the movement.
-Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent.
+While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three subjects showed no detectable response.
+The absence of a response to a large force disturbance within an optimal feedback control framework and is instead consistent with adherence to a trajectory planned before any disturbance.
+Alternatively,  control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof.
 It remains to be seen if and how change in intent might be triggered.
@@
 The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
-However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
+However, by substituting an approximation, reverberating loop, or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
 In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.
@@
 Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
+
+While we use the terms \textit{intent}, $x_d$, and \textit{equilibrium trajectory of the arm} interchangeably, they are not always identical.
+We assume that action does not deviate from intent in the absence of external disturbance.
+For this to be true, $x_d$ must be the intent.
+Not all action need arise from a continuous intent trajectory, however.
+One might volitionally twitch a muscle without planning its effect on the limb.
+In the context of reaching, our results and others' \cite{flash1991arm, bizzi1984posture, shadmehr1994adaptive, rohrer2004submovements, fortney2012computational, mistry2013optimal} suggest a continuous pre-planned trajectory with kinematic constraints, perhaps composed in an intermittent fashion from some motor primitive.
+Outside of this context, the word \textit{intent} may often be inappropriate.
 

### 2015-03-23 Paper1Editted/paper1Beditted.tex
New figure + new manuscript

@@
 While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
-Even if the process model was a reverberating loop \cite{fortney2012computational} driven by a reference intent, we might still recover that reference intent by inverting the loop.
+Even if the process model was a reverberating loop \cite{fortney2012computational} driven by a reference intent, we might still recover that reference intent by simulating the loop.
 We are not limited to dynamic models as this might be extended to a lookup table.
@@
 To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
-The spans between departure of the hand and estimated intent were compared against 120 milliseconds using the Wilcoxon signed rank test.
+The span between departure of the hand and estimated intent was compared against 120 milliseconds using the sign test.
 The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.
@@
 Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance.
-While some subjects' intents showed no significant from baseline despite disturbance, when intent deviated it did so 150-255 milliseconds after the hand deviated (Figure 3B).
+Despite the disturbance, three subjects' intents did not depart significantly from baseline.
+The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3B).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
 When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3A).
-Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 4, orange region).
-Hand movements returned to undisturbed levels of error within 600 milliseconds, but intention did not always return to undisturbed levels within one second.
+Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, orange region).
+This counteraction is consistent with both change of intent and stiffening in response to the disturbance.
 
@@
 The proposed approach could explain the discovery by Bizzi et al.
-\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging.
+\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement.
 By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.
@@
 While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three subjects showed no detectable response.
-The absence of a response to a large force disturbance within an optimal feedback control framework and is instead consistent with adherence to a trajectory planned before any disturbance.
-Alternatively,  control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof.
+The absence of response to a large force disturbance seems inconsistent with optimal feedback control framework and is instead consistent with adherence to a trajectory planned before the disturbance.
+Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof.
 It remains to be seen if and how change in intent might be triggered.
@@
 We assume that action does not deviate from intent in the absence of external disturbance.
-For this to be true, $x_d$ must be the intent.
+For this to be true, the equilibrium trajectory of the arm must be the intent.
 Not all action need arise from a continuous intent trajectory, however.
@@
 In the context of reaching, our results and others' \cite{flash1991arm, bizzi1984posture, shadmehr1994adaptive, rohrer2004submovements, fortney2012computational, mistry2013optimal} suggest a continuous pre-planned trajectory with kinematic constraints, perhaps composed in an intermittent fashion from some motor primitive.
-Outside of this context, the word \textit{intent} may often be inappropriate.
+This planned trajectory may not itself be intended and might arise from an intent to minimize cost or effort.
 

### 2015-03-23 Paper1Editted/paper1Beditted.tex

@@
 We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay.
-In both pulse timing conditions, the intended trajectory remained straight towards the target, even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.
+In both pulse timing conditions, the intended trajectory remained straight towards the target in excess of 120 milliseconds ($p<0.001$), even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.
 
@@
 Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance.
-Despite the disturbance, three subjects' intents did not depart significantly from baseline.
-The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3B).
+Despite the disturbance, three subjects' intents did not depart significantly from baseline (Figure 3, left portion).
+The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3, right portion).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3A).
+When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected.
 Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, orange region).
@@
 While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three subjects showed no detectable response.
-The absence of response to a large force disturbance seems inconsistent with optimal feedback control framework and is instead consistent with adherence to a trajectory planned before the disturbance.
+The absence of response to a large force disturbance is instead consistent with adherence to a trajectory planned before the disturbance.
 Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof.

### 2015-03-24 Paper1Editted/paper1Beditted.tex
updated derivation

--- 2015-03-23 Paper1Editted/paper1Beditted.tex
@@
 The largest sensitivity arose mainly from uncertainty in joint stiffnesses.
-Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 ms following the onset of a disturbance.
-As expected, the intended trajectory showed no change from undisturbed reaching for more than 150 ms after the disturbance onset.
-Beyond this point, however, we sometimes detected a change in intent, surprisingly even when the hand is already near the target.
+Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 milliseconds following the onset of a disturbance.
+As expected, the intended trajectory showed no change from undisturbed reaching for more than 150 milliseconds after the disturbance onset.
+Beyond this point, however, we detected a change in intent in five out of eight subjects, surprisingly even when the hand is already near the target.
 Knowing such an intent signal is broadly applicable: enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and complex systems that embody planning such as brain machine interfaces, team sports, crowds, or swarms.
@@
 Therefore, it could not well-represent the intent of a simple reaching movement.
-Unlike the equilibrium point of muscle, the equilibrium point of the entire arm matches undisturbed movement.
+Unlike the equilibrium point of muscle, the equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance.
 In the presence of disturbance, the arm might be deflected from its equilibrium.
-It is presently unknown if or how the arm's equilibrium point might change due to disturbances.
+It is presently unknown if or how this arm equilibrium point might change due to disturbances.
 
@@
 
-It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators.
-Importantly, this has been claimed to be the intended movement\cite{feldman1995origin}.
-To consider this, we also generalize our approach to any system with a moving intention that governs the control law.
-This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $q_e$ (sometimes represented as $\lambda$).
-The feedforward component must use and share this actuator and hence use the actuator's equilibrium.
-Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some action space, $x$,
+To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we must take a series of counter-intuitive steps.
+First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$.
+Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components.
+Third, we use power series expansions to model these components in a very general sense,
 \begin{equation} \label{eq:generalPlant}
@@
 \end{equation}
-where $x_e$ is the equilibrium trajectory of the actuator.
-The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ such that $x$ will track a desired path $x_d$.
+where $x_e$ is the equilibrium trajectory of the actuator. Fourth, we describe the actuator control law using the same expansions,
 \begin{equation}
@@
 \end{equation}
-We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives
+where an internal model, $\hat{P}$ and $\hat{E}$, predicts system dynamics and external disturbances in order to determine the actuator equilibrium $x_e$ such that $x$ will track a desired path $x_d$.
+The feedforward component must use the actuator and hence it shares the actuator's equilibrium with the plant.
+This allows us to finally relate the physical system and its control law by solving for $x_e$ and substituting into \eqref{eq:generalPlant},
 \begin{equation}
@@
 \end{equation}
-wherein $x_e$ vanishes, recovering our familiar model.
-While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory.
+wherein $x_e$ vanishes recovering our familiar model.
+While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating this equilibrium and intended trajectory.
 Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero.
+This is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}.
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
-In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
+If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
 
 While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
-Even if the process model was a reverberating loop \cite{fortney2012computational} driven by a reference intent, we might still recover that reference intent by simulating the loop.
-We are not limited to dynamic models as this might be extended to a lookup table.
+Instead of a dynamic model, it could be a lookup table.
 For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
@@
 \begin{equation}
-E=c(K_p\Delta q+K_d \Delta \dot{q})+I(q)\Delta\ddot{q}
+c\big(K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})\big)=\big(\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)\big)-\big(M(q)\ddot{q}+G(q,\dot{q})\big)+E
 \end{equation}
@@
 To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
-The span between departure of the hand and estimated intent was compared against 120 milliseconds using the sign test.
+The time span between departure of the hand and estimated intent was compared against 120 milliseconds using the nonparametric \textit{sign test}.
 The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.
@@
 \subsection*{Human Intent Trajectories}
-Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3A).
+Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3).
 If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands.
 We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay.
-In both pulse timing conditions, the intended trajectory remained straight towards the target, even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.
+In both pulse timing conditions, the intended trajectory remained straight for longer than the hypothesized 120 milliseconds ($p<0.001$), even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.
 
@@
 Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance.
-Despite the disturbance, three subjects' intents did not depart significantly from baseline.
+Despite the disturbance, three of the eight subjects' intents did not depart significantly from baseline (Figure 3A).
 The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3B).
 When disturbances occurred early in reaching, intent showed some signs of correction mid-reach.
-When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3A).
-Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, orange region).
-This counteraction is consistent with both change of intent and stiffening in response to the disturbance.
+When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected.
+Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, bottom panels).
 
@@
 Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated \cite{shadmehr1994adaptive, franklin2003adaptation, burdet2006stability}.
-Demonstrating the validity and plausibility of mathematically transforming those filters to instead convert measured hand trajectories to intent adds a novel method to the arsenal of analytical methods for exploring human motor control.
+Demonstrating the validity and plausibility of mathematically transforming those filters to instead convert measured hand trajectories to intent adds a novel method to the arsenal of analytic methods for exploring human motor control.
 
@@
 \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement.
-By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.
+By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to the virtual trajectory and can be used to determine the muscle equilibrium of a $\lambda$ model.
 
-Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibration.
+Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibrium.
 Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of the mass term in the intention formulation -- reaching intent cannot jump abruptly.
@@
 In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}.
-While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three subjects showed no detectable response.
-The absence of response to a large force disturbance seems inconsistent with optimal feedback control framework and is instead consistent with adherence to a trajectory planned before the disturbance.
-Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof.
+While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three of our eight subjects showed no detectable response.
+The absence of response to a large force disturbance is instead consistent with adherence to a trajectory planned (perhaps entirely) before the disturbance.
+Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain such observations.
 It remains to be seen if and how change in intent might be triggered.
@@
 The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
-However, by substituting an approximation, reverberating loop, or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
+However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
 In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.
@@
 
-While we use the terms \textit{intent}, $x_d$, and \textit{equilibrium trajectory of the arm} interchangeably, they are not always identical.
-We assume that action does not deviate from intent in the absence of external disturbance.
-For this to be true, the equilibrium trajectory of the arm must be the intent.
-Not all action need arise from a continuous intent trajectory, however.
-One might volitionally twitch a muscle without planning its effect on the limb.
-In the context of reaching, our results and others' \cite{flash1991arm, bizzi1984posture, shadmehr1994adaptive, rohrer2004submovements, fortney2012computational, mistry2013optimal} suggest a continuous pre-planned trajectory with kinematic constraints, perhaps composed in an intermittent fashion from some motor primitive.
-This planned trajectory may not itself be intended and might arise from an intent to minimize cost or effort.
+The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature.
+Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances.
+A rich history of research has suggested many ways that an equilibrium trajectory might explain how the central nervous system \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive} can explain action.
+Here we suggest that one should directly define intent in the context of a hypothesized control model.
+Our work here assumed the model of Burdet \cite{burdet2006stability}.
+However, other, more sophisticated, models may be deemed more appropriate in the future.
+What is clear is that given some model of control, this intent extraction approach is able to invert the process to estimate intent.
+
+It is important to mention that it is not necessary for the brain to represent intent explicitly to generate action.
+It is possible that the nervous system can simply learn the relationship between muscle activations and the accomplishment of goals.
+For example, one might simply evaluate a motion outcome and adjust descending signals to the muscles.
+At the same time, one may not deny the fact that there are tasks that require a explicit representation of a trajectory, such as performing a dance, drawing a picture, or conducting an orchestra.
+For these tasks, intent is likely represented at  some level in the nervous system.
+Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not.
+Here our goal was more modest, which was to answer, "where would the arm have gone?"
 
@@
 {\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied.
-This subject's extracted desired trajectories, in orange, do not significantly differ from baseline during force application, but show some reaction afterwards.
+Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent (orange) region (mean $\pm$ 95\% confidence interval on the mean) not departing the undisturbed movement (blue) region.
+The time at which the hand (black) region departed from undisturbed movement is marked with a green dots.
+The time of hand departure for the two subjects with similar trajectories is also marked with green dots.
+Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B).
+The time of intent departure from the undisturbed trajectories is marked with red dots.
 }
@@
 
-\begin{figure}
-\caption{
-{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ 95\% confidence interval on the mean) can no longer be explained by undisturbed hand deviation (blue band).
-The specific time at which this occurs is marked with a green line for each subject.
-Movement intention (orange dots, orange band is mean $\pm$ 95\% confidence interval on the mean) does not significantly deviate until at least 300 ms after the onset of disturbance.
-The specific time at which this occurs is marked with a red line for each subject.
-}
-\label{fig:grouptrends}
-\end{figure}
 \pagebreak

### 2015-03-25 Paper1Editted/paper1Beditted.tex
Paper updates to match rebuttal letter

@@
 $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$.
-This $q_d$ signifies the unknown desired trajectory that we seek to discover.
+This $q_d$ has a dual meaning in that it signifies both the unknown desired trajectory that we seek to discover and also the moving equilibrium trajectory of the arm.
 
@@
 Even as the model is insensitive to normal variations, a human can halve or double their stiffness \cite{franklin2007endpoint} significantly reducing determination accuracy.
+This stiffening may even explain the reported deviation of intent from the baseline trajectories.
 These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations.
@@
 Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not.
-Here our goal was more modest, which was to answer, "where would the arm have gone?"
+Here our goal was more modest, which was to answer, ``where would the arm have gone?''
 
@@
 Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent (orange) region (mean $\pm$ 95\% confidence interval on the mean) not departing the undisturbed movement (blue) region.
-The time at which the hand (black) region departed from undisturbed movement is marked with a green dots.
-The time of hand departure for the two subjects with similar trajectories is also marked with green dots.
+The time at which the hand (black) region departed from undisturbed movement is marked with a green dot for each subject.
 Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B).
-The time of intent departure from the undisturbed trajectories is marked with red dots.
+The time of intent departure from the undisturbed trajectories is marked with a red dot for each subject.
 }

### 2015-03-25 Paper1Editted/paper1Beditted.tex
Updated to cite khalil

@@
 Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components.
-Third, we use power series expansions to model these components in a very general sense,
+Third, we model these components in a very general sense as time-varying linear systems,
 \begin{equation} \label{eq:generalPlant}
@@
 \end{equation}
-where $x_e$ is the equilibrium trajectory of the actuator. Fourth, we describe the actuator control law using the same expansions,
+where $x_e$ is the equilibrium trajectory of the actuator.
+Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and Lipschitz continuous\cite[p.~161]{khalil2002nonlinear}.
+Fourth, we describe the actuator control law using the same expansions,
 \begin{equation}
@@
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
-If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
+If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
 
-While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model.
+While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model so long as its input is a stable equilbrium.
 Instead of a dynamic model, it could be a lookup table.
@@
 As long as some bidirectional relationship exists between state and outcome (even if determined empirically), intended outcome can be determined.
-This allows determination of intent in many situations of interest, even where the process is irreducibly complex.
+This allows determination of intent in many situations of interest, even where the process is otherwise irreducibly complex.
 
@@
 Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not.
-Here our goal was more modest, which was to answer, ``where would the arm have gone?''
+Here our goal was to answer a more modest question, ``where would the arm have gone?''
 

### 2015-04-15 Paper1Editted/paper1Beditted.tex
Not very clear on what I'm committing

@@
 
-% PLoS doesn't support/use the next package, but we get issues if we don't use this to generate the PDF. Comment this line out when submitting.
 \usepackage[section]{placeins}
@@
 Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.
+In other words, we operationally define \textit{intent} as a subjunctive -- where would the motion have gone had it not been disturbed?
 
@@
 By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would.
-Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
+Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this muscle equilibrium point, and not the equilibrium of the whole arm, represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would.
+Upon Gomi and Kawato's inspection of the muscle equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
 Therefore, it could not well-represent the intent of a simple reaching movement.
@@
 This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance.
-If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero.
+If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$; otherwise $\hat{E}$ is set to zero.
 $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$.
@@
 
-To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we must take a series of counter-intuitive steps.
+To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we take a series of steps to leverage a relationship between the equilibria of a nonlinear system and the equilibria of linearizations of that system.
 First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$.
@@
 where $x_e$ is the equilibrium trajectory of the actuator.
-Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and Lipschitz continuous\cite[p.~161]{khalil2002nonlinear}.
-Fourth, we describe the actuator control law using the same expansions,
+Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and has a bounded time derivative (Lipschitz continuous)\cite[p.~161]{khalil2002nonlinear}.
+Fourth, we describe a  control law for the actuator using the same expansions,
 \begin{equation}
@@
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
-If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.
+If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$, revealing the intended trajectory.
 
@@
 \subsection*{Experimental Design}
-We chose 15 centimeters long simulated reaches of the right arm, beginning at 38 centimeters out from and 5.7 centimeters left of the right shoulder and ending at 38 centimeters out from and 10.7 centimeters right of the right shoulder.
+We chose 15 centimeter long simulated reaches of the right arm, beginning at 38 centimeters out from and 5.7 centimeters left of the right shoulder and ending at 38 centimeters out from and 10.7 centimeters right of the right shoulder.
 We used the following two types of perturbing forces for each combination of distance and direction:
@@
 Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories.
-To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
-The time span between departure of the hand and estimated intent was compared against 120 milliseconds using the nonparametric \textit{sign test}.
+To detect this departure, we use the one-tailed Student's t-Test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories.
+The median time span between departure of the hand and departure of the estimated intent was compared against 120 milliseconds using the nonparametric Sign Test.
 The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.
@@
 Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances.
-A rich history of research has suggested many ways that an equilibrium trajectory might explain how the central nervous system \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive} can explain action.
+A rich history of research has suggested many ways that an equilibrium trajectory can explain action through feedback responses \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive}.
 Here we suggest that one should directly define intent in the context of a hypothesized control model.
@@
 \pagebreak
+
 \section*{Figure Legends}
@@
 {\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied.
-Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent (orange) region (mean $\pm$ 95\% confidence interval on the mean) not departing the undisturbed movement (blue) region.
-The time at which the hand (black) region departed from undisturbed movement is marked with a green dot for each subject.
+Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent region not departing the undisturbed movement (blue) region.
+Shaded regions represent mean $\pm$ 95\% confidence interval on the mean.
+The first time at which the hand (black) region departed from undisturbed movement is marked with a green dot for each subject.
 Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B).
@@
 \end{figure}
-
 \pagebreak

### 2015-04-15 Paper 3/2Dfreeexploration/sparseMJPfitResid.m

@@ -1,0 +1,72 @@
+function cost=sparseMJPfitResid(P)
+
+global xdot t
+
+%% Step 0 P->w,ts
+
+lP5=length(P)/5;
+
+%This three defacto modes:
+%1) lP5=2, two fit+one already known
+%2) lP5=3, three fit
+%3) lP5>3, trying to fit something like a whole movement at once.
+
+r=reshape(P,5,lP5);
+
+w=r(1:3,:);
+tc=r(4,:);
+ts=abs(r(5,:));
+
+%% Step 1 cost....sum error.^2 over tc(1)+/-ts(1)
+
+uppers=tc+ts/2;
+lowers=tc-ts/2;
+
+inds=find((t>lowers(1))&(t<uppers(1)));
+tinds=t(inds);
+
+ydot=zeros(length(inds),lP5);
+kerns=zeros(length(inds),lP5);
+
+for k=1:lP5
+    ta=(tinds-tc(k))/ts(k)+.5;
+    kerns(:,k)=(30*ta.^2-60*ta.^3+30*ta.^4)/ts(k);
+    kerns(tinds<=lowers(k),k)=0;
+    kerns(tinds>=uppers(k),k)=0;
+    ydot=ydot+kerns(:,k)*(w(:,k)');
+end
+
+if exist('Pknown','var')
+   %Contribute to ydot and calc overlap from this to 2 
+else
+    overlap=max(0, min(uppers(2),uppers(3))-max(lowers(2),lowers(3)));
+    overlapdir=sign(tc(2)-tc(3));
+end
+
+fit_error=ydot-xdot(inds,:);
+gradw=zeros(lP5,3);
+gradtc=zeros(lP5,1);
+gradts=zeros(lP5,1);
+
+for k=1:lP5
+    gradw(k,:)=sum([fit_error(:,1).*kerns(:,k) fit_error(:,2).*kerns(:,k) fit_error(:,3).*kerns(:,k)]); %After summing, should be a row vector
+
+    ta=(tinds-tc(k))/ts(k)+.5;
+    tcdiff=(-60*ta+180*ta.^2-120*ta.^3)/(ts(k)^2);
+    gradtc(k)=sum(sum([w(1,k)*fit_error(:,1).*tcdiff, w(2,k)*fit_error(:,2).*tcdiff, w(3,k)*fit_error(:,3).*tcdiff]));
+    
[…]

### 2015-04-20 Paper1rev2/paper1rev1.tex
updated for rev2

--- 2015-04-15 Paper1Editted/paper1Beditted.tex
@@
 \usepackage[section]{placeins}
+\usepackage{graphics} % for pdf, bitmapped graphics files
 
@@
 \begin{equation}
-\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
+\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripetal}}}_{\text{Plant}}+E=0
 \end{equation}
@@
 \begin{equation}
-\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
+\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripetal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripetal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impedance, Feedback Controller}}
 \end{equation}
@@
 
-To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we take a series of steps to leverage a relationship between the equilibria of a nonlinear system and the equilibria of linearizations of that system.
+To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we take a series of steps.
 First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$.
 Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components.
-Third, we model these components in a very general sense as time-varying linear systems,
+Third, we model these components in a very general sense as operations,
 \begin{equation} \label{eq:generalPlant}
@@
 \end{equation}
-where $x_e$ is the equilibrium trajectory of the actuator.
-Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and has a bounded time derivative (Lipschitz continuous)\cite[p.~161]{khalil2002nonlinear}.
-Fourth, we describe a  control law for the actuator using the same expansions,
+where $x_e$ is the equilibrium trajectory of the actuator. Fourth, we describe a  control law for the actuator using the same expansions,
 \begin{equation}
@@
 Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero.
-This is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}.
+
+The fact that the actuator's equilibrium $x_e$ is not the process's equilibrium is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}.
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
-If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$, revealing the intended trajectory.
+If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the system is stable (e.g., the stiffness and damping are positive), it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$, revealing the intended trajectory.
 
-While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model so long as its input is a stable equilbrium.
-Instead of a dynamic model, it could be a lookup table.
+While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model, simply because any hypothesis can be modeled and used by our approach to determine where the model would have gone had it not been disturbed.
+
+Note that instead of a dynamic model, any model may be possible, such as lookup table.
 For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism.
@@
 \subsection*{Indices of variance-based sensitivity}
-To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
+To ensure a properly spaced and efficient evaluation of sensitivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density.
 Nominal parameter values were mostly taken from Dempster \cite{dempster1955space} and Burdet et al.
@@
 It remains to be seen if there are methods to further reduce the mis-estimation due to force error or eliminate the need for force estimation altogether.
+It will be also interesting to determine whether relaxing constraints and operating in 3D would produce the same levels of accuracy found here.
 
@@
 This stiffening may even explain the reported deviation of intent from the baseline trajectories.
-These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations.
-Additionally, real-time determination of intent may allow real-time determination of stiffness.
+Of particular concern are the large hooks at the end of the motions.
+These limitations may be overcome by detecting muscle co-contraction or using additional modeling assumptions to allow real-time estimation of stiffness.
 
@@
 % Do NOT remove this, even if you are not including acknowledgments.
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 Research supported by US NIH NINDS1 R01 NS053606.
@@
 %
+
 \pagebreak
@@
 %\end{figure}
-\begin{figure}
+\begin{figure}[!b]
+\includegraphics{fig1.eps}
 \caption{
@@
 \begin{figure}
+\includegraphics{fig2.eps}
 \caption{
@@
 \begin{figure}
+\includegraphics{fig3.eps}
 \caption{

### 2015-04-20 Paper1rev2/paper1rev2.tex
updated for rev2

--- 2015-04-20 Paper1rev2/paper1rev1.tex
@@
 The motion behavior changes with the addition of feedforward and/or feedback controllers,
-\begin{equation}
+\begin{equation} \label{eq:controlLaw}
 \underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripetal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripetal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impedance, Feedback Controller}}
@@
 Intent determination takes the novel approach of instead solving for $\ddot{q}_d$,
-\begin{equation}
+\begin{equation} \label{eq:solveDDQ}
 \ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})]\right\}
@@
 \cite{burdet2006stability}.
-We fit a subject-specific constant scale factor for impedance, $c$, relative to this model to account for any task-dependence of the subjects' impedance\cite{gomi1998task}.
-\begin{equation}
-c\big(K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})\big)=\big(\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)\big)-\big(M(q)\ddot{q}+G(q,\dot{q})\big)+E
-\end{equation}
-This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that task-perpendicular torque and displacement from the first 150 ms of movement from noise-disturbed trials is used in place of a stationary task.
-
+We fit a subject-specific constant scale factor for impedance relative to this model to account for any task-dependence of the subjects' impedance\cite{gomi1998task} as described in the next section.
 Desired trajectory in time was idealized as a typical minimum jerk, 5th order polynomial of duration 700 ms starting and ending with zero velocity and acceleration \cite{flash1985coordination}.
@@
 Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.
+
+\subsection*{Torque Calculations}
+The impedance properties of the human arm are known to be task-dependent\cite{gomi1998task}, but we need to model them in the context of our task without assuming our conclusions in the premises.
+To accomplish this, we fit a linear scale factor to the torque output an established model using an intermittently, unpredictably presented secondary task.
+This unpredictability enables us to leverage the ergodic assumption: our tasks should at least initally share the same motor plan and impedance properties.
+Both portions of this assumption are important.
+First, we fit some expectation of the motor plan, $x_d$, by using a Gaussian weighted average ($\sigma=22$ milliseconds) of each subject's hand trajectory as it evolves in time during undisturbed movement.
+Separate expectations were fit for each pair of movement start and end points.
+We then use inverse kinematics to convert these expected hand trajectories to expected joint angle trajectories, $q_d$.
+Taking time derivates of $q_d$ then allows us to form an expectation of the feedforward torque, $\tau_{ff}$, that will be present early in movement as in \eqref{eq:controlLaw},
+\begin{equation}
+\tau_{ff}=\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)+\hat{E}
+\end{equation}
+which then allows us to construct the full torque balance for the joints,
+\begin{equation} \label{eq:torqueBalance}
+\tau_P+E=\tau_{ff}+c\tau_{fb}
+\end{equation}
+where $\tau_P$ is calculated from the arm's dynamics, $\tau_{fb}$ is calculated from the impedance model of Burdet et al.\cite{burdet2006stability}, and $c$ is the scaling parameter to be fit.
+We calculate $\tau_P$ using measured limb segment lengths, $L_1$ and $L_2$ and self-reported body mass, $m_g$.
+We converted from these gross measurements to specific parameters using the nominal ratios provided in Table~\ref{tab:parameters},
+\begin{align}
+m_1 &=0.028 m_g & L_{m1}&=0.426 L_1 & J_1&=m_1(0.322 L_1)^2 \\  
+m_2 &=0.022 m_g & L_{m2}&=0.682 L_2 & J_2&=m_2(0.468 L_2)^2
+\end{align}
+where $m_1$ and $m_2$ are segment masses, $L_{m1}$ and $L_{m2}$ are limb centers of mass, and $J_1$ and $J_2$ are mass moments of inertia.
+We then calculate $M$, $G$, and $\tau_P$ using these subject-specific parameters and measured values for the joint trajectory, $q$, and its time derivatives from reaches that received the white noise force disturbance,
+\begin{align}
+M&=\begin{bmatrix}J_1+J_2+m_1 L_{m1}^2+m_2(L_1^2+L_{m2}^2+2 L_1 L_{m2} \cos{q_2}) & J_2+m_2(L_{m2}^2+L_1 L_{m2} \cos{q_2}) \\
+J_2+m_2(L_{m2}^2+L_1 L_{m2} \cos{q_2}) & J_2+m_2L_{m2}^2
+\end{bmatrix} \\
+G&=\begin{bmatrix}
+m_2 L_1 L_{m2} \dot{q}_2(2\dot{q}_1+\dot{q}_2)\sin{q_2} \\
+m_2 L_1 L_{m2} \dot{q}_1^2\sin{q_2}
+\end{bmatrix} \\
+\tau_P&=M(q)\ddot{q}+G(q,\dot{q})
+\end{align}
+for the first 150 milliseconds following the onset of movement.
+$\tau_{ff}$ was calculated in the same fashion $\tau_{ff}$ except that $\hat{M}$ and $\hat{G}$ were calculated by replacing all instances of $q$ with $q_d$.
+We then calculated $\tau_{fb}$ as in Burdet et al.\cite{burdet2006stability} while additionally leveraging the dual identity of the muscle torque, $\tau_m$,
+\begin{align}
+|\tau_m|&=|\tau_{ff}+\tau_{fb}|=|\tau_P+E| \\
+K&=\begin{bmatrix}
+10.8+3.18|\tau_{m1}| & 2.83+2.15|\tau_{m2}| \\
+2.51+2.34|\tau_{m2}| & 8.67+6.18|\tau_{m2}|
+\end{bmatrix}\\
+\tau_{fb}(t)&=K\big(q_d(t)-q(t)+\frac{\dot{q}_d(t)-q(t)}{12}+\frac{q_d(t-\phi)-q(t-\phi)+2(\dot{q}_d(t-\phi)-q(t-\phi))}{50}\big)
+\end{align}
+where $\phi$ is a delay of 60 milliseconds. Finally, constrained optimization is performed to find the value of $c$ that minimizes the expression,
+\begin{equation}
+\tau_{ff}+c\tau_{fb}-\tau_P-E
+\end{equation}
+for each subject as the sum of all torques acting on the joints should ideally be zero.
+The optimization is constrained such that $c>0.15$ both to avoid unphysiological values and numerical instabilities during simulation.
+The value of $c$ is then used when extracting intended trajectories from pulse-disturbed movements as in \eqref{eq:solveDDQ},
+\begin{equation}
+\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+c\tau_{fb}]\right\}
+\end{equation}
+and integrated numerically as described. This provides a check on the fit value $c$ as an inappropriate value would cause $q_d$ to deviate from reasonable values.
 
@@
 Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not.
-Here our goal was to answer a more modest question, ``where would the arm have gone?''
+Here our goal was to answer a more modest question, ``\textit{when} movement intent is a changing equilibrium, can we recover it despite external disturbances?''
 
@@
 %\end{figure}
-\begin{figure}[!b]
-\includegraphics{fig1.eps}
+\begin{figure}
 \caption{
@@
 
-\begin{figure}
-\includegraphics{fig2.eps}
+\begin{figure}[!]
 \caption{
@@
 
-\begin{figure}
-\includegraphics{fig3.eps}
+\begin{figure}[!]
 \caption{

### 2015-04-22 Paper1rev2/paper1rev2.tex
updated for work from home

@@
 
-\subsection*{Torque Calculations}
+\subsection*{Torque Calculations for Intent Extraction}
 The impedance properties of the human arm are known to be task-dependent\cite{gomi1998task}, but we need to model them in the context of our task without assuming our conclusions in the premises.
@@
 \begin{equation}
-\tau_{ff}=\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)+\hat{E}
+\tau_{ff}=\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)
 \end{equation}
@@
 \end{equation}
-where $\tau_P$ is calculated from the arm's dynamics, $\tau_{fb}$ is calculated from the impedance model of Burdet et al.\cite{burdet2006stability}, and $c$ is the scaling parameter to be fit.
+where $E$ is the measured force on the hand converted to joint torques via the Jacobian, $\tau_P$ is calculated from the arm's dynamics, $\tau_{fb}$ is calculated from the impedance model of Burdet et al.\cite{burdet2006stability}, and $c$ is a scalar constant to be fit.
 We calculate $\tau_P$ using measured limb segment lengths, $L_1$ and $L_2$ and self-reported body mass, $m_g$.
+Terms followed by a subscript 1 indicate the upper arm or shoulder joint while terms followed by a subscript 2 indicate the forearm or elbow joint.
 We converted from these gross measurements to specific parameters using the nominal ratios provided in Table~\ref{tab:parameters},
@@
 \end{equation}
-and integrated numerically as described. This provides a check on the fit value $c$ as an inappropriate value would cause $q_d$ to deviate from reasonable values.
+and integrated numerically as described.
+By scaling an established model using a force disturbance other than our disturbance of interest, we avoided presuming the intent trajectory we hoped to recover.
 

### 2015-04-29 Paper2frontiers/frontiers.tex
Updated for paper2 rewrite

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use frontiers.cls nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontier after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.0 Generated 2014/12/19 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics articles

%\setcitestyle{square}
\usepackage{url,lineno}
\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Horowitz {et~al.}} %use et al only if is more than 1 author
\def\Authors{Justin Horowitz\,$^1$, Co-Author\,$^{2}$ and Co-Author\,$^{2,*}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{James Patton}
\def\corrAddress{Laboratory X, Institute X, Department X, Organization X, Street X, City X , State XX (only USA, Canada and Australia), Zip Code, X Country X}
\def\corrEmail{pattonj@uic.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[Running Title]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
\topic{}% If your article is part of a Research Topic, please indicate here which.

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% The sections below are for reference only.
%%%
%%% For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the
%%% following sections or their equivalents for your field:
%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
%%% Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
%%%
%%%For information about Clinical Trial Registration, please go to http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
%%%
%%% For Clinical Case Studies the following sections are mandatory: Abstract, Introduction, Background, Discussion, and Concluding Remarks.
%%%
%%% For all other article types there are no mandatory sections.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

%%% Leave the Abstract empty if your article falls under any of the following categories: Editorial Book Review, Commentary, Field Grand Challenge, Opinion or specialty Grand Challenge.
\section{}
%As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.
Refer to \url{http://www.frontiersin.org/about/AuthorGuidelines} \\ or \textbf{Table\ref{Tab:01}} for abstract requirement and length according to article type.

\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
Humans often interact with uncertain and complicated environments, such as turbulence, crowds, and traffic. Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning. In such an environment, planning a specific might seem nonsensical. Instead, a continuous cycle of reaction to the environment seems intuitive. We recently developed an algorithm that could recover the path a person intended to take even if they were forced off that path by environmental disturbances \citep{horowitz2015determining}. Surprisingly, some subjects did not alter their intended even following exposure to unexpected pulses of force. If people continuously reacted to their environment, we should always see some response.

%\begin{methods}
\section{Material \& Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{Neuro2013} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text. \citep{Gene2012} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{Neurobot2013} might want to know about  text text text text

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:01} Table \arabic{table}.}{ Maximum size of the Manuscript }

\processtable{ }
{\begin{tabular}{lllll}\toprule
 & Abstract max. legth (incl. spaces) & Figures or tables & Manuscript max. length & Final PDF length\\\midrule
Clinical Case Study & & & &\\
Clinical Trial & & & &\\
Hypothesis and Theory & & & &\\
Methods & 2000 characters  & 15 & 12000 words & 12 pages\\
Original Research & & & &\\
Review & & & &\\
Technology Report & & & &\\
Focused Review & 2000 characters & 5 & 5000 words & 5 pages\\
CPC &  1250 characters& 6 & 2500 words & 4 pages\\
Perspective & 1250 characters & 2 & 3000 words & 3 pages\\
Mini Review & & & &\\
Classification & 1250 characters & 10 & 2000 words & 12 pages\\
Editorial & none & none & 1000 words & 1 page \\
Book review & & & &\\
Frontiers Commentary & none & 1 & 1000 words & 1 page\\
General Commentary & & & &\\
Field Grand Challenge & & & &\\
Opinion & none & 1 & 2000 words & 2 pages\\
Specialty Grand Challenge& & & &\\\botrule
\end{tabular}}{}
\end{table}

Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space. These tables will be published as supplementary material on the online article abstract page at the time of acceptance. The author will notified during the typesetting of the final article if this is the case. A link in the final PDF will direct to the online material.

\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}

For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the following sections or their equivalents for your field:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Succinct, with no subheadings.
\item Materials and Methods: This section may be divided by subheadings. This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
\item Discussion: This section may be divided by subheadings. Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
\end{itemize}

Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\subsection{Clinical Case Studies}

For Clinical Case Studies the following sections are mandatory:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Include symptoms at presentation, physical exams and lab results.
\item Background: This section may be divided by subheadings. Include history and review of similar cases.
\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
\item Concluding Remarks
\end{itemize}

%\end{methods}

\section{Results}

Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions. Figures which are not according to the guidelines will cause substantial delay during the production process.

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{ Resolution Requirements for the figures}

\processtable{}
{\begin{tabular}{lllll}\toprule
Image Type & Description & Format & Color Mode & Resolution\\\midrule
Line Art & An image composed of lines and text,  & TIFF, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
           & which does not contain tonal or shaded areas.& & &\\
           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
Combination & Image contains halftone + text or line art elements. & TIFF, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
\end{tabular}}{}
\end{table}

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\textbf{Table\ref{Tab:02}} shows the resolution requirements for the figures. The figures must be legible:
\begin{enumerate}
\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
\item Solid lines are not broken up.
\item Image areas are not pixelated or stair stepped.
\item Text is legible and of high quality.
\item Any lines in the graphic are no smaller than 2 points width.
\item The actual size of the figure must be of at least 8.5 cm.
\end{enumerate}

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
Additional Requirements:
\subsection{Corrections}

If you need to communicate important changes to a published article please submit a General Commentary. Submit the article with the title “Corrigendum: Original Title of Article”.

\subsection{Commentaries on Articles}

At the beginning of your manuscript provide the citation of the article commented on.

\subsection{Focused Reviews}

For Tier 2 invited Focused Reviews the sections Introduction, Material and Methods, Results, and Discussion are recommended. In addition the authors must submit a short biography of the corresponding author(s). This short biography has a maximum of 600 characters, including spaces.

A picture (5 x 5 cm, in *.tif or *.jpg, min 300 dpi) must be submitted along with the biography in the manuscript and separately during figure upload.
Focused Reviews highlight and explain key concepts of your work. Please highlight a minimum of four and a maximum of ten key concepts in bold in your manuscript and provide the definitions/explanations at the end of your manuscript under “Key Concepts”. Each definition has a maximum of 400 characters, including spaces.

\subsection{Human Search and Animal Research}

All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations. In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards. For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects. In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.

\subsection{Clinical Trial Registration}

Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers. Trials must be registered at or before the start of patient enrollment. A clinical trial is defined as"any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\url{www.who.int/ictrp/en}). A list of acceptable registries can be found at \url{www.who.int/ictrp/en and www.icmje.org}.

\subsection{Inclusion of Proteomics Data}

Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches. Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.

Authors should provide as supplementary material information used to identify proteins and/or peptides. This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.

For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.

\subsection{Data Sharing}

Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research. In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.

\section*{Disclosure/Conflict-of-Interest Statement}
%Frontiers follows the recommendations by the International Committee of Medical Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require that all financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest. When disclosing the potential conflict of interest, the authors need to address the following points:
%•	Did you or your institution at any time receive payment or services from a third party for any aspect of the submitted work?
%•	Please declare financial relationships with entities that could be perceived to influence, or that give the appearance of potentially influencing, what you wrote in the submitted work.
%•	Please declare patents and copyrights, whether pending, issued, licensed and/or receiving royalties relevant to the work.
%•	Please state other relationships or activities that readers could perceive to have influenced, or that give the appearance of potentially influencing, what you wrote in the submitted work.

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
%When determining authorship the following criteria should be observed:
%•	Substantial contributions to the conception or design of the work; or the acquisition, analysis, or interpretation of data for the work; AND
%•	Drafting the work or revising it critically for important intellectual content; AND
%•	Final approval of the version to be published ; AND
%•	Agreement to be accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
%Contributors who meet fewer than all 4 of the above criteria for authorship should not be listed as authors, but they should be acknowledged. (http://www.icmje.org/roles_a.html)

The statement about the authors and contributors can be up to several sentences long, describing the tasks of individual authors referred to by their initials and should be included at the end of the manuscript before the References section.

\section*{Acknowledgments}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.

\section*{Supplemental Data}
Supplementary Material should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

%\bibliographystyle{frontiersinSCNS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{paper2.bib}

%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file

\section*{Figures}

%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.jpg and logo2.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{logo1}% This is a *.jpg file
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}

%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.

%%% Frontiers will add the figures at the end of the provisional pdf automatically %%%

%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}

### 2015-05-04 Paper2frontiers/frontiers.tex
updated for writing from home

@@
 
-\title[Running Title]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence}
+\title[Running Title]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Forces}
 \author[\firstAuthorLast ]{\Authors}
@@
 Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning.
-In such an environment, planning a specific might seem nonsensical.
-Instead, a continuous cycle of reaction to the environment seems intuitive.
-We recently developed an algorithm that could recover the path a person intended to take even if they were forced off that path by environmental disturbances \citep{horowitz2015determining}.
-Surprisingly, some subjects did not alter their intended even following exposure to unexpected pulses of force.
-If people continuously reacted to their environment, we should always see some response.
+In such an environment, planning a specific route might seem nonsensical.
+Instead, a continuous cycle of reaction to the environment with no specific intended route might seem more intuitive.
+We explore the counter-intuitive possibility that a specific route is constructed even when the environment is unstable and unpredictable.
+
+If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended path using optimal feedback control\citep{todorov2002optimal}.
+This modeling strategy is not necessarily a scientific hypothesis as a measure of cost can always be found \text{a posteriori} that reproduces the behaviors observed and no hypothesized cost function has been claimed as sufficient to describe all reaching.
+This   but they predict reaction to disturbance.
+We recently developed an algorithm that could recover the route a person intended to take even if they were forced off of it by environmental disturbances \citep{horowitz2015determining}.
+Surprisingly, some subjects did not alter their intended route even following exposure to unexpected pulses of force.
+If humans a plan a route, even if they might change that route in response to disturbance, can they use feedback?
+Is a route really what they want?
+Does it take learning?
 

### 2015-05-06 Paper2frontiers/frontiers.tex
updated with intro stuff

--- 2015-04-29 Paper2frontiers/frontiers.tex
@@
 
-\title[Running Title]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite Turbulence}
+\title[Realtime Feedback of Intended Trajectory]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Forces}
 \author[\firstAuthorLast ]{\Authors}
@@
 % For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
-Humans often interact with uncertain and complicated environments, such as turbulence, crowds, and traffic.
+Humans often interact with machines in uncertain and complicated environments, such as crowds and traffic, where they must contend with turbulence, moving obstacles, distractions, and disturbances.
 Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning.
-In such an environment, planning a specific might seem nonsensical.
-Instead, a continuous cycle of reaction to the environment seems intuitive.
-We recently developed an algorithm that could recover the path a person intended to take even if they were forced off that path by environmental disturbances \citep{horowitz2015determining}.
-Surprisingly, some subjects did not alter their intended even following exposure to unexpected pulses of force.
-If people continuously reacted to their environment, we should always see some response.
+Sensory delays, sensory fusion, delays in sensory processing, and delays in action further confound our ability to behave reactively.
+The nervous system intelligently solves many problems by planning ahead \citep{belen1967control}, yet in the face of uncertainties we often cannot adequately prevent errors.
+There is the possibility, however, to exploit additional information from instruments that may be available -- particularly fast and accurate force sensors -- that can measure human machine interactions.
+Combining sensors with filtering techniques makes it possible to determine a person’s underlying intent -- the motion they would have done if they had not been disturbed.
+This intent provides new ways to understand the nature of control and provide novel feedback.
+
+Recent work in our laboratory has attempted to outline the methodology for obtaining estimates of intent \citep{horowitz2015determining}.
+This method assumes a model of the dynamics and control of the human arm, and then through a manipulation of the equations of motion it integrates to find a unique estimate of intent.
+The algorithm recovers the trajectory a person intended to take, even if they were forced away from it due to environmental disturbances.
+This analysis has enabled us to show how subjects might sometimes alter their intent following exposure to unexpected force pulses.
+
+A new question that arises is whether seeing one’s own intent, rather that what what actually happens, may be useful.
+The intent extraction method can be abriged to allow for real-time estimations and presented to the subject as a cursor.
+It is possible that performance using this estimated intention may compare quite favorably to the actual outcome in the presence of unexpected disturbances.
+If so, such a method holds great promise in any situation where humans and machines interact as it enables the machine to give the human operator what they want.
+This human-machine collaboration might be able to outperform what a person can do alone.
+
+In contrast, displaying anything other than what truthfully happens is a distortion and a lie.
+Like many other visual distortion experiments \citep{miles1980long, pine1996learning}, intent feedback (IF) introduces a visuomotor discrepancy that may be confusing to the nervous system.
+Preplanning a specific route may not necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal.
+If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended route using optimal feedback control\citep{todorov2002optimal}.
+While this modeling strategy has been very successful at explaining data, it fundamentally assumes that corrective actions will be taken in response to relevant disturbances.
+Goals can also be reached at minimal \textit{expected} cost by constructing -- and possibly updating -- a specific intended route.
+IF may be troublesome if the nervous system cannot recognize it because no particular trajectory was intended.
+
+While performance is the best indicator of IF's value, changes in arm stiffness can provide secondary confirmation that subjects are actually getting what they want.
+Arm stiffness is known to increase during exposure to instability \cite{franklin2003adaptation} and uncertainty \cite{takahashi2001impedance}.
+We anticipated that these change might also be modulated by presentation of IF.
+Reduces the effects of environmental instability and presenting the subject with a signal already known to them may make the arm relax and become more compliant.
+We hypothesized that any elevated arm stiffness from noisy disturbances would decrease when subjects received IF.
+
+In this paper we describe our abridged method for real-time IF, and then expose subjects to IF in an unstable and unpredictable environment to compare their ability to perform goal-directed reaching using visual feedback of their hand position against visual feedback of their estimated intended position.
+We hypothesized that IF should lead to better performance in the presence of force-based disturbances.
+Accordingly, we hypothesized that during random disturbances, the intent trajectory should deviate less than the hand.
 

### 2015-05-06 Paper2frontiers/frontiers.tex
mild intro updates

@@
 
-\title[Realtime Feedback of Intended Trajectory]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Forces}
+\title[Realtime Feedback of Intended Trajectory]{Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Environment}
 \author[\firstAuthorLast ]{\Authors}
@@
 
-While performance is the best indicator of IF's value, changes in arm stiffness can provide secondary confirmation that subjects are actually getting what they want.
+While performance is the best indicator of IF's value, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want.
 Arm stiffness is known to increase during exposure to instability \cite{franklin2003adaptation} and uncertainty \cite{takahashi2001impedance}.
 We anticipated that these change might also be modulated by presentation of IF.
-Reduces the effects of environmental instability and presenting the subject with a signal already known to them may make the arm relax and become more compliant.
+Reducing the effects of environmental instability and presenting the subject with a signal already known to them should make the arm relax and become more compliant.
 We hypothesized that any elevated arm stiffness from noisy disturbances would decrease when subjects received IF.
@@
 
-%\begin{methods}
+\begin{methods}
 \section{Material \& Methods}

### 2015-05-11 Paper2frontiers/frontiers.tex
updated a smidge. Needs more work.

@@
 %As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.
-Refer to \url{http://www.frontiersin.org/about/AuthorGuidelines} \\ or \textbf{Table\ref{Tab:01}} for abstract requirement and length according to article type.
+Unstable and uncertain environments are an everyday problem that affects piloting and input devices in human-machine interaction.
+Here we experimentally evaluate intent feedback (IF), which estimates and displays the human operator’s underlying intent in real-time.
+IF is a filter that combines a model of the arm with position and force data to determine the intended position.
+Subjects performed targeted reaching motions while seeing either their hand position or their intent estimate as a cursor while they experienced white noise forces rendered by a robotic handle.
+We found significantly better reaching performance during force exposure using the estimated intent.
+Additionally, IF reduced estimated arm stiffness to about half that without IF, indicating a more relaxed state of operation.
+While visual distortions typically degrade performance and require an adaptation period to overcome, this particular distortion immediately enhances performance.
+In the future, this method could provide novel insights into the nature of control.
+IF might also be applied in driving and piloting applications to best follow a person’s desire in unpredictable or turbulent conditions.
 
@@
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\cite{Neuro2013} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\citep{Gene2012} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\cite{Neurobot2013} might want to know about  text text text text
+\subsection{Intent Extraction}
+The well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} relates arm trajectory, $q$, to desired arm trajectory, $q_d$, and any external disturbance, $E$ using physical parameters of the arm.
+To show how this model can be algebraically inverted to instead describe desired arm trajectory as a function of arm trajectory and external disturbance, we write it as a torque balance:
+\begin{equation}
+\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=
+\underbrace{\overbrace{\tau_{ff}}^{\text{Feedforward}}+\overbrace{\tau_{fb}}^{\text{Feedback}}}_\text{Controller}
+\end{equation}
+Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocities, $\ddot{q}$ is joint angular accelerations, and $G$ contains both Coriolis and centripetal effects.
+Typical applications solve this torque balance for $\ddot{q}$ and use a numerical differential equation solver to predict arm trajectory in the context of a disturbance of interest, a feedback model, and feedforward torques determined by inverse dynamics.
+Rather than test hypotheses regarding the learning, production, or composition of this feedforward torque, we instead solve for it:
+\begin{equation}
+\tau_{ff}=M(q)\ddot{q}+G(q,\dot{q})+E-\tau_{fb}
+\end{equation}
+Then we note that feedforward torque can have a one-to-one correspondence with desired acceleration, $\ddot{q}_d$:
+\begin{equation}
+\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)+\hat{E}=\tau_{ff}
+\end{equation}
+Hats ( $\hat{}$ ) denote the nervous system's best estimate of a physical quantity. Combining these expressions, suppressing state dependencies, and solving for $\ddot{q}_d$:
+\begin{equation}
+\ddot{q}_d=\hat{M}^{-1} \{M\ddot{q}+G-\hat{G}+E-\hat{E}-\tau_{fb}\}
+\end{equation}
+In this form, a differential equation solver can determine $q_d$ as it evolves in time if a few assumptions are made and conditions are met.
+First, $\hat{E}$ must be modeled or assumed, so we choose $\hat{E}=0$.
+In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures\cite{scheidt2001learning}.
+Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{gomi1998task} and may vary over the course of a reach\cite{niu2010temporal}.
+With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
+The experiment was repeated with a lower stiffness estimate (experiment 2) to explore any dependence on this assumption.
 
-\begin{table}[!t]
-\textbf{\refstepcounter{table}\label{Tab:01} Table \arabic{table}.}{ Maximum size of the Manuscript }
+\subsection{Apparatus}
+A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass.
+The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
+Visual feedback of cursor position was performed at 60 Hz using OpenGL.
+Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
 
-\processtable{ }
-{\begin{tabular}{lllll}\toprule
- & Abstract max. legth (incl. spaces) & Figures or tables & Manuscript max. length & Final PDF length\\\midrule
-Clinical Case Study & & & &\\
-Clinical Trial & & & &\\
-Hypothesis and Theory & & & &\\
-Methods & 2000 characters  & 15 & 12000 words & 12 pages\\
-Original Research & & & &\\
-Review & & & &\\
-Technology Report & & & &\\
-Focused Review & 2000 characters & 5 & 5000 words & 5 pages\\
-CPC &  1250 characters& 6 & 2500 words & 4 pages\\
-Perspective & 1250 characters & 2 & 3000 words & 3 pages\\
-Mini Review & & & &\\
-Classification & 1250 characters & 10 & 2000 words & 12 pages\\
-Editorial & none & none & 1000 words & 1 page \\
-Book review & & & &\\
-Frontiers Commentary & none & 1 & 1000 words & 1 page\\
-General Commentary & & & &\\
-Field Grand Challenge & & & &\\
-Opinion & none & 1 & 2000 words & 2 pages\\
-Specialty Grand Challenge& & & &\\\botrule
-\end{tabular}}{}
-\end{table}
+\subsection{Human Subjects}
+The human data trajectories analyzed here are drawn from nine subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
+Seven male and one female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
+Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
 
-Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
-These tables will be published as supplementary material on the online article abstract page at the time of acceptance.
-The author will notified during the typesetting of the final article if this is the case.
-A link in the final PDF will direct to the online material.
+\subsection{Experimental Design}
+We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at 60 degree intervals about that central point.
+Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of 5 blocks of 96 reaches each.
+During blocks 2 through 4, subjects experienced filtered white noise forces drawn from a white noise generator at 1000 Hertz with flat power spectral density of 1 Newton, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ radians per second.
+During block 3, cursor position indicated deduced intent.
+During all other blocks, the cursor position indicated hand position (Figure 2A).
 
-\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}
+\subsection{Dynamic simulation of arm and intended trajectories}
+Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} were used to estimate relationships between body mass and limb mass, limb length and limb center of mass, and limb mass and length and moment of inertia.
+Viscosity parameters were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
+Stiffness parameters were either taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} ($K_{P1}$, experiment 1) or estimated ($K_{P2}$, experiment 2).
+Expressed in Newton-Meters per Radian:
+\begin{equation}
+K_{P1}=
+\begin{bmatrix}
+15 & 6 \\
+6 & 16
+\end{bmatrix}
+\quad
+K_{P2}=
+\begin{bmatrix}
+8 & 2 \\
+2 & 5
+\end{bmatrix}
+\end{equation}
+Numerical simulation was performed in real time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\cite{gough2009gnu}.
 
-For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself.
-It is recommended to organize your manuscript in the following sections or their equivalents for your field:
-
-\begin{itemize}
-%for bulleted list, use itemize
-\item Introduction: Succinct, with no subheadings.
-\item Materials and Methods: This section may be divided by subheadings.
-This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
-\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
-\item Discussion: This section may be divided by subheadings.
-Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
-\end{itemize}
-
-Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
-
-\subsection{Clinical Case Studies}
-
-For Clinical Case Studies the following sections are mandatory:
-
-\begin{itemize}
-%for bulleted list, use itemize
-\item Introduction: Include symptoms at presentation, physical exams and lab results.
-\item Background: This section may be divided by subheadings. Include history and review of similar cases.
-\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
-\item Concluding Remarks
-\end{itemize}
-
-%\end{methods}
+\subsection{Metrics and Statistical Analysis}
+Trajectories and forces were rotated such that movement and force parallel to the line connecting the previous target (the reach origin) and the presented target lie along a \textit{progress} axis while perpendicular movement and force lie along an \textit{error} axis.
+Reach onset was detected as the moment the cursor's distance from the center of the previous target first exceeded 1 centimeter.
+Maximum perpendicular error for a trajectory was the largest error magnitude within 250 milliseconds of reach onset.
+A scalar stiffness, $k$, was calculated for the \textit{error} axis during this same 250 millisecond time span by linear regression:
+\begin{equation}
+F_e=m\ddot{e}+b\dot{e}+ke+F_O
+\end{equation}
+Force ($F_e$) and state ($\ddot{e}$, $\dot{e}$, $e$) were known.
+Mass ($m$), viscosity ($b$), and stiffness offset ($F_0$) terms were calculated, but discarded.
+While stiffness is better described as a matrix, a scalar was calculated to facilitate statistical comparison between treatments and blocks.
+The paired Wilcoxon signed rank test was used to detect differences in maximum perpendicular error and stiffness between blocks and treatments at the 5\% significance level using the MATLAB statistics toolbox package \cite{MATLAB:2014}.
+\end{methods}
 
 \section{Results}
+As expected, the model was able to deduce an intended trajectory and all subjects were able to utilize this estimate of their intent to perform targeted reaching while experiencing turbulent forces (Figure 1).
+The following sections address the quality of this reaching.
 
-Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
-Figures will then be automatically embedded at the bottom of the submitted manuscript.
-Kindly ensure that each table and figure is mentioned in the text and in numerical order.
-Permission must be obtained for use of copyrighted material from other sources (including the web).
-Please note that it is compulsory to follow figure instructions.
-Figures which are not according to the guidelines will cause substantial delay during the production process.
+\begin{figure}[!b]
+\centering
+\framebox{\parbox{3in}{\includegraphics{Yplots.eps}
+}}
+\caption{Typical subjects' made center-out targeted reaching motions under experimentally varied force and feedback conditions.
+Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
+While both were recorded under all conditions, only one was shown on screen at a time.
+Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
+The force disturbance was designed to be turbulent and unpredictable to minimize any effect of learning.}
+\label{Yplots}
+\end{figure}
 
-\begin{table}[!t]
-\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{ Resolution Requirements for the figures}
+\begin{figure}[t]
+\centering
+\framebox{\parbox{3in}{\includegraphics{error.eps}
+}}
+\caption{Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback (A).
+Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance degraded reaching performance in both hand (B) and intent (C).
+Comparisons within treatment conditions revealed a significant advantage for the hand when force disturbance is not present and a significant advantage for intent when feedback of intent was used during exposure to turbulent forces.
+Performance using intent during turbulence was significantly better than performance using the hand (D).
+Significance among comparisons is denoted by an asterisk.
+Error did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
+\label{error}
+\end{figure}
 
-\processtable{}
-{\begin{tabular}{lllll}\toprule
-Image Type & Description & Format & Color Mode & Resolution\\\midrule
-Line Art & An image composed of lines and text,  & TIFF, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
-           & which does not contain tonal or shaded areas.& & &\\
-           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
-Combination & Image contains halftone + text or line art elements. & TIFF, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
-\end{tabular}}{}
-\end{table}
+\subsection*{Reaching Accuracy}
+Even if subjects were capable of performing a reaching task using their estimated intent, a poor estimate would have lead to poor performance.
+Indeed, the hand significantly outperformed our intent estimate when the force disturbance is not present in block 1 (Figure 2D).
+Model and sensor inaccuracies degrade performance and result in an error of approximately 5 millimetres.
+In the presence of a force disturbance in block 2, the hand loses its advantage as model and sensor inaccuracies are balanced by some degree of cancellation of the disturbance.
+In block 3 where visual feedback of the intent is given instead of hand, the intent significantly outperforms the hand.
+For this intent estimate to be useful, subjects should perform better when using feedback of the intent than when using feedback of the hand.
+Comparing the hand's performance in block 2 to the intent's performance in block 3, subjects performed significantly better using their estimated intent.
+Finally, no significant differences were present across the two experiments.
+Performance was relatively insensitive to this change in the stiffness model used to deduce the intent.
 
-\begin{equation}
-\sum x+ y =Z\label{eq:01}
-\end{equation}
+\subsection*{Endpoint Stiffness}
+While accuracy is desirable, it can be attained by cocontracting muscles to increase arm stiffness at the cost of metabolic effort \cite{gribble2003role}\cite{takahashi2001impedance}.
+Subjects stiffened significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3C).
+Subjects' stiffness significantly decreased when feedback of the deduced intent replaced feedback of the hand, as seen by comparing blocks 2 and 3.
+This decrease does not return stiffness to undisturbed levels, and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
+Stiffness may have decreased with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
+Subjects did not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
 
-\textbf{Table\ref{Tab:02}} shows the resolution requirements for the figures. The figures must be legible:
-\begin{enumerate}
-\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
-\item Solid lines are not broken up.
-\item Image areas are not pixelated or stair stepped.
-\item Text is legible and of high quality.
-\item Any lines in the graphic are no smaller than 2 points width.
-\item The actual size of the figure must be of at least 8.5 cm.
-\end{enumerate}
+\begin{figure}[t]
+\centering
+\framebox{\parbox{3in}{\includegraphics{stiffness.eps}
+}}
+\caption{Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
+Arm stiffness calculated by linear regression during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance increased arm stiffness (B).
+Comparisons between treatment conditions revealed that exposure to turbulent forces caused by arm to stiffen significantly, but feedback of intent could significantly reduce this increase (C).
+Stiffness did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
+\label{stiffness}
+\end{figure}
 
 \section{Discussion}
+While other visual distortions typically degrade performance and require an adaptation period to overcome, IF immediately enhances performance.
+The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
-Additional Requirements:
-\subsection{Corrections}
+This is scientifically relevant as performance enhancement is an active of research.
+In addition, this finding supports the notion that a particular trajectory is calculated even when explict trajectory constraints are not present.
 
-If you need to communicate important changes to a published article please submit a General Commentary. Submit the article with the title “Corrigendum: Original Title of Article”.
+Stiffness decreased, which provides secondary evidence of lower instability and unpredictability.
+Well, maybe it's just because of the numbers we chose.
+If we crank them down, can we get the stiffness even lower?
+Not so much.
+So clearly getting the stiffness right is very important.
 
-\subsection{Commentaries on Articles}
+While this is pretty awesome, getting the stiffness right is an enormous limitation.
+Moreover, the stiffness model breaks down in the vicinity of the target for reasons unknown.
+The nature of control might be changing there and if so, we don't really know the specifics, but Shapiro has something to say about this.
 
-At the beginning of your manuscript provide the citation of the article commented on.
-
-\subsection{Focused Reviews}
-
-For Tier 2 invited Focused Reviews the sections Introduction, Material and Methods, Results, and Discussion are recommended.
-In addition the authors must submit a short biography of the corresponding author(s).
-This short biography has a maximum of 600 characters, including spaces.
-
-A picture (5 x 5 cm, in *.tif or *.jpg, min 300 dpi) must be submitted along with the biography in the manuscript and separately during figure upload.
-Focused Reviews highlight and explain key concepts of your work.
-Please highlight a minimum of four and a maximum of ten key concepts in bold in your manuscript and provide the definitions/explanations at the end of your manuscript under “Key Concepts”.
-Each definition has a maximum of 400 characters, including spaces.
-
-\subsection{Human Search and Animal Research}
-
-All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations.
-In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards.
-For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects.
-In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.
-
-\subsection{Clinical Trial Registration}
-
-Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers.
-Trials must be registered at or before the start of patient enrollment.
-A clinical trial is defined as"any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\url{www.who.int/ictrp/en}).
-A list of acceptable registries can be found at \url{www.who.int/ictrp/en and www.icmje.org}.
-
-\subsection{Inclusion of Proteomics Data}
-
-Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches.
-Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.
-
-Authors should provide as supplementary material information used to identify proteins and/or peptides.
-This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
-For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.
-
-For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
-Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.
-
-\subsection{Data Sharing}
-
-Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research.
-In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.
+Overall, this is something of a breakthrough
 

### 2015-05-26 Paper2frontiers/frontiers.tex
updates from work

@@
 While performance is the best indicator of IF's value, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want.
-Arm stiffness is known to increase during exposure to instability \cite{franklin2003adaptation} and uncertainty \cite{takahashi2001impedance}.
-We anticipated that these change might also be modulated by presentation of IF.
+Arm stiffness is known to increase during exposure to instability \citep{franklin2003adaptation} and uncertainty \citep{takahashi2001impedance}.
+We anticipated that these changes might also be modulated by presentation of IF.
 Reducing the effects of environmental instability and presenting the subject with a signal already known to them should make the arm relax and become more compliant.
@@
 First, $\hat{E}$ must be modeled or assumed, so we choose $\hat{E}=0$.
-In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures\cite{scheidt2001learning}.
+In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures\citep{scheidt2001learning}.
 Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
-Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{gomi1998task} and may vary over the course of a reach\cite{niu2010temporal}.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\citep{gomi1998task} and may vary over the course of a reach\citep{niu2010temporal}.
 With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
@@
 \subsection{Apparatus}
-A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass.
-The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
+A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to minimize any friction or mass.
+The MATLAB XPC-TARGET package \citep{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
 Visual feedback of cursor position was performed at 60 Hz using OpenGL.
 Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
+Numerical simulation was performed in real-time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\cite{gough2009gnu}.
 
@@
 Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} were used to estimate relationships between body mass and limb mass, limb length and limb center of mass, and limb mass and length and moment of inertia.
-Viscosity parameters were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
+Viscosity parameters, $K_d$, were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
 Stiffness parameters were either taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} ($K_{P1}$, experiment 1) or estimated ($K_{P2}$, experiment 2).
@@
 \end{equation}
-Numerical simulation was performed in real time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\cite{gough2009gnu}.
+To estimate this reduced stiffness, a pilot subject was asked to intend to remain still on the center target, $q_d$, while co-contracting as little as possible.
+$K_{P2}$ was then calculated from one minute of white noise forces, $E$, and joint angle traces, $q$ as the least squares solution to the system:
+\begin{equation}
+K_{P2}\big(q_d(t)-q(t)\big)=M\big(q(t)\big)\ddot{q}(t)+G(q(t),\dot{q}(t))+E(t)+K_D\dot{q}(t)
+\end{equation}
+where $K_{P2}$ is a 2-by-2 matrix while the state difference and torque are 2-by-60000 matrices.
 
@@
 Mass ($m$), viscosity ($b$), and stiffness offset ($F_0$) terms were calculated, but discarded.
-While stiffness is better described as a matrix, a scalar was calculated to facilitate statistical comparison between treatments and blocks.
+While joint stiffness is usually described as a matrix, instantaneous endpoint stiffness in only the error direction is a scalar.
+This metric isolates stiffness in the error direction and facilitates statistical comparison between treatments and blocks.
 The paired Wilcoxon signed rank test was used to detect differences in maximum perpendicular error and stiffness between blocks and treatments at the 5\% significance level using the MATLAB statistics toolbox package \cite{MATLAB:2014}.
@@
 
-\begin{figure}[!b]
+\begin{figure}[h]
 \centering
@@
 }}
-\caption{Typical subjects' made center-out targeted reaching motions under experimentally varied force and feedback conditions.
+\caption{Typical subjects made center-out targeted reaching motions under experimentally varied force and feedback conditions.
 Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
-While both were recorded under all conditions, only one was shown on screen at a time.
 Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
-The force disturbance was designed to be turbulent and unpredictable to minimize any effect of learning.}
+The white noise force disturbance was designed to be unpredictable in order to minimize any effect of learning.}
 \label{Yplots}
@@
 \subsection*{Endpoint Stiffness}
-While accuracy is desirable, it can be attained by cocontracting muscles to increase arm stiffness at the cost of metabolic effort \cite{gribble2003role}\cite{takahashi2001impedance}.
+While accuracy is desirable, it can be attained by co-contracting muscles to increase arm stiffness at the cost of metabolic effort \citep{gribble2003role, takahashi2001impedance}.
 Subjects stiffened significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3C).
@@
 This decrease does not return stiffness to undisturbed levels, and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
-Stiffness may have decreased with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
+Stiffness may have decreased with duration of exposure to uncertain disturbances\citep{takahashi2001impedance}.
 Subjects did not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
@@
 \section{Discussion}
+
+Although exposure to random forces damaged subjects’ reaching accuracy and increased their arm stiffness, replacing the veridical feedback with IF alleviated accuracy and stiffness.
 While other visual distortions typically degrade performance and require an adaptation period to overcome, IF immediately enhances performance.
 The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
+In addition to its promise in performance enhancement, IF represents a novel means of revealing and studying the mechanisms of planning and control.
 
-This is scientifically relevant as performance enhancement is an active of research.
-In addition, this finding supports the notion that a particular trajectory is calculated even when explict trajectory constraints are not present.
+While IF alleviated the increased stiffness caused by exposure to random forces, stiffness remained significantly above baseline levels.
+Many explanations are reasonable.
+In particular, we might have hypothesized that the stiffness model used to determine intent might bias a person's arm stiffness, but the data did not support this conclusion.
+Alternatively, we might hypothesize that inaccuracy and incompleteness of the models used results in an estimate of intent that less accurate than subjects might desire.
+Since noise and performance inaccuracy can both be addressed by co-contraction, this may account for the remaining additional stiffness.
+Finally, it is possible that subjects co-contracted simply to counteract any residual perceived instability due to our slight lag in processing and presenting feedback.
+
+Specific planned state trajectory even in the absence of explicit kinematic constraints.
+
+Of particular scientific interest, "strict invariant" which has been described as opposite optimal feedback control.
 
@@
 So clearly getting the stiffness right is very important.
+In the absence of disturbing forces, estimated intent accuracy was lower than
+that of the hand, which suggests the need for model improvements.
+
+ The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
+If so, IF r Moreover, this immediate performance enhancement during human-machine interaction implies a means by which robots can productively interact with a human without impeding them.
 

### 2015-05-27 Paper1rev3/paper1rev3.tex
updates from work
slightly better

--- 2015-04-22 Paper1rev2/paper1rev2.tex
@@
 By scaling an established model using a force disturbance other than our disturbance of interest, we avoided presuming the intent trajectory we hoped to recover.
+This also serves as a test of model generalization since the training and testing data are drawn from different force disturbances.
+If the structure of the model is sufficiently accurate, the value of $c$ fit using data from the white noise forces will also be effective, at least initially, for the pulse force disturbances.
 

### 2015-05-27 Paper2frontiers/frontiers.tex
getting closer

@@
 
-Specific planned state trajectory even in the absence of explicit kinematic constraints.
+The ease with which subjects could make use of their estimated intent provides strong preliminary evidence that a specific desired state trajectory is computed even when moving in a highly variable environment.
+While the nature of the force disturbance and its high trial-to-trial variability prevent clear characterization of learning, subjects appear to have learned to use their intent within the first exposure.
+While recent work has identified kinematic constraints irrelevant to the task \citep{mistry2013optimal}, this is the first direct evidence that the entire trajectory is controlled even in the absence of specific instructions or constraints.
+It remains unclear when the intended trajectory is computed, but variability in the desired trajectory in phases 2 and 4 suggests that the trajectory is planed or at least can be replanned mid-movement.
+Certainly, it is not finalized before the onset of movement and is not a "strict invariant." Intent determination provides us with a novel and useful window into the nature of motor control.
 
-Of particular scientific interest, "strict invariant" which has been described as opposite optimal feedback control.
+IF can facilitate human-machine collaboration and artificial performance augmentation by enabling the machine to preserve an operators intent while cancelling unexpected disturbanes from the environment.
+This would reduce the demands on the human operator increase performance, especially in environments with rapidly changing conditions that can be quickly and accurately measured with electronic instrumentation.
+This assistance goes beyond environment cancellation by also accounting for any corrections the operator might make based on their expectation of disturbance.
 
-Stiffness decreased, which provides secondary evidence of lower instability and unpredictability.
-Well, maybe it's just because of the numbers we chose.
-If we crank them down, can we get the stiffness even lower?
-Not so much.
-So clearly getting the stiffness right is very important.
-In the absence of disturbing forces, estimated intent accuracy was lower than
-that of the hand, which suggests the need for model improvements.
+Limitatinos. The stiffness model is shit-tastic. Reaching trajectories, but not tracking, nonlinear, or 3D.
 
- The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
-If so, IF r Moreover, this immediate performance enhancement during human-machine interaction implies a means by which robots can productively interact with a human without impeding them.
-
-While this is pretty awesome, getting the stiffness right is an enormous limitation.
-Moreover, the stiffness model breaks down in the vicinity of the target for reasons unknown.
-The nature of control might be changing there and if so, we don't really know the specifics, but Shapiro has something to say about this.
-
-Overall, this is something of a breakthrough
+The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.
+Sensitivity analysis for this approach applied to the human arm demonstrated favorable conditions for determination of intent during force-disturbed reaching.
+Reaching intent became significantly different from baseline reaching only after enough time had elapsed for the disturbance to be processed and descending motor commands to change.
+This window into intent should allow advances in arm modeling, motor training, and human-machine interaction.
 

### 2015-05-29 Paper2frontiers/frontiers.tex
updated with tejas

--- 2015-05-26 Paper2frontiers/frontiers.tex
@@
 \def\firstAuthorLast{Horowitz {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Justin Horowitz\,$^1$, Co-Author\,$^{2}$ and Co-Author\,$^{2,*}$}
+\def\Authors{Justin Horowitz\,$^1$, Tejas Madhavan\,$^1$, Christine Massie\,$^1$ and James Patton\,$^{1,2,*}$}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
 % If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
-$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
+\def\Address{$^{1}$Neuro-Machine Interaction Laboratory, University of Illinois at Chicago, Bionengineering, Chicago, Illinois, United States of America \\
+$^{2}$Robotics Lab, Rehabilitation Institute of Chicago, Sensory Motor Performance Program, Chicago, Illinois, United States of America }
 % The Corresponding Author should be marked with an asterisk
@@
 \def\corrAuthor{James Patton}
-\def\corrAddress{Laboratory X, Institute X, Department X, Organization X, Street X, City X , State XX (only USA, Canada and Australia), Zip Code, X Country X}
+\def\corrAddress{Neuro-Machine Interaction Laboratory, University of Illinois at Chicago, Bionengineering, 851 S. Morgan Street, Chicago, Illinois, 60607-7052, United States of America}
 \def\corrEmail{pattonj@uic.edu}
@@
 Preplanning a specific route may not necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal.
-If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended route using optimal feedback control\citep{todorov2002optimal}.
+If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended route using optimal feedback control \citep{todorov2002optimal}.
 While this modeling strategy has been very successful at explaining data, it fundamentally assumes that corrective actions will be taken in response to relevant disturbances.
@@
 \subsection{Intent Extraction}
-The well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} relates arm trajectory, $q$, to desired arm trajectory, $q_d$, and any external disturbance, $E$ using physical parameters of the arm.
+The well-known motion control structure of \cite{shadmehr1994adaptive} relates arm trajectory, $q$, to desired arm trajectory, $q_d$, and any external disturbance, $E$ using physical parameters of the arm.
 To show how this model can be algebraically inverted to instead describe desired arm trajectory as a function of arm trajectory and external disturbance, we write it as a torque balance:
@@
 First, $\hat{E}$ must be modeled or assumed, so we choose $\hat{E}=0$.
-In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures\citep{scheidt2001learning}.
+In the presence of a zero mean white noise force disturbance its mean should be zero, but it is unlikely to be exactly zero and may reflect an average of only the last few exposures \citep{scheidt2001learning}.
 Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
-Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\citep{gomi1998task} and may vary over the course of a reach\citep{niu2010temporal}.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent \citep{gomi1998task} and may vary over the course of a reach \citep{niu2010temporal}.
 With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
@@
 Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
-Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
+Intent was estimated using either the stiffness model of \cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
 The white noise force disturbance was designed to be unpredictable in order to minimize any effect of learning.}
@@
 Significance among comparisons is denoted by an asterisk.
-Error did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
+Error did not signficantly depend on the choice between stiffness model of \cite{shadmehr1994adaptive} (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
 \label{error}
@@
 \subsection*{Reaching Accuracy}
-Even if subjects were capable of performing a reaching task using their estimated intent, a poor estimate would have lead to poor performance.
+Even if subjects were capable of using IF to perform a reaching task, a poor estimate would have lead to poor performance.
 Indeed, the hand significantly outperformed our intent estimate when the force disturbance is not present in block 1 (Figure 2D).
@@
 Subjects stiffened significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3C).
-Subjects' stiffness significantly decreased when feedback of the deduced intent replaced feedback of the hand, as seen by comparing blocks 2 and 3.
+Subjects' stiffness significantly decreased when using IF replaced the hand position as their cursor, as seen by comparing blocks 2 and 3.
 This decrease does not return stiffness to undisturbed levels, and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
-Stiffness may have decreased with duration of exposure to uncertain disturbances\citep{takahashi2001impedance}.
+Stiffness may have decreased with duration of exposure to uncertain disturbances \citep{takahashi2001impedance}.
 Subjects did not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
-
-\begin{figure}[t]
-\centering
-\framebox{\parbox{3in}{\includegraphics{stiffness.eps}
-}}
-\caption{Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
-Arm stiffness calculated by linear regression during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance increased arm stiffness (B).
-Comparisons between treatment conditions revealed that exposure to turbulent forces caused by arm to stiffen significantly, but feedback of intent could significantly reduce this increase (C).
-Stiffness did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
-\label{stiffness}
-\end{figure}
 
@@
 
-Although exposure to random forces damaged subjects’ reaching accuracy and increased their arm stiffness, replacing the veridical feedback with IF alleviated accuracy and stiffness.
-While other visual distortions typically degrade performance and require an adaptation period to overcome, IF immediately enhances performance.
-The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
-In addition to its promise in performance enhancement, IF represents a novel means of revealing and studying the mechanisms of planning and control.
+Although exposure to random forces hindered subjects’ reaching accuracy and increased their arm stiffness, replacing the veridical feedback with IF improved accuracy and decreased stiffness.
+While other visual distortions typically degrade performance and require an adaptation period to overcome, IF immediately enhanced performance.
+The performance variability inherent in white noise force disturbances complicates observation of learning, but there were no obvious differences in performance between the first and final exposures to IF.The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
+In addition to its promise in performance enhancement, IF represents a novel means of revealing and studying the mechanisms of motor planning and motor control.
 
@@
 Many explanations are reasonable.
-In particular, we might have hypothesized that the stiffness model used to determine intent might bias a person's arm stiffness, but the data did not support this conclusion.
-Alternatively, we might hypothesize that inaccuracy and incompleteness of the models used results in an estimate of intent that less accurate than subjects might desire.
-Since noise and performance inaccuracy can both be addressed by co-contraction, this may account for the remaining additional stiffness.
-Finally, it is possible that subjects co-contracted simply to counteract any residual perceived instability due to our slight lag in processing and presenting feedback.
+In particular, we might have hypothesized that subjects would adapt their own arm stiffness to decrease conflict with the stiffness model used to estimate their intent and thereby increase the accuracy of the estimate, but the data did not support this conclusion.
+Alternatively, inaccuracy and incompleteness of the models used might have resulted in an estimate of intent less accurate than subjects might desired.
+Since noise and performance inaccuracy can both be addressed by co-contraction, this may account for the residual stiffness.
+Finally, as subjects were not cued regarding the onset or removal of IF, the residual co-contraction may have been a precaution against the resumption of veridical feedback.
 
-Specific planned state trajectory even in the absence of explicit kinematic constraints.
+The ease with which subjects could make use of their estimated intent provides strong preliminary evidence that a specific intended trajectory is computed for the hand even when reaching in a highly variable environment.
+While recent work has identified kinematic constraints unnecessary for a task \citep{mistry2013optimal}, this is the first direct evidence that the entire trajectory is controlled even in the absence of specific instructions or constraints.
+It remains unclear when the intended trajectory is computed, but variability in the desired trajectory in phases 2 and 4 suggests that the trajectory is planned or at least can be replanned mid-movement.
+Certainly, it is not finalized before the onset of movement and is not a "strict invariant."
 
-Of particular scientific interest, "strict invariant" which has been described as opposite optimal feedback control.
+IF can facilitate human-machine collaboration and artificial performance augmentation by enabling the machine to preserve an operator's intent while cancelling unexpected disturbances from the environment.
+This would reduce the demand on the human operator and increase performance -- especially in environments with rapidly changing conditions.
+This assistance goes beyond environment cancellation by also accounting for any errors the operator might make based on their expectation of disturbance.
 
-Stiffness decreased, which provides secondary evidence of lower instability and unpredictability.
-Well, maybe it's just because of the numbers we chose.
-If we crank them down, can we get the stiffness even lower?
-Not so much.
-So clearly getting the stiffness right is very important.
-In the absence of disturbing forces, estimated intent accuracy was lower than
-that of the hand, which suggests the need for model improvements.
+While IF holds tremendous promise, it also has strong limitations.
+IF is entirely dependent on the accuracy of the models used.
+While we were able to leverage measurements of cadaver anthropometrics, average tendencies do not capture individual variability.
+Similarly, the stiffness model of \cite{shadmehr1994adaptive} appears to have accurately captured the mean tendency (figure 3, bottom left) without accounting for variation among individuals or variation over time.
+Techniques that could estimate stiffness trajectories, perhaps even in real-time, would greatly increase the accuracy and utility of IF.
 
- The simplest explanation for this is that IF approximates a signal already known to the brain: the path planned for the hand.
-If so, IF r Moreover, this immediate performance enhancement during human-machine interaction implies a means by which robots can productively interact with a human without impeding them.
-
-While this is pretty awesome, getting the stiffness right is an enormous limitation.
-Moreover, the stiffness model breaks down in the vicinity of the target for reasons unknown.
-The nature of control might be changing there and if so, we don't really know the specifics, but Shapiro has something to say about this.
-
-Overall, this is something of a breakthrough
+IF benefited subjects moving in a changing, uncertain environment.
+In addition to increasing subjects' accuracy, IF may have allowed subjects to reach their goals with less effort as arm stiffness decreased.
+While IF is limited by the accuracy of the models used, many candidate models are available and may outperform the simple model we investigated here.
+IF provides a novel form of feedback that may facilitate new insights into the nature of motor control and allows a machine to collaborate more effectively with a human user.
 
@@
 
-The statement about the authors and contributors can be up to several sentences long, describing the tasks of individual authors referred to by their initials and should be included at the end of the manuscript before the References section.
+All authors designed the work, analyzed the data, drafted and revised the manuscript, approved the final version, and agree to be held accountable for all aspects. TM and CM collected the data.
 
 \section*{Acknowledgments}
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
+The authors wish to thank Yazan Abdel Majeed for his review of the preliminary manuscript.
+We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
 
-\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.
+\paragraph{Funding\textcolon} Funded By NIH R01-NS053606 and NIDRR H133E120010.
 
-\section*{Supplemental Data}
-Supplementary Material should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
-
-%\bibliographystyle{frontiersinSCNS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
+\bibliographystyle{frontiersinSCNS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
 \bibliography{paper2.bib}
@@
 
+\begin{figure}[t]
+\centering
+\framebox{\parbox{3in}{\includegraphics{stiffness.eps}
+}}
+\caption{Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
+Arm stiffness calculated by linear regression during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance increased arm stiffness (B).
+Comparisons between treatment conditions revealed that exposure to turbulent forces caused by arm to stiffen significantly, but feedback of intent could significantly reduce this increase (C).
+Stiffness did not signficantly depend on the choice between the stiffness model of \cite{shadmehr1994adaptive} (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
+\label{stiffness}
+\end{figure}
+
 %%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.

### 2015-06-01 Paper2frontiers/frontiers.tex
Updated with interns

@@
 Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning.
-Sensory delays, sensory fusion, delays in sensory processing, and delays in action further confound our ability to behave reactively.
+Sensory delays, sensory fusion, delays in sensory processing, and motor delays further confound our reactions.
 The nervous system intelligently solves many problems by planning ahead \citep{belen1967control}, yet in the face of uncertainties we often cannot adequately prevent errors.
-There is the possibility, however, to exploit additional information from instruments that may be available -- particularly fast and accurate force sensors -- that can measure human machine interactions.
-Combining sensors with filtering techniques makes it possible to determine a person’s underlying intent -- the motion they would have done if they had not been disturbed.
+There is the possibility, however, to exploit additional information from instruments -- particularly fast and accurate force sensors -- that can measure human machine interactions.
+Combining sensors with filtering techniques makes it possible to determine a person’s underlying intent -- the motion they would have done had they not been disturbed.
 This intent provides new ways to understand the nature of control and provide novel feedback.
 
-Recent work in our laboratory has attempted to outline the methodology for obtaining estimates of intent \citep{horowitz2015determining}.
-This method assumes a model of the dynamics and control of the human arm, and then through a manipulation of the equations of motion it integrates to find a unique estimate of intent.
+Recent work in our laboratory has attempted to outline a methodology for obtaining estimates of intent \citep{horowitz2015determining}.
+This method assumes a model of the dynamics and control of the human arm.
+Following manipulations of the equations of motion, the method integrates to find a unique estimate of intent.
 The algorithm recovers the trajectory a person intended to take, even if they were forced away from it due to environmental disturbances.
-This analysis has enabled us to show how subjects might sometimes alter their intent following exposure to unexpected force pulses.
+This analysis has enabled us to show how some subjects sometimes alter their intent following exposure to unexpected force pulses.
 
-A new question that arises is whether seeing one’s own intent, rather that what what actually happens, may be useful.
-The intent extraction method can be abriged to allow for real-time estimations and presented to the subject as a cursor.
-It is possible that performance using this estimated intention may compare quite favorably to the actual outcome in the presence of unexpected disturbances.
+A new question that arises is whether seeing one’s own intent, rather than what actually happens, may be useful.
+The intent extraction method can be streamlined to allow for real-time estimations of intent that can be presented to the subject as a cursor.
+Estimated intention may outperform the movement accuracy in the presence of unexpected disturbances.
 If so, such a method holds great promise in any situation where humans and machines interact as it enables the machine to give the human operator what they want.
-This human-machine collaboration might be able to outperform what a person can do alone.
+This human-machine collaboration could outperform what a person can do alone.
 
-In contrast, displaying anything other than what truthfully happens is a distortion and a lie.
+Displaying anything other than what truthfully happens is a distortion and a deceit.
 Like many other visual distortion experiments \citep{miles1980long, pine1996learning}, intent feedback (IF) introduces a visuomotor discrepancy that may be confusing to the nervous system.
-Preplanning a specific route may not necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal.
-If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended route using optimal feedback control \citep{todorov2002optimal}.
+Preplanning a specific route may not be necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal.
+If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment \citep{todorov2002optimal}.
+No specific intended route is needed when using optimal feedback control.
 While this modeling strategy has been very successful at explaining data, it fundamentally assumes that corrective actions will be taken in response to relevant disturbances.
 Goals can also be reached at minimal \textit{expected} cost by constructing -- and possibly updating -- a specific intended route.
-IF may be troublesome if the nervous system cannot recognize it because no particular trajectory was intended.
+If no particular trajectory is intended, the nervous system could be unable to recognize IF.
 
-While performance is the best indicator of IF's value, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want.
+While performance is the best indicator of IF's worth, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want.
 Arm stiffness is known to increase during exposure to instability \citep{franklin2003adaptation} and uncertainty \citep{takahashi2001impedance}.
-We anticipated that these changes might also be modulated by presentation of IF.
-Reducing the effects of environmental instability and presenting the subject with a signal already known to them should make the arm relax and become more compliant.
-We hypothesized that any elevated arm stiffness from noisy disturbances would decrease when subjects received IF.
+We anticipated that these changes might also be modulated by the presence of IF.
+Reducing the effects of environmental instability and presenting the subject with a signal already known to them should relax their arm and make it more compliant.
+We hypothesized that any elevated arm stiffness from noisy disturbances would decrease while subjects received IF.
 
-In this paper we describe our abridged method for real-time IF, and then expose subjects to IF in an unstable and unpredictable environment to compare their ability to perform goal-directed reaching using visual feedback of their hand position against visual feedback of their estimated intended position.
+In this paper we describe our streamlined method for real-time IF and then we expose subjects to IF in an unstable and unpredictable environment.
+We then compare their ability to perform goal-directed reaching using visual feedback of their hand position against IF.
 We hypothesized that IF should lead to better performance in the presence of force-based disturbances.
-Accordingly, we hypothesized that during random disturbances, the intent trajectory should deviate less than the hand.
+Accordingly, we hypothesized that during random disturbances, the intent trajectory should deviate less than the hand trajectory.
 

### 2015-06-02 Paper1rev3/paper1rev3.tex
tiny update

--- 2015-04-22 Paper1rev2/paper1rev2.tex
@@
 By scaling an established model using a force disturbance other than our disturbance of interest, we avoided presuming the intent trajectory we hoped to recover.
+This also serves as a test of model generalization since the training and testing data are drawn from different force disturbances.
+If the structure of the model is not sufficiently accurate, the value of $c$ fit using data from the white noise forces will be ineffective for the pulse force disturbances.
 

### 2015-06-03 Paper2frontiers/frontiers.tex
mild updates + fixed figure format

@@
 Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent \citep{gomi1998task} and may vary over the course of a reach \citep{niu2010temporal}.
-With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
+With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of \cite{shadmehr1994adaptive} anticipating that it is sufficiently accurate or easy to learn (experiment 1).
 The experiment was repeated with a lower stiffness estimate (experiment 2) to explore any dependence on this assumption.
@@
 \subsection{Apparatus}
-A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to minimize any friction or mass.
+A planar manipulandum (described in \cite{patton2004robot}) was programmed to minimize any friction or mass.
 The MATLAB XPC-TARGET package \citep{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.
@@
 Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.
-Numerical simulation was performed in real-time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\cite{gough2009gnu}.
+Numerical simulation was performed in real-time using the GNU Scientific Library's odeiv2 driver with Runge-Kutta-Fehlberg (4,5) stepping\citep{gough2009gnu}.
 
 \subsection{Human Subjects}
-The human data trajectories analyzed here are drawn from nine subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
-Seven male and one female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
+The human data trajectories analyzed here are drawn from sixteen subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
+fourteen male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
 Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
@@
 \subsection{Experimental Design}
-We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at 60 degree intervals about that central point.
-Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of 5 blocks of 96 reaches each.
+We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at $60^\circ$ intervals about that central point.
+Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of five blocks of 96 reaches each.
 During blocks 2 through 4, subjects experienced filtered white noise forces drawn from a white noise generator at 1000 Hertz with flat power spectral density of 1 Newton, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ radians per second.
@@
 \subsection{Dynamic simulation of arm and intended trajectories}
-Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} were used to estimate relationships between body mass and limb mass, limb length and limb center of mass, and limb mass and length and moment of inertia.
+Anatomical landmarks and values from \cite{dempster1955space} and \cite{winter2009biomechanics} were used to estimate relationships between body mass and limb mass, limb length and limb center of mass, and limb mass and length and moment of inertia.
 Viscosity parameters, $K_d$, were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
-Stiffness parameters were either taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} ($K_{P1}$, experiment 1) or estimated ($K_{P2}$, experiment 2).
+Stiffness parameters were either taken from \cite{shadmehr1994adaptive} ($K_{P1}$, experiment 1) or estimated ($K_{P2}$, experiment 2).
 Expressed in Newton-Meters per Radian:
@@
 \subsection{Metrics and Statistical Analysis}
-Trajectories and forces were rotated such that movement and force parallel to the line connecting the previous target (the reach origin) and the presented target lie along a \textit{progress} axis while perpendicular movement and force lie along an \textit{error} axis.
+Trajectories and forces were rotated such that movement and force parallel to the line connecting the previous target (the reach origin) and the presented target were along a \textit{progress} axis while perpendicular movement and force were along an \textit{error} axis.
 Reach onset was detected as the moment the cursor's distance from the center of the previous target first exceeded 1 centimeter.
 Maximum perpendicular error for a trajectory was the largest error magnitude within 250 milliseconds of reach onset.
-A scalar stiffness, $k$, was calculated for the \textit{error} axis during this same 250 millisecond time span by linear regression:
+A scalar stiffness, $k$, was calculated for the \textit{error} axis during this same 250 milliseconds time span by linear regression:
 \begin{equation}
@@
 The following sections address the quality of this reaching.
-
-\begin{figure}[h]
-\centering
-\framebox{\parbox{3in}{\includegraphics{Yplots.eps}
-}}
-\caption{Typical subjects made center-out targeted reaching motions under experimentally varied force and feedback conditions.
-Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
-Intent was estimated using either the stiffness model of \cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
-The white noise force disturbance was designed to be unpredictable in order to minimize any effect of learning.}
-\label{Yplots}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\framebox{\parbox{3in}{\includegraphics{error.eps}
-}}
-\caption{Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback (A).
-Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance degraded reaching performance in both hand (B) and intent (C).
-Comparisons within treatment conditions revealed a significant advantage for the hand when force disturbance is not present and a significant advantage for intent when feedback of intent was used during exposure to turbulent forces.
-Performance using intent during turbulence was significantly better than performance using the hand (D).
-Significance among comparisons is denoted by an asterisk.
-Error did not signficantly depend on the choice between stiffness model of \cite{shadmehr1994adaptive} (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
-\label{error}
-\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=10cm]{logo1}% This is a *.jpg file
+\includegraphics[width=10cm]{Yplots.eps}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
+\textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Typical subjects made center-out targeted reaching motions under experimentally varied force and feedback conditions.
+Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
+Intent was estimated using either the stiffness model of \cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
+The white noise force disturbance was designed to be unpredictable in order to minimize any effect of learning.
+}
 \end{figure}
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=10cm]{logo2}% This is an *.eps file
-%\end{center}
-%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
-%\end{figure}
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=10cm]{error.eps}
+\end{center}
+\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback (A).
+Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance degraded reaching performance in both hand (B) and intent (C).
+Comparisons within treatment conditions revealed a significant advantage for the hand when force disturbance is not present and a significant advantage for intent when feedback of intent was used during exposure to turbulent forces.
+Performance using intent during turbulence was significantly better than performance using the hand (D).
+Significance among comparisons is denoted by an asterisk.
+Error did not signficantly depend on the choice between stiffness model of \cite{shadmehr1994adaptive} (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
+\end{figure}
 
-\begin{figure}[t]
-\centering
-\framebox{\parbox{3in}{\includegraphics{stiffness.eps}
-}}
-\caption{Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=10cm]{stiffness.eps}
+\end{center}
+\textbf{\refstepcounter{figure}\label{fig:03} Figure \arabic{figure}.}{ Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
 Arm stiffness calculated by linear regression during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance increased arm stiffness (B).
@@
 Stiffness did not signficantly depend on the choice between the stiffness model of \cite{shadmehr1994adaptive} (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
-\label{stiffness}
 \end{figure}

### 2015-06-08 Paper1rev3/paper1rev3.tex
Updated for submitted version

--- 2015-04-22 Paper1rev2/paper1rev2.tex
@@
 
+Since trajectories were not controlled experimentally, our tests rely heavily on the assumption that disturbed intentions initially mirror undisturbed intentions.
+Therefore, it is not completely certain that subjects intended to move along a straight line to the target.
+However, this study examined straight line motions because it is well understood that people tend to repetitively attempt to reach in a straight line towards a target \cite{hodgson2000model}.
+Also, even when subjects are persistently perturbed by a force field, they recover their motions to a straight line with repetitive experience \cite{shadmehr1994adaptive}.
+Nevertheless, it may be better to employ tasks with a more specific trajectory in order to validate the intended path.
+Unfortunately, even such an intended trajectory may not be strictly invariant during external disturbance because some displacements may cause intent to be recalculated by the nervous system.
+
 A key assumption in this study is that the joint impedance (stiffness) is linearly related to torque.
@@
 More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
-Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.
+Like others who have examined large force disturbances, we find lower stiffnesses (our $c$ parameter) than might be expected \cite{crevecoeur2014beyond} (Table in S1 Table).
+This may be due to differences in tasks \cite{gomi1998task} and/or the dependency of muscle impedance on perturbation size \cite{van2010vivo}.
+Our results also point to the possibility of the impedance changing during motion, perhaps as a function of time.
 
@@
 
-%\section*{Supporting Information Legends}
-%
-% Please enter your Supporting Information captions below in the following format:
-%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
-% 
-%\begin{description}
-%\item {\bf}
-%\item {\bf}
-%\end{description}
+\section*{Supporting Information Legends}
+
+\begin{description}
+\item{\bf Table S1.
+Subject-specific parameters.} Gross body mass was self-reported.
+Upper arm length, forearm length, and shoulder position in the robot's coordinate system were measured \textit{in situ}.
+Feedback torque gain was fit as described in our methods section using the intermittently presented white noise force disturbance.
+This gain is relative to the total feedback torque output (stiffness, damping, and reflexes) of the model of Burdet et al.\cite{burdet2006stability}.
+\end{description}
 

### 2015-06-08 Paper2frontiers/frontiers.tex
Updated for submitted version

@@
 The human data trajectories analyzed here are drawn from sixteen subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki.
-fourteen male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
+Fourteen male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated.
 Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.
@@
 \subsection{Experimental Design}
-We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at $60^\circ$ intervals about that central point.
+We chose center-out reaches of the right arm to visually-present targets 15 centimeters from the center chosen at $60^\circ$ intervals.
 Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of five blocks of 96 reaches each.

### 2015-06-10 Paper2frontiers/frontiers.tex
update with jim

@@
 \section{Results}
-As expected, the model was able to deduce an intended trajectory and all subjects were able to utilize this estimate of their intent to perform targeted reaching while experiencing turbulent forces (Figure 1).
+As expected, the model was able to deduce an intended trajectory and all subjects were able to use this estimate of their intent to perform targeted reaching while experiencing turbulent forces (Figure 1).
 The following sections address the quality of this reaching.
@@
 \subsection*{Reaching Accuracy}
-Even if subjects were capable of using IF to perform a reaching task, a poor estimate would have lead to poor performance.
+The results do indeed show that performance using IF feedback exceeds that of truthful feedback of the hand.
+Even if subjects were capable of using IF to perform a reaching task, a poor estimate would have led to poor performance.
 Indeed, the hand significantly outperformed our intent estimate when the force disturbance is not present in block 1 (Figure 2D).

### 2015-07-20 Paper1rev4/paper1rev4.tex
kk

--- 2015-06-08 Paper1rev3/paper1rev3.tex
@@
 
-It is important to distinguish intended action from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
-In typical motor control studies, subjects are motivated to complete an experiment in a timely fashion, and are usually explicitly provided targets to reach to.
-Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of discrete actions (walking/standing/flexion/extension/etc).
-While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves.
-Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.
-In other words, we operationally define \textit{intent} as a subjunctive -- where would the motion have gone had it not been disturbed?
+It is not necessary for the brain to represent intent explicitly to generate action.
+Instead the nervous system might simply learn the relationship between muscle activations and the accomplishment of goals.
+For example, one might simply evaluate a motion outcome and adjust descending signals to the muscles.
+At the same time, one may not deny the fact that there are tasks that require an explicit representation of a trajectory, such as performing a dance, drawing a picture, or conducting an orchestra.
+For these tasks, intent is likely represented at some level in the nervous system.
+Nevertheless, the mechanics of the body dictate the existence of an equilibrium whether intent is explicitly represented or not.
+Here our goal was to answer a modest question, ``\textit{when} movement intent is a changing equilibrium, can we recover it despite external disturbances?''
 
-Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles.
-Springs produce a force according to their impedance and stretch.
-By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
-Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this muscle equilibrium point, and not the equilibrium of the whole arm, represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would.
-Upon Gomi and Kawato's inspection of the muscle equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
-Therefore, it could not well-represent the intent of a simple reaching movement.
-Unlike the equilibrium point of muscle, the equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance.
-In the presence of disturbance, the arm might be deflected from its equilibrium.
-It is presently unknown if or how this arm equilibrium point might change due to disturbances.
+The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature (\nameref{S1_Text}).
+Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances.
+We recover this intent through a mathematical procedure that only requires that intent be \textit{stable}.
+Two intuitive conditions need to be met:
+\begin{enumerate}
+\item{Action will match intent in the absence of any unexpected disturbances.}
+\item{Action grows closer to intent over time following a sufficiently small disturbance.}
+\end{enumerate}
+When intent estimation fails, it will most often be due to the first condition.
+Consider a person attempting to lift Stonehenge's Heel Stone, which weighs approximately 35 tons.
+We could not determine how far the person intends to lift the stone as they are not strong enough to lift it at all.
+The second condition will seldom fail in practice as it amounts to an often-dangerous loss of control.
+For example, this condition fails temporarily while tripping or during a seizure.
+The rigidity of bone and elasticity of muscle usually ensure that this condition is met.
 
-Here we propose a revised and more general method, \textit{intent extraction}, which calculates the intended trajectory even in the face of disturbances by relating environmental interaction and state to intent.
+Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance.
+The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
+However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
+In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.
+
+Here we propose a general method, \textit{intent extraction}, which calculates the intended trajectory even in the face of disturbances by relating environmental interaction and state to intent.
 The method is a class of filters that can infer the intended trajectory from the disturbance (turbulence), process (dynamic equations), actuator (muscles), and motion.
@@
 \end{equation}
+
 where $x_e$ is the equilibrium trajectory of the actuator. Fourth, we describe a control law for the actuator using the same expansions,
@@
 
-The fact that the actuator's equilibrium $x_e$ is not the process's equilibrium is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}.
+The fact that the actuator's equilibrium $x_e$ is not the process's equilibrium is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin} (\nameref{S1_Text}).
 This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.
@@
 The impedance properties of the human arm are known to be task-dependent\cite{gomi1998task}, but we need to model them in the context of our task without assuming our conclusions in the premises.
-To accomplish this, we fit a linear scale factor to the torque output an established model using an intermittently, unpredictably presented secondary task.
-This unpredictability enables us to leverage the ergodic assumption: our tasks should at least initally share the same motor plan and impedance properties.
+We intermittently and unexpectedly presented a secondary task, which we used to fit a linear scale factor to the torque output of an established model.
+This addresses any concern that model parameters could be tuned to favor at a particular outcome.
+Additionally, unpredictability enables us to leverage the ergodic assumption: our tasks should at least initially share the same motor plan and impedance properties.
 Both portions of this assumption are important.
@@
 Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to mis-estimation of stiffness and changes in shoulder position from trial to trial.
-Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate, and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty.
+Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate, and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty(I found this sentence pretty long and hard to comprehend).
 Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.
@@
 
-The proposed approach could explain the discovery by Bizzi et al.
-\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement.
-By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to the virtual trajectory and can be used to determine the muscle equilibrium of a $\lambda$ model.
-
 Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibrium.
@@
 
-The process model does not need to be explicitly formulated.
-Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance.
-The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above.
-However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed.
-In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.
-
 Nevertheless, there are some limitations to these methods.
@@
 Since trajectories were not controlled experimentally, our tests rely heavily on the assumption that disturbed intentions initially mirror undisturbed intentions.
-Therefore, it is not completely certain that subjects intended to move along a straight line to the target.
-However, this study examined straight line motions because it is well understood that people tend to repetitively attempt to reach in a straight line towards a target \cite{hodgson2000model}.
+This study examined straight line motions because it is well understood that people tend to repetitively attempt to reach in a straight line towards a target \cite{hodgson2000model}.
 Also, even when subjects are persistently perturbed by a force field, they recover their motions to a straight line with repetitive experience \cite{shadmehr1994adaptive}.
-Nevertheless, it may be better to employ tasks with a more specific trajectory in order to validate the intended path.
+Future studies might employ tasks with a more specific and/or explicit trajectory to facilitate comparison with the estimate intent's path.
 Unfortunately, even such an intended trajectory may not be strictly invariant during external disturbance because some displacements may cause intent to be recalculated by the nervous system.
 
-A key assumption in this study is that the joint impedance (stiffness) is linearly related to torque.
+Another key assumption in this study is that the joint impedance (stiffness) is linearly related to torque.
 While other approaches tend to construct paradigms that indicate or presume that the intention is static in order to derive impedance -- the so called family of system identification applications, this study makes a constrained assumption of stiffness in order to determine intent.
@@
 More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result.
-Like others who have examined large force disturbances, we find lower stiffnesses (our $c$ parameter) than might be expected \cite{crevecoeur2014beyond} (Table in S1 Table).
-This may be due to differences in tasks \cite{gomi1998task} and/or the dependency of muscle impedance on perturbation size \cite{van2010vivo}.
+Like others who have examined large force disturbances, we find lower stiffnesses (our $c$ parameter) than might be expected \cite{crevecoeur2014beyond} (\nameref{S1_Table}).
+Differences in tasks studied \cite{gomi1998task} and/or the dependency of muscle impedance on perturbation size \cite{van2010vivo} may explain this stiffness reduction.
 Our results also point to the possibility of the impedance changing during motion, perhaps as a function of time.
-
-The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature.
-Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances.
-A rich history of research has suggested many ways that an equilibrium trajectory can explain action through feedback responses \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive}.
-Here we suggest that one should directly define intent in the context of a hypothesized control model.
-Our work here assumed the model of Burdet \cite{burdet2006stability}.
-However, other, more sophisticated, models may be deemed more appropriate in the future.
-What is clear is that given some model of control, this intent extraction approach is able to invert the process to estimate intent.
-
-It is important to mention that it is not necessary for the brain to represent intent explicitly to generate action.
-It is possible that the nervous system can simply learn the relationship between muscle activations and the accomplishment of goals.
-For example, one might simply evaluate a motion outcome and adjust descending signals to the muscles.
-At the same time, one may not deny the fact that there are tasks that require a explicit representation of a trajectory, such as performing a dance, drawing a picture, or conducting an orchestra.
-For these tasks, intent is likely represented at  some level in the nervous system.
-Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not.
-Here our goal was to answer a more modest question, ``\textit{when} movement intent is a changing equilibrium, can we recover it despite external disturbances?''
 
@@
 Research supported by US NIH NINDS1 R01 NS053606.
-We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, and Mohi Ahmed for their review of preliminary manuscript.
+We thank Felix Huang, Jagan Jimmy, Christine Massie, Tejas Madhavan, Mohi Ahmed, and Yazan Abdel Majeed for their review of preliminary manuscript.
 We also thank the community  of the Robotics Lab at the Rehabilitation Institute of Chicago for critical commentary during the genesis of this paper and the work leading up to it.
@@
 
-\section*{Supporting Information Legends}
+\section*{Supporting Information}
 
-\begin{description}
-\item{\bf Table S1.
-Subject-specific parameters.} Gross body mass was self-reported.
+\subsection*{S1 Table}
+\label{S1_Table}
+{\bf Subject-specific parameters.} Gross body mass was self-reported.
 Upper arm length, forearm length, and shoulder position in the robot's coordinate system were measured \textit{in situ}.
@@
 This gain is relative to the total feedback torque output (stiffness, damping, and reflexes) of the model of Burdet et al.\cite{burdet2006stability}.
-\end{description}
+
+\subsection*{S1 Text}
+\label{S1_Text}
+{\bf What intent is not.} It is important to distinguish intended action from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}.
+In typical motor control studies, subjects are motivated to complete an experiment in a timely fashion, and are usually explicitly provided targets to reach to.
+Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of discrete actions (walking/standing/flexion/extension/etc).
+While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves.
+Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.
+In other words, we operationally define \textit{intent} as a subjunctive -- where would the motion have gone had it not been disturbed?
+
+Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles.
+Springs produce a force according to their impedance and stretch.
+By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point.
+Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this muscle equilibrium point, and not the equilibrium of the whole arm, represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would.
+Upon Gomi and Kawato's inspection of the muscle equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable.
+Therefore, it could not well-represent the intent of a simple reaching movement.
+Unlike the equilibrium point of muscle, the a pre-planned equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance \cite{shadmehr1994adaptive}.
+In the presence of disturbance, the arm might be deflected from its equilibrium.
+It is presently unknown if or how this arm equilibrium point might change due to disturbances.
+
+The proposed approach could explain the discovery by Bizzi et al.
+\cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- intent and $\lambda$ are equivalent only when intent is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement.
+By deriving the technique in a general form, we discovered that our intent is equivalent to the virtual trajectory and can be used to determine the muscle equilibrium of a $\lambda$ model.
 

### 2015-07-20 matlab/armdynamicsInvertedBurdetReflexes.m

@@ -1,6 +1,6 @@
 function [dqi,torque_fb,torque_inertial]=armdynamicsInvertedBurdetReflexes(t,qi)
 
-global measuredVals measuredTime errorVals errorTime kpgain massgain
+global measuredVals measuredTime errorVals errorTime kpgain massgain reflexcontrib
 
 interped=twoNearestNeighbor(measuredVals,measuredTime,t);
 theta_real=interped(1:2)';
@@ -27,7 +27,8 @@ kp=kp0+kp1*diag(joint_torques);
 
 kp=kpgain*kp;
 
-torque_fb=kp*((theta_real-theta_desired) + (1/12)*(omega_real-omega_desired))+(kp/50)*(reflexE+2*reflexV);
+%should be (kp/#50#)*(reflexE+2*reflexV)
+torque_fb=(1-reflexcontrib)*kp*((theta_real-theta_desired) + (1/12)*(omega_real-omega_desired))+(reflexcontrib*kp)*(reflexE+2*reflexV);
 
 
 % Update the change in desired state

### 2015-07-20 matlab/figure3merged.m

@@ -2,7 +2,7 @@ clc
 clear all
 
 figure(37)
-set(gcf,'position',[8.8268    3.2713   26.8753   20.0790])
+%set(gcf,'position',[8.8268    3.2713   26.8753   20.0790])
 clf
 hold on
 
@@ -24,7 +24,8 @@ black=[0 0 0];
 
 yoff=[2 4; 1 3]*.1;
 xbase=[0 -.4];
-leftright=[6 8];
+%leftright=[6 8];
+leftright=[8 4];
 
 for LR=1:2
     specialestsub=leftright(LR);
@@ -194,7 +195,8 @@ rangemids=tnorm*plotMids;
 
 %plot(onset+[0 0],lscale*mP*[-1 1]+yoff,'color',[1 0 0],'linewidth',1.5)
 %plot(handonset+[0 0],lscale*mP*[-1 1]+yoff,'color',[0 1 0],'linewidth',1.5)
-lineorder={[6 1:4],[8 5 7]};
+%lineorder={[6 1:4],[8 5 7]};
+lineorder={[8 1:3 5:6],[4 7]};
 
 for LR=1:2
     specialestsub=leftright(LR);

### 2015-07-20 matlab/fullWorkFlow.m

@@ -2,7 +2,7 @@ clc
 clear all
 
 n=1:8;
-
+return
 %Add differential optimization just in case it's needed
 addpath ../../DeOpt/
 

### 2015-08-10 Paper1rev3/paper1rev2.tex
will this work?

--- 2015-06-02 Paper1rev3/paper1rev3.tex
@@
 By scaling an established model using a force disturbance other than our disturbance of interest, we avoided presuming the intent trajectory we hoped to recover.
-This also serves as a test of model generalization since the training and testing data are drawn from different force disturbances.
-If the structure of the model is not sufficiently accurate, the value of $c$ fit using data from the white noise forces will be ineffective for the pulse force disturbances.
 

### 2015-08-10 Paper 3/2Dfreeexploration/batch_boltzmann.m

@@ -2,7 +2,14 @@ clc
 clear all
 close all
 
-sublets={'JL','JT','LP','MF','YM'};
+sublets={'JL'} %,'JT','LP','MF','YM'};
+
+figure(4000)
+clf
+subplot(2,1,1)
+hold on
+subplot(2,1,2)
+hold on
 
 for k=1:length(sublets)
     fname=['free_exp_',sublets{k},'.mat'];
@@ -44,13 +51,19 @@ for k=1:length(sublets)
     
     figure(k)
     clf
-    outs(k)=boltzmannsubjectplots(t,x,v,a)
+    [outs(k),P,E]=boltzmannsubjectplots(t,x,v,a)
 
     suplabel(['Healthy Subject ',sublets{k}],'t')
-    %set(findall(gcf,'type','text'),'fontSize',12)
+    set(findall(gcf,'type','text'),'fontname','Times','fontSize',12)
     set(gcf,'position',[76 11 1195 925])
     print('-dtiff','-r300',['summary',sublets{k},'.tiff'])
     print('-dpng','-r300',['summary',sublets{k},'.png'])
+    
+    figure(4000)
+    subplot(2,1,1)
+    plot(P(:,2),P(:,1),'b')
+    subplot(2,1,2)
+    plot(E(:,2),E(:,1),'b')
 end
 
 save('healthy.mat','outs')

### 2015-08-10 Paper 3/2Dfreeexploration/boltzmannsubjectplots.m

@@ -1,4 +1,4 @@
-function ofinterest=boltzmannsubjectplots(t,x,v,a)
+function [ofinterest,Pnormecdf,Enormecdf]=boltzmannsubjectplots(t,x,v,a)
 
 global fitmecounts fitmebins
 
@@ -36,9 +36,11 @@ bar(centers,counts)
 fitmecounts=counts;
 fitmebins=edges;
 
-W=expfit(abs(Pt));
+W=fminbnd(@fitexpcdf,0,10);
 
 ofinterest.W=W;
+
+Pnormecdf=[counts centers/W];
 
 cdf=expcdf(fitmebins,W);
 fit=cdf(2:end)-cdf(1:end-1);
@@ -92,9 +94,7 @@ bar(centers,counts)
 
 fitmecounts=counts;
 fitmebins=edges;
-nT=gamfit(Tt);
-ofinterest.n=nT(1);
-ofinterest.T=nT(2);
+nT=fmincon(@fitgamcdf,[1.5; 1],-eye(2),-[1; 0.01]);
 
 cdf=gamcdf(fitmebins,nT(1),nT(2));
 fit=cdf(2:end)-cdf(1:end-1);
@@ -130,9 +130,18 @@ bar(centers,counts)
 fitmecounts=counts;
 fitmebins=.5*m*edges.^2;
 
+nT=fmincon(@fitgamcdf,[1.5; 1],-eye(2),-[1; 0.01]);
 cdf=gamcdf(fitmebins,nT(1),nT(2));
 fit=cdf(2:end)-cdf(1:end-1);
+ofinterest.n=nT(1);
+ofinterest.T=nT(2);
 ofinterest.speed_rmse=sqrt(mean((fit-counts).^2));
+
+Enormecdf=[counts centers/nT(2)];
+
+cdf=gamcdf(fitmebins,nT(1),nT(2));
+fit=cdf(2:end)-cdf(1:end-1);
+
 plot(centers,fit,'r','linewidth',3)
 
 title(['RMS Error = ',num2str(ofinterest.speed_rmse,3)])

### 2015-08-10 Paper 3/2Dfreeexploration/fitexpcdf.m

@@ -1,0 +1,7 @@
+function [cost,fit]=fitexpcdf(lambda)
+
+global fitmecounts fitmebins
+
+ecdf=expcdf(fitmebins,lambda);
+fit=ecdf(2:end)-ecdf(1:end-1);
+cost=sum((fitmecounts-fit).^2);

### 2015-08-10 Paper 3/2Dfreeexploration/gamma_example.m

@@ -1,0 +1,28 @@
+clc
+clear all
+
+figure(26)
+clf
+hold on
+plot(-1*[1 1],-1*[1 1],'b-')
+plot(-1*[1 1],-1*[1 1],'r-')
+
+E=0:.01:1;
+
+for n=[1 1.5 2 3]
+    PE=gampdf(E,n,.15);
+    plot(E,PE)
+    text(E(40),PE(40),['n=',num2str(n)],'horizontalalignment','left')
+end
+
+for n=[1 1.5 2 3]
+    PE=gampdf(E,n,.5);
+    plot(E,PE,'r')
+    text(E(20),PE(20),['n=',num2str(n)],'horizontalalignment','right')
+end
+
+xlim([0 1])
+ylim([0 3.5])
+xlabel('Energy, J')
+ylabel('Pr(Energy)')
+legend('T=.35','T=.7')

### 2015-08-10 Paper 3/2Dfreeexploration/smokinggun.m

@@ -1,12 +1,85 @@
 clc
 clear all
+close all
 
 load('free_exp_05stroke.mat')
+%load('free_exp_MF.mat')
 
 en=dot(v',v');
+figure(74)
+clf
+inds=1:20:size(x,1);
+plot3(x(inds,1),x(inds,2),en(inds),'.')
+
+mu=expfit(en);
 
 figure(75)
 clf
+<<<<<<< HEAD
 sx1=size(x,1);
 ds=1:150:sx1;
 plot3(x(ds,1),x(ds,2),en(ds),'.')
+=======
+
+lower=min(x);
+upper=max(x);
+range=upper-lower;
+nbins=64;
+selectors=[(x(:,1)-lower(1))/range(1), (x(:,2)-lower(2))/range(2)];
+selectors=floor(selectors*64)+1;
+for k=1:2
+    selectors(selectors(:,k)==65,k)=64;
+end
+
+counts=zeros(nbins);
+for k=1:nbins
+    f=find(selectors(:,1)==k);
+    for kk=1:nbins
+        counts(k,kk)=sum(selectors(f,2)==kk);
+    end
+    counts(k,counts(k,:)==0)=1;
+end
+
+U=-log(counts)*mu;
+Unorm=U-min(U(:));
+xbins=linspace(lower(1),upper(1),nbins);
+ybins=linspace(lower(2),upper(2),nbins);
+imagesc(xbins,ybins,Unorm)
+axis equal
+hold on
+plot(-.25+[0 .05],-.6*[1 1],'k-')
+text(-.25+.025, -.6,'5 cm','HorizontalAlignment','center','verticalalignment','top')
+axis off
+
+
+figure(76)
+clf
+hold on
+[xgrid,ygrid]=meshgrid(xbins,ybins);
+mesh(xgrid',ygrid',Unorm,'facealpha',0)
+inds=1:15:size(x,1);
+enNorm=en;
+for k=1:length(en)
+    enNorm(k)=en(k)+Unorm(selectors(k,1),selectors(k,2));
+end
+plot3(x(inds,1),x(inds,2),enNorm(inds),'.','markersize',10)
+zlabel('Total Energy, J')
+ylabel('Robot y, m')
+xlabel('Robot x, m')
+
+figure(77)
+clf
+hold on
+[sun,ord]=sort(Unorm(:));
+for k=1:length(ord)
+        XR=mod(ord(k)-1,nbins)+1;
+        YR=floor(ord(k)/nbins)+1;
[…]

### 2015-08-10 internStuff/attenuation.m

@@ -1,40 +1,0 @@
-clc
-clear all
-close all
-
-subs=[2:8; 23:29]';
-blocksetup=[0 0 1;1 0 2; 1 1 3; 1 0 2; 0 0 1];
-colors='krg';
-markers={'-','-.','--'};
-
-for k=2 %:7
-    figure(k)
-    clf
-    hold on
-    for kk=1:3
-        plot(0,0,colors(kk))
-    end
-    
-    for exnum=[2]
-        S=subs(k,exnum);
-        load(['./Data/output',num2str(S),'.mat'])
-        for T=5:480
-            T
-            B=floor(T/96)+1;
-            if trials(T).targcat~=0
-                if B==5
-                    continue
-                end
-                t=trials(T).t-trials(T).t(1);
-                inds=1:min(350,length(t));
-                shand=sign(trials(T).yrot(inds,2));
-                plot(t(inds),log10(abs(trials(T).xrot(inds,2))./abs(trials(T).yrot(inds,2))),[colors(blocksetup(B,3)),'.'],'markersize',.001);
-            end
-        end
-    end
-end
-%axis equal
-%legend('No forces','Forces','Forces+Intent')
-ylabel('log(F/a)')
-xlabel('time, s')
-

### 2015-08-10 internStuff/christine_stiffness.m

@@ -1,74 +1,0 @@
-clc
-clear all
-
-phases=zeros(96*5,1);
-for k=1:5
-    phases(96*(k-1)+1:96*k)=k;
-end
-
-load './Data/christine2.mat'
-
-global K
-
-set2dGlobals(params.l1, params.l2, params.shoulder,params.mass) %Make sure mass actually in kg
-
-F=find(phases==3);
-
-for k=1:length(F)
-    xvaf=[trials(F(k)).x trials(F(k)).v trials(F(k)).a trials(F(k)).f];
-    qt=xvaf2arm(xvaf);
-
-    x0=xvaf(1,1:2); %orig;
-    x1=trials(F(k)).targ;
-    x=[xvaf(:,1)-x0(1), xvaf(:,2)-x0(2)];
-    M=x1-x0;
-
-    ft=find((trials(F(k)).t-trials(F(k)).t(1))<.5,1,'last');
-    
-    xRhumb=xvaf(1:ft,1:2);
-    qRhumb=xRhumb;
-    xdat(k).t=qt(1:ft,7:8);
-    
-    for kk=1:ft
-        xRhumb(kk,:)=xRhumb(kk,:)-(x(kk,1:2)-(dot(x(kk,:),M)/dot(M,M))*M);
-        qRhumb(kk,:)=ikin(xRhumb(kk,:));
-        
-        %[D,C]=computeDC(qt(kk,1:2),qt(kk,3:4));
-        %xdat(k).t(kk,:)=qt(kk,7:8); %+(D*qt(kk,7:8)'-C)'-;
-    end
-    
-    xdat(k).q=qRhumb(:,1:2)-qt(1:ft,1:2);
-
-end
-
-q=vertcat(xdat.q);
-t=vertcat(xdat.t);
-
-K=q\t
-
-figure(10)
-clf
-hold on
-
-for k=1:length(F)
-    xvaf=[trials(F(k)).x trials(F(k)).v trials(F(k)).a trials(F(k)).f];
-    trials(F(k)).yfixed=extract(trials(F(k)).t,xvaf,@armdynamics_inverted);
-    plot(xvaf(:,1),xvaf(:,2),'k',trials(F(k)).y(:,1),trials(F(k)).y(:,2),'r',trials(F(k)).yfixed(:,1),trials(F(k)).yfixed(:,2),'b')
-end
-axis equal
-
-
-
-figure(1)
-clf
-plot(q(:,1),t(:,1),'.','markersize',1e-6)
[…]

### 2015-08-10 internStuff/frontiersFig1.m

@@ -1,0 +1,99 @@
+clc
+clear all
+
+%% Set constants and initialize
+
+%Measurements in cm. Movements are 15 cm long
+
+[hand,intent]=getColorScheme;
+
+H=20;
+W=18; %Do NOT adjust
+
+yw=15*cosd(30);
+
+leftMargin=60;
+legendThickness=2;
+legendSpacing=7;
+legendp=[-yw 61];
+center=[0 48];
+spacing=[40 30];
+font='Arial';
+fontsize=12;
+interpreter='tex';
+handfudge=0;
+intentfudge=0;
+
+snums=[4 28]; %Exp1=2-9, Exp2=23-30; 4 and 28 are the easiest visually
+
+f=figure(1);
+clf
+set(f,'units','centimeters');
+set(f,'Position',[3,3,18,8])
+
+hold on
+
+tlist=(1:480)';
+b=floor((tlist-1)/96)+1;
+
+%% Draw Y's
+for E=1:2
+    load(['./Data/output',num2str(snums(E)),'.mat'])
+    for T=3:480
+        if trials(T).targcat==0
+            continue
+        end
+        B=b(T);
+        x=100*trials(T).x;
+        y=100*trials(T).y;
+        
+        x=[x(:,1)-center(1)+(B-1)*spacing(1), -x(:,2)+center(2)+(2-E)*spacing(2)];
+        plot(x(:,1),x(:,2),'color',hand)
+        if B==3
+            y=[y(:,1)-center(1)+(B-1)*spacing(1), -y(:,2)+center(2)+(2-E)*spacing(2)];
+            plot(y(:,1),y(:,2),'color',intent)
+        end
+        
+    end
+end
+
+%% Text labels
+
+textInvariant=['\fontname{',font,'} \fontsize{',num2str(fontsize),'} '];
+%Forces
+txt=[textInvariant,'Forces On'];
+text(legendp(1),legendp(2),txt,'HorizontalAlignment','right','VerticalAlignment','middle','interpreter','tex')
+plot([-yw+spacing(1)*1,yw+spacing(1)*3],legendp(2)+[0 0],'k','linewidth',legendThickness)
+
+%Cursor=hand
+txt=[textInvariant,'Cursor = \color[rgb]{',num2str(hand(1)),',',num2str(hand(2)),',',num2str(hand(3)),'}Hand'];
[…]

### 2015-08-10 internStuff/miniplot.m

@@ -1,8 +1,9 @@
-function miniplot(x,y,xoff,marker,msize,rn,hh,hsize)
+function p=miniplot(x,y,xoff,marker,msize,rn,hh,hsize)
 
 plot(xoff+rn,x-y,marker,'markersize',msize)
 
-[~,h]=signrank(x,y);
+%[p,h]=signrank(x,y);
+[h,p]=ttest(x,y);
 if h
     plot(xoff,hh,'k*','markersize',hsize)
 end

### 2015-08-10 matlab/addSubjectPulse.m

@@ -1,6 +1,6 @@
 function addSubjectPulse(k)
 
-subnums=[324 789 300 301];
+subnums=[324 789 300 301 5:8];
 
 name=num2str(subnums(k));
 input=load(['../Data/Data_pulse/input',name,'.dat']);

### 2015-08-10 matlab/cart2model.m

@@ -1,59 +1,49 @@
-function [SNE,OUT,BOUT,oT]=cart2model(X,Y)
+function [TBMP,TSP,TNP]=cart2model(X,Y)
 
 % Assume that forces perpendicular to the direction of movement don't
 % *quickly* effect parallel progress
 % (X is x,y and derivatives measured; Y is y-component desired/typical)
 
-% SNE=sum of non-error-dependent terms
-% Ep=position error
-% Ev=velocity error
-% Tm=muscle torque magnitude vector, needed for Perrault-Franklin style fitting
-% BOUT=[Kp0*(Ep+Ev/12) Kp1*(Tm.*(Ep+Ev/12))]
-% OUT=[Ep Ev Ep.*Tm Ev.*Tm]
-% In theory, Kp_burdet = BOUT\SNE, Kp_burdet = [Kp0 Kp1]
+% TBMP=Body Mass Proportional Torque
+% TSP=Stiffness Proportional Torque
+% TNP=Torque Proportional to Neither of the above
 
 global fJ getAlpha
 
-SNE=zeros(2,size(X,1));
-oT=zeros(6,size(X,1));
-Ep=SNE;
-Ev=SNE;
-Tm=SNE;
+TBMP=zeros(size(X,1),2);
+TSP=zeros(size(X,1),2);
+TNP=zeros(size(X,1),2);
 
 kp0=[10.8 2.83; 2.51 8.67];
 kp1=[3.18 2.15; 2.34 6.18];
 
 for k=1:size(X,1)
+    %Compute real q and tau
     q=ikin(X(k,1:2));
     fJq=fJ(q);
     qdot=fJq\X(k,3:4)';
     qddot=getAlpha(q,qdot,X(k,5:6)');
-    torque=-fJq'*X(k,7:8)'; %This line could scrooge it... with the -sign
+    torque=-fJq'*X(k,7:8)'; %Negative sign is consistent with extract.m
     qreal=[q; qdot; qddot];
     [D_real,C_real]=computeDC(qreal(1:2),qreal(3:4));
 
-    q=ikin([X(k,1) Y(k,1)]);
+    %Compute desired q and tau
+    q=ikin(Y(k,1:2));
     fJq=fJ(q);
-    qdot=fJq\[X(k,3) Y(k,2)]';
[…]

### 2015-08-10 matlab/computeDC.m

@@ -11,8 +11,8 @@ d12=I2+m2*(lc2^2+l1*lc2*c2);
 d21=d12;
 d22=I2+m2*lc2^2;
 
-h=m2*l1*lc2*sin(theta(2));
-C=[h*omega(2)*(2*omega(1)+omega(2)); %This represents a sign change.
+h=m2*l1*lc2*sin(theta(2)); %sign change. should be -? - reflects coordinate difference. + reflects burdet 2006
+C=[h*omega(2)*(2*omega(1)+omega(2));
     h*omega(1)^2];
 
 D=[d11, d12;

### 2015-08-10 matlab/computeDC_opt.m

@@ -1,0 +1,24 @@
+function [D,C]=computeDC_opt(theta, omega,l1,paramvec)
+
+lc1=paramvec(1);
+lc2=paramvec(2);
+m1=paramvec(3);
+m2=paramvec(4);
+I1=paramvec(5);
+I2=paramvec(6);
+
+%Compute alpha to torque relationship, eq. 7.87 in Spong's Robot Control
+%and Modeling: pg 262
+
+c2=cos(theta(2));
+d11=I1+I2+m1*lc1^2+m2*(l1^2+lc2^2+2*l1*lc2*c2);
+d12=I2+m2*(lc2^2+l1*lc2*c2);
+d21=d12;
+d22=I2+m2*lc2^2;
+
+h=-m2*l1*lc2*sin(theta(2)); %sign change. should be -? - reflects coordinate difference. + reflects burdet 2006
+C=[h*omega(2)*(2*omega(1)+omega(2));
+    h*omega(1)^2];
+
+D=[d11, d12;
+    d21, d22];

### 2015-08-10 matlab/deoptImpedanceParams.m

@@ -1,0 +1,83 @@
+function params=deoptImpedanceParams(qd,qr,tau_a,l1,nominal)
+
+% F_VTR		"Value To Reach" (stop when ofunc < F_VTR)
+F_VTR = 0.01; 
+
+% I_D		number of parameters of the objective function 
+I_D = 14; 
+
+% FVr_minbound,FVr_maxbound   vector of lower and bounds of initial population
+%    		the algorithm seems to work especially well if [FVr_minbound,FVr_maxbound] 
+%    		covers the region where the global minimum is expected
+%               *** note: these are no bound constraints!! ***
+      FVr_minbound = .5*nominal; 
+      FVr_maxbound = 1.5*nominal; 
+      I_bnd_constr = 0;  %1: use bounds as bound constraints, 0: no bound constraints      
+            
+% I_NP            number of population members
+		I_NP = 10; 
+
+% I_itermax       maximum number of iterations (generations)
+		I_itermax = 5000; %Was 5000
+       
+% F_weight        DE-stepsize F_weight ex [0, 2]
+		F_weight = 0.85; 
+
+% F_CR            crossover probabililty constant ex [0, 1]
+		F_CR = 1; 
+
+% I_strategy     1 --> DE/rand/1:
+%                      the classical version of DE.
+%                2 --> DE/local-to-best/1:
+%                      a version which has been used by quite a number
+%                      of scientists. Attempts a balance between robustness
+%                      and fast convergence.
+%                3 --> DE/best/1 with jitter:
+%                      taylored for small population sizes and fast convergence.
+%                      Dimensionality should not be too high.
+%                4 --> DE/rand/1 with per-vector-dither:
+%                      Classical DE with dither to become even more robust.
+%                5 --> DE/rand/1 with per-generation-dither:
+%                      Classical DE with dither to become even more robust.
+%                      Choosing F_weight = 0.3 is a good start here.
+%                6 --> DE/rand/1 either-or-algorithm:
+%                      Alternates between differential mutation and three-point-
+%                      recombination.           
+
+		I_strategy = 3;
+
+% I_refresh     intermediate output will be produced after "I_refresh"
+%               iterations. No intermediate output will be produced
+%               if I_refresh is < 1
+      I_refresh = 0;
+
+% I_plotting    Will use plotting if set to 1. Will skip plotting otherwise.
[…]

### 2015-08-10 matlab/extractPulses.m

@@ -1,32 +1,22 @@
 function extractPulses(k)
 
-global kpgain
+global kpgain massgain reflexcontrib
+
+massgain=1;
+ 
+kg=[.18];
+rc=[.07 .1 .1 .1 .1 .1 .1 .1];
+kpgain=kg(k);
+reflexcontrib=rc(k);
 
 figure(k)
 clf
 hold on
 axis equal
 
-scale=.8;
-
-load(['../Data/Data_pulse/pulse',num2str(k),'.mat'])
 load(['../Data/Data_pulse/pulse',num2str(k),'W.mat'])
-params
-%Subject 1
-% params.mass=50;
-% params.shoulder(2)=.48;
-% params.l1=.28;
-% params.l2=.3;
-
-%Subject 2
-kpgain=.16
-
-%Subject 3 and 4
-%kpgain=scale*W(end,1)
 
 set2dGlobals(params.l1,params.l2,params.origin,params.shoulder,params.mass)
-
-offsetForce=onset;
 
 dcats=[trials.disturbcat];
 F=find((dcats>0)&(dcats<5)); %Just like title implies, only the pulses
@@ -35,14 +25,14 @@ for c=1:length(F)
     c/length(F)
 
     kk=F(c);
-    onset=find(vecmag(trials(kk).v)>.05,1,'first');
-    start=max(onset-35,1);
+    start=onsetDetector(trials(kk));
     onset2=find(vecmag(trials(kk+1).v)>.1,1,'first');
     xvaf1=[trials(kk).x trials(kk).v trials(kk).a trials(kk).f];
     xvaf2=[trials(kk+1).x trials(kk+1).v trials(kk+1).a trials(kk+1).f];
     
     t=[trials(kk).t(start:end); trials(kk+1).t(1:onset2)]';
     xvaf=[xvaf1(start:end,:); xvaf2(1:onset2,:)];
+    
     y=extract(t,xvaf,'reflex');
     trials(kk).y=y;
     trials(kk).ty=t;

### 2015-08-10 matlab/figure3merged.m

@@ -1,0 +1,302 @@
+clc
+clear all
+
+figure(37)
+%set(gcf,'position',[8.8268    3.2713   26.8753   20.0790])
+clf
+hold on
+
+%% Fig3 stuff
+
+lw=1.5;
+
+SKIP=9;
+qscale=.002;
+fade=.7;
+
+gray=.5*[1 1 1];
+green=[.1 .15 .7];
+red=[1 .5 .1];
+pink=[1 .8 .8];
+darkpink=.8*[.8 1 .8];
+blue=[.8 .8 1];
+black=[0 0 0];
+
+yoff=[2 4; 1 3]*.1;
+xbase=[0 -.4];
+leftright=[6 8];
+%leftright=[8 4];
+
+for LR=1:2
+    specialestsub=leftright(LR);
+    xoff=[.15 .15;0 0]+xbase(LR);
+    
+    S=specialestsub;
+    load(['../Data/Data_pulse/pulse',num2str(S),'W.mat'])
+    load(['../Data/Data_pulse/pulse',num2str(S),'Y.mat'])
+    %load(['../Data/Data_pulse/pulse',num2str(S),'U.mat'])
+    
+    starts=[trialInfo.startcat];
+    ends=[trialInfo.endcat];
+    dcat=[trials.disturbcat];
+    clean=[trialInfo.clean];
+    
+    f=find(clean);
+    for c=2:length(f)
+        if rand<.9
+            continue
+        end
+        k=f(c);
+        if (trials(k).t(end)-trials(k).t(1))>2
+            continue
+        end
+        X=[trials(k).x(1:SKIP:end,1)-means(starts(k)),trials(k).x(1:SKIP:end,2)-.5];
+        rlength=abs(starts(k)-ends(k));
+        rdir=sign(ends(k)-starts(k));
+        if rdir<0
+            X=-X;
+        end
+        for earlylate=1:2
+            xo=xoff(rlength,earlylate);
+            yo=yoff(rlength,earlylate);
+            plot(X(:,1)+xo,X(:,2)+yo,'color',green,'linewidth',.5)
+        end
+    end
+    
+    for k=2:length(trials)
+        if rand<0 %.7
+            continue
+        end
+        %First decide length class
+        rlength=abs(starts(k)-ends(k));
+        rdir=sign(ends(k)-starts(k));
+        switch dcat(k)
+            case {1 2}
+                earlylate=1;
+            case {3 4}
+                earlylate=2;
+            otherwise
[…]

### 2015-08-10 matlab/finddelay.m

@@ -1,0 +1,53 @@
+clc
+clear all
+
+figure(46)
+clf
+hold on
+
+fit=zeros(10,2,8);
+
+for S=1:8 %why not
+    
+    prefix='pulse';
+    
+    load(['../Data/Data_pulse/',prefix,num2str(S),'.mat'])
+    
+    a=vertcat(trials(3:479).a);
+    f=vertcat(trials(3:479).f);
+    
+    l=size(a,1);
+    
+    subplot(4,2,S)
+    hold on
+    
+    for blah=3:12
+        ind=floor(blah*l/16):floor((blah+1)*l/16);
+        pairs=zeros(31,2);
+        for k=-30:30
+            mdl=fitlm(a(ind,1),f(ind+k,1));
+            pairs(31+k,:)=[k*.005 mdl.Rsquared.Ordinary];
+        end
+        [v,i]=max(pairs(:,2));
+        plot(1000*pairs(:,1),pairs(:,2))
+        plot(1000*pairs(i,1),pairs(i,2),'rx')
+        fit(blah-2,:,S)=[1000*pairs(i,1),pairs(i,2)];
+    end
+    ave=dot(fit(:,1,S),fit(:,2,S))/sum(fit(:,2,S));
+    yl=ylim;
+    plot(ave*[1 1],[0 yl(2)],'m-')
+    xlabel('Force Sensor Lag, ms')
+    ylabel('R^2, regression of acceleration and force')
+    
+end
+
+for S=1:8
+    subplot(4,2,S)
+    if S~=3
+            ylabel('')
+    end
+    title(num2str(S))
+    forcelag(S)=dot(fit(:,1,S),fit(:,2,S))/sum(fit(:,2,S));
+end
+
+save('forcelag.mat','forcelag')

### 2015-08-10 matlab/fitAllParams.m

@@ -1,0 +1,65 @@
+function [fitparams,nominal]=fitAllParams(S)
+
+global fJ getAlpha
+
+load(['../Data/Data_pulse/pulse',num2str(S),'W.mat'])
+
+set2dGlobals(params.l1, params.l2, params.origin, params.shoulder, params.mass)
+
+%% Discard the empties
+ss={storeme.starts};
+notempty=zeros(length(ss),1);
+for k=1:length(ss)
+    notempty(k)=~isempty(storeme(k).starts);
+end
+
+storeme=storeme(notempty==1);
+
+%% Compute q, qd, and tau_ext
+
+subinds=41:80;
+
+for SM=1:length(storeme)
+    X=storeme(SM).X(subinds,:);
+    Y=storeme(SM).Y(subinds,:);
+    storeme(SM).qreal=zeros(6,size(X,1));
+    storeme(SM).qdes=zeros(6,size(X,1));
+    storeme(SM).tau_app=zeros(2,size(X,1));
+
+    for k=1:size(X,1)
+        %Qreal and torque
+        q=ikin(X(k,1:2));
+        fJq=fJ(q);
+        qdot=fJq\X(k,3:4)';
+        qddot=getAlpha(q,qdot,X(k,5:6)');
+        storeme(SM).tau_app(:,k)=-fJq'*X(k,7:8)'; %Negative sign is consistent with extract.m
+        storeme(SM).qreal(:,k)=[q; qdot; qddot];
+
+        %Qdes
+        q=ikin(Y(k,1:2));
+        fJq=fJ(q);
+        qdot=fJq\Y(k,3:4)';
+        qddot=getAlpha(q,qdot,Y(k,5:6)');
+        storeme(SM).qdes(:,k)=[q; qdot; qddot];
+    end
+end
+
+%% Since we're ignoring time shifts, just concatenate
+
+qr=[storeme.qreal];
+qd=[storeme.qdes];
+tau_a=[storeme.tau_app];
+
+%% Plug everything into an optimizer
+l1=params.l1;
+l2=params.l2;
+mass=params.mass;
+m1=.028*mass;
+m2=.022*mass;
+
+kp0=[10.8 2.83; 2.51 8.67];
+kp1=[3.18 2.15; 2.34 6.18];
+
+nominal=[.436*l1,.682*l2,m1,m2,m1*(.322*l1)^2,m2*(.468*l2)^2,kp0(:)',kp1(:)'];
[…]

### 2015-08-10 matlab/fullWorkFlow.m

@@ -1,7 +1,10 @@
 clc
 clear all
 
-n=1:4;
+n=1:8;
+return
+%Add differential optimization just in case it's needed
+addpath ../../DeOpt/
 
 % workflow={'','addSubjectPulse';
 %     'W','findKpGain';
@@ -11,13 +14,14 @@ n=1:4;
 %     'NEVERMATCH','finalfig4'
 %     };
 
-workflow={'','addSubjectPulse';
-    'W','findKpGain';
-    'Y','extractPulses';
-    'U','extractUndisturbed';
+% workflow={'','addSubjectPulse';
+%     'W','findWhiteKpGain';
+%     'Y','extractPulses'
+%     };
+
+workflow={'W','findWhiteKpGain';
+    'Y','extractPulses'
     };
-
-
 
 mfiles=dir('.');
 datafiles=dir('../Data/Data_pulse/');

### 2015-08-10 matlab/getSubjectParams.m

@@ -1,50 +1,59 @@
 function params=getSubjectParams(num)
 
 switch num
-    case {'1', '500', '21', '31','32','55','729'} %me
-        l1=.33;
-        l2=.34;
-        mass=175;
-        shoulder=[0 .45];
-    case {'300'} %Yazan
+    case '300'
         l1=.31;
         l2=.34;
         mass=210;
         shoulder=[0 .51];
-    case {'301'} %Guy
+        origin=[-0.0095, 0.4250];
+    case '301'
         l1=.33;
         l2=.345;
         mass=183;
         shoulder=[0 .49];
-    case {'3','9','10'} %Jim
-        l1=.36;
-        l2=.31;
-        mass=225;
-        shoulder=[0 .55];
-    case {'11','12'} %Tes
-        l1=.3;
-        l2=.32;
-        mass=213;
-        shoulder=[0 .52];
-    case {'324'}
+        origin=[-0.0095, 0.4250];
+    case '324'
         l1=.28;
         l2=.32;
         shoulder=[0 .53];
         mass=213;
-    case {'789'}
+        origin=[-0.0095, 0.4250];
+    case '789'
         l1=.28;
         l2=.3;
         shoulder=[-.03 .48];
         mass=120;
+        origin=[-0.0095, 0.4250];
+    case '5'
+        l1=.34;
+        l2=.34;
+        shoulder=[-0.02, .55];
+        mass=208;
+        origin=[0, .44];
+    case '6'
+        l1=.37;
+        l2=.38;
+        shoulder=[-0.03, .51];
+        mass=190;
+        origin=[0, .44];
+    case '7'
+        l1=.28;
+        l2=.31;
+        shoulder=[-.05, .48];
+        mass=140;
+        origin=[0, .44];
+    case '8'
+        l1=.29;
+        l2=.33;
+        shoulder=[0, .5];
+        mass=160;
+        origin=[0, .44];
 end
 
 params.l1=l1;
 params.l2=l2;
 params.shoulder=shoulder;
-params.origin=[-0.0095, 0.4250];
+params.origin=origin;
 
-if exist('mass','var')
-    params.mass=0.453592*mass;
-else
-    params.mass=((l1+l2)/.77)*91; %Shitty locally linear scaling of weight with height, remember that 200 lbs = 91 kg.
-end
+params.mass=0.453592*mass; %Kg vs lbs

### 2015-08-10 matlab/objImpedanceParams.m

@@ -1,0 +1,51 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Function:         S_MSE= objfun(FVr_temp, S_struct)
+% Author:           Rainer Storn
+% Description:      Implements the cost function to be minimized.
+% Parameters:       FVr_temp     (I)    Paramter vector
+%                   S_Struct     (I)    Contains a variety of parameters.
+%                                       For details see Rundeopt.m
+% Return value:     S_MSE.I_nc   (O)    Number of constraints
+%                   S_MSE.FVr_ca (O)    Constraint values. 0 means the constraints
+%                                       are met. Values > 0 measure the distance
+%                                       to a particular constraint.
+%                   S_MSE.I_no   (O)    Number of objectives.
+%                   S_MSE.FVr_oa (O)    Objective function values.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+function S_MSE= objImpedanceParams(FVr_temp, S_struct)
+
+%---Check the passband first------------------------------------
+
+qr = S_struct.qr;
+qd = S_struct.qd;
+tau_a = S_struct.tau_a;
+l1 = S_struct.l1;
+
+mpvec = FVr_temp(1:6);
+kp0 = reshape(FVr_temp(7:10),2,2);
+kp1 = reshape(FVr_temp(11:14),2,2);
+
+tau_pred=zeros(size(tau_a));
+
+for k=1:size(tau_a,1)
+    [Dr,Cr]=computeDC_opt(qr(1:2,k),qr(3:4,k),l1,mpvec);
+    [Dd,Cd]=computeDC_opt(qd(1:2,k),qd(3:4,k),l1,mpvec);
+
+    Tff=Dd*qd(5:6,k)+Cd;
+    Tin=Dr*qr(5:6,k)+Cr;
+    Tmuscle=abs(Tin+tau_a(:,k));
+    Ep=qr(1:2,k)-qd(1:2,k);
+    Ev=qr(3:4,k)-qd(3:4,k);
+    
+    tau_pred(:,k)=Tff-Tin-(kp0+kp1*diag(Tmuscle))*(Ep+Ev/12);
+end
+
+F_cost  = sum(sum((tau_a-tau_pred).^2));
+
+%   
[…]

### 2015-08-10 matlab/onsetDetector.m

@@ -1,0 +1,4 @@
+function start=onsetDetector(trial)
+
+onset=find(vecmag(trial.v)>.025,1,'first');
+start=max(onset-35,1);

### 2015-08-10 tsDescribe/fitTs.m

@@ -1,0 +1,78 @@
+function fitTs(SUB)
+
+load(['../Data/Data_pulse/pulse',num2str(SUB),'.mat'])
+
+%% Categorize by start/end pair
+clustermeS=zeros(length(trials)-1,1); %#ok<NODEF>
+clustermeE=clustermeS;
+for t=1:length(trials)-1
+    clustermeS(t)=trials(t+1).x(1,1);
+    clustermeE(t)=trials(t+1).x(end,1);
+end
+[trash,means]=kmeans([clustermeE; clustermeE],3,'emptyaction','singleton','start',[-0.175;-0.026;0.125]);
+means=sort(means);
+
+starts=zeros(length(trials),1);
+ends=starts;
+for t=1:length(trials)
+    [trash,starts(t)]=min(abs(trials(t).x(1,1)-means));
+    [trash,ends(t)]=min(abs(trials(t).x(end,1)-means));
+    trialInfo(t).startcat=starts(t);
+    trialInfo(t).endcat=ends(t);
+    trialInfo(t).x0=means(starts(t));
+    trialInfo(t).xf=means(ends(t));
+end
+starts(1)=-inf; %never use this reach...
+trialInfo(1).startcat=-inf;
+
+reachcat=10*starts+ends;
+urc=unique(reachcat);
+urc=urc(2:end);
+
+%% Use undisturbed examples only
+dcats=[trials.disturbcat];
+
+nclean=4;
+clean=0*dcats;
+for kk=1:nclean
+    clean=clean+[zeros(1,kk-1) dcats(1:end-(kk-1))];
+end
+clean=(clean==0);
+
+for t=1:length(trials)
+    trialInfo(t).clean=clean(t);
+    trialInfo(t).reachcat=find(reachcat(t)==urc);
+end
+
+%% For each trial, fit a 5th order poly to velocity
+
+tesses=-1*ones(size(trials));
+
+for t=2:length(trials)
+    if trialInfo(t).clean
+        tesses(t)=getTsMetric(trials(t).x,trials(t).t,[means(starts(t)) .5],[means(ends(t)) .5]);
+    end
+end
+
+figure(SUB)
+clf
+subplot(1,2,1)
+hold on
+tessesfixed=abs(tesses(tesses~=-1));
[…]

### 2015-08-10 RealtimeFeedbackPaper/realtimefeedback.tex.BASE.22759.tex

--- 2015-03-13 RealtimeFeedbackPaper/realtimefeedback.tex
@@
 Next, the matrix $\hat{M}(q_d)$ must be invertible, but we can ensure this through our choice of workspace.
-Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{gomi1998task} and may vary over the course of a reach\cite{niu2010temporal}.
+Finally, feedback torque requires a model of arm impedance, which is known to be task-dependent\cite{} and to vary over the course of a reach\cite{shapiro}.
 With no prior knowledge of arm impedance for this task-disturbance combination, we presumed the feedback torque model of Shadmehr and Mussa-Ivaldi anticipating that it is sufficiently accurate or easy to learn (experiment 1).
@@
 While our choice of stiffness model is clearly inaccurate, we had good reason to believe that our subjects would readily adapt to a constant stiffness, but this does not seems to be the case.
-Moreover, difficulties acquiring and holding the target during intent feedback support claims that different mechanisms are at work during that stage of movement\cite{niu2010temporal}.
-Improving modeling assumptions address limb mechanics, force learning, and impedance can be improved in future work.
+Moreover, difficulties acquiring and holding the target support claims that different mechanisms are at work during that stage of movement\cite{niu2010temporal}.
 
-Future directions include extending this modeling approach to more complex plants such as motor vehicles, touch screen devices, and reaching movements in three dimensions.
-This may potentially improve human-robot interaction by giving roboticists a real-time estimate of human movement intent despite disturbances and incongruous interaction on the part of the robot.
-In order to extend this work to situations in which learning occurs, a learning model will be necessary or additional assumptions regarding the form and production of intent will be required.
-Intent estimation may be useful tool for revealing previously unknown features of planning and adaptation.
-
-\begin{figure}[thpb]
-      \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Typical subjects' center-out targeted reaching motions under experimentally varied force and feedback conditions.
-Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
-While both were recorded under all conditions, only one was shown on screen.
-Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
-The force disturbance was designed to be turbulent and unpredictable to minimize any effect of learning.}
-      \label{typicalsubject}
-   \end{figure}
-
-   \begin{figure}[thpb]
-      \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback.
-Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement reveals that turbulent force disturbance degrade reaching performance.
-}
-      \label{figurelabel}
-   \end{figure}
-
-   \begin{figure}[thpb]
-      \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Inductance of oscillation winding on amorphous
-       magnetic core versus DC bias magnetic field}
-      \label{figurelabel}
-   \end{figure}
+Future directions include
 
 \section*{ACKNOWLEDGMENT}
+
+%\begin{thebibliography}{99}
 \bibliographystyle{IEEEtran}
@@
 
+%\end{thebibliography}
+
 \end{document}

### 2015-08-10 RealtimeFeedbackPaper/realtimefeedback.tex.REMOTE.22759.tex

--- 2015-01-27 RealtimeFeedbackPaper/realtimefeedback.tex
@@
 We found that reaching performance during turbulence was significantly better using the estimated intent.
-Interestingly, when viewing intent, the subjects' estimated arm stiffness became lower, indicating a more relaxed state of operation when viewing their estimated intent.
+Interestingly, when viewing intent, the subjects' estimated arm stiffness became lower, indicating a more relaxed state of operation.
 In the future, this method could be abstracted and applied to situations where the human arm is disturbed, such as in vehicle turbulence, tremor, and other noisy human-machine interactions.
@@
 Often the operator has new information (such as a change in priorities or plan) that may still require high-level piloting commands.
-Instead of taking the operator out of the control loop, we recently developed an algorithm \cite{horowitz2015determining} that can determine the operator's underlying intent in spite of disturbances.
+Instead of taking the operators (and their intelligence) out of the control loop, there are numerous reasons to keep them in.
+We recently developed such an algorithm \cite{horowitz2015determining} that can determine the operator's underlying \textit{intent} in spite of disturbances.
 One question might be if the operator could see their own instantaneous intent, would they be able to recognize, appreciate, and make use of it?
@@
 Gomi and Kawato\cite{gomi1997human} performed these measurements and reported highly complex and often not anatomically realizable trajectories that could not well-represent the intent of a simple reaching movement.
+More recently our approach \cite{horowitz2015determining} builds on this method to include the important component of internal models in feedforward planning for reaching actions.
 
 Since we lack a fully accepted model, we turn to reaching accuracy and stiffness to test whether or not our deduced intended trajectory is reasonable.
-First, if this is indeed an estimate of intent, subjects should be able to perform targeted reaching using it even in the presence of a force-based disturbance.
-Second, we should expect that during an unpredictable disturbance, such as turbulent forces, the intended trajectory will deviate less on average than the hand path during the pre-planned, early phase of movement.
+First, if this is indeed an estimate of intent, then a feedback that displays intent rather than the actual endpoint location as the cursor \textit{intent feedback} should lead to better performance.
+This should especially be true in the presence of force-based disturbances.
+Accordingly, in this paper we hypothesized that during random turbulent forces, the intended trajectory should deviate less than the hand.
 Finally, arm stiffness is known to increase during exposure to instability\cite{franklin2003adaptation} and uncertainty\cite{takahashi2001impedance}.
-If subjects reach in an uncertain and unstable force environment, we should expect their arm stiffness to increase from baseline.
-If subjects are then asked to reach using visual feedback of their intent instead of their hand, even if our estimate is not perfect, the decrease in uncertainty and instability should lead to a decrease in stiffness.
+We also hypothesized that any elevated arm stiffness from turbulence would decrease if subjects were given intent feedback.
 
@@
 \subsection*{Dynamic simulation of arm and intended trajectories}
-Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia.
+Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} were used to estimate relationships between body mass and limb mass, limb length and limb center of mass, and limb mass and length and moment of inertia.
 Viscosity parameters were taken from Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive}.
@@
 \end{equation}
-Force ($F_e$) and state ($\ddot{e}$, $\dot{e}$, $e$) are known.
+Force ($F_e$) and state ($\ddot{e}$, $\dot{e}$, $e$) were known.
 Mass ($m$), viscosity ($b$), and stiffness offset ($F_0$) terms were calculated, but discarded.
@@
 
-\begin{figure}[h]
+\begin{figure}[!b]
 \centering
-	\includegraphics{error.eps}      
-	\caption{This is a figure caption.}
+\framebox{\parbox{3in}{\includegraphics{Yplots.eps}
+}}
+\caption{Typical subjects' made center-out targeted reaching motions under experimentally varied force and feedback conditions.
+Subjects used feedback of either hand motions (black lines) or estimated intent (red lines) to complete these reaches.
+While both were recorded under all conditions, only one was shown on screen at a time.
+Intent was estimated using either the stiffness model of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} (Experiment 1) or a reduced stiffness (Experiment 2) to explore any dependence of reaching stiffness or accuracy on this assumption.
+The force disturbance was designed to be turbulent and unpredictable to minimize any effect of learning.}
+\label{Yplots}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\framebox{\parbox{3in}{\includegraphics{error.eps}
+}}
+\caption{Subjects' reaching accuracy depended on the presence of force disturbance and the contents of visual feedback (A).
+Maximum deviation from straight-line reaching calculated during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance degraded reaching performance in both hand (B) and intent (C).
+Comparisons within treatment conditions revealed a significant advantage for the hand when force disturbance is not present and a significant advantage for intent when feedback of intent was used during exposure to turbulent forces.
+Performance using intent during turbulence was significantly better than performance using the hand (D).
+Significance among comparisons is denoted by an asterisk.
+Error did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
 \label{error}
@@
 \subsection*{Reaching Accuracy}
-Even if subjects are capable of performing a reaching task using their estimated intent, a poor estimate should lead to poor performance.
-Indeed, the hand significantly outperforms our intent estimate when the force disturbance is not present in block 1 (Figure 2B).
-Model and sensor inaccuracies degrade performance and result in an error of approximately 5 millimeters.
+Even if subjects were capable of performing a reaching task using their estimated intent, a poor estimate would have lead to poor performance.
+Indeed, the hand significantly outperformed our intent estimate when the force disturbance is not present in block 1 (Figure 2D).
+Model and sensor inaccuracies degrade performance and result in an error of approximately 5 millimetres.
 In the presence of a force disturbance in block 2, the hand loses its advantage as model and sensor inaccuracies are balanced by some degree of cancellation of the disturbance.
@@
 For this intent estimate to be useful, subjects should perform better when using feedback of the intent than when using feedback of the hand.
-Comparing the hand's performance in block 2 to the intent's performance in block 3, subjects perform significantly better using their estimated intent.
+Comparing the hand's performance in block 2 to the intent's performance in block 3, subjects performed significantly better using their estimated intent.
 Finally, no significant differences were present across the two experiments.
-Performance appears relatively insensitive to this change in the stiffness model used to deduce the intent.
+Performance was relatively insensitive to this change in the stiffness model used to deduce the intent.
 
 \subsection*{Endpoint Stiffness}
-While accuracy is desirable, it can be attained by cocontracting the muscles to increase arm stiffness at the cost of metabolic effort \cite{gribble2003role}\cite{takahashi2001impedance}.
-Subjects stiffen significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3B).
-Subjects' stiffness significantly decreases when feedback of the deduced intent replaces feedback of the hand as seen by comparing blocks 2 and 3.
-This decrease does not return stiffness to undisturbed levels and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
-Stiffness may decrease with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
-Subjects do not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
+While accuracy is desirable, it can be attained by cocontracting muscles to increase arm stiffness at the cost of metabolic effort \cite{gribble2003role}\cite{takahashi2001impedance}.
+Subjects stiffened significantly in response to white noise forces as seen by comparing stiffness in blocks 1 and 2 (Figure 3C).
+Subjects' stiffness significantly decreased when feedback of the deduced intent replaced feedback of the hand, as seen by comparing blocks 2 and 3.
+This decrease does not return stiffness to undisturbed levels, and it remains even after feedback of the hand resumes as seen by comparing blocks 3 and 4.
+Stiffness may have decreased with duration of exposure to uncertain disturbances\cite{takahashi2001impedance}.
+Subjects did not appear to have adjusted their stiffness to match the stiffness used in the estimation as subject and model stiffnesses were significantly different in experiment 2.
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
-	\includegraphics{stiffness.eps}      
-	\caption{This is a figure caption.}
+\framebox{\parbox{3in}{\includegraphics{stiffness.eps}
+}}
+\caption{Subjects' arm stiffness depended on the presence of force disturbance and the contents of visual feedback (A).
+Arm stiffness calculated by linear regression during the first 250 milliseconds after the onset of movement revealed that turbulent force disturbance increased arm stiffness (B).
+Comparisons between treatment conditions revealed that exposure to turbulent forces caused by arm to stiffen significantly, but feedback of intent could significantly reduce this increase (C).
+Stiffness did not signficantly depend on the choice between Shadmehr and Mussa-Ivaldi's\cite{shadmehr1994adaptive} stiffness model (experiment 1, blue) or a reduced stiffness (experiment 2, red) to determine intent.}
 \label{stiffness}
@@
 \section{Discussion}
-Exposure to turbulent forces damaged subjects' reaching accuracy and increased their arm stiffness, but replacing feedback of a subject's hand with feedback of their estimated intent alleviated accuracy and stiffness.
+This study experimentally evaluated an alternative to turbulence-cancellation, which estimated and displayed the human operator's underlying intent in real-time.
+The evidence here showed that subjects exploited this new form of feedback.
+Although exposure to turbulent forces damaged subjects' reaching accuracy and increased their arm stiffness, when we replaced the veridical feedback with estimated intent alleviated accuracy and stiffness.
 In the absence of disturbing forces, estimated intent accuracy was lower than that of the hand, which suggests the need for model improvements.
@@
 Moreover, difficulties acquiring and holding at the target during intent feedback support claims that different mechanisms are at work during that stage of movement\cite{niu2010temporal}.
-Improving modeling assumptions to better estimate limb mechanics, force learning, and impedance may improve our estimate in future work.
+Improving modeling assumptions to better estimate limb mechanics, force learning, and impedance may improve our estimates and hence operator performance.
 
